### 2019-01-13 Write_up/Writing/Introduction_outline.Rmd
Introduction outline

---
title: "Introduction_outline"
author: "Amieroh Abrahams"
date: "13 January 2019"
output: html_document
---

### Introduction

Seawater temperature is a key indicator of environmental change in both terrestrial and marine ecosystems (Pearce et al., 2006), which directly influences the earth’s natural ecological functioning. Climate change is widely believed to result in a rise in the earth’s average surface temperatures over time (Stocker et al., 2013) and this is usually associated with an increase in the intensity and duration of extreme weather events such as cold-spells, heatwaves, floods and storms. Consequently, these extreme climatic events can often have profound short-term effects on both humans as well as many ecosystems (Easterling et al., 2000).

‘Cold spells’ are usually referred to as a phenomenon where abnormally cold water remains on the surface of the ocean for extended periods of time. However, several other definitions exist based on the statistical properties of the temperature record in association with the location of temperature collections and the time of year .  Studies that have quantified the impacts of extreme temperatures have provided evidence of demonstrably negative effects on  associated surrounding regions as well as the specific regions within which these climatic events occur. While extensive research has been done on other extreme weather events such as floods, wind storms and marine heat waves, minimal research has been done on marine cold spells (MCS). It is known that warm-water anomalies may result in severe coral bleaching and is one of the factors resulting in the demise of coral reefs. Additionally, anomalous warm water events are known to have negative impacts on species diversity and mortality as well as the distribution of organisms. However, recent studies done by Lirman et al. explains that corals living in regions or surrounding regions of unusually prolonged anonymously cold waters may be just as detrimental; this was illustrated in January of 2010 when the reefs of the Florida Reef Track (FRT) were exposed to an unusual cold water event that resulted in large scale coral mortality. This climatic event showed clear association with North Atlantic Oscillation (NAO) index that produced northerly surface wind anomalies and the southward advection of the cold Arctic air. Similarly, research by Shinn et al. (1962) reported high mortality of Acropora cervicornis in shallow reefs as temperatures dropped. Likewise, studies conducted by Hudson et al.  also documented the mortality of Montastraea colonies in near shore patch reefs due to cold fronts between 1969–1970 and 1977–1978, respectively.

MCSs are lethal to all marine organisms and are known to have caused  mass extinctions of several fish and invertebrate species (Woodward. 1987 & Gunter. 1941, 1951 & Holt and Holt. 1983, Gunter. 1951 & Crisp. 1964). Temperature has long been considered one of the primary factors controlling species distribution patterns by limiting their distribution north or southward towards the high latitudes (Firth et al., 2011). Studies done by Firth et al. (2011, 2015) have demonstrated the range shifts of engineer species such as mussels as a result of MCSs. Thus, providing a clearer understanding of the influence of oceanic and atmospheric processes on sea surface temperatures (SST) within the upper mixed layer, however minimal research has been done on the influences of local and broad scale impacts on seawater temperature.

Research by Grunter (1941) and Firth et al. (2011) explains that some MCSs that are believed to have influenced species populations were caused by atmospheric cold spells which had ultimately affected the local intertidal and coastal biota. MCSs have been proven to have a great effect on ecosystem change, thus understanding these climatic drivers may provide useful tools for conservation and management efforts. This study will aim to develop a better understanding of the impacts of weak and strong upwelling events and how they may be influencing the intensity and duration of MCSs. This study applied the MCSs definitions to datasets at a local-scale and an offshore broad scale.

Eastern Boundary Upwelling Systems (EBUS) are among the most productive regions of the world. These upwelling systems include the; California, Humboldt, Canary and Benguela currents, and are important for biological productivities by supporting commercial fisheries (Pauly and Christensen, 1995) thus forming an essential component to these region’s economies, as well as providing recreational services to the human population along their coasts. One of the five continental margin upwelling systems is the Benguela Current; an eastern boundary current of the south Atlantic subtropical gyre which flows towards the equator approximately 300 km offshore of the coast of southwest Africa before splitting into the Benguela Coastal Current (BCC) and Benguela Oceanic Current (BOC). Shoreward of the main stream of the Benguela current there is a slow flow along the continental margin into the Angola Basin. In this slow flow the prevailing south and south eastern winds drive coastal upwelling of cold, nutrient rich South Atlantic Central Water (SACW).

The SACW originates by sinking at the Subtropical Front, known as the Subtropical Convergence (Schell, 1970 & Nelson, 1989). Beneath the SACW lies cold, nutrient-rich Antarctic Intermediate Water (AAIW) which originates by sinking at the Polar Front. Throughout each of the seasons, upwelling of nutrient rich waters covers most of the continental shelf and much of the continental slope (Shannon, 1985), thus creating what we refer to as the Benguela Current Upwelling System. Research studies have shown that these cold waters move north over the shelf in the Benguela Coastal Current (Hay and Brock, 1992). Offshore they are separated from the warmer waters of the subtropical Atlantic by a well-developed thermal front (Shannon, 1985). This quasistationary region is often linked with the shelf edge, although it is distorted by eddies, plumes and jets which extend west into the SE Atlantic (Lutmharms and Stockton, 1987; Waldron and Probyn, 1992). These plumes largely expand the region of high productivity as a result of upwelling by rapidly advecting cold, nutrient-rich water from the coastal upwelling sites into the open ocean (Lutjeiarms et al, 1991). Northward, the Benguela Current Upwelling system is linked by the Angola Benguela Front which converge between the Benguela current and the Benguela coastal current as well as the warm saline waters of the Angola Current (Peterson and Stramma, 1991). Southward, the coastal upwelling system is bounded by the Agulhas Bank. South of the African coast the warm Agulhas Current flows in from the Indian Ocean and returns eastwards while generating warm eddies that spin off in a north-westerly direction in the Benguela current (Duncombe, 1991). Research studies done by Shannon et al. (1985) explains that coastal upwelling varies on a seasonal basis and that during winter and spring, water of approximately 16°C extends along the coast. Upwelling is coldest and most persistent off the Namibian coast in the Lüderitz region at 25°S and weakest at 17°S (Lutjeharms and Meeuwis, 1987). Lutjeharms and Meeuwis, (1987) also found a strong relationship between the intensity of upwelling events and the strength and direction of prevailing winds.  It is also hypothesised that upwelling is more intense where deep water is closer to the coast.

EBUS mesoscale and sub-mesoscale oceanic movements are known to influence marine ecosystems in that it enhances primary production (McGillicuddy et al., 1998; Oschlies and Garçon, 1998) and is key in managing the exchange and of transporting of organic resources and marine organisms from the coastal zone to the open ocean. Large-scale atmospheric pressure systems that favour along-shore, equatorward winds, result in the high levels of productivity in the EBUS, which in combination with the Coriolis effect, affects surface water offshore. This results in a process whereby cold, nutrient rich, with low pH and low oxygen concentration, originates from deep within the ocean and is pushed to the surface where they, in the form of phytoplankton blooms, fuel primary production. As a result of the intensity and duration of the upwelling event; regions affected by these cold water events are exposed to distinct ecological conditions and tend to differ from its surrounding regions (Bosman et al., 1987; Barshis et al., 2011). Significant research evidence suggests that SST have increased, however some regions still show a lower or non-significant increase, and in some cases cooling trends are evident. Inshore regions and embayment’s that are protected from wind action may be subjected to intense episodes of primary production (Graham and Largier, 1997). Vander Wouder et al. (2006) found that these sites represent the highest levels of chlorophyll concentration.

Atmospheric and oceanographic processes results in upwelling, while the amplitude and timing of upwelling-favourable winds are sensitive to climate variability (Montecinos et al., 2003; Macias et al., 2012), and are highly likely to be affected by global warming (Barros et al., 2014; Bakun et al., 2015). It is hypothesised that global warming will influence atmospheric pressure gradients, ultimately affecting the amplitude and timing of coastal winds that cause upwelling (Bakun, 1990; Bakun et al., 2015) and more recent observations (Rykaczewski et al., 2015; Wanget al., 2015) by means of global model projections strongly suggests intensification of coastal winds. Understanding the influence of changing winds on upwelling events is highly important. For example, weak winds may result in weak upwelling, which in turn may limit nutrient concentration; resulting in potentially negative impacts on primary production (Chhak and Di Lorenzo, 2007). In contrast, stronger upwelling winds, could theoretically result in stronger upwelling events, thereby increasing the nutrient concentration which may increase offshore transport (Cury and Roy, 1989). Increased wind action may also result in increased turbulence (Cury and Roy, 1989); and may even affect the processes associated with ocean acidification, creating widespread changes within ecosystems and ultimately impacting regional productivity (Gruber, 2011; Doney et al. 2012).

**** A better understanding of the impact of global warming on wind variability is important in that it allows for better management of fisheries and other marine resources in the EBUS which could be improved with a better understanding of how nutrient abundance and primary production are associated with upwelling events. This study aims to emphasize the drivers of upwelling-favourable winds, as well as the factors influencing nutrient enrichment, and potential changes in biogeochemistry*****

The Benguela current, forming part of the EBUS, experiences cross-shore atmospheric pressure gradients, which results in alongshore, equatorial winds. These winds are known to drive coastal upwelling events. With this in mind, research by Bakun (1990) suggests that an increase in greenhouse gas emissions would exacerbate warm-season upwelling. The understanding behind this mechanism was based on the idea that air above land would warm rapidly as compared to air above the ocean, and that the deepening of the continental thermal low pressure system (CTLPS) would increase the cross-shore pressure gradient which drives upwelling-favourable winds. Several research studies examining the trend between upwelling events and wind patterns, have however indicated contrasting results. With more than 20 years of observational and modelled data; Sydeman et al. (2014a) discovered the most coherent pattern of intensifying upwelling-favourable winds in the Humboldt and California upwelling systems. It has proven difficult to attribute coastal wind intensification in EBUS to global warming, despite the fact that the meta-analysis done by Sydeman et al. (2014a) generally supports Bakun’s (1990) proposition.

This study cannot discount the role of multi-decadal climate variability in the observed trends (Chhak and Di Lorenzo, 2007; Narayan et al., 2010; Pérez et al., 2010; Macias et al., 2012; Santos et al., 2012; Cropper et al., 2014; Jacox et al., 2014), nor is it a test of the Bakun (1990) hypothesis mechanism. Given the broad range of contradicting studies, the only conclusion that can be made is that upwelling-favourable winds tend to be the most intense during the warmer months of the year (Huyer, 1983; Nelson and Hutchings, 1983). 
The timing and the intensity of upwelling favourable winds are highly variable within the EBUS, therefore creating much uncertainty (Chavez and Messié, 2009). Understanding the factors contributing to the intensity and duration of upwelling events is important because variations in the timing of upwelling strongly influences ecosystem productivity. Since the late 1940s studies have shown an increase in upwelling favourable winds associated with the California current, however, less variability has been observed in the strength of upwelling winds through the year in the central Benguela and Canary Currents (Shannon and Nelson, 1996; Varela et al., 2015). It is important to note that the lack of research conducted may obscure the validity of any conclusion drawn here. This study further researched the changes in the timing and variability of upwelling within the Benguela Current and how these phenomena influence associated ecosystems.

As a result of climate change, global ocean temperatures have increased at an accelerating rate (Stocker et al., 2013). Sufficient research suggests that coastal sea surface temperatures (SST) in EBUS such as the Benguela current are cooler than offshore regions as a result of the near shore upwelling waters. There is a large but contradicting amount of evidence on the SST trends in EBUS (Mendelssohn and Schwing, 2002; Arístegui et al., 2009; Belkin, 2009; Demarcq, 2009; Lebassi et al., 2009; García-Reyes and Largier, 2010; Rouault et al., 2010; Gutiérrez et al., 2011; Seo et al., 2012; Barton et al., 2013; Barros et al., 2014; Sydeman et al., 2014b; Salvanes et al., 2015). The contradictions amongst studies exist because of varying resolutions of SST data used, thus resulting in the capturing of different aspects of the upwelling processes, showing different trends that exist based on the resolutions of the different SST datasets as well as other intrinsic factors. Generally, because of the resolution of SST datasets it is often difficult to separate near shore upwelling temperatures from regional, broad scale and eastern boundary current temperatures. Lima and Wethey’s (2012) found that high resolution data represented clear cooling trends predominantly during summer and at high latitudes in most of the EBUSs. Additionally several research studies examined the difference between coastal and offshore temperature trends, revealing decreasing or increasing trends in near shore SST, compared with increasing trends offshore in the Benguela compared with increasing trends offshore in the Benguela (Rouault et al.,2010; Santos et al., 2012; Salvanes et al., 2015). In this study however, will utilise a range of spatio-temporal resolutions of SST data and in situ collected seawater temperature measurements for the Benguela upwelling system. The aim of this is to demonstrate the variable upwelling events based on the different resolutions at which temperature is being collected.

Generally, regional ocean climate oscillations like the Benguela Niño for the Benguela Current System impose considerable variability on temperatures and complicate attempts to assess long-term trends (Demarcq,2009). It is predicted that these climate oscillations may increase the variance with climate change (Timmermann et al., 1999; Kuzmina et al., 2005; Sydeman et al., 2013; Cai et al., 2015) which will further result in unusual trends.

Research studies provide significant evidence on the increase in warming of the ocean surface. This warming slowly propagates down through the water column and in turn increases the thermal stratification (Gruber, 2011). Strong stratification limits the depth at which water is up welled and hence controls the amount of nutrients brought up (Chhak and Di Lorenzo, 2007; Jacox and Edwards, 2011; Jacox et al., 2015). The interplay between upwelling winds and stratification determine the amount of nutrients present within the system and thus is strongly linked to biological productivity. Globally; as a result of climate change, stratification of the water-column between 0 and 200 m has increased by an estimated 4% during the period 1971 to 2010 (Stocker et al., 2013). Variability in the biogeochemical properties of up welled waters influences species distribution, ecosystem structure and productivity within the Benguela upwelling system. The deep waters being pushed to the surface through the upwelling process is high in nutrients and dissolved CO2, a consequence of the long period of isolation at a depth from the euphotic zone (Feely et al., 2008). Increased ocean stratification as a result of surface warming may reduce the ventilation of deep-water masses. As the deep water is being up welled to the surface, decreases in oxygen and pH levels are predicted along with an increase in nutrient levels (Rykaczewski and Dunne, 2010).

Biotic systems within the Benguela upwelling system may be negatively influenced by the changes in oxygen levels given their adaption to natural variable conditions (Capone and Hutchins, 2013).  This study will seek to examine the impact of both strong and weak upwelling winds on the intensity and duration of the upwelling event; additionally it will also aim to examine how chlorophyll concentrations vary with regard to contrasting upwelling conditions as well as the impact this has on species distribution and habitats.

##### Ideas #####

-chnaging the structure it to describe MCS, explain the dangers of these phenomena on marine life (, the implications of this (eg distributions being affected), and the causes of MCS.

-explain what info is unknown about MCS and why this is a problem in areas where they can occur

- link this to the EBUS and Benguela stuff

- clearly mention the aims, objectives and approach to the study.

### 2019-01-13 Write_up/Writing/Introduction_outline.Rmd
edited

@@
 Seawater temperature is a key indicator of environmental change in both terrestrial and marine ecosystems (Pearce et al., 2006), which directly influences the earth’s natural ecological functioning.
-Climate change is widely believed to result in a rise in the earth’s average surface temperatures over time (Stocker et al., 2013) and this is usually associated with an increase in the intensity and duration of extreme weather events such as cold-spells, heatwaves, floods and storms.
+Climate change is widely believed to result in a rise in the earth’s average surface temperatures over time (Stocker et al., 2013) and this is usually associated with an increase in the intensity and duration of extreme events such as cold-spells, heatwaves, floods and storms.
 Consequently, these extreme climatic events can often have profound short-term effects on both humans as well as many ecosystems (Easterling et al., 2000).

### 2019-01-18 Write_up/Papers/AJ_papers_references_15jan/funcs_Intro-copy.Rmd
reading_2.0

### Writing your own functions

Writing your code into functions is key for writing neat R code and prevents repitition of code. In my experience, young biologists are reluctant to use functions in their code and often when functions are being used they are abused and created to do too much at once. RStudio has several built in functions, and with more packages installed, more functions are available. Throughout this course we have continuously been making use of functions. The **tidyverse** package in itself contain several functions, such as the `mean()`, `sqrt()` and **ggplot2** functions.

A typical example of a function would be `sqrt()`. Where the argument must be a number, and the return value is the square root of that number. Executing a function is called *calling* the function. Here we *call* a built in R function.

```{r, eval=FALSE, purl=FALSE}
b <- sqrt(a)
```

Here, the value of `a` is given to the `sqrt()` function, the `sqrt()` function calculates the square root, and returns the value which is then assigned to the object `b`. This function is very simple, because it takes just one argument.

Now we learn to write our own basic function. The process in which we write any function, regardless of the size, is similar and involves three key steps:

- Make use of the term `function`
- Specify the arguements in that function
- Specify the body of the function

```{r, eval=FALSE}
my_func <- function(arg1, arg2) {
  body
}
```

`my_func` is the function name. This can be any variable name, but as previously mentioned throughout this workshop, we should avoid using names that are used in built in R functions such as `mean`, `length`, *etc*., as this may often lead to confusion and ultimately result in unnessasary errors.

`arg1` and `arg2` are the *arguements* of the function. You can write a function with any number of arguments, and some arguments may have default values specified.

Above, 'body' denotes the function's body. This is all the code found within the { ... } brackets. The size of this code may vary depending on the required outcome.

Once we start putting things in functions so that we can re-use them, we need to start testing that those functions are working correctly. To see how to do this, lets write a function. As mentioned earlier some arguments may have default values specified and so here I specify `y` to have a default value of 1. So, in the function call, when no value is assigned to `y` the default value will be 1.

```{r}
add <- function(x , y = 1) {
  x + y
}
```

Here we created a function called `add()`. This function takes two arguments, `x` and `y`, and returns the sum of `x` and `y`. Since `y` already has a default value assigned to it, it is necessary to only specify `x` in the function call; specifying `y` is optional. Since the code was 'activated' by running it through the console, the function is now available, like any other built-in functions in R, for you to use.

Now lets put our function to test:

```{r}
add(15) # When no y value is given it is taken as 1, as specified in the function

add(15, 5)
```

For some practice lets write a function where we calculate the sum of the squares:

```{r}
sum_squares <- function(x, y) {
  x^2 + y^2
}
```

Testing the function

```{r}
sum_squares(9, 20)
```

## Basic plotting using functions

```{r}
library(tidyverse)
chicks <- datasets::ChickWeight

Diet1 <- chicks %>% 
  filter(Diet == "1")
Diet2 <- chicks %>% 
  filter(Diet == "2")
Diet3 <- chicks %>% 
  filter(Diet == "3")
```

```{r}
func_plot <- function(df) {
  plot1 <- df %>% 
    ggplot(aes(x = Time, y = weight)) +
    geom_point() +
    geom_smooth(method = "lm") +
    labs(x = "Days", y = "Weight (gm)", colour = "diet type") 
}
```

```{r}
(diet1 <- func_plot(df = Diet1))
(diet2 <- func_plot(df = Diet2))
(diet3 <- func_plot(df = Diet3))
```

### Summary statistics func.

```{r}
sum_stats <- function(df) {
  new_data <- df %>% 
    summarise(mean_wt = mean(weight),
              sd_wt = sd(weight),
              var_wt = var(weight)) 
}
```

```{r}
(diet1 <- sum_stats(df = Diet1))
(diet2 <- sum_stats(df = Diet2))
(diet3 <- sum_stats(df = Diet3))
```

## Exercise:

### 2019-01-18 Write_up/Writing/Introduction_outline.Rmd
Improved intro

@@
 
-### Introduction
+Chapter 1: General Introduction
 
-Seawater temperature is a key indicator of environmental change in both terrestrial and marine ecosystems (Pearce et al., 2006), which directly influences the earth’s natural ecological functioning.
-Climate change is widely believed to result in a rise in the earth’s average surface temperatures over time (Stocker et al., 2013) and this is usually associated with an increase in the intensity and duration of extreme events such as cold-spells, heatwaves, floods and storms.
-Consequently, these extreme climatic events can often have profound short-term effects on both humans as well as many ecosystems (Easterling et al., 2000).
+## Eastern boundary upwelling systems:
 
-‘Cold spells’ are usually referred to as a phenomenon where abnormally cold water remains on the surface of the ocean for extended periods of time.
-However, several other definitions exist based on the statistical properties of the temperature record in association with the location of temperature collections and the time of year .
-Studies that have quantified the impacts of extreme temperatures have provided evidence of demonstrably negative effects on  associated surrounding regions as well as the specific regions within which these climatic events occur.
-While extensive research has been done on other extreme weather events such as floods, wind storms and marine heat waves, minimal research has been done on marine cold spells (MCS).
-It is known that warm-water anomalies may result in severe coral bleaching and is one of the factors resulting in the demise of coral reefs.
-Additionally, anomalous warm water events are known to have negative impacts on species diversity and mortality as well as the distribution of organisms.
-However, recent studies done by Lirman et al.
-explains that corals living in regions or surrounding regions of unusually prolonged anonymously cold waters may be just as detrimental; this was illustrated in January of 2010 when the reefs of the Florida Reef Track (FRT) were exposed to an unusual cold water event that resulted in large scale coral mortality.
-This climatic event showed clear association with North Atlantic Oscillation (NAO) index that produced northerly surface wind anomalies and the southward advection of the cold Arctic air.
-Similarly, research by Shinn et al.
-(1962) reported high mortality of Acropora cervicornis in shallow reefs as temperatures dropped.
-Likewise, studies conducted by Hudson et al.
-also documented the mortality of Montastraea colonies in near shore patch reefs due to cold fronts between 1969–1970 and 1977–1978, respectively.
+Globally, there are four major eastern boundary upwelling systems (EBUS).
+These are the Benguela Current off south-western Africa, the Canary Current off north-western Africa including its northern extension off the Iberian Peninsula of south-western Europe, the Peru-Humboldt Current off western South America, and the California Current off the western continental USA and north-western Mexico (Bakun et al, 2015; Pauly and Christensen, 1995).
+Each of these systems are characterized as vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans.
+EBUS cover only 1% of the world’s ocean surface, but despite these are among the most productive regions of the world.
+This is because these systems are subject to coastal upwelling, which is defined as the process whereby cold, nutrient rich, high CO2 concentrated, low pH, and low oxygen waters are pushed to the surface as a result of alongshore winds interacting with the earth’s rotation.
+As a result of upwelling, diverse marine fauna and flora are able to thrive in these areas because of the abundance of nutrients and favourable temperatures.
 
-MCSs are lethal to all marine organisms and are known to have caused  mass extinctions of several fish and invertebrate species (Woodward.
-1987 & Gunter.
-1941, 1951 & Holt and Holt.
-1983, Gunter.
-1951 & Crisp.
-1964).
-Temperature has long been considered one of the primary factors controlling species distribution patterns by limiting their distribution north or southward towards the high latitudes (Firth et al., 2011).
-Studies done by Firth et al.
-(2011, 2015) have demonstrated the range shifts of engineer species such as mussels as a result of MCSs.
-Thus, providing a clearer understanding of the influence of oceanic and atmospheric processes on sea surface temperatures (SST) within the upper mixed layer, however minimal research has been done on the influences of local and broad scale impacts on seawater temperature.
+Upwelling is primarily caused by alongshore, equator ward winds.
+These winds are caused by cross-shore atmospheric pressure gradients, and these gradients occur predominantly during heating periods.
+Bakun then proposed that if there is an increase in greenhouse gases, it will cause more heating during the day and more cooling during the night which cause the temperature gradient to increase and would result in a strong pressure gradient, and as we know the pressure gradient will affect the wind and therefore will affect upwelling.
+Several people have tried to model this using different datasets of SST and wind pattern data at various resolution and initially found that increases in gases, such as CO2 causes a reduction in trade winds, and therefore a reduction in upwelling.
+However, these models were based on coarse data and when higher resolutions are used, such as for the Benguela EBUS, the models have shown that upwelling will intensify in accordance to the upwelling intensification hypotheses (eg, Hsieh and Boer 1992, Mote and Mantua 2002, Lima and Wethey’s 2012).
 
-Research by Grunter (1941) and Firth et al.
-(2011) explains that some MCSs that are believed to have influenced species populations were caused by atmospheric cold spells which had ultimately affected the local intertidal and coastal biota.
-MCSs have been proven to have a great effect on ecosystem change, thus understanding these climatic drivers may provide useful tools for conservation and management efforts.
-This study will aim to develop a better understanding of the impacts of weak and strong upwelling events and how they may be influencing the intensity and duration of MCSs.
-This study applied the MCSs definitions to datasets at a local-scale and an offshore broad scale.
+### Ecology and primary production within upwelling zones:
 
-Eastern Boundary Upwelling Systems (EBUS) are among the most productive regions of the world.
-These upwelling systems include the; California, Humboldt, Canary and Benguela currents, and are important for biological productivities by supporting commercial fisheries (Pauly and Christensen, 1995) thus forming an essential component to these region’s economies, as well as providing recreational services to the human population along their coasts.
-One of the five continental margin upwelling systems is the Benguela Current; an eastern boundary current of the south Atlantic subtropical gyre which flows towards the equator approximately 300 km offshore of the coast of southwest Africa before splitting into the Benguela Coastal Current (BCC) and Benguela Oceanic Current (BOC).
-Shoreward of the main stream of the Benguela current there is a slow flow along the continental margin into the Angola Basin.
-In this slow flow the prevailing south and south eastern winds drive coastal upwelling of cold, nutrient rich South Atlantic Central Water (SACW).
-
-The SACW originates by sinking at the Subtropical Front, known as the Subtropical Convergence (Schell, 1970 & Nelson, 1989).
-Beneath the SACW lies cold, nutrient-rich Antarctic Intermediate Water (AAIW) which originates by sinking at the Polar Front.
-Throughout each of the seasons, upwelling of nutrient rich waters covers most of the continental shelf and much of the continental slope (Shannon, 1985), thus creating what we refer to as the Benguela Current Upwelling System.
-Research studies have shown that these cold waters move north over the shelf in the Benguela Coastal Current (Hay and Brock, 1992).
-Offshore they are separated from the warmer waters of the subtropical Atlantic by a well-developed thermal front (Shannon, 1985).
-This quasistationary region is often linked with the shelf edge, although it is distorted by eddies, plumes and jets which extend west into the SE Atlantic (Lutmharms and Stockton, 1987; Waldron and Probyn, 1992).
-These plumes largely expand the region of high productivity as a result of upwelling by rapidly advecting cold, nutrient-rich water from the coastal upwelling sites into the open ocean (Lutjeiarms et al, 1991).
-Northward, the Benguela Current Upwelling system is linked by the Angola Benguela Front which converge between the Benguela current and the Benguela coastal current as well as the warm saline waters of the Angola Current (Peterson and Stramma, 1991).
-Southward, the coastal upwelling system is bounded by the Agulhas Bank.
-South of the African coast the warm Agulhas Current flows in from the Indian Ocean and returns eastwards while generating warm eddies that spin off in a north-westerly direction in the Benguela current (Duncombe, 1991).
-Research studies done by Shannon et al.
-(1985) explains that coastal upwelling varies on a seasonal basis and that during winter and spring, water of approximately 16°C extends along the coast.
-Upwelling is coldest and most persistent off the Namibian coast in the Lüderitz region at 25°S and weakest at 17°S (Lutjeharms and Meeuwis, 1987).
-Lutjeharms and Meeuwis, (1987) also found a strong relationship between the intensity of upwelling events and the strength and direction of prevailing winds.
-It is also hypothesised that upwelling is more intense where deep water is closer to the coast.
-
-EBUS mesoscale and sub-mesoscale oceanic movements are known to influence marine ecosystems in that it enhances primary production (McGillicuddy et al., 1998; Oschlies and Garçon, 1998) and is key in managing the exchange and of transporting of organic resources and marine organisms from the coastal zone to the open ocean.
-Large-scale atmospheric pressure systems that favour along-shore, equatorward winds, result in the high levels of productivity in the EBUS, which in combination with the Coriolis effect, affects surface water offshore.
-This results in a process whereby cold, nutrient rich, with low pH and low oxygen concentration, originates from deep within the ocean and is pushed to the surface where they, in the form of phytoplankton blooms, fuel primary production.
-As a result of the intensity and duration of the upwelling event; regions affected by these cold water events are exposed to distinct ecological conditions and tend to differ from its surrounding regions (Bosman et al., 1987; Barshis et al., 2011).
-Significant research evidence suggests that SST have increased, however some regions still show a lower or non-significant increase, and in some cases cooling trends are evident.
-Inshore regions and embayment’s that are protected from wind action may be subjected to intense episodes of primary production (Graham and Largier, 1997).
+The duration and intensity of coastal upwelling are known to have a critical role in the phenology of key marine ecosystem processes.
+Similarly, changes in the characteristics of upwelling are shown to have cause large disturbances to ecosystems at several trophic levels (Wang et al., 2015).
+Cold, nutrient rich, low pH, and low oxygen upwelled waters fuels primary production as a consequence of the development of phytoplankton blooms (Ryther 1969, Cushing 1971, Payne 1987, McGillicuddy et al., 1998; Oschlies and Garçon, 1998).
+In these upwelling systems, the high productivity supports vast and diverse marine populations.
+Mesoscale and sub-mesoscale oceanic movements are key in managing the exchange and transportation of organic resources and marine organisms from the coastal zone to the open ocean.
+Most of the world’s major fishing industries are dependent on upwelling, because this process brings nutrient rich waters into the euphotic zone, where it forms the foundation of a localized food web that ecologically differs from surrounding, non-upwelled regions (Bosman et al., 1987; Barshis et al., 2011).
+Whilst nutrients play a primary role in the survival of these systems, the cool temperatures of these waters are critical as well.
+There is an increasingly large body of research that has provided evidence showing that sea surface temperatures (SST) have and are continuing to increase.
+However, some regions still show a lower or non-significant increase, and in some cases cooling trends are evident (REFERENCES).
+Upwelling occurring along the Cape south coast is proven to have less of an economic importance in contrast to the west coast.
+Inshore regions and embayments that are protected from wind action may be subjected to intense episodes of primary production (Graham and Largier, 1997).
 Vander Wouder et al.
 (2006) found that these sites represent the highest levels of chlorophyll concentration.
+Biotic systems within the Benguela upwelling system may be negatively influenced by the changes in oxygen levels given their adaption to natural variable conditions (Capone and Hutchins, 2013).
 
-Atmospheric and oceanographic processes results in upwelling, while the amplitude and timing of upwelling-favourable winds are sensitive to climate variability (Montecinos et al., 2003; Macias et al., 2012), and are highly likely to be affected by global warming (Barros et al., 2014; Bakun et al., 2015).
-It is hypothesised that global warming will influence atmospheric pressure gradients, ultimately affecting the amplitude and timing of coastal winds that cause upwelling (Bakun, 1990; Bakun et al., 2015) and more recent observations (Rykaczewski et al., 2015; Wanget al., 2015) by means of global model projections strongly suggests intensification of coastal winds.
-Understanding the influence of changing winds on upwelling events is highly important.
-For example, weak winds may result in weak upwelling, which in turn may limit nutrient concentration; resulting in potentially negative impacts on primary production (Chhak and Di Lorenzo, 2007).
-In contrast, stronger upwelling winds, could theoretically result in stronger upwelling events, thereby increasing the nutrient concentration which may increase offshore transport (Cury and Roy, 1989).
-Increased wind action may also result in increased turbulence (Cury and Roy, 1989); and may even affect the processes associated with ocean acidification, creating widespread changes within ecosystems and ultimately impacting regional productivity (Gruber, 2011; Doney et al.
-2012).
+For several decades, regime shifts have become an important factor of marine research.
+Regime shifts are often defined as a shift from one stable ecosystem state to another.
+This process affects several species within a short period of time (deYoung et al., 2004; Cury and Shannon, 2004; Jarre et al., 2006, Blamey 2012).
+Regime shifts are thought to be driven by abiotic processes such as climate change (Polovina, 2005, Blamey 2012).
+The southern Benguela upwelling system supports a wide variety of commercially-exploited fish species and exhibits substantial variability in both oceanographic and biological components.
+Over the past 3 three decades, a number of species occurring within this region have undergone shifts in their distributions and Roy et al.
+(2007) linked these shifts to cooling waters along the inner shelf of the Agulhas bank on the south coast, east of Cape Agulhas (Blamey, 2012).
+This coast is associated with high levels of harmful toxic algal blooms, which negatively impact commercial and recreational interests along the coastal region.
+These harmful blooms are often attributed to flagellated species with the red tide representing dinoflagellate blooms (Pitcher et al.
+1993A, Pitcher 1998).
+Formations of red tides are often related to the prevailing winds of the southern Benguela which govern most hydrodynamic processes on the continental shelf.
+Red tide events are greatest during the latter part of the upwelling season during transitions in the synoptic weather patterns which result in diminished upwelling activity and increased thermal stratification (Pitcher et al.
+1993a).
 
-**** A better understanding of the impact of global warming on wind variability is important in that it allows for better management of fisheries and other marine resources in the EBUS which could be improved with a better understanding of how nutrient abundance and primary production are associated with upwelling events.
-This study aims to emphasize the drivers of upwelling-favourable winds, as well as the factors influencing nutrient enrichment, and potential changes in biogeochemistry*****
+### Atmosphere:
 
-The Benguela current, forming part of the EBUS, experiences cross-shore atmospheric pressure gradients, which results in alongshore, equatorial winds.
-These winds are known to drive coastal upwelling events.
-With this in mind, research by Bakun (1990) suggests that an increase in greenhouse gas emissions would exacerbate warm-season upwelling.
-The understanding behind this mechanism was based on the idea that air above land would warm rapidly as compared to air above the ocean, and that the deepening of the continental thermal low pressure system (CTLPS) would increase the cross-shore pressure gradient which drives upwelling-favourable winds.
-Several research studies examining the trend between upwelling events and wind patterns, have however indicated contrasting results.
-With more than 20 years of observational and modelled data; Sydeman et al.
-(2014a) discovered the most coherent pattern of intensifying upwelling-favourable winds in the Humboldt and California upwelling systems.
-It has proven difficult to attribute coastal wind intensification in EBUS to global warming, despite the fact that the meta-analysis done by Sydeman et al.
-(2014a) generally supports Bakun’s (1990) proposition.
+Throughout region surrounding the Benguela EBUS, the atmosphere is dominated by anticyclonic high pressure cells with quasi-stationary positions.
+The South Atlantic Ocean High is situated along the west, drawing cool, dry air onto the west of the subcontinent (Van Heerden and Hurry, 1998).
+Solar heating during summer may result in the development of low pressure cells, known as heat lows.
+These cells are absent during winter time (Tyson and Preston-Whyte, 2000).
+When the anticyclones shift north, during winter months, the cold westerlies have a substantial impact on the weather of the southern tip of the South African subcontinent (Van Heerden and Hurry, 1998).
+The atmospheric temperatures along the coastline are largely influenced by the Benguela Current (Van Heerden and Hurry, 1998).
+However, upwelling systems may be forced by various atmospheric conditions (Lutjeharms et al., 2003; Roberts, 2005; Hutchings et al., 2009).
 
-This study cannot discount the role of multi-decadal climate variability in the observed trends (Chhak and Di Lorenzo, 2007; Narayan et al., 2010; Pérez et al., 2010; Macias et al., 2012; Santos et al., 2012; Cropper et al., 2014; Jacox et al., 2014), nor is it a test of the Bakun (1990) hypothesis mechanism.
-Given the broad range of contradicting studies, the only conclusion that can be made is that upwelling-favourable winds tend to be the most intense during the warmer months of the year (Huyer, 1983; Nelson and Hutchings, 1983).
-The timing and the intensity of upwelling favourable winds are highly variable within the EBUS, therefore creating much uncertainty (Chavez and Messié, 2009).
-Understanding the factors contributing to the intensity and duration of upwelling events is important because variations in the timing of upwelling strongly influences ecosystem productivity.
-Since the late 1940s studies have shown an increase in upwelling favourable winds associated with the California current, however, less variability has been observed in the strength of upwelling winds through the year in the central Benguela and Canary Currents (Shannon and Nelson, 1996; Varela et al., 2015).
-It is important to note that the lack of research conducted may obscure the validity of any conclusion drawn here.
-This study further researched the changes in the timing and variability of upwelling within the Benguela Current and how these phenomena influence associated ecosystems.
+### Sea surface temperature and data collection:
 
-As a result of climate change, global ocean temperatures have increased at an accelerating rate (Stocker et al., 2013).
-Sufficient research suggests that coastal sea surface temperatures (SST) in EBUS such as the Benguela current are cooler than offshore regions as a result of the near shore upwelling waters.
-There is a large but contradicting amount of evidence on the SST trends in EBUS (Mendelssohn and Schwing, 2002; Arístegui et al., 2009; Belkin, 2009; Demarcq, 2009; Lebassi et al., 2009; García-Reyes and Largier, 2010; Rouault et al., 2010; Gutiérrez et al., 2011; Seo et al., 2012; Barton et al., 2013; Barros et al., 2014; Sydeman et al., 2014b; Salvanes et al., 2015).
-The contradictions amongst studies exist because of varying resolutions of SST data used, thus resulting in the capturing of different aspects of the upwelling processes, showing different trends that exist based on the resolutions of the different SST datasets as well as other intrinsic factors.
-Generally, because of the resolution of SST datasets it is often difficult to separate near shore upwelling temperatures from regional, broad scale and eastern boundary current temperatures.
-Lima and Wethey’s (2012) found that high resolution data represented clear cooling trends predominantly during summer and at high latitudes in most of the EBUSs.
-Additionally several research studies examined the difference between coastal and offshore temperature trends, revealing decreasing or increasing trends in near shore SST, compared with increasing trends offshore in the Benguela compared with increasing trends offshore in the Benguela (Rouault et al.,2010; Santos et al., 2012; Salvanes et al., 2015).
-In this study however, will utilise a range of spatio-temporal resolutions of SST data and in situ collected seawater temperature measurements for the Benguela upwelling system.
-The aim of this is to demonstrate the variable upwelling events based on the different resolutions at which temperature is being collected.
+Water temperature affects the survival, abundance, distribution and growth of marine species (REFERENCE).
+As such, a clear understanding of the relationship between marine ecosystems and their thermal tolerance is necessary for conservation and management planning.
+Future predictions in biogeographical patterns and ecosystem functioning may also be estimated based on various climatic behaviours.
+Regardless of the objectives, whether it be studies based on climate change, extreme events or ecosystem functioning, it is important that the research is conducted through accurate, reliable datasets.
+Unfortunately, not all data are created equally and under the same conditions and as such not all datasets are usable for a particular task.
 
-Generally, regional ocean climate oscillations like the Benguela Niño for the Benguela Current System impose considerable variability on temperatures and complicate attempts to assess long-term trends (Demarcq,2009).
+Remotely-sensed satellite SST products are produced and maintained by a number of organisations worldwide.
+In order to obtain these SST data, advanced technology is launched into space, in order to obtain infra-red light being emitted by the ocean.
+This irradiance is then taken as a proxy for SST.
+Due to the distance at which these SST are being collected, certain coarseness within the datasets are unavoidable.
+Satellite SST products undergo a range of quality control processes that are well researched and tested (Reynolds and Smith, 1994; Brown et al., 1999; Martin et al., 2012), making their use more reliable.
+However, Smit et al.
+(2013) have shown that SST data are not particularly reliable for coastal regions and should be used cautiously as they can be prone to both warm and cold bias.
+Nevertheless, satellite data products are extremely useful when investigating the ocean at broad scale and can provide important SST data.
+<will add more here on the different types of resolution used in this thesis>
+
+### Global warming and climate change:
+
+Climate change as a result of anthropogenic activities is now established as a global concern for both scientific and political importance.
+As the human population grows, the rate of global warming increases, resulting in profound effects on marine and other ecosystems.
+Climbing temperatures create a host of additional changes to marine systems, such as rising sea levels, increased ocean stratification, decreased sea-ice extent, and altered patterns of ocean circulation and precipitation (Doney et al., 2011).
+Furthermore, direct effects on ocean temperature changes and chemistry may alter the functioning and behaviour or marine organisms.
+In turn, these changes lead to a change in species integration and trophic pathways.
+
+Significant research studies accept the climate change and variability have great impacts on marine ecosystems (Stenseth et al.
+2002, Harley et al.
+2006, Hoegh-Guldberg & Bruno 2010).
+The Intergovernmental Panel on Climate Change (IPCC) predicts future projections with a mean increase in air temperature of 1.1°C by the 2020’s and 1.8°C by the 2040’s as a result of greenhouse gas concentrations (Solomon et al.
+2007).
+A large amount of this heat will be absorbed by the world’s ocean.
+In current years, as a result of changes in air temperature, the heat content of the ocean has increased largely.
+Given ocean warming in association with their physical changes such as sea level rise and declines in sea ice, marine life is expected to intensify in future decades.
+Global warming represents distinct warming and cooling periods throughout the 20th and 21st century (IPCC, 2007).
+Global warming however is not evenly distributed across the global ocean thus presenting regions warming much faster or slower than then the global average (Levitus et al., 2000; Palttridge and Woodruff, 1981; Santos 2012).
+In particular, the Atlantic Ocean contributes most to the increase of the heat content (Nerem et al., 1999; Levitus et al., 2000; Strong et al., 2000).
+
+As a result of climate change, research studies provide significant evidence on the increase in warming of the ocean surface.
+This warming slowly propagates down through the water column and in turn increases the thermal stratification (Gruber, 2011).
+Strong stratification limits the depth at which water is upwelled and hence controls the amount of nutrients brought up (Chhak and Di Lorenzo, 2007; Jacox and Edwards, 2011; Jacox et al., 2015).
+The interplay between upwelling winds and stratification determine the amount of nutrients present within the system and thus is strongly linked to biological productivity.
+Globally, as a result of climate change, stratification of the water-column between 0 and 200 m has increased by an estimated 4% between 1971 to 2010 (Stocker et al., 2013).
+Increased ocean stratification as a result of surface warming may reduce the ventilation of deep-water masses.
+As the deep water is being up welled to the surface, decreases in oxygen and pH levels are predicted along with an increase in nutrient levels (Rykaczewski and Dunne, 2010).
+Generally, regional ocean climate oscillations like the Benguela Niño for the Benguela Current System impose considerable variability on temperatures and complicate attempts to assess long-term trends (Demarcq, 2009).
 It is predicted that these climate oscillations may increase the variance with climate change (Timmermann et al., 1999; Kuzmina et al., 2005; Sydeman et al., 2013; Cai et al., 2015) which will further result in unusual trends.
 
-Research studies provide significant evidence on the increase in warming of the ocean surface.
-This warming slowly propagates down through the water column and in turn increases the thermal stratification (Gruber, 2011).
-Strong stratification limits the depth at which water is up welled and hence controls the amount of nutrients brought up (Chhak and Di Lorenzo, 2007; Jacox and Edwards, 2011; Jacox et al., 2015).
-The interplay between upwelling winds and stratification determine the amount of nutrients present within the system and thus is strongly linked to biological productivity.
-Globally; as a result of climate change, stratification of the water-column between 0 and 200 m has increased by an estimated 4% during the period 1971 to 2010 (Stocker et al., 2013).
-Variability in the biogeochemical properties of up welled waters influences species distribution, ecosystem structure and productivity within the Benguela upwelling system.
-The deep waters being pushed to the surface through the upwelling process is high in nutrients and dissolved CO2, a consequence of the long period of isolation at a depth from the euphotic zone (Feely et al., 2008).
-Increased ocean stratification as a result of surface warming may reduce the ventilation of deep-water masses.
-As the deep water is being up welled to the surface, decreases in oxygen and pH levels are predicted along with an increase in nutrient levels (Rykaczewski and Dunne, 2010).
+As previously mentioned, upwelling winds in EBUSs are generated by pressure differences between continental thermal lows and oceanic highs.
+As warming intensifies, pressure gradients between land and sea increases, resulting in a greater intensity of upwelling winds (Bakun et al.
+2010).
+In rare cases, projected upwelling increases may overcome the countervailing effects of upper-ocean warming and stratification to cause regional cooling (Auad et al.2006).
+Research studies reveal declines in ocean oxygen levels as a result of ocean warming and reduced ventilation from stratification and circulation changes (Keeling et al.
+2010).
+Similarly, low oxygen levels are also possible because of shifts in the composition of sinking organic material (Hofmann & Schellnhuber 2009, Doney et al., 2011).
 
-Biotic systems within the Benguela upwelling system may be negatively influenced by the changes in oxygen levels given their adaption to natural variable conditions (Capone and Hutchins, 2013).
-This study will seek to examine the impact of both strong and weak upwelling winds on the intensity and duration of the upwelling event; additionally it will also aim to examine how chlorophyll concentrations vary with regard to contrasting upwelling conditions as well as the impact this has on species distribution and habitats.
+## Problem statement:
+With global warming continuing at a rapid pace the amount of greenhouse gases within the atmosphere are going to increase in the future.
+This increase will undoubtedly have an impact on the world’s oceans and the marine life within, both directly and indirectly.
+The oceanic process of coastal upwelling is one aspect that is likely to be affected as changes in global temperature will affect temperature and pressure gradients, and in turn affect alongshore winds which are responsible for driving this process within EBUS regions.
+Because upwelling is important for productivity within coastal systems, a decrease in this process will negatively impact on marine fauna and flora, and on human activities such as fishing.
+To assess changes in upwelling events, here I aim to use SST, wind, chlorophyll data etc.
+to monitor durations and intensities of upwelling within one of the EBUS.....
 
-##### Ideas #####
+## Study Area:
 
--chnaging the structure it to describe MCS, explain the dangers of these phenomena on marine life (, the implications of this (eg distributions being affected), and the causes of MCS.
+This study focusses on the Benguela EBUS.
+The Benguela EBUS is unique in that it is bounded at both the equator ward and pole ward ends by warm water regimes.
+These currents are forced locally by the wind stress field off South-West Africa (Nelson and Hutchings 1983; Fennel 1997).
+This upwelling system stretches along the south-western coast of Africa from southern Angola to Cape Agulhas (Cole, 1999).
+Water is coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S (Lutjeharms and Meeuwis, 1987).
+The main upwelling period in the Benguela north of 25°S extends mainly from March to November, peaking during August (Shannon, 1985), but upwelling persists throughout the year.
+Upwelling season in the northern Benguela extends between September and March, with very little upwelling taking place during winter (June-August).
+In the central Benguela and also at Cape Frio upwelling is continuous throughout the year (Shannon, 1985).
+The cold Benguela waters are bound in the south by the warm Agulus retroflexion region and in the north by the southward flowing Angola current.
+The flow of the Benguela is often considered to be topographically guided (Nelson and Hutchings, 1938; Barange et al.
+1992).
+The Benguela system is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region within the Namibian region (Shannon 1985; Cole 1999).
 
--explain what info is unknown about MCS and why this is a problem in areas where they can occur
+Upwelling in the northern Benguela is in phase with the seasonal insulation cycle, whereas in the southern Benguela it is not.
+The result is that there is a pronounced seasonal cycle in SST in the north while in the south there is no clear SST cycle.
+The seasonal warming occurring during the late summer season off northern and central Namibia is a regular occurrence and is associated with warmer, higher salinity water of Angolan origin into the region (O'Toole, 1980; Shannon, 1985).
+There have been a number of warm and cool periods documented in the northern and southern Benguela (e.g.
+Stander and De Decker, 1969; Walker et al.1984; Shannon et al., 1984; Boyd and Thomas, 1984; Shelton et al.
+1985; McLain et al.: 1985).
+Walter (1937) found a major climatic event in the northern Benguela region of anonymously warm water.
+This unusual warm water may have resulted due to reduced upwelling or an increase southward advection of warm water.
+Another warming event occurred during the summer and autumn months of 1984 (Shannon et al, 1986).
+The presence of this anomalous warm water during 1984 followed two cooler than normal years in the northern Benguela.
+Shannon et al.
+(1986) illustrated the warm (21-23°C) water near Walvis Bay as a large warm water patch offshore west of Luderitz.
+The absence of upwelling between Walvis Bay and north of Luderitz is evident however, the upwelling centre in the Benguela system (Shannon, 1985), and upwelling was minimal.
+Interestingly, in the southern Benguela region, there was no evidence of any unusual patterns in temperature distributions.
+The seasonal temperature pattern within the northern Benguela is somewhat similar to that in the eastern equatorial Atlantic and the regular southward intrusion of warm water into the northern Benguela during the summer season.
+During these warming periods the equator ward wind stress over the northern Benguela region was stronger than normal; yet upwelling was suppressed by the southward movement of warm saline waters from the Angolan region.
 
-- link this to the EBUS and Benguela stuff
+Northward, the Benguela Current Upwelling system is linked by the Angola Benguela Front which converge between the Benguela current and the Benguela coastal current as well as the warm saline waters of the Angola Current (Peterson and Stramma, 1991).
+Research done by Yamagata and Iizuka (1995) discovered slight indication that coastal Kelvin waves, originating from equatorial Kelvin waves, carry the down welling process southward.
+Southward, the coastal upwelling system is bounded by the Agulhas Bank.
+South of the African coast the warm Agulhas Current, western boundary current, flows in from the Indian Ocean and returns eastwards while generating warm eddies that spin off in a north-westerly direction in the Benguela current (Duncombe, 1991).
+The region surrounding the Benguela experiences a persistent along shore wind.
+This wind is associated with the St.
+Helena high pressure system.
+Lutjeharms and Meeuwis, (1987) also found a strong relationship between the intensity of upwelling events and the strength and direction of prevailing winds.
+These favourable upwelling equatorial winds represent a maximum at 25°S and decrease towards the northern and southern boundaries of the Benguela upwelling system.
+Continental shelf bathymetry and upwelling favourable winds provide a large-scale upwelling mechanism in the southern Benguela, whereas local topography and meteorology create an alternating pattern of active and passive upwelling circulations along the coast.
+Winds occurring within the southern regions represent a seasonal variation, reaching a maximum during spring and summer months (Boyd 1987; Shannon and Nelson 1996; Fennel 1997).
+Winds occurring within this region tend to show increasing patterns away from the coast whereas, winds occurring in the northern region exhibits relatively less seasonal variation (Shannon 1985.
+Pitcher et al.
+1992).
+Upwelling occurring along the South Coast of the Cape Province within South Africa, specifically along the coastline is generated by local wind.
+The wind increases somewhat away from the coast.
+This facilitated the idea of wind stress curl, as discussed in Bakun and Nelson (1991) and Shannon and Nelson (1996).
 
-- clearly mention the aims, objectives and approach to the study.
+## Extreme events
+
+### Cold spells
+- Definition
+- Impact
+-Major cold spells
+
+## Aims and approach:
+In this thesis I aim to…

### 2019-01-20 Methods.Rmd
Updates

---
title: "Methods"
author: "Amieroh Abrahams"
date: "19 January 2019"
output: html_document
---

# Libraries

First I need to find, install and load various packages. These packages will be available on CRAN and can be accessed and installed in the usual way.

```{r prelim_opts, echo=FALSE}
knitr::opts_chunk$set(
  comment = "R>",
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(plyr)
library(lubridate)
library(ggpubr)
library(zoo)
library(lubridate)
library(FNN)
library(forecast)
library(astrochron)
library(WaveletComp)
library(data.table)
```

setwd("~/Documents/MastersProject")

### Load site list SACTN data

Now to get to the data. The first step involves the loading of the site list. The statistical properties of the seawater temperature representing the South African coastline, such as the mean, minimum and maximum temperatures. These values vary among coastal sections due to the influence of the cold Benguala and warm Agulhas currents. Here we will only focus on the temperature data found along the west coast (wc) (i.e. sites influenced by the Benguela current, EBUS). The SACTN dataset comprise of 129 *in situ* coastal seawater temperatures derived from daily measurements over up to 40 years. The SACTN temperature dataset was compiled by measuring coastal temperatures at 129 sites along the coast of South Africa, daily from 1972 until 2017.

```{r load_files1, include=TRUE}
load("Data/site_list_v4.2.RData")
load("Data/SACTN_daily_v4.2.RData")
```

Now we select only the sites occuring along the west coast. Thereafter, I include all the sites which had a time series of 10years or greater. This yield a 2 rows of data for Paternoster. I removed the row of data obtained by the DEA, thus ending up with all of the data being obtained from SAWS in this dataset. I then exclude the sites that do not fall within the upwelling cell. Once those sites were removed, we have a total of 9 sites. These are: Port Nolloth, Hondeklip baai, Doring baai, Lamberts Bay, Elands Bay, St Helena Bay, Paternoster, Saldanna and Yzerfontein.

```{r}
site_list_sub <- site_list %>% 
  filter(coast == "wc")

SACTN_daily_wc <- site_list_sub %>% 
  filter(length >= 3650)

upwelling_SACTN_daily_wc <- SACTN_daily_wc[-6,] # Here we had two rows of data for Paternoster; i removed the DEA row: All data here are now collected by SAWS 
upwelling_SACTN_daily_wc <- upwelling_SACTN_daily_wc[-10:-14,] # Here i remove the sites that does not fall within the upwelling cell

SACTN_daily_temps <- SACTN_daily_v4.2 %>% 
  left_join(site_list[,c(4,13)], by = "index") %>% 
  filter(index %in% upwelling_SACTN_daily_wc$index)

SACTN_daily_temps_indexsplit <- SACTN_daily_temps %>% 
  separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>% 
  dplyr::rename(insitu_temp = temp)
```

Here I add monthly and yearly column as well as a seasonal column. This is done as it allows us to compare temperature variation on a seasonal basis. Research studies done suggests that the intensity and duration of upwelling varies on a seasonal basis. Here we test this....... #Could plot on a seasonal basis##

```{r}
SACTN_daily_temps_monthly <- SACTN_daily_temps %>% 
  mutate(month = month(date, abbr = T, label = T),
         year = year(date)) %>%
  group_by(index, month, year) %>%
    mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",        
                           ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
                                ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
                                       ifelse(month %in% c("Oct", "Nov", "Dec"),
                                              "Spring","Error")))))
```

In this thesis, different remotely sensed SST datasets are used. Hence, these datasets all represent different resolutions at which data was obtained. We will make use of five different remotely sensed SST datasets, these include; the  MUR, AVHRR OISST, Pathfinder and G1SST data.
Once the SST data were downloaded with a netCDF format i needed to convert this netCDfs to CSV

Here I load the AVHRR CSV dataset. Large dataset and slows down the laptop. Here i split the data into xaa and xab. Here xaa and xab only include temperature collected for two years.
```{r}
# AVHRR <- read_csv("data/csvavhrr-only-v2-19810901-20180630.csv")

OISSTDir <- "~/Documents/Masters_2019/MastersProject/Data"
OISST <- fread(paste0(OISSTDir, "/csvavhrr-only-v2-19810901-20180630.csv"),
            col.names = c("lon", "lat", "temp", "date"))

#OISST <- as.tibble(fread("~/Documents/Masters_2019/MastersProject/Data/csvavhrr-only-v2-19810901-20180630.csv"))
##################
# xaa <- read_csv("Data/xaa.csv",  col_names = c("lon", "lat", "temp", "date"))
# xab <- read_csv("Data/xab.csv",  col_names = c("lon", "lat", "temp", "date"))
# combined_OISST <- rbind(xaa, xab)
# save(combined_OISST, file = "data/combined_OISST.RData")
```

# Quick visualisation of the data

To explore the data i visualise the min temperatures along the South african coastline.

```{r}
OISST %>%
  filter(date == min(date)) %>%
  ggplot(aes(x = lon, y = lat)) +
  geom_raster(aes(fill = temp))
ggsave(plot = OISST, filename = "Figures/OISST")
```

## Here i find the nearest 5 SST pixels

Now we apply the FNN (Fast Nearest Neighbor) package to determine the nearesr SST pixel to the insitu collected sites. In some cases however our sites are located close to eachother and as such more than one sites may have the same SST record. Here we use a pixel of 1.

```{r}
unique_pixel <- OISST %>%
  select(lon, lat) %>%
  unique()

# Select nearest 1 pixels (k = 1)
# here we use knnx to find the closes 1 pixels to the insitu sites
match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
                          query = as.matrix(upwelling_SACTN_daily_wc[,5:6]), k = 1)

# Select SST pixels nearest to insitu sites
pixel_match <- unique_pixel[match_index,] %>%
  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
  mutate(site = upwelling_SACTN_daily_wc$site)
```

## Subsetting the OISST data to match the upwelling sites within the in situ collected temperature data

```{r}
OISST_match <- OISST %>%
  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
  filter(combi %in% pixel_match$combi)
```

## Visualisation

```{r}
ggplot(data = OISST_match, aes(x = date, y = temp)) +
  geom_line() +
  facet_wrap(~combi)
ggsave(plot = OISST_match, filename = "Figures/combined_OISST_match")
```

```{r}
OISST_match %>%
  filter(date == max(date)) %>%
  ggplot(aes(x = lon, y = lat)) +
  geom_raster(aes(fill = temp))
```

```{r}
OISST_sites <- OISST_match %>%
  left_join(pixel_match, by = c("combi", "lon", "lat")) %>% 
  dplyr::rename(temp_OISST =temp)

# Check that the pixel that is shared is duplicated
# unique(combined_OISST_sites$site)
# Visualise time series
ggplot(data = OISST_sites, aes(x = date, y = temp)) +
  geom_line() +
  facet_wrap(~site)

save(OISST_sites, file = "Data/OISST_sites.RData")
```

```{r}
load("~/Documents/MastersProject/Data/combined_OISST_sites.RData")
```

## Matching the Insitu data with the OISST SST data
- Here there are many NA values because the OISST have many values collected per day where as the insitu temp only has 1

```{r}
matched_SST <- OISST_sites %>% 
  left_join(SACTN_daily_temps_indexsplit, by = c("site", "date")) %>% 
    na.trim()
```

## MUR dataset

```{r}
# MUR_Doringbaai <- read_csv("Data/MUR_nearest5pixels/MUR_Doringbaai_SST_timeseries_5nearest.csv")
# MUR_Elands_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Elands Bay_SST_timeseries_5nearest.csv")
# MUR_Hondeklipbaai <- read_csv("Data/MUR_nearest5pixels/MUR_Hondeklipbaai_SST_timeseries_5nearest.csv")
# MUR_Lamberts_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Lamberts Bay_SST_timeseries_5nearest.csv")
# MUR_Paternoster <- read_csv("Data/MUR_nearest5pixels/MUR_Paternoster_SST_timeseries_5nearest.csv")
# MUR_Port_Nolloth <- read_csv("Data/MUR_nearest5pixels/MUR_Port Nolloth_SST_timeseries_5nearest.csv")
# MUR_Saldanha_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Saldanha Bay_SST_timeseries_5nearest.csv")
# MUR_St_Helena_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_St Helena Bay_SST_timeseries_5nearest.csv")
# MUR_Yzerfontein <- read_csv("Data/MUR_nearest5pixels/MUR_Yzerfontein_SST_timeseries_5nearest.csv")
# 
# MUR_SST <- rbind(MUR_Doringbaai,MUR_Elands_Bay, MUR_Hondeklipbaai,MUR_Lamberts_Bay,MUR_Yzerfontein,
#                  MUR_Paternoster, MUR_Port_Nolloth,MUR_Saldanha_Bay, MUR_St_Helena_Bay) %>% 
#   dplyr::rename(site = station)
# save(MUR_SST, file = "Data/MUR_SST.RData")
```

```{r}
load("~/Documents/Masters_2019/MastersProject/Data/MUR_SST.RData")
```

hee1 <- try1 %>% 
  dplyr::group_by(site, date) %>% 
  dplyr::summarise(temp = mean(temp, na.rm = TRUE)) %>% 
  dplyr::mutate(no = seq(1:n())) %>%
  dplyr::ungroup() %>% 
  dplyr::select(site, temp, date)

ggplot(data = thee1, aes(x = date, y = temp)) +
  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ site, nrow = 4) +
  theme_bw() +
  theme(strip.text.x = element_text(size = 3))

### 2019-01-20 Write_up/Writing/Introduction_outline.Rmd
Updates

@@
 Each of these systems are characterized as vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans.
-EBUS cover only 1% of the world’s ocean surface, but despite these are among the most productive regions of the world.
-This is because these systems are subject to coastal upwelling, which is defined as the process whereby cold, nutrient rich, high CO2 concentrated, low pH, and low oxygen waters are pushed to the surface as a result of alongshore winds interacting with the earth’s rotation.
+EBUS cover only 1% of the world’s ocean surface, but despite these are among the most productive regions of the world (Pauly and Christensen, 1995).
+This is because these systems are subject to coastal upwelling, which is defined as the process whereby cold, nutrient rich, high CO2 concentrated, low pH, and low oxygen waters are pushed to the surface as a result of alongshore winds interacting with the earth’s rotation (Huyer 1983).
 As a result of upwelling, diverse marine fauna and flora are able to thrive in these areas because of the abundance of nutrients and favourable temperatures.
@@
 Bakun then proposed that if there is an increase in greenhouse gases, it will cause more heating during the day and more cooling during the night which cause the temperature gradient to increase and would result in a strong pressure gradient, and as we know the pressure gradient will affect the wind and therefore will affect upwelling.
-Several people have tried to model this using different datasets of SST and wind pattern data at various resolution and initially found that increases in gases, such as CO2 causes a reduction in trade winds, and therefore a reduction in upwelling.
+Several people have tried to model this using different datasets of SST and wind pattern data at various resolution and initially found that increases in gases, such as CO^2 causes a reduction in trade winds, and therefore a reduction in upwelling.
 However, these models were based on coarse data and when higher resolutions are used, such as for the Benguela EBUS, the models have shown that upwelling will intensify in accordance to the upwelling intensification hypotheses (eg, Hsieh and Boer 1992, Mote and Mantua 2002, Lima and Wethey’s 2012).
@@
 The Benguela EBUS is unique in that it is bounded at both the equator ward and pole ward ends by warm water regimes.
-These currents are forced locally by the wind stress field off South-West Africa (Nelson and Hutchings 1983; Fennel 1997).
+These currents are forced locally by the wind stress field off South-West Africa (Nelson and Hutchings 1983; Fennel 1999).
 This upwelling system stretches along the south-western coast of Africa from southern Angola to Cape Agulhas (Cole, 1999).
@@
 The cold Benguela waters are bound in the south by the warm Agulus retroflexion region and in the north by the southward flowing Angola current.
-The flow of the Benguela is often considered to be topographically guided (Nelson and Hutchings, 1938; Barange et al.
-1992).
+The flow of the Benguela is often considered to be topographically guided (Nelson and Hutchings, 1983; Barange et al.
+1991a and 1991b).
 The Benguela system is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region within the Namibian region (Shannon 1985; Cole 1999).
@@
 Continental shelf bathymetry and upwelling favourable winds provide a large-scale upwelling mechanism in the southern Benguela, whereas local topography and meteorology create an alternating pattern of active and passive upwelling circulations along the coast.
-Winds occurring within the southern regions represent a seasonal variation, reaching a maximum during spring and summer months (Boyd 1987; Shannon and Nelson 1996; Fennel 1997).
+Winds occurring within the southern regions represent a seasonal variation, reaching a maximum during spring and summer months (Boyd 1987; Shannon and Nelson 1996; Fennel 1999).
 Winds occurring within this region tend to show increasing patterns away from the coast whereas, winds occurring in the northern region exhibits relatively less seasonal variation (Shannon 1985.

### 2019-01-21 Methods.Rmd
Updated commit

@@
 ```
-
-setwd("~/Documents/MastersProject")
 
@@
 
-upwelling_SACTN_daily_wc <- SACTN_daily_wc[-6,] # Here we had two rows of data for Paternoster; i removed the DEA row: All data here are now collected by SAWS 
-upwelling_SACTN_daily_wc <- upwelling_SACTN_daily_wc[-10:-14,] # Here i remove the sites that does not fall within the upwelling cell
+upwelling_sites <- SACTN_daily_wc[-6,] # Here we had two rows of data for Paternoster; i removed the DEA row: All data here are now collected by SAWS 
+upwelling_sites <- upwelling_sites[-10:-14,] # Here i remove the sites that does not fall within the upwelling cell
 
@@
   left_join(site_list[,c(4,13)], by = "index") %>% 
-  filter(index %in% upwelling_SACTN_daily_wc$index)
+  filter(index %in% upwelling_sites$index)
+
+# save(SACTN_daily_temps, file = "SACTN_daily_temps.RData")
 
@@
   dplyr::rename(insitu_temp = temp)
+
+save(SACTN_daily_temps_indexsplit, file = "SACTN_daily_temps_indexsplit.RData")
 ```
@@
 ```{r}
-SACTN_daily_temps_monthly <- SACTN_daily_temps %>% 
+SACTN_monthly <- function(df) {
+  SACTN_split_monthly <- df %>%
   mutate(month = month(date, abbr = T, label = T),
@@
                                 ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
-                                       ifelse(month %in% c("Oct", "Nov", "Dec"),
-                                              "Spring","Error")))))
+                                       ifelse(month %in% c("Oct", "Nov", "Dec"), "Spring","Error")))))
+  return(SACTN_split_monthly)
+}
+
+# SACTN_daily_temps_indexsplit <- SACTN_monthly(df = SACTN_daily_temps_indexsplit)
+# save(SACTN_daily_temps_indexsplit, file = "SACTN_daily_temps_indexsplit.RData")
+```
+
+Find where different sites have overlapping time series within this `SACTN_daily_temps_indexsplit` dataset. A time series with a length of 3years is long enough.
+
+```{r}
+load("~/Documents/Masters_2019/MastersProject/Data/SACTN_daily_temps_indexsplit.RData")
+SACTN_3yr <- SACTN_daily_temps_indexsplit %>% 
+  filter(year(date) %in% seq(1998, 1999, 2000))
+```
+
+```{r}
+ggplot(data = SACTN_3yr , aes(x = date, y = insitu_temp)) +
+  geom_hline(aes(yintercept = mean(insitu_temp)), colour = "salmon") +
+  geom_line() +
+  facet_wrap(~ site, nrow = 4) +
+  theme_bw()
 ```
@@
 
-Here I load the AVHRR CSV dataset. Large dataset and slows down the laptop. Here i split the data into xaa and xab. Here xaa and xab only include temperature collected for two years.
+## MUR dataset
+
+```{r}
+MUR_Doringbaai <- read_csv("Data/MUR_nearest5pixels/MUR_Doringbaai_SST_timeseries_5nearest.csv")
+MUR_Elands_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Elands Bay_SST_timeseries_5nearest.csv")
+MUR_Hondeklipbaai <- read_csv("Data/MUR_nearest5pixels/MUR_Hondeklipbaai_SST_timeseries_5nearest.csv")
+MUR_Lamberts_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Lamberts Bay_SST_timeseries_5nearest.csv")
+MUR_Paternoster <- read_csv("Data/MUR_nearest5pixels/MUR_Paternoster_SST_timeseries_5nearest.csv")
+MUR_Port_Nolloth <- read_csv("Data/MUR_nearest5pixels/MUR_Port Nolloth_SST_timeseries_5nearest.csv")
+MUR_Saldanha_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Saldanha Bay_SST_timeseries_5nearest.csv")
+MUR_St_Helena_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_St Helena Bay_SST_timeseries_5nearest.csv")
+MUR_Yzerfontein <- read_csv("Data/MUR_nearest5pixels/MUR_Yzerfontein_SST_timeseries_5nearest.csv")
+
+MUR_SST <- rbind(MUR_Doringbaai,MUR_Elands_Bay, MUR_Hondeklipbaai,MUR_Lamberts_Bay,MUR_Yzerfontein,
+                 MUR_Paternoster, MUR_Port_Nolloth,MUR_Saldanha_Bay, MUR_St_Helena_Bay) %>%
+  dplyr::rename(site = station)
+
+MUR_SST$date <- (ymd(MUR_SST$date))
+
+save(MUR_SST, file = "Data/MUR_SST.RData")
+```
+
+```{r}
+load("~/Documents/Masters_2019/MastersProject/Data/MUR_SST.RData")
+```
+
+```{r}
+MUR_SST <- MUR_SST %>% 
+  dplyr::rename(temp = nearest1)
+
+match_func <- function(df){
+  match <- SACTN_daily_temps_indexsplit  %>% 
+  left_join(df, by = c("site", "date")) %>% 
+  na.trim()
+  return(match)
+}
+
+insitu_MUR <- match_func(df = MUR_SST)
+```
+
+```{r}
+match_plot <- function(df){
+  plot1 <- df %>% 
+ggplot(aes(x = date, y = temp)) +
+  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
+  geom_line() +
+  facet_wrap(~ site, nrow = 4) +
+  theme_bw()
+  return(plot1)
+}
+
+MUR_plot <- match_plot(df = MUR_SST)
+```
+
+################################
+
+-  Need to plot a line graph with all SST dat and the insitu data on one graph
+
+################################
+
+Here I load the AVHRR CSV dataset.
+
 ```{r}
@@
 #OISST <- as.tibble(fread("~/Documents/Masters_2019/MastersProject/Data/csvavhrr-only-v2-19810901-20180630.csv"))
-##################
-# xaa <- read_csv("Data/xaa.csv",  col_names = c("lon", "lat", "temp", "date"))
-# xab <- read_csv("Data/xab.csv",  col_names = c("lon", "lat", "temp", "date"))
-# combined_OISST <- rbind(xaa, xab)
-# save(combined_OISST, file = "data/combined_OISST.RData")
 ```
@@
 match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
-                          query = as.matrix(upwelling_SACTN_daily_wc[,5:6]), k = 1)
+                          query = as.matrix(upwelling_sites[,5:6]), k = 1)
 
@@
   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
-  mutate(site = upwelling_SACTN_daily_wc$site)
+  mutate(site = upwelling_sites$site)
 ```
@@
 ```{r}
-matched_SST <- OISST_sites %>% 
-  left_join(SACTN_daily_temps_indexsplit, by = c("site", "date")) %>% 
-    na.trim()
-```
-
-## MUR dataset
-
-```{r}
-# MUR_Doringbaai <- read_csv("Data/MUR_nearest5pixels/MUR_Doringbaai_SST_timeseries_5nearest.csv")
-# MUR_Elands_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Elands Bay_SST_timeseries_5nearest.csv")
-# MUR_Hondeklipbaai <- read_csv("Data/MUR_nearest5pixels/MUR_Hondeklipbaai_SST_timeseries_5nearest.csv")
-# MUR_Lamberts_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Lamberts Bay_SST_timeseries_5nearest.csv")
-# MUR_Paternoster <- read_csv("Data/MUR_nearest5pixels/MUR_Paternoster_SST_timeseries_5nearest.csv")
-# MUR_Port_Nolloth <- read_csv("Data/MUR_nearest5pixels/MUR_Port Nolloth_SST_timeseries_5nearest.csv")
-# MUR_Saldanha_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Saldanha Bay_SST_timeseries_5nearest.csv")
-# MUR_St_Helena_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_St Helena Bay_SST_timeseries_5nearest.csv")
-# MUR_Yzerfontein <- read_csv("Data/MUR_nearest5pixels/MUR_Yzerfontein_SST_timeseries_5nearest.csv")
-# 
-# MUR_SST <- rbind(MUR_Doringbaai,MUR_Elands_Bay, MUR_Hondeklipbaai,MUR_Lamberts_Bay,MUR_Yzerfontein,
-#                  MUR_Paternoster, MUR_Port_Nolloth,MUR_Saldanha_Bay, MUR_St_Helena_Bay) %>% 
-#   dplyr::rename(site = station)
-# save(MUR_SST, file = "Data/MUR_SST.RData")
+insitu_OISST <- match_func(df = OISST_sites)
 ```
@@
 ```{r}
-load("~/Documents/Masters_2019/MastersProject/Data/MUR_SST.RData")
+OISST_plot <- match_plot(df = OISST_sites)
 ```
-
-hee1 <- try1 %>% 
-  dplyr::group_by(site, date) %>% 
-  dplyr::summarise(temp = mean(temp, na.rm = TRUE)) %>% 
-  dplyr::mutate(no = seq(1:n())) %>%
-  dplyr::ungroup() %>% 
-  dplyr::select(site, temp, date)
-
-ggplot(data = thee1, aes(x = date, y = temp)) +
-  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
-  geom_line() +
-  facet_wrap(~ site, nrow = 4) +
-  theme_bw() +
-  theme(strip.text.x = element_text(size = 3))

### 2019-01-21 Write_up/Writing/Introduction_outline.Rmd
Updated commit

@@
 date: "13 January 2019"
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 ---
@@
 <will add more here on the different types of resolution used in this thesis>
+
+### In situ data
+
+Having coastal, in situ seawater temperature data is a rare privilege to researchers.
+Most biological and ecological research requires coastal temperature data.
+The South African Coastal Temperature Network (SACTN).
+The availability of the in situ collected coastal temperature data products for several decades and provides a reliable source of accurate coastal seawater temperature data (Smit et al.
+2013).
+SACTN has collected coastal seawater temperature along the South African coastline from as early as 1972, with contributions from a total of seven different organisations and governmental departments.
+At present, this dataset contained temperature of 129 sites along the three distinct coasts of the South African coastline.
+Temperature recordings are still ongoing at many of these sites, but unfortunately, the durations and extents of these are uneve per site.
+With the longest time series in the dataset being that of Gordons Bay, recorded by SAWS.
+Data collected for this region started on 13 September 1972 and concluded on 26 January 2017, with recordings still continuing daily.
+During the 1970s, a total of 11 time series began recording.
+A further 53 entries were added during the 1980s, 34 entries were added during the 1990s, and 18 entries were added during the 2000s.
+
+The uneven and inconsistent nature of the SATCN dataset can be attributed to several factors.
+For instance, these may be as a result of variances in the equipment used by the various organisations and governmental bodies who recorded the data.
+One problematic outcome of this sampling methodology is that due to specific work of the South African Weather Service (SAWS) and Kwazulu Natal Sharks Board (KZNSB), two thirds of the SACTN data that were sampled were obtained with hand-held thermometers and records at a data precision of 0.500 ◦ C.
+However, the current global standard is to make use of underwater temperature recorders (UTRs), which have a data precision of 0.001◦ C (Jarraud, 2008).
+Because of these inconsistencies, this dataset required verification from an independent source before its use within research.
+Smit et al.
+(2013) thoroughly compared the temperature data within the SACTN dataset to those of several remotely-sensed SST products and found bias results.
+Using these in situ collected temperature data along with remotely sensed SST data requires a clear understanding in the characteristics of the contributing data sources in order to understand how accurate, reliable and useful these long term measurements are for use in climate change studies.
+
+It must also be acknowledged that most of the time series within the SACTN dataset do not have meta-data records that meet the international climate change research requirements (Aguilar et al., 2003).
+Specifically uncertainty lies in the drift of the instrument, who was sampling and where the exact sampling locations were collected throughout the collection period.
+We are however certain that all of the thermometer time series were collected by thermometers and that the UTR time series were collected by UTRs.
+Given these issues associated with the mete-data, recent work has been started to develop a nation wide standard.
+
+For this analysis, the seawater temperature data, obtained from the seven different sources, were combined and formatted into standardized comma delineated value (CSV) files which allowed for a fixed methodology to be used across the entire dataset.
+Prior to data analysis, all data points exceeding 35°C and/or below 0°C were removed as these were considered as outliers.
+These data points were then changed to NA (not available) so as to not interfere with analysis.
+All analyses were conducted in R software version 3.4.2 (http://www.r-project.org/).
+The data used within this study and comprehensive script used for data analyses, and production of figures can be found at https://github.com/AmierohAbrahams/MASTERSPROJECT.
 
@@
 
-## Extreme events
+### Extreme events
+
+Extreme thermal events have a range of broad scale definitions, but they can each be classified into two categories, namely, cold-spells (e.g.
+Gunter, 1941; Lirman et al., 2011; Boucek et al., 2016) or heatwaves (e.g.
+Gordon et al., 1988; Stott et al., 2004; Perkins-Kirkpatrick et al., 2016).
+These events occur rapidly, and unexpectedly and they profoundly impact on species diversity, distributions, growth and survival where they take place.
+These events are often undetected but studies have found that extreme events have occurred as recently as approximately four decades ago.
+Jentsch et al.
+(2007) were the first to discuss this mode of climate change and brought it to focus.
+Steps have since been taken towards understanding the drivers, impacts and consequences of these events.
+Subsequently, research has predicted an increase in the occurrence of these extreme events as a result of long-term warming of the planet via the process of global warming (REFERENCE).
+Extreme events are often associated with toxic algal blooms, shifts in species distribution, and the bleaching of coral reefs (McCabe et al., 2016; Harley and Paine, 2009; Schoepf et al., 2015).
+
+Little is known about extreme events occurring within the South African region.
+Very little South African researchers spend their time and careers on studyig extreme events within this region.
+During the 1980s, a study done by Birkett and Cook was the only study done where they anaylsed the impact of extreme warming events on the life cycle of white muscles,Donax serra Röding.
+However, more recently studies done by Schlegel et al, (2017,2018) examined the relationship between the nearshore and offshore occurances of marine heatwaves and cold-spells as well as the influence of predominant atmosphetic and oceanic patterns during Coastal Marine Heatwaves.
+It is now becoming accepted that extreme events are a threat to several ecosystem (Thompson et al., 2013).
 

### 2019-01-21 Methods.Rmd
updated

@@
 
-save(SACTN_daily_temps_indexsplit, file = "SACTN_daily_temps_indexsplit.RData")
+# save(SACTN_daily_temps_indexsplit, file = "Data/SACTN_daily_temps_indexsplit.RData")
 ```
@@
 
-# SACTN_daily_temps_indexsplit <- SACTN_monthly(df = SACTN_daily_temps_indexsplit)
-# save(SACTN_daily_temps_indexsplit, file = "SACTN_daily_temps_indexsplit.RData")
+SACTN_split <- SACTN_monthly(df = SACTN_daily_temps_indexsplit)
+save(SACTN_split, file = "SACTN_split.RData")
 ```
@@
 ```{r}
-load("~/Documents/Masters_2019/MastersProject/Data/SACTN_daily_temps_indexsplit.RData")
-SACTN_3yr <- SACTN_daily_temps_indexsplit %>% 
+load("~/Documents/Masters_2019/MastersProject/Data/SACTN_split.RData")
+
+SACTN_3yr <- SACTN_split %>% 
   filter(year(date) %in% seq(1998, 1999, 2000))
@@
 match_func <- function(df){
-  match <- SACTN_daily_temps_indexsplit  %>% 
+  match <- SACTN_split  %>% 
   left_join(df, by = c("site", "date")) %>% 

### 2019-01-21 Methods.Rmd
updated...

@@
 
-SACTN_split <- SACTN_monthly(df = SACTN_daily_temps_indexsplit)
-save(SACTN_split, file = "SACTN_split.RData")
+# SACTN_split <- SACTN_monthly(df = SACTN_daily_temps_indexsplit)
+# save(SACTN_split, file = "SACTN_split.RData")
 ```
 
-Find where different sites have overlapping time series within this `SACTN_daily_temps_indexsplit` dataset. A time series with a length of 3years is long enough.
+Find where different sites have overlapping time series within this `SACTN_split` dataset.
+A time series with a length of 3years is long enough.
+The only problem is that the SST data do not overlap during these years (1998, 1999, 2000)
 
@@
 Once the SST data were downloaded with a netCDF format i needed to convert this netCDfs to CSV
+
+Here i am suppose to overlap the MUR data with the insitu collected data but the insitu collected data is from 199-2000
+- find other overlapping more recent time with the in situ data?
+-
 
@@
 
-OISSTDir <- "~/Documents/Masters_2019/MastersProject/Data"
+OISSTDir <- "~/Documents/Masters_2019/Data_SST"
 OISST <- fread(paste0(OISSTDir, "/csvavhrr-only-v2-19810901-20180630.csv"),

### 2019-01-22 Methods.Rmd
some_changes

@@
 I then exclude the sites that do not fall within the upwelling cell.
-Once those sites were removed, we have a total of 9 sites.
-These are: Port Nolloth, Hondeklip baai, Doring baai, Lamberts Bay, Elands Bay, St Helena Bay, Paternoster, Saldanna and Yzerfontein.
+Once those sites were removed, we have a total of 4 sites.
+These are: Port Nolloth, Lamberts Bay, Saldanna and Yzerfontein.
 
@@
 
+# Here I find a time series with the data occuring in the most recent years
+upwelling_sites <- upwelling_sites[c(-2, -3, -5, -6, -8),] 
+# save(upwelling_sites, file = "Data/upwelling_sites.RData")
+
 SACTN_daily_temps <- SACTN_daily_v4.2 %>% 
@@
   filter(index %in% upwelling_sites$index)
-
 # save(SACTN_daily_temps, file = "SACTN_daily_temps.RData")
@@
   dplyr::rename(insitu_temp = temp)
-
 # save(SACTN_daily_temps_indexsplit, file = "Data/SACTN_daily_temps_indexsplit.RData")
@@
 SACTN_3yr <- SACTN_split %>% 
-  filter(year(date) %in% seq(1998, 1999, 2000))
+  filter(year(date) %in% seq(2015, 2016))
 ```
@@
   geom_line() +
-  facet_wrap(~ site, nrow = 4) +
+  facet_wrap(~ site, nrow = 2) +
   theme_bw()
@@
 ```{r}
-MUR_Doringbaai <- read_csv("Data/MUR_nearest5pixels/MUR_Doringbaai_SST_timeseries_5nearest.csv")
-MUR_Elands_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Elands Bay_SST_timeseries_5nearest.csv")
-MUR_Hondeklipbaai <- read_csv("Data/MUR_nearest5pixels/MUR_Hondeklipbaai_SST_timeseries_5nearest.csv")
 MUR_Lamberts_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Lamberts Bay_SST_timeseries_5nearest.csv")
-MUR_Paternoster <- read_csv("Data/MUR_nearest5pixels/MUR_Paternoster_SST_timeseries_5nearest.csv")
 MUR_Port_Nolloth <- read_csv("Data/MUR_nearest5pixels/MUR_Port Nolloth_SST_timeseries_5nearest.csv")
 MUR_Saldanha_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Saldanha Bay_SST_timeseries_5nearest.csv")
-MUR_St_Helena_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_St Helena Bay_SST_timeseries_5nearest.csv")
 MUR_Yzerfontein <- read_csv("Data/MUR_nearest5pixels/MUR_Yzerfontein_SST_timeseries_5nearest.csv")
 
-MUR_SST <- rbind(MUR_Doringbaai,MUR_Elands_Bay, MUR_Hondeklipbaai,MUR_Lamberts_Bay,MUR_Yzerfontein,
-                 MUR_Paternoster, MUR_Port_Nolloth,MUR_Saldanha_Bay, MUR_St_Helena_Bay) %>%
+MUR_SST <- rbind(MUR_Lamberts_Bay,MUR_Yzerfontein, MUR_Port_Nolloth,MUR_Saldanha_Bay) %>%
   dplyr::rename(site = station)
@@
 match_func <- function(df){
-  match <- SACTN_split  %>%  
+  match <- SACTN_3yr  %>%  
   left_join(df, by = c("site", "date")) %>% 
@@
 
-MUR_plot <- match_plot(df = MUR_SST)
+MUR_plot <- match_plot(df = insitu_MUR)
 ```
 
-################################
-
--  Need to plot a line graph with all SST dat and the insitu data on one graph
-
-################################
-
-Here I load the AVHRR CSV dataset.
-
+Here I load the AVHRR OISST CSV dataset. 
 ```{r}
@@
 ```{r}
-load("~/Documents/MastersProject/Data/combined_OISST_sites.RData")
+# load("~/Documents/MastersProject/Data/combined_OISST_sites.RData")
 ```
@@
 ## Matching the Insitu data with the OISST SST data
-- Here there are many NA values because the OISST have many values collected per day where as the insitu temp only has 1
 
@@
 ```
+
+##

### 2019-01-22 Write_up/Writing/Introduction_outline.Rmd
refs_fixed

@@
 Globally, there are four major eastern boundary upwelling systems (EBUS).
-These are the Benguela Current off south-western Africa, the Canary Current off north-western Africa including its northern extension off the Iberian Peninsula of south-western Europe, the Peru-Humboldt Current off western South America, and the California Current off the western continental USA and north-western Mexico (Bakun et al, 2015; Pauly and Christensen, 1995).
+These are the Benguela Current off south-western Africa, the Canary Current off north-western Africa including its northern extension off the Iberian Peninsula of south-western Europe, the Peru-Humboldt Current off western South America, and the California Current off the western continental USA and north-western Mexico [@bakun ; @pauly].
 Each of these systems are characterized as vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans.
-EBUS cover only 1% of the world’s ocean surface, but despite these are among the most productive regions of the world (Pauly and Christensen, 1995).
-This is because these systems are subject to coastal upwelling, which is defined as the process whereby cold, nutrient rich, high CO2 concentrated, low pH, and low oxygen waters are pushed to the surface as a result of alongshore winds interacting with the earth’s rotation (Huyer 1983).
+EBUS cover only 1% of the world’s ocean surface, but despite these are among the most productive regions of the world [@pauly].
+This is because these systems are subject to coastal upwelling, which is defined as the process whereby cold, nutrient rich, high CO2 concentrated, low pH, and low oxygen waters are pushed to the surface as a result of alongshore winds interacting with the earth’s rotation [@huyer].
 As a result of upwelling, diverse marine fauna and flora are able to thrive in these areas because of the abundance of nutrients and favourable temperatures.
@@
 Several people have tried to model this using different datasets of SST and wind pattern data at various resolution and initially found that increases in gases, such as CO^2 causes a reduction in trade winds, and therefore a reduction in upwelling.
-However, these models were based on coarse data and when higher resolutions are used, such as for the Benguela EBUS, the models have shown that upwelling will intensify in accordance to the upwelling intensification hypotheses (eg, Hsieh and Boer 1992, Mote and Mantua 2002, Lima and Wethey’s 2012).
+However, these models were based on coarse data and when higher resolutions are used, such as for the Benguela EBUS, the models have shown that upwelling will intensify in accordance to the upwelling intensification hypotheses [@hsieh; @mote; @lima].
 
@@
 The duration and intensity of coastal upwelling are known to have a critical role in the phenology of key marine ecosystem processes.
-Similarly, changes in the characteristics of upwelling are shown to have cause large disturbances to ecosystems at several trophic levels (Wang et al., 2015).
-Cold, nutrient rich, low pH, and low oxygen upwelled waters fuels primary production as a consequence of the development of phytoplankton blooms (Ryther 1969, Cushing 1971, Payne 1987, McGillicuddy et al., 1998; Oschlies and Garçon, 1998).
+Similarly, changes in the characteristics of upwelling are shown to have cause large disturbances to ecosystems at several trophic levels [@wang].
+Cold, nutrient rich, low pH, and low oxygen upwelled waters fuels primary production as a consequence of the development of phytoplankton blooms [@Ryther; @Cushing; @mcGillicuddy; @oschlies].
 In these upwelling systems, the high productivity supports vast and diverse marine populations.
 Mesoscale and sub-mesoscale oceanic movements are key in managing the exchange and transportation of organic resources and marine organisms from the coastal zone to the open ocean.
-Most of the world’s major fishing industries are dependent on upwelling, because this process brings nutrient rich waters into the euphotic zone, where it forms the foundation of a localized food web that ecologically differs from surrounding, non-upwelled regions (Bosman et al., 1987; Barshis et al., 2011).
+Most of the world’s major fishing industries are dependent on upwelling, because this process brings nutrient rich waters into the euphotic zone, where it forms the foundation of a localized food web that ecologically differs from surrounding, non-upwelled regions [@bosman; @barshis].
 Whilst nutrients play a primary role in the survival of these systems, the cool temperatures of these waters are critical as well.
@@
 Upwelling occurring along the Cape south coast is proven to have less of an economic importance in contrast to the west coast.
-Inshore regions and embayments that are protected from wind action may be subjected to intense episodes of primary production (Graham and Largier, 1997).
+Inshore regions and embayments that are protected from wind action may be subjected to intense episodes of primary production [@graham].
 Vander Wouder et al.
 (2006) found that these sites represent the highest levels of chlorophyll concentration.
-Biotic systems within the Benguela upwelling system may be negatively influenced by the changes in oxygen levels given their adaption to natural variable conditions (Capone and Hutchins, 2013).
+Biotic systems within the Benguela upwelling system may be negatively influenced by the changes in oxygen levels given their adaption to natural variable conditions [@capone].
 
@@
 Regime shifts are often defined as a shift from one stable ecosystem state to another.
-This process affects several species within a short period of time (deYoung et al., 2004; Cury and Shannon, 2004; Jarre et al., 2006, Blamey 2012).
-Regime shifts are thought to be driven by abiotic processes such as climate change (Polovina, 2005, Blamey 2012).
+This process affects several species within a short period of time [@cury; @jarre; @blamey).
+Regime shifts are thought to be driven by abiotic processes such as climate change [@polovina; @blamey].
 The southern Benguela upwelling system supports a wide variety of commercially-exploited fish species and exhibits substantial variability in both oceanographic and biological components.
 Over the past 3 three decades, a number of species occurring within this region have undergone shifts in their distributions and Roy et al.
-(2007) linked these shifts to cooling waters along the inner shelf of the Agulhas bank on the south coast, east of Cape Agulhas (Blamey, 2012).
+(2007) linked these shifts to cooling waters along the inner shelf of the Agulhas bank on the south coast, east of Cape Agulhas [@blamey].
 This coast is associated with high levels of harmful toxic algal blooms, which negatively impact commercial and recreational interests along the coastal region.
 These harmful blooms are often attributed to flagellated species with the red tide representing dinoflagellate blooms (Pitcher et al.
-1993A, Pitcher 1998).
+1993A, @pitcher1998].
 Formations of red tides are often related to the prevailing winds of the southern Benguela which govern most hydrodynamic processes on the continental shelf.
@@
 Throughout region surrounding the Benguela EBUS, the atmosphere is dominated by anticyclonic high pressure cells with quasi-stationary positions.
-The South Atlantic Ocean High is situated along the west, drawing cool, dry air onto the west of the subcontinent (Van Heerden and Hurry, 1998).
+The South Atlantic Ocean High is situated along the west, drawing cool, dry air onto the west of the subcontinent [@van].
 Solar heating during summer may result in the development of low pressure cells, known as heat lows.
 These cells are absent during winter time (Tyson and Preston-Whyte, 2000).
-When the anticyclones shift north, during winter months, the cold westerlies have a substantial impact on the weather of the southern tip of the South African subcontinent (Van Heerden and Hurry, 1998).
-The atmospheric temperatures along the coastline are largely influenced by the Benguela Current (Van Heerden and Hurry, 1998).
-However, upwelling systems may be forced by various atmospheric conditions (Lutjeharms et al., 2003; Roberts, 2005; Hutchings et al., 2009).
+When the anticyclones shift north, during winter months, the cold westerlies have a substantial impact on the weather of the southern tip of the South African subcontinent [@van].
+The atmospheric temperatures along the coastline are largely influenced by the Benguela Current [@van].
+However, upwelling systems may be forced by various atmospheric conditions [@Lutjeharms; @roberts; @hutchings].
 
@@
 The South African Coastal Temperature Network (SACTN).
-The availability of the in situ collected coastal temperature data products for several decades and provides a reliable source of accurate coastal seawater temperature data (Smit et al.
-2013).
+The availability of the in situ collected coastal temperature data products for several decades and provides a reliable source of accurate coastal seawater temperature data [@smit].
 SACTN has collected coastal seawater temperature along the South African coastline from as early as 1972, with contributions from a total of seven different organisations and governmental departments.
@@
 One problematic outcome of this sampling methodology is that due to specific work of the South African Weather Service (SAWS) and Kwazulu Natal Sharks Board (KZNSB), two thirds of the SACTN data that were sampled were obtained with hand-held thermometers and records at a data precision of 0.500 ◦ C.
-However, the current global standard is to make use of underwater temperature recorders (UTRs), which have a data precision of 0.001◦ C (Jarraud, 2008).
+However, the current global standard is to make use of underwater temperature recorders (UTRs), which have a data precision of 0.001◦ C [@jarraud].
 Because of these inconsistencies, this dataset required verification from an independent source before its use within research.
@@
 
-It must also be acknowledged that most of the time series within the SACTN dataset do not have meta-data records that meet the international climate change research requirements (Aguilar et al., 2003).
+It must also be acknowledged that most of the time series within the SACTN dataset do not have meta-data records that meet the international climate change research requirements [@aguilar].
 Specifically uncertainty lies in the drift of the instrument, who was sampling and where the exact sampling locations were collected throughout the collection period.
@@
 
-### Global warming and climate change:
+### Climate change:
 
@@
 As the human population grows, the rate of global warming increases, resulting in profound effects on marine and other ecosystems.
-Climbing temperatures create a host of additional changes to marine systems, such as rising sea levels, increased ocean stratification, decreased sea-ice extent, and altered patterns of ocean circulation and precipitation (Doney et al., 2011).
+Climbing temperatures create a host of additional changes to marine systems, such as rising sea levels, increased ocean stratification, decreased sea-ice extent, and altered patterns of ocean circulation and precipitation [@doney].
 Furthermore, direct effects on ocean temperature changes and chemistry may alter the functioning and behaviour or marine organisms.
@@
 
-Significant research studies accept the climate change and variability have great impacts on marine ecosystems (Stenseth et al.
-2002, Harley et al.
-2006, Hoegh-Guldberg & Bruno 2010).
-The Intergovernmental Panel on Climate Change (IPCC) predicts future projections with a mean increase in air temperature of 1.1°C by the 2020’s and 1.8°C by the 2040’s as a result of greenhouse gas concentrations (Solomon et al.
-2007).
+Significant research studies accept the climate change and variability have great impacts on marine ecosystems [@Stenseth; @harley; @hoegh].
+The Intergovernmental Panel on Climate Change (IPCC) predicts future projections with a mean increase in air temperature of 1.1°C by the 2020’s and 1.8°C by the 2040’s as a result of greenhouse gas concentrations [@solomon].
 A large amount of this heat will be absorbed by the world’s ocean.
@@
 Global warming represents distinct warming and cooling periods throughout the 20th and 21st century (IPCC, 2007).
-Global warming however is not evenly distributed across the global ocean thus presenting regions warming much faster or slower than then the global average (Levitus et al., 2000; Palttridge and Woodruff, 1981; Santos 2012).
-In particular, the Atlantic Ocean contributes most to the increase of the heat content (Nerem et al., 1999; Levitus et al., 2000; Strong et al., 2000).
+Global warming however is not evenly distributed across the global ocean thus presenting regions warming much faster or slower than then the global average [@levitus; @santos].
+In particular, the Atlantic Ocean contributes most to the increase of the heat content [@nerem; @levitus].
 
 As a result of climate change, research studies provide significant evidence on the increase in warming of the ocean surface.
-This warming slowly propagates down through the water column and in turn increases the thermal stratification (Gruber, 2011).
-Strong stratification limits the depth at which water is upwelled and hence controls the amount of nutrients brought up (Chhak and Di Lorenzo, 2007; Jacox and Edwards, 2011; Jacox et al., 2015).
+This warming slowly propagates down through the water column and in turn increases the thermal stratification [@gruber].
+Strong stratification limits the depth at which water is upwelled and hence controls the amount of nutrients brought up [@chhak; @jacox2011; @jocox2016].
 The interplay between upwelling winds and stratification determine the amount of nutrients present within the system and thus is strongly linked to biological productivity.
-Globally, as a result of climate change, stratification of the water-column between 0 and 200 m has increased by an estimated 4% between 1971 to 2010 (Stocker et al., 2013).
+Globally, as a result of climate change, stratification of the water-column between 0 and 200 m has increased by an estimated 4% between 1971 to 2010 [@Stocker].
 Increased ocean stratification as a result of surface warming may reduce the ventilation of deep-water masses.
-As the deep water is being up welled to the surface, decreases in oxygen and pH levels are predicted along with an increase in nutrient levels (Rykaczewski and Dunne, 2010).
-Generally, regional ocean climate oscillations like the Benguela Niño for the Benguela Current System impose considerable variability on temperatures and complicate attempts to assess long-term trends (Demarcq, 2009).
-It is predicted that these climate oscillations may increase the variance with climate change (Timmermann et al., 1999; Kuzmina et al., 2005; Sydeman et al., 2013; Cai et al., 2015) which will further result in unusual trends.
+As the deep water is being up welled to the surface, decreases in oxygen and pH levels are predicted along with an increase in nutrient levels [@Rykaczewski).
+Generally, regional ocean climate oscillations like the Benguela Niño for the Benguela Current System impose considerable variability on temperatures and complicate attempts to assess long-term trends [@demarcq].
+It is predicted that these climate oscillations may increase the variance with climate change [@timmermann; @kuzmina; @sydeman] which will further result in unusual trends.
 
 As previously mentioned, upwelling winds in EBUSs are generated by pressure differences between continental thermal lows and oceanic highs.
-As warming intensifies, pressure gradients between land and sea increases, resulting in a greater intensity of upwelling winds (Bakun et al.
-2010).
-In rare cases, projected upwelling increases may overcome the countervailing effects of upper-ocean warming and stratification to cause regional cooling (Auad et al.2006).
+As warming intensifies, pressure gradients between land and sea increases, resulting in a greater intensity of upwelling winds [@bakun2010].
+In rare cases, projected upwelling increases may overcome the countervailing effects of upper-ocean warming and stratification to cause regional cooling [@auad].
 Research studies reveal declines in ocean oxygen levels as a result of ocean warming and reduced ventilation from stratification and circulation changes (Keeling et al.
 2010).
-Similarly, low oxygen levels are also possible because of shifts in the composition of sinking organic material (Hofmann & Schellnhuber 2009, Doney et al., 2011).
+Similarly, low oxygen levels are also possible because of shifts in the composition of sinking organic material [@hofmann; @doney).
 
 ## Problem statement:
+
 With global warming continuing at a rapid pace the amount of greenhouse gases within the atmosphere are going to increase in the future.
@@
 
-## Study Area:
+## Benguela :
 
@@
 The Benguela EBUS is unique in that it is bounded at both the equator ward and pole ward ends by warm water regimes.
-These currents are forced locally by the wind stress field off South-West Africa (Nelson and Hutchings 1983; Fennel 1999).
-This upwelling system stretches along the south-western coast of Africa from southern Angola to Cape Agulhas (Cole, 1999).
-Water is coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S (Lutjeharms and Meeuwis, 1987).
-The main upwelling period in the Benguela north of 25°S extends mainly from March to November, peaking during August (Shannon, 1985), but upwelling persists throughout the year.
+These currents are forced locally by the wind stress field off South-West Africa [@nelson ; @fennel].
+This upwelling system stretches along the south-western coast of Africa from southern Angola to Cape Agulhas [@cole, 1999).
+Water is coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S [@Lutjeharms].
+The main upwelling period in the Benguela north of 25°S extends mainly from March to November, peaking during August [@shannon1985], but upwelling persists throughout the year.
 Upwelling season in the northern Benguela extends between September and March, with very little upwelling taking place during winter (June-August).
-In the central Benguela and also at Cape Frio upwelling is continuous throughout the year (Shannon, 1985).
+In the central Benguela and also at Cape Frio upwelling is continuous throughout the year [@shannon1985].
 The cold Benguela waters are bound in the south by the warm Agulus retroflexion region and in the north by the southward flowing Angola current.
-The flow of the Benguela is often considered to be topographically guided (Nelson and Hutchings, 1983; Barange et al.
-1991a and 1991b).
-The Benguela system is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region within the Namibian region (Shannon 1985; Cole 1999).
+The flow of the Benguela is often considered to be topographically guided [@nelson; @barange; @barange.
+The Benguela system is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region within the Namibian region [@shannon; Cole 1999).
 
@@
 The result is that there is a pronounced seasonal cycle in SST in the north while in the south there is no clear SST cycle.
-The seasonal warming occurring during the late summer season off northern and central Namibia is a regular occurrence and is associated with warmer, higher salinity water of Angolan origin into the region (O'Toole, 1980; Shannon, 1985).
-There have been a number of warm and cool periods documented in the northern and southern Benguela (e.g.
-Stander and De Decker, 1969; Walker et al.1984; Shannon et al., 1984; Boyd and Thomas, 1984; Shelton et al.
-1985; McLain et al.: 1985).
+The seasonal warming occurring during the late summer season off northern and central Namibia is a regular occurrence and is associated with warmer, higher salinity water of Angolan origin into the region [@o; @shannon].
+There have been a number of warm and cool periods documented in the northern and southern Benguela [e.g.
+@stander; @walker; @shannon1984; @boyd; @brainard].
 Walter (1937) found a major climatic event in the northern Benguela region of anonymously warm water.
 This unusual warm water may have resulted due to reduced upwelling or an increase southward advection of warm water.
-Another warming event occurred during the summer and autumn months of 1984 (Shannon et al, 1986).
+Another warming event occurred during the summer and autumn months of 1984 [@shannon1986].
 The presence of this anomalous warm water during 1984 followed two cooler than normal years in the northern Benguela.
@@
 (1986) illustrated the warm (21-23°C) water near Walvis Bay as a large warm water patch offshore west of Luderitz.
-The absence of upwelling between Walvis Bay and north of Luderitz is evident however, the upwelling centre in the Benguela system (Shannon, 1985), and upwelling was minimal.
+The absence of upwelling between Walvis Bay and north of Luderitz is evident however, the upwelling centre in the Benguela system [@shannon1986], and upwelling was minimal.
 Interestingly, in the southern Benguela region, there was no evidence of any unusual patterns in temperature distributions.
@@
 
-Northward, the Benguela Current Upwelling system is linked by the Angola Benguela Front which converge between the Benguela current and the Benguela coastal current as well as the warm saline waters of the Angola Current (Peterson and Stramma, 1991).
+Northward, the Benguela Current Upwelling system is linked by the Angola Benguela Front which converge between the Benguela current and the Benguela coastal current as well as the warm saline waters of the Angola Current [@peterson].
 Research done by Yamagata and Iizuka (1995) discovered slight indication that coastal Kelvin waves, originating from equatorial Kelvin waves, carry the down welling process southward.
 Southward, the coastal upwelling system is bounded by the Agulhas Bank.
-South of the African coast the warm Agulhas Current, western boundary current, flows in from the Indian Ocean and returns eastwards while generating warm eddies that spin off in a north-westerly direction in the Benguela current (Duncombe, 1991).
+South of the African coast the warm Agulhas Current, western boundary current, flows in from the Indian Ocean and returns eastwards while generating warm eddies that spin off in a north-westerly direction in the Benguela current [@duncombe].
 The region surrounding the Benguela experiences a persistent along shore wind.
@@
 Continental shelf bathymetry and upwelling favourable winds provide a large-scale upwelling mechanism in the southern Benguela, whereas local topography and meteorology create an alternating pattern of active and passive upwelling circulations along the coast.
-Winds occurring within the southern regions represent a seasonal variation, reaching a maximum during spring and summer months (Boyd 1987; Shannon and Nelson 1996; Fennel 1999).
-Winds occurring within this region tend to show increasing patterns away from the coast whereas, winds occurring in the northern region exhibits relatively less seasonal variation (Shannon 1985.
-Pitcher et al.
-1992).
+Winds occurring within the southern regions represent a seasonal variation, reaching a maximum during spring and summer months [@boyd; @shannon1996; @fennel].
+Winds occurring within this region tend to show increasing patterns away from the coast whereas, winds occurring in the northern region exhibits relatively less seasonal variation [@shannon1985; @pitcher).
 Upwelling occurring along the South Coast of the Cape Province within South Africa, specifically along the coastline is generated by local wind.
@@
 
-Extreme thermal events have a range of broad scale definitions, but they can each be classified into two categories, namely, cold-spells (e.g.
-Gunter, 1941; Lirman et al., 2011; Boucek et al., 2016) or heatwaves (e.g.
-Gordon et al., 1988; Stott et al., 2004; Perkins-Kirkpatrick et al., 2016).
+Extreme thermal events have a range of broad scale definitions, but they can each be classified into two categories, namely, cold-spells [e.g.
+@gunter; @lirman; @boucek] or heatwaves (e.g.
+[@gordon; @stott; @perkins].
 These events occur rapidly, and unexpectedly and they profoundly impact on species diversity, distributions, growth and survival where they take place.
@@
 Subsequently, research has predicted an increase in the occurrence of these extreme events as a result of long-term warming of the planet via the process of global warming (REFERENCE).
-Extreme events are often associated with toxic algal blooms, shifts in species distribution, and the bleaching of coral reefs (McCabe et al., 2016; Harley and Paine, 2009; Schoepf et al., 2015).
+Extreme events are often associated with toxic algal blooms, shifts in species distribution, and the bleaching of coral reefs [@mccabe; @harley; @schoepf].
 
@@
 However, more recently studies done by Schlegel et al, (2017,2018) examined the relationship between the nearshore and offshore occurances of marine heatwaves and cold-spells as well as the influence of predominant atmosphetic and oceanic patterns during Coastal Marine Heatwaves.
-It is now becoming accepted that extreme events are a threat to several ecosystem (Thompson et al., 2013).
+It is now becoming accepted that extreme events are a threat to several ecosystem [@thompson].
 
 ### Cold spells
+
 - Definition

### 2019-01-22 Write_up/Writing/Introduction_outline.Rmd
minor_edits_refs

@@
 Similarly, changes in the characteristics of upwelling are shown to have cause large disturbances to ecosystems at several trophic levels [@wang].
-Cold, nutrient rich, low pH, and low oxygen upwelled waters fuels primary production as a consequence of the development of phytoplankton blooms [@Ryther; @Cushing; @mcGillicuddy; @oschlies].
+Cold, nutrient rich, low pH, and low oxygen upwelled waters fuels primary production as a consequence of the development of phytoplankton blooms [@ryther; @Cushing; @mcGillicuddy; @oschlies].
 In these upwelling systems, the high productivity supports vast and diverse marine populations.
@@
 Solar heating during summer may result in the development of low pressure cells, known as heat lows.
-These cells are absent during winter time (Tyson and Preston-Whyte, 2000).
+These cells are absent during winter time [@tyson].
 When the anticyclones shift north, during winter months, the cold westerlies have a substantial impact on the weather of the southern tip of the South African subcontinent [@van].
@@
 Due to the distance at which these SST are being collected, certain coarseness within the datasets are unavoidable.
-Satellite SST products undergo a range of quality control processes that are well researched and tested (Reynolds and Smith, 1994; Brown et al., 1999; Martin et al., 2012), making their use more reliable.
+Satellite SST products undergo a range of quality control processes that are well researched and tested [@reynolds], making their use more reliable.
 However, Smit et al.
@@
 
-Significant research studies accept the climate change and variability have great impacts on marine ecosystems [@Stenseth; @harley; @hoegh].
+Significant research studies accept the climate change and variability have great impacts on marine ecosystems [@stenseth; @harley; @hoegh].
 The Intergovernmental Panel on Climate Change (IPCC) predicts future projections with a mean increase in air temperature of 1.1°C by the 2020’s and 1.8°C by the 2040’s as a result of greenhouse gas concentrations [@solomon].
@@
 2010).
-Similarly, low oxygen levels are also possible because of shifts in the composition of sinking organic material [@hofmann; @doney).
+Similarly, low oxygen levels are also possible because of shifts in the composition of sinking organic material [@hofmann; @doney].
 
@@
 Extreme thermal events have a range of broad scale definitions, but they can each be classified into two categories, namely, cold-spells [e.g.
-@gunter; @lirman; @boucek] or heatwaves (e.g.
-[@gordon; @stott; @perkins].
+@gunter; @lirman; @boucek] or heatwaves [e.g.
+@gordon; @stott; @perkins].
 These events occur rapidly, and unexpectedly and they profoundly impact on species diversity, distributions, growth and survival where they take place.

### 2019-01-23 Methods.Rmd
testing bits OISST/MUR

@@
   dplyr::rename(insitu_temp = temp)
+
 # save(SACTN_daily_temps_indexsplit, file = "Data/SACTN_daily_temps_indexsplit.RData")
+```
+
+This code is testing out the wavelet analyses for the in situ temperature data
+
+```{r}
+SACTN_wavelet <- SACTN_daily_temps_indexsplit %>% 
+  filter(year(date) %in% seq(2015, 2016))
+
+SACTN_wavelet.d <- SACTN_wavelet %>% 
+  dplyr::group_by(site, date) %>% 
+  dplyr::summarise(temp = mean(insitu_temp, na.rm = TRUE)) %>% 
+  dplyr::mutate(no = seq(1:n())) %>%
+  dplyr::ungroup() %>% 
+  dplyr::select(site, no, temp, date)
+
+ggplot(data = SACTN_wavelet, aes(x = date, y = insitu_temp)) +
+  geom_hline(aes(yintercept = mean(insitu_temp)), colour = "salmon") +
+  geom_line() +
+  facet_wrap(~ site, nrow = 4) +
+  theme_bw()
+
+auto.arima(dplyr::filter(SACTN_wavelet, site == "Port Nolloth")[, "insitu_temp"], max.p = 3, max.q = 3, stationary = FALSE, 
+           seasonal = FALSE)
+
+PN <- SACTN_wavelet.d %>% 
+  dplyr::filter(site == "Port Nolloth") %>% 
+  dplyr::select(temp, -date, -site) %>% 
+  dplyr::mutate(no = seq(1:n())) %>% 
+  dplyr::select(no, temp)
 ```
@@
 SACTN_3yr <- SACTN_split %>% 
-  filter(year(date) %in% seq(2015, 2016))
-```
-
-```{r}
-ggplot(data = SACTN_3yr , aes(x = date, y = insitu_temp)) +
-  geom_hline(aes(yintercept = mean(insitu_temp)), colour = "salmon") +
-  geom_line() +
-  facet_wrap(~ site, nrow = 2) +
-  theme_bw()
+  filter(year(date) %in% seq(2015, 2016)) %>% 
+  drop_na() # Removing NA values within the dataset
 ```
@@
 Once the SST data were downloaded with a netCDF format i needed to convert this netCDfs to CSV
-
-Here i am suppose to overlap the MUR data with the insitu collected data but the insitu collected data is from 199-2000
-- find other overlapping more recent time with the in situ data?
--
 
@@
 
-save(MUR_SST, file = "Data/MUR_SST.RData")
+# save(MUR_SST, file = "Data/MUR_SST.RData")
 ```
@@
 load("~/Documents/Masters_2019/MastersProject/Data/MUR_SST.RData")
+
+MUR_SST <- MUR_SST %>% 
+  drop_na()
 ```
@@
 
-insitu_MUR <- match_func(df = MUR_SST)
+insitu_MUR <- match_func(df = MUR_SST) %>%
+  drop_na()
 ```
@@
 ```{r}
-ggplot(data = OISST_match, aes(x = date, y = temp)) +
+ggplot(data = OISST_match, aes(x = date, y = temp_OISST)) +
   geom_line() +
@@
 
-save(OISST_sites, file = "Data/OISST_sites.RData")
+# save(OISST_sites, file = "Data/OISST_sites.RData")
 ```
@@
 ```{r}
-# load("~/Documents/MastersProject/Data/combined_OISST_sites.RData")
+load("~/Documents/MastersProject/Data/OISST_sites.RData")
+
+OISST_sites <- OISST_sites %>% 
+  dplyr::rename(temp = temp_OISST)
 ```

### 2019-01-23 Methods.Rmd
working_bits

@@
 
-This code is testing out the wavelet analyses for the in situ temperature data
-
-```{r}
-SACTN_wavelet <- SACTN_daily_temps_indexsplit %>% 
-  filter(year(date) %in% seq(2015, 2016))
-
-SACTN_wavelet.d <- SACTN_wavelet %>% 
-  dplyr::group_by(site, date) %>% 
-  dplyr::summarise(temp = mean(insitu_temp, na.rm = TRUE)) %>% 
-  dplyr::mutate(no = seq(1:n())) %>%
-  dplyr::ungroup() %>% 
-  dplyr::select(site, no, temp, date)
-
-ggplot(data = SACTN_wavelet, aes(x = date, y = insitu_temp)) +
-  geom_hline(aes(yintercept = mean(insitu_temp)), colour = "salmon") +
-  geom_line() +
-  facet_wrap(~ site, nrow = 4) +
-  theme_bw()
-
-auto.arima(dplyr::filter(SACTN_wavelet, site == "Port Nolloth")[, "insitu_temp"], max.p = 3, max.q = 3, stationary = FALSE, 
-           seasonal = FALSE)
-
-PN <- SACTN_wavelet.d %>% 
-  dplyr::filter(site == "Port Nolloth") %>% 
-  dplyr::select(temp, -date, -site) %>% 
-  dplyr::mutate(no = seq(1:n())) %>% 
-  dplyr::select(no, temp)
-```
-
 Here I add monthly and yearly column as well as a seasonal column.
@@
   drop_na() # Removing NA values within the dataset
+```
+
+```{r}
+ggplot(data = SACTN_3yr, aes(x = date, y = insitu_temp, colour = site)) +
+geom_line(aes(group = site)) +
+labs(x = "", y = "Temperature (°C)") +
+theme(axis.text.x = element_text(angle = 45)) +
+theme_linedraw() +
+theme(legend.position = "top")
 ```
@@
   drop_na()
+
+insitu_MUR <- match_func(df = MUR_SST) %>%
+  drop_na()
 ```
@@
 
+```{r}
+ggplot(data = insitu_MUR, aes(x = date, y = temp, colour = site)) +
+geom_line(aes(group = site)) +
+labs(x = "", y = "Temperature (°C)") +
+theme(axis.text.x = element_text(angle = 45)) +
+theme_linedraw() +
+theme(legend.position = "top")
+```
+
 Here I load the AVHRR OISST CSV dataset. 
@@
 
-To explore the data i visualise the min temperatures along the South african coastline.
+To explore the data I visualise the min temperatures along the South african coastline.
 
@@
 OISST_sites <- OISST_sites %>% 
-  dplyr::rename(temp = temp_OISST)
+  dplyr::rename(temp = temp_OISST) %>%
+  dplyr::mutate(date = as.Date(date)) %>% 
+  drop_na()
 ```
@@
 ```{r}
-OISST_plot <- match_plot(df = OISST_sites)
+OISST_plot <- match_plot(df = insitu_OISST)
 ```
 
-##
+```{r}
+ggplot(data = insitu_OISST, aes(x = date, y = temp, colour = site)) +
+geom_line(aes(group = site)) +
+labs(x = "", y = "Temperature (°C)") +
+theme(axis.text.x = element_text(angle = 45)) +
+theme_linedraw() +
+theme(legend.position = "top")
+```

### 2019-01-23 Methods.Rmd
some_bits

@@
   drop_na() # Removing NA values within the dataset
+
+save(SACTN_3yr, file = "Data/SACTN_3yr.RData")
 ```
@@
 ```
+
+## CMC dataset
+
+```{r}
+CMC_Lamberts_Bay <- read_csv("Data/CMC/CMC_Lamberts Bay_SST_timeseries_5nearest.csv")
+CMC_Port_Nolloth <- read_csv("Data/CMC/CMC_Port Nolloth_SST_timeseries_5nearest.csv")
+CMC_Saldanha_Bay <- read_csv("Data/CMC/CMC_Saldanha Bay_SST_timeseries_5nearest.csv")
+CMC_Yzerfontein <- read_csv("Data/CMC/CMC_Yzerfontein_SST_timeseries_5nearest.csv")
+
+CMC_SST <- rbind(CMC_Lamberts_Bay,CMC_Port_Nolloth, CMC_Saldanha_Bay,CMC_Yzerfontein) %>%
+  dplyr::rename(site = station)
+
+CMC_SST$date <- (ymd(CMC_SST$date))
+
+# save(CMC_SST, file = "Data/CMC_SST.RData")
+```
+
+```{r}
+load("Data/CMC_SST.RData")
+
+CMC_SST <- CMC_SST %>% 
+  drop_na()
+
+CMC_SST <- CMC_SST %>% 
+  dplyr::rename(temp = nearest1)
+
+insitu_CMC <- match_func(df = CMC_SST) %>%
+  drop_na()
+CMC_plot <- match_plot(df = insitu_CMC)
+
+ggplot(data = insitu_CMC, aes(x = date, y = temp, colour = site)) +
+geom_line(aes(group = site)) +
+labs(x = "", y = "Temperature (°C)") +
+theme(axis.text.x = element_text(angle = 45)) +
+theme_linedraw() +
+theme(legend.position = "top")
+```
+
+This code is testing out the wavelet analyses for the in situ temperature data. This is for the insitu temperature only. This plot is a wavelet analyses for all of the insitu collected data only
+
+```{r}
+load("Data/SACTN_3yr.RData")
+
+temp.d <- SACTN_3yr %>% 
+  dplyr::group_by(site, date) %>% 
+  dplyr::summarise(temp = mean(insitu_temp, na.rm = T)) %>%
+  dplyr::mutate(no = seq(1:n())) %>%
+  dplyr::ungroup() %>% 
+  dplyr::select(site, no, temp, date)
+
+auto.arima(dplyr::filter(SACTN_3yr, site == "Port Nolloth")[, "insitu_temp"], max.p = 3, max.q = 3, stationary = FALSE, 
+           seasonal = FALSE)
+
+PN <- temp.d %>% 
+  dplyr::filter(site == "Port Nolloth") %>% 
+  dplyr::select(temp, -date, -site) %>% 
+  dplyr::mutate(no = seq(1:n())) %>% 
+  dplyr::select(no, temp)
+
+temp.d <- temp.d %>% 
+  select(site, no, temp, date)
+
+prewhite_fun <- function(x) {
+  df <- x[, 2:3]
+  out <- prewhiteAR(df, order = 3, method = "mle", aic = TRUE,
+             genplot = FALSE, verbose = FALSE)
+  colnames(out) <- c("no", "temp")
+  return(out)
+  }
+PN_prewhite <- as.tibble(ddply(temp.d, .(site), prewhite_fun))
+
+ggplot(PN_prewhite, aes(x = no, y = temp)) +
+  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
+  geom_line() +
+  facet_wrap(~ site, nrow = 4) +
+  theme_bw()
+
+wl.fun <- function(x) {
+  analyze.wavelet(x, "temp", loess.span = 0, dt = 1,
+                   dj = 1/50, lowerPeriod = 2, make.pval = TRUE, n.sim = 50, 
+                   method = "white.noise", verbose = FALSE)
+}
+
+PN_wave <- dlply(PN_prewhite, .(site), wl.fun)
+
+site.names <- list("Port Nolloth", "Lamberts Bay", "Saldanha Bay", "Yzerfontein")
+for (i in 1:length(PN_wave)) {
+  attributes(PN_wave[[i]]) <- c(attributes(PN_wave[[i]]), ref = site.names[i])
+  }
+
+plot_fun <- function(x) {
+  wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
+           timelab = "Days", verbose = FALSE, useRaster = TRUE,
+           periodlab = "Period", lwd = 1, graphics.reset = FALSE,
+           main = attributes(x)$ref)
+  }
+
+# ... apply the function
+#op <- par(no.readonly = TRUE)
+par(mfrow = c(2, 2),
+    mar = c(2.5, 3.5, 2.5, 0.5),
+    mgp = c(2.5, 1, 0),
+    oma = c(0, 0, 1, 0))
+l_ply(PN_wave, plot_fun)
+#par(op)
+
+# Save each of the graph seperate- graph PN, LB, saldanha Bay and Yzerfontein
+```

### 2019-01-24 Methods.Rmd
Individual plots

@@
 
-site.names <- list("Port Nolloth", "Lamberts Bay", "Saldanha Bay", "Yzerfontein")
+# RWS: The order of these sites was incorrect
+# Rather get the names from the pre-existing data
+# site.names <- list("Port Nolloth", "Lamberts Bay", "Saldanha Bay", "Yzerfontein")
+
 for (i in 1:length(PN_wave)) {
-  attributes(PN_wave[[i]]) <- c(attributes(PN_wave[[i]]), ref = site.names[i])
+  attributes(PN_wave[[i]]) <- c(attributes(PN_wave[[i]]), ref = names(PN_wave)[i])
 }
 
-plot_fun <- function(x) {
+# x <- PN_wave[[3]]
+
+plot_fun <- function(x, plot_name = attributes(x)$ref) {
+  png(filename = paste0("Figures/",plot_name,"_wavelet.png"),
+      width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
   wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
@@
            periodlab = "Period", lwd = 1, graphics.reset = FALSE,
-           main = attributes(x)$ref)
+           main = plot_name)
+  dev.off()
 }
+
+ldply(PN_wave, plot_fun)
 
@@
 #op <- par(no.readonly = TRUE)
-par(mfrow = c(2, 2),
-    mar = c(2.5, 3.5, 2.5, 0.5),
-    mgp = c(2.5, 1, 0),
-    oma = c(0, 0, 1, 0))
-l_ply(PN_wave, plot_fun)
+# par(mfrow = c(2, 2),
+#     mar = c(2.5, 3.5, 2.5, 0.5),
+#     mgp = c(2.5, 1, 0),
+#     oma = c(0, 0, 1, 0))
 #par(op)
@@
 # Save each of the graph seperate- graph PN, LB, saldanha Bay and Yzerfontein
+# l_ply(PN_wave, plot_fun)
 ```

### 2019-01-25 Write_up/Writing/Introduction_outline.Rmd
updated

@@
 
-Climate change as a result of anthropogenic activities is now established as a global concern for both scientific and political importance.
+Climate change as a result of anthropogenic activities is  an established global concern of both scientific and political importance.
 As the human population grows, the rate of global warming increases, resulting in profound effects on marine and other ecosystems.
@@
 Global warming represents distinct warming and cooling periods throughout the 20th and 21st century (IPCC, 2007).
-Global warming however is not evenly distributed across the global ocean thus presenting regions warming much faster or slower than then the global average [@levitus; @santos].
+Global warming however is not evenly distributed across the global ocean, and thus several regions are warming faster or slower than then García-Reyes et al., 2015 the global average [@levitus; @santos].
 In particular, the Atlantic Ocean contributes most to the increase of the heat content [@nerem; @levitus].
 
-As a result of climate change, research studies provide significant evidence on the increase in warming of the ocean surface.
+As a result of climate change, research studies provide significant evidence relating to the increase in the warming of the ocean surface.
 This warming slowly propagates down through the water column and in turn increases the thermal stratification [@gruber].
 Strong stratification limits the depth at which water is upwelled and hence controls the amount of nutrients brought up [@chhak; @jacox2011; @jocox2016].
-The interplay between upwelling winds and stratification determine the amount of nutrients present within the system and thus is strongly linked to biological productivity.
+The interplay between upwelling winds and stratification determines the amount of nutrients present within the system and is therfore strongly linked to biological productivity.
 Globally, as a result of climate change, stratification of the water-column between 0 and 200 m has increased by an estimated 4% between 1971 to 2010 [@Stocker].
@@
 
+Climate change has severe ecological consequences on both plants and animals (Bradley et al.
+1999; Walther et al.
+2002).
+For example, in the past two decades, sea surface temperature has risen in many regions.
+Seasonal rise in temperature can act as an important cue in species survival, reproduction and abundance (Cushing 1982; Cushing 1990).
+The differing responses of species as a result of climate change are important to understand as these changes affect, among many things, the trophic dynamics and fisheries yield of several regional ecosystems (Murawski 1993).
+
+Understanding the relationship between global warming and climate change as a result of anthropogenic activity is a vast and challenging process.
+In a bid to further this understanding, Rockström et al.
+(2009) introduced the concept of planetary boundaries.
+According to this concept, within each of these boundaries exists a level of safety that should not be exceeded.
+For example, the planetary boundary for biodiversity loss, has been surpassed (Steffen et al., 2015).
+Similarly, the safety boundary for warming of the globe by greenhouse gas emission has also been surpassed.
+The idea of planetary boundaries is essential in that it helps our understanding of the threats we cause via the continued destruction of the natural environment.
+Consequently, studies of anthropogenic influences on the ecosystem are generally divided into land or sea.
+Terrestrial ecosystems are more familiar to the general populace; whereas marine ecosystems are associated with high levels of primary production (Chavez et al., 2011) and are highly at risk as a result of climate change (Österblom et al., 2017).
+This has been proven to decrease primary production thus affecting the trophic levels and leading to trophic turbulence and destruction of marine fauna (Lewandowska et al., 2014; Dell et al., 2014, McCauley et al., 2015).
+
+Climate change similarly affects coastal species such as oysters, corals and, sea grass.
+Lord (2017) also found evidence that suggests that a change in climate as a result of rising sea temperatures may increase the spread of invasive species and increase the chances of an outbreak of diseases (Altizer et al., 2013).
+Research by Mead et al (2013) found that coastal ecosystems are at high risk as a result of climate change, later, Whitfield et al (2016) confirmed that coastal bio-diversity have been observed to be affected.
+Given the importance of primary productivity, it was determined that the Benguela upwelling systerm within the South African region should become the focus of this thesis.
+Here the increase in global temperatures are one of the main focus points along with the chlorophyll concentration.
+By making use of these two measurements, I am able to analyses how chlorophyll concentration vary with a change in temperature.
+
 ## Problem statement:
 
-With global warming continuing at a rapid pace the amount of greenhouse gases within the atmosphere are going to increase in the future.
+With global warming continuing at a rapid pace, the amount of greenhouse gases within the atmosphere are going to increase in the future.
 This increase will undoubtedly have an impact on the world’s oceans and the marine life within, both directly and indirectly.
@@
 Because upwelling is important for productivity within coastal systems, a decrease in this process will negatively impact on marine fauna and flora, and on human activities such as fishing.
-To assess changes in upwelling events, here I aim to use SST, wind, chlorophyll data etc.
-to monitor durations and intensities of upwelling within one of the EBUS.....
+To assess changes in upwelling events, here I aim to use SST, wind, and chlorophyll data to monitor the durations and intensities of upwelling events within the Benguela EBUS and identify wether this resulted in an extreme anomalous cold event.
+Atmospheric-ocean coupling is known to be affected by climate change, likely changing wind patterns and leading to upwelling favourable conditions (García-Reyes et al., 2015).
+Here we test if an increase in upwelling may lead to a detection of more MCSs simply because these cold temperatures caused by aseasonal upwelling will reflect in historic climatologies.
 
@@
 
+Benguela Niños are defined as anomalous warm events occurring between the southward flowing Angola current and the Benguela upwelling system off south western Africa (Shannon et al.
+1986).
+These warm events profoundly impact on fisheries and the climate of this region.
+They are proven to induce unexpected rainfall events and drastically influence fish abundance and distribution (Rouault et al.
+2003, Boyer et al.
+2001).
+Evidently, Benguela Niños are expressed as regions of abnormal, persistent high sea surface temperatures.
+Many researchers often associate the equatorial interannual variability pattern in the Atlantic to the ENSO phenomenon found within the Pacific region (Sutton et al.
+2000; Zebiak 1993).
+Benguela Niños have been recorded to occur in 1934, 1949, 1963, 1984 and 1995 (Shannon et al.
+1986; Gammelsrød et al.
+1998).
+Picaut (1985) suggested that eastern tropic oceans are governed by remote wind stress effects through equatorial wave dynamics.
+Anomalies in trade winds may result in Kelvin waves that propagate eastward along the equator, inducing a deepening or a lifting of the thermocline.
+It is possible that equatorial Kelvin waves may play a role in the evolution of Benguela Niños.
+
 ### Extreme events
@@
 These events occur rapidly, and unexpectedly and they profoundly impact on species diversity, distributions, growth and survival where they take place.
-These events are often undetected but studies have found that extreme events have occurred as recently as approximately four decades ago.
+These events are often undetected, but studies have found that extreme events have occurred as recently as approximately four decades ago.
 Jentsch et al.
-(2007) were the first to discuss this mode of climate change and brought it to focus.
+(2007) were the first to discuss this mode of climate change and brought it to greater focus.
 Steps have since been taken towards understanding the drivers, impacts and consequences of these events.
-Subsequently, research has predicted an increase in the occurrence of these extreme events as a result of long-term warming of the planet via the process of global warming (REFERENCE).
-Extreme events are often associated with toxic algal blooms, shifts in species distribution, and the bleaching of coral reefs [@mccabe; @harley; @schoepf].
+Subsequently, research has predicted an increase in the occurrence of these extreme events as a result of long-term warming of the planet via the process of global warming [@mccabe; @harley; @schoepf].
 
 Little is known about extreme events occurring within the South African region.
-Very little South African researchers spend their time and careers on studyig extreme events within this region.
-During the 1980s, a study done by Birkett and Cook was the only study done where they anaylsed the impact of extreme warming events on the life cycle of white muscles,Donax serra Röding.
-However, more recently studies done by Schlegel et al, (2017,2018) examined the relationship between the nearshore and offshore occurances of marine heatwaves and cold-spells as well as the influence of predominant atmosphetic and oceanic patterns during Coastal Marine Heatwaves.
-It is now becoming accepted that extreme events are a threat to several ecosystem [@thompson].
-
-### Cold spells
-
-- Definition
-- Impact
--Major cold spells
+Few South African researchers have specifically focussed on investigating this topic, and published literature are sparse.
+During the 1980’s however, Birkett and Cook analysed the impacts of extreme warming events on the life cycle of white muscles, Donax serra Röding.
+Futher studies on extreme events occurred within the Florida Reef Tract, where anomolously cold events resulted in the mass mortality of coral reefs and fish.
+Acropora cervicornis, was found to be particularly susceptible to low temperatures and experienced extensive mortality (Shinn 1966; Davis 1982).
+While these phenomenons are not as common as warming events, they have a dramtic and long-lasting effect on coral communities.
+However, studies in January 2010 reveal that the Florida Reef Tract experienced a cold-water anomaly, marine cold spells (MCSs) with temperatures below 16 ◦ C recorded for up to 14 days (Lirman et al.
+2011).
+Species population distributon are often set by cold distribution boundaries.
+More recently, Schlegel et al, (2017, 2018) examined the relationship between the nearshore and offshore occurrences of marine heatwaves and cold-spells as well as the influence of predominant atmospheric and oceanic patterns during Coastal Marine Heatwaves.
+It is now becoming accepted that extreme events are a threat to several ecosystems [@thompson].
+In a warming planet one would assume that the occurrence and intensity of the severity of marine cold spells (MCSs) are likely to decrease.
+This thesis aims to address this assumption.
 
 ## Aims and approach:
-In this thesis I aim to…
+
+The warming of the global climate as a result of anthropogenic activity, has negatively influenced both terrestrial and marine ecosystems, and ultimately inhibits humanity and ecological functioning (McCauley et al., 2015).
+This climate variability, manifested as increases in frequency and duration of extreme events is regarded as a great threat (Thompsonet al., 2013).
+Very few studies have addressed the occurrence of cold temperature spells or cold events within the Benguela region.
+This investigation examines recent trends and variation in the frequency and duration of cold spells occurring in this region.
+Due to the lack of in situ collected data, cold spells are mainly examined between the period 2014-2017 for four sites along the South African coastline falling within the Benguela upwelling cell.
+Here we aim to assess whether the intensity and duration of these extreme events may show any changes in frequency and intensity on a seasonal basis.
+In this research project I make use of various satellite derived observations of sea surface temperature (SST) and wind stress to further investigate the Benguela upwelling system along the South African coastline, whilst considering cold events and warm events occurring within this region

### 2019-01-25 Methods_tidy.Rmd
temporary_push

---
title: "Methods_tidy"
author: "Amieroh Abrahams"
date: "25 January 2019"
output: html_document
---

# Libraries

First I need to find, install and load various packages. These packages will be available on CRAN and can be accessed and installed in the usual way.

```{r prelim_opts, echo=FALSE}
knitr::opts_chunk$set(
  comment = "R>",
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(plyr)
library(lubridate)
library(ggpubr)
library(zoo)
library(lubridate)
library(FNN)
library(forecast)
library(astrochron)
library(WaveletComp)
library(data.table)
```

### Load site list SACTN data

Now to get to the data. The first step involves the loading of the site list. The statistical properties of the seawater temperature representing the South African coastline, such as the mean, minimum and maximum temperatures. These values vary among coastal sections due to the influence of the cold Benguala and warm Agulhas currents. Here we will only focus on the temperature data found along the west coast (wc) (i.e. sites influenced by the Benguela current, EBUS). The SACTN dataset comprise of 129 *in situ* coastal seawater temperatures derived from daily measurements over up to 40 years. The SACTN temperature dataset was compiled by measuring coastal temperatures at 129 sites along the coast of South Africa, daily from 1972 until 2017.

```{r load_files1, include=TRUE}
load("Data/site_list_v4.2.RData")
load("Data/SACTN_daily_v4.2.RData")
```

Now we select only the sites occuring along the west coast. Thereafter, I include all the sites which had a time series of 10 years or greater. This yield a 2 rows of data for Paternoster. I removed the row of data obtained by the DEA, thus ending up with all of the data being obtained from SAWS in this dataset. I then exclude the sites that do not fall within the upwelling cell. Once those sites were removed, we have a total of 4 sites. These are: Port Nolloth, Lamberts Bay, Saldanna and Yzerfontein.

```{r}
site_list_sub <- site_list %>% 
  filter(coast == "wc")

SACTN_daily_wc <- site_list_sub %>% 
  filter(length >= 3650)

upwelling_sites <- SACTN_daily_wc[-6,] # Here we had two rows of data for Paternoster; i removed the DEA row: All data here are now collected by SAWS 
upwelling_sites <- upwelling_sites[-10:-14,] # Here i remove the sites that does not fall within the upwelling cell

# Here I find a time series with the data occuring in the most recent years
upwelling_sites <- upwelling_sites[c(-2, -3, -5, -6, -8),] 
# save(upwelling_sites, file = "Data/upwelling_sites.RData")

SACTN_daily_temps <- SACTN_daily_v4.2 %>% 
  left_join(site_list[,c(4,13)], by = "index") %>% 
  filter(index %in% upwelling_sites$index)
# save(SACTN_daily_temps, file = "SACTN_daily_temps.RData")

SACTN_daily_temps_indexsplit <- SACTN_daily_temps %>% 
  separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>% 
  dplyr::rename(insitu_temp = temp)

# save(SACTN_daily_temps_indexsplit, file = "Data/SACTN_daily_temps_indexsplit.RData")
```

Here I add monthly, yearly and seasonal column. This allows us to compare temperature variation on a seasonal basis. Research studies done suggests that the intensity and duration of upwelling varies on a seasonal basis.

```{r}
SACTN_monthly <- function(df) {
  SACTN_split_monthly <- df %>%
  mutate(month = month(date, abbr = T, label = T),
         year = year(date)) %>%
  group_by(index, month, year) %>%
    mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",       
                           ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
                                ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
                                       ifelse(month %in% c("Oct", "Nov", "Dec"), "Spring","Error")))))
  return(SACTN_split_monthly)
}

# SACTN_split <- SACTN_monthly(df = SACTN_daily_temps_indexsplit)
# save(SACTN_split, file = "SACTN_split.RData")
```

Find where different sites have overlapping time series within this `SACTN_split` dataset.

```{r}
load("~/Documents/Masters_2019/MastersProject/Data/SACTN_split.RData")

SACTN_3yr <- SACTN_split %>% 
  filter(year(date) %in% seq(2014, 2015)) %>% 
  drop_na() # Removing NA values within the dataset

save(SACTN_3yr, file = "Data/SACTN_3yr.RData")

ggplot(data = SACTN_3yr, aes(x = date, y = insitu_temp, colour = site)) +
geom_line(aes(group = site)) +
labs(x = "", y = "Temperature (°C)") +
theme(axis.text.x = element_text(angle = 45)) +
theme_linedraw() +
theme(legend.position = "top")
```

In this thesis, different remotely-sensed SST datasets are used. Hence, these datasets represent different resolutions at which data was obtained. We will make use of five remotely-sensed SST datasets, these include; the  MUR, AVHRR OISST, Pathfinder and G1SST data.

## MUR dataset

```{r}
MUR_Lamberts_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Lamberts Bay_SST_timeseries_5nearest.csv")
MUR_Port_Nolloth <- read_csv("Data/MUR_nearest5pixels/MUR_Port Nolloth_SST_timeseries_5nearest.csv")
MUR_Saldanha_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Saldanha Bay_SST_timeseries_5nearest.csv")
MUR_Yzerfontein <- read_csv("Data/MUR_nearest5pixels/MUR_Yzerfontein_SST_timeseries_5nearest.csv")

MUR_SST <- rbind(MUR_Lamberts_Bay,MUR_Yzerfontein, MUR_Port_Nolloth,MUR_Saldanha_Bay) %>%
  dplyr::rename(site = station)

MUR_SST$date <- (ymd(MUR_SST$date))
# save(MUR_SST, file = "Data/MUR_SST.RData")

load("~/Documents/Masters_2019/MastersProject/Data/MUR_SST.RData")
MUR_SST <- MUR_SST %>% 
  drop_na()
```

Here I write a function to match the remotely-sensed SST to the *in-situ* collected SST

```{r}
MUR_SST <- MUR_SST %>% 
  dplyr::rename(temp = nearest1) # nearest1- 1 pixel as applied to the OISST data

match_func <- function(df){
  match <- SACTN_3yr  %>%  
  left_join(df, by = c("site", "date")) %>% 
  na.trim()
  return(match)
}

insitu_MUR <- match_func(df = MUR_SST) %>%
  drop_na()

match_plot <- function(df){
  plot1 <- df %>% 
ggplot(aes(x = date, y = temp)) +
  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ site, nrow = 2) +
  theme_bw()
  return(plot1)
}

MUR_plot <- match_plot(df = insitu_MUR)
ggplot(data = insitu_MUR, aes(x = date, y = temp, colour = site)) +
geom_line(aes(group = site)) +
labs(x = "", y = "Temperature (°C)") +
theme(axis.text.x = element_text(angle = 45)) +
theme_linedraw() +
theme(legend.position = "top")
```

## OISST dataset

```{r}
# AVHRR <- read_csv("data/csvavhrr-only-v2-19810901-20180630.csv")

OISSTDir <- "~/Documents/Masters_2019/Data_SST"
OISST <- fread(paste0(OISSTDir, "/csvavhrr-only-v2-19810901-20180630.csv"),
            col.names = c("lon", "lat", "temp", "date"))

#OISST <- as.tibble(fread("~/Documents/Masters_2019/MastersProject/Data/csvavhrr-only-v2-19810901-20180630.csv"))
# Quick visualisation of the data
# To explore the data I visualise the min temperatures along the South african coastline.

OISST %>%
  filter(date == min(date)) %>%
  ggplot(aes(x = lon, y = lat)) +
  geom_raster(aes(fill = temp))
ggsave(plot = OISST, filename = "Figures/OISST")
```

## Here I find the nearest SST pixels

Now we apply the FNN (Fast Nearest Neighbor) package to determine the nearesr SST pixel to the insitu collected sites.

```{r}
unique_pixel <- OISST %>%
  select(lon, lat) %>%
  unique()

# Select nearest 1 pixels (k = 1)
# here we use knnx to find the closes 1 pixels to the insitu sites
match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
                          query = as.matrix(upwelling_sites[,5:6]), k = 1)

# Select SST pixels nearest to insitu sites
pixel_match <- unique_pixel[match_index,] %>%
  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
  mutate(site = upwelling_sites$site)

## Subsetting the OISST data to match the upwelling sites within the in situ collected temperature data
OISST_match <- OISST %>%
  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
  filter(combi %in% pixel_match$combi)

## Visualisation
ggplot(data = OISST_match, aes(x = date, y = temp_OISST)) +
  geom_line() +
  facet_wrap(~combi)
ggsave(plot = OISST_match, filename = "Figures/combined_OISST_match")

OISST_match %>%
  filter(date == max(date)) %>%
  ggplot(aes(x = lon, y = lat)) +
  geom_raster(aes(fill = temp))
```

```{r}
OISST_sites <- OISST_match %>%
  left_join(pixel_match, by = c("combi", "lon", "lat")) %>% 
  dplyr::rename(temp_OISST =temp)

# Check that the pixel that is shared is duplicated
# unique(combined_OISST_sites$site)
# Visualise time series
ggplot(data = OISST_sites, aes(x = date, y = temp)) +
  geom_line() +
  facet_wrap(~site)

# save(OISST_sites, file = "Data/OISST_sites.RData")
# load("~/Documents/MastersProject/Data/OISST_sites.RData")

OISST_sites <- OISST_sites %>% 
  dplyr::rename(temp = temp_OISST) %>%
  dplyr::mutate(date = as.Date(date)) %>% 
  drop_na()
```

## Matching the Insitu data with the OISST SST data

```{r}
insitu_OISST <- match_func(df = OISST_sites)
OISST_plot <- match_plot(df = insitu_OISST)

ggplot(data = insitu_OISST, aes(x = date, y = temp, colour = site)) +
geom_line(aes(group = site)) +
labs(x = "", y = "Temperature (°C)") +
theme(axis.text.x = element_text(angle = 45)) +
theme_linedraw() +
theme(legend.position = "top")
```

## G1SST dataset

```{r}
Lamberts_Bay <- read_csv("Data/G1SST_sub/Lamberts Bay_SST_timeseries_5nearest.csv")
Port_Nolloth <- read_csv("Data/G1SST_sub/Port Nolloth_SST_timeseries_5nearest.csv")
Saldanha_Bay <- read_csv("Data/G1SST_sub/Saldanha Bay_SST_timeseries_5nearest.csv")
Yzerfontein <- read_csv("Data/G1SST_sub/Yzerfontein_SST_timeseries_5nearest.csv")

G1SSTsub_SST <- rbind(Lamberts_Bay,Port_Nolloth, Saldanha_Bay, Yzerfontein) %>%
  dplyr::rename(site = station)

G1SSTsub_SST$date <- (ymd(G1SSTsub_SST$date))

# save(G1SSTsub_SST, file = "Data/G1SSTsub_SST.RData")
load("Data/G1SSTsub_SST.RData")

G1SSTsub_SST <- G1SSTsub_SST %>% 
  drop_na()

G1SSTsub_SST <- G1SSTsub_SST %>% 
  dplyr::rename(temp = nearest1)

insitu_G1SST <- match_func(df = G1SSTsub_SST) %>%
  drop_na()
G1SST_plot <- match_plot(df = insitu_G1SST )

ggplot(data = insitu_G1SST, aes(x = date, y = temp, colour = site)) +
geom_line(aes(group = site)) +
labs(x = "", y = "Temperature (°C)") +
theme(axis.text.x = element_text(angle = 45)) +
theme_linedraw() +
theme(legend.position = "top")
```

```{r}
load("Data/SACTN_3yr.RData")
temp.d <- SACTN_3yr %>% 
  dplyr::group_by(site, date) %>% 
  dplyr::summarise(temp = mean(insitu_temp, na.rm = T)) %>%
  dplyr::mutate(no = seq(1:n())) %>%
  dplyr::ungroup() %>% 
  dplyr::select(site, no, temp, date)
auto.arima(dplyr::filter(SACTN_3yr, site == "Port Nolloth")[, "insitu_temp"], max.p = 3, max.q = 3, stationary = FALSE, 
           seasonal = FALSE)
PN <- temp.d %>% 
  dplyr::filter(site == "Port Nolloth") %>% 
  dplyr::select(temp, -date, -site) %>% 
  dplyr::mutate(no = seq(1:n())) %>% 
  dplyr::select(no, temp)
temp.d <- temp.d %>% 
  select(site, no, temp, date)
prewhite_fun <- function(x) {
  df <- x[, 2:3]
  out <- prewhiteAR(df, order = 3, method = "mle", aic = TRUE,
             genplot = FALSE, verbose = FALSE)
  colnames(out) <- c("no", "temp")
  return(out)
  }
PN_prewhite <- as.tibble(ddply(temp.d, .(site), prewhite_fun))
ggplot(PN_prewhite, aes(x = no, y = temp)) +
  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ site, nrow = 4) +
  theme_bw()
wl.fun <- function(x) {
  analyze.wavelet(x, "temp", loess.span = 0, dt = 1,
                   dj = 1/50, lowerPeriod = 2, make.pval = TRUE, n.sim = 50, 
                   method = "white.noise", verbose = FALSE)
}
PN_wave <- dlply(PN_prewhite, .(site), wl.fun)
# RWS: The order of these sites was incorrect
# Rather get the names from the pre-existing data
# site.names <- list("Port Nolloth", "Lamberts Bay", "Saldanha Bay", "Yzerfontein")
for (i in 1:length(PN_wave)) {
  attributes(PN_wave[[i]]) <- c(attributes(PN_wave[[i]]), ref = names(PN_wave)[i])
}
# x <- PN_wave[[3]]
plot_fun <- function(x, plot_name = attributes(x)$ref) {
  png(filename = paste0("Figures/Insitu/",plot_name,"_wavelet.png"),
      width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
  wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
           timelab = "Days", verbose = FALSE, useRaster = TRUE,
           periodlab = "Period", lwd = 1, graphics.reset = FALSE,
           main = plot_name)
  dev.off()
}
ldply(PN_wave, plot_fun)
# ... apply the function
#op <- par(no.readonly = TRUE)
# par(mfrow = c(2, 2),
#     mar = c(2.5, 3.5, 2.5, 0.5),
#     mgp = c(2.5, 1, 0),
#     oma = c(0, 0, 1, 0))
#par(op)
# Save each of the graph seperate- graph PN, LB, saldanha Bay and Yzerfontein
# l_ply(PN_wave, plot_fun)
```

```{r}
# OISST
# load("Data/OISST_sites.RData")

temp.d <- insitu_OISST %>% 
  dplyr::group_by(site, date) %>% 
  dplyr::summarise(temp = mean(temp, na.rm = T)) %>%
  dplyr::mutate(no = seq(1:n())) %>%
  dplyr::ungroup() %>% 
  dplyr::select(site, no, temp, date)

auto.arima(dplyr::filter(insitu_OISST, site == "Port Nolloth")[, "temp"], max.p = 3, max.q = 3, stationary = FALSE, 
           seasonal = FALSE)

PN <- temp.d %>% 
  dplyr::filter(site == "Port Nolloth") %>% 
  dplyr::select(temp, -date, -site) %>% 
  dplyr::mutate(no = seq(1:n())) %>% 
  dplyr::select(no, temp)

temp.d <- temp.d %>% 
  select(site, no, temp, date)

prewhite_fun <- function(x) {
  df <- x[, 2:3]
  out <- prewhiteAR(df, order = 3, method = "mle", aic = TRUE,
             genplot = FALSE, verbose = FALSE)
  colnames(out) <- c("no", "temp")
  return(out)
  }
PN_prewhite <- as.tibble(ddply(temp.d, .(site), prewhite_fun))

ggplot(PN_prewhite, aes(x = no, y = temp)) +
  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ site, nrow = 4) +
  theme_bw()

wl.fun <- function(x) {
  analyze.wavelet(x, "temp", loess.span = 0, dt = 1,
                   dj = 1/50, lowerPeriod = 2, make.pval = TRUE, n.sim = 50, 
                   method = "white.noise", verbose = FALSE)
}

PN_wave <- dlply(PN_prewhite, .(site), wl.fun)

for (i in 1:length(PN_wave)) {
  attributes(PN_wave[[i]]) <- c(attributes(PN_wave[[i]]), ref = names(PN_wave)[i])
}
# x <- PN_wave[[3]]
plot_fun <- function(x, plot_name = attributes(x)$ref) {
  png(filename = paste0("Figures/OISST/",plot_name,"_wavelet.png"),
      width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
  wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
           timelab = "Days", verbose = FALSE, useRaster = TRUE,
           periodlab = "Period", lwd = 1, graphics.reset = FALSE,
           main = plot_name)
  dev.off()
}
ldply(PN_wave, plot_fun)
```

```{r}

# MUR
# load("Data/insitu_MUR.RData")

temp.d <- insitu_MUR %>% 
  dplyr::group_by(site, date) %>% 
  dplyr::summarise(temp = mean(temp, na.rm = T)) %>%
  dplyr::mutate(no = seq(1:n())) %>%
  dplyr::ungroup() %>% 
  dplyr::select(site, no, temp, date)

auto.arima(dplyr::filter(insitu_MUR, site == "Port Nolloth")[, "temp"], max.p = 3, max.q = 3, stationary = FALSE, 
           seasonal = FALSE)

PN <- temp.d %>% 
  dplyr::filter(site == "Port Nolloth") %>% 
  dplyr::select(temp, -date, -site) %>% 
  dplyr::mutate(no = seq(1:n())) %>% 
  dplyr::select(no, temp)

temp.d <- temp.d %>% 
  select(site, no, temp, date)

prewhite_fun <- function(x) {
  df <- x[, 2:3]
  out <- prewhiteAR(df, order = 3, method = "mle", aic = TRUE,
             genplot = FALSE, verbose = FALSE)
  colnames(out) <- c("no", "temp")
  return(out)
  }
PN_prewhite <- as.tibble(ddply(temp.d, .(site), prewhite_fun))

ggplot(PN_prewhite, aes(x = no, y = temp)) +
  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ site, nrow = 4) +
  theme_bw()

wl.fun <- function(x) {
  analyze.wavelet(x, "temp", loess.span = 0, dt = 1,
                   dj = 1/50, lowerPeriod = 2, make.pval = TRUE, n.sim = 50, 
                   method = "white.noise", verbose = FALSE)
}

PN_wave <- dlply(PN_prewhite, .(site), wl.fun)

for (i in 1:length(PN_wave)) {
  attributes(PN_wave[[i]]) <- c(attributes(PN_wave[[i]]), ref = names(PN_wave)[i])
}
# x <- PN_wave[[3]]
plot_fun <- function(x, plot_name = attributes(x)$ref) {
  png(filename = paste0("Figures/MUR/",plot_name,"_wavelet.png"),
      width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
  wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
           timelab = "Days", verbose = FALSE, useRaster = TRUE,
           periodlab = "Period", lwd = 1, graphics.reset = FALSE,
           main = plot_name)
  dev.off()
}
ldply(PN_wave, plot_fun)
```

```{r}
# load("Data/insitu_G1SST.RData")

temp.d <- insitu_G1SST %>% 
  dplyr::group_by(site, date) %>% 
  dplyr::summarise(temp = mean(temp, na.rm = T)) %>%
  dplyr::mutate(no = seq(1:n())) %>%
  dplyr::ungroup() %>% 
  dplyr::select(site, no, temp, date)

auto.arima(dplyr::filter(insitu_G1SST, site == "Port Nolloth")[, "temp"], max.p = 3, max.q = 3, stationary = FALSE, 
           seasonal = FALSE)

PN <- temp.d %>% 
  dplyr::filter(site == "Port Nolloth") %>% 
  dplyr::select(temp, -date, -site) %>% 
  dplyr::mutate(no = seq(1:n())) %>% 
  dplyr::select(no, temp)

temp.d <- temp.d %>% 
  select(site, no, temp, date)

prewhite_fun <- function(x) {
  df <- x[, 2:3]
  out <- prewhiteAR(df, order = 3, method = "mle", aic = TRUE,
             genplot = FALSE, verbose = FALSE)
  colnames(out) <- c("no", "temp")
  return(out)
  }
PN_prewhite <- as.tibble(ddply(temp.d, .(site), prewhite_fun))

ggplot(PN_prewhite, aes(x = no, y = temp)) +
  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ site, nrow = 4) +
  theme_bw()

wl.fun <- function(x) {
  analyze.wavelet(x, "temp", loess.span = 0, dt = 1,
                   dj = 1/50, lowerPeriod = 2, make.pval = TRUE, n.sim = 50, 
                   method = "white.noise", verbose = FALSE)
}

PN_wave <- dlply(PN_prewhite, .(site), wl.fun)

for (i in 1:length(PN_wave)) {
  attributes(PN_wave[[i]]) <- c(attributes(PN_wave[[i]]), ref = names(PN_wave)[i])
}
# x <- PN_wave[[3]]
plot_fun <- function(x, plot_name = attributes(x)$ref) {
  png(filename = paste0("Figures/G1SST/",plot_name,"_wavelet.png"),
      width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
  wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
           timelab = "Days", verbose = FALSE, useRaster = TRUE,
           periodlab = "Period", lwd = 1, graphics.reset = FALSE,
           main = plot_name)
  dev.off()
}
ldply(PN_wave, plot_fun)
```

### 2019-01-25 Write_up/Writing/Introduction_outline.Rmd
references

@@
 In rare cases, projected upwelling increases may overcome the countervailing effects of upper-ocean warming and stratification to cause regional cooling [@auad].
-Research studies reveal declines in ocean oxygen levels as a result of ocean warming and reduced ventilation from stratification and circulation changes (Keeling et al.
-2010).
+Research studies reveal declines in ocean oxygen levels as a result of ocean warming and reduced ventilation from stratification and circulation changes [@Keeling].
 Similarly, low oxygen levels are also possible because of shifts in the composition of sinking organic material [@hofmann; @doney].
@@
 According to this concept, within each of these boundaries exists a level of safety that should not be exceeded.
-For example, the planetary boundary for biodiversity loss, has been surpassed (Steffen et al., 2015).
+For example, the planetary boundary for biodiversity loss, has been surpassed [@steffen].
 Similarly, the safety boundary for warming of the globe by greenhouse gas emission has also been surpassed.
@@
 Consequently, studies of anthropogenic influences on the ecosystem are generally divided into land or sea.
-Terrestrial ecosystems are more familiar to the general populace; whereas marine ecosystems are associated with high levels of primary production (Chavez et al., 2011) and are highly at risk as a result of climate change (Österblom et al., 2017).
-This has been proven to decrease primary production thus affecting the trophic levels and leading to trophic turbulence and destruction of marine fauna (Lewandowska et al., 2014; Dell et al., 2014, McCauley et al., 2015).
+Terrestrial ecosystems are more familiar to the general populace; whereas marine ecosystems are associated with high levels of primary production [@Chavez] and are highly at risk as a result of climate change [@osterblom].
+This has been proven to decrease primary production thus affecting the trophic levels and leading to trophic turbulence and destruction of marine fauna [@lewandowska; @dell, @mcCauley].
 
 Climate change similarly affects coastal species such as oysters, corals and, sea grass.
-Lord (2017) also found evidence that suggests that a change in climate as a result of rising sea temperatures may increase the spread of invasive species and increase the chances of an outbreak of diseases (Altizer et al., 2013).
+Lord (2017) also found evidence that suggests that a change in climate as a result of rising sea temperatures may increase the spread of invasive species and increase the chances of an outbreak of diseases [@altizer].
 Research by Mead et al (2013) found that coastal ecosystems are at high risk as a result of climate change, later, Whitfield et al (2016) confirmed that coastal bio-diversity have been observed to be affected.
@@
 
-Benguela Niños are defined as anomalous warm events occurring between the southward flowing Angola current and the Benguela upwelling system off south western Africa (Shannon et al.
-1986).
+Benguela Niños are defined as anomalous warm events occurring between the southward flowing Angola current and the Benguela upwelling system off south western Africa [@shannon1986].
 These warm events profoundly impact on fisheries and the climate of this region.
-They are proven to induce unexpected rainfall events and drastically influence fish abundance and distribution (Rouault et al.
-2003, Boyer et al.
-2001).
+They are proven to induce unexpected rainfall events and drastically influence fish abundance and distribution [@rouault; @boyer].
 Evidently, Benguela Niños are expressed as regions of abnormal, persistent high sea surface temperatures.
@@
 2000; Zebiak 1993).
-Benguela Niños have been recorded to occur in 1934, 1949, 1963, 1984 and 1995 (Shannon et al.
-1986; Gammelsrød et al.
-1998).
+Benguela Niños have been recorded to occur in 1934, 1949, 1963, 1984 and 1995 [@shannon; @gammelsrod].
 Picaut (1985) suggested that eastern tropic oceans are governed by remote wind stress effects through equatorial wave dynamics.
@@
 In this research project I make use of various satellite derived observations of sea surface temperature (SST) and wind stress to further investigate the Benguela upwelling system along the South African coastline, whilst considering cold events and warm events occurring within this region
+
+Picaut, J., 1985: Major dynamics affecting the eastern tropical At-
+lantic and Pacific Oceans. CalCOFI Report, Vol 26.

### 2019-01-25 Write_up/Writing/Introduction_outline.Rmd
updated_writing

@@
   html_document: default
+fontsize: 10pt
+geometry: margin=1in
+language: Australian
+mainfont: Minion Pro
+monofont: Courier
+papersize: A4
+sansfont: Myriad Pro
+tables: yes
+bibliography: ../bib_master.bib
 ---
@@
 Nevertheless, satellite data products are extremely useful when investigating the ocean at broad scale and can provide important SST data.
-<will add more here on the different types of resolution used in this thesis>
+<Will add more on the different types of data and the varying resolutions>
 
@@
 The interplay between upwelling winds and stratification determines the amount of nutrients present within the system and is therfore strongly linked to biological productivity.
-Globally, as a result of climate change, stratification of the water-column between 0 and 200 m has increased by an estimated 4% between 1971 to 2010 [@Stocker].
+Globally, as a result of climate change, stratification of the water-column between 0 and 200 m has increased by an estimated 4% between 1971 to 2010 [@stocker].
 Increased ocean stratification as a result of surface warming may reduce the ventilation of deep-water masses.
-As the deep water is being up welled to the surface, decreases in oxygen and pH levels are predicted along with an increase in nutrient levels [@Rykaczewski).
+As the deep water is being up welled to the surface, decreases in oxygen and pH levels are predicted along with an increase in nutrient levels [@rykaczewski].
 Generally, regional ocean climate oscillations like the Benguela Niño for the Benguela Current System impose considerable variability on temperatures and complicate attempts to assess long-term trends [@demarcq].
@@
 In rare cases, projected upwelling increases may overcome the countervailing effects of upper-ocean warming and stratification to cause regional cooling [@auad].
-Research studies reveal declines in ocean oxygen levels as a result of ocean warming and reduced ventilation from stratification and circulation changes [@Keeling].
+Research studies reveal declines in ocean oxygen levels as a result of ocean warming and reduced ventilation from stratification and circulation changes [@keeling].
 Similarly, low oxygen levels are also possible because of shifts in the composition of sinking organic material [@hofmann; @doney].
@@
 The cold Benguela waters are bound in the south by the warm Agulus retroflexion region and in the north by the southward flowing Angola current.
-The flow of the Benguela is often considered to be topographically guided [@nelson; @barange; @barange.
-The Benguela system is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region within the Namibian region [@shannon; Cole 1999).
+The flow of the Benguela is often considered to be topographically guided [@nelson; @barange; @barange].
+The Benguela system is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region within the Namibian region [@shannon; @cole].
 
@@
 Winds occurring within the southern regions represent a seasonal variation, reaching a maximum during spring and summer months [@boyd; @shannon1996; @fennel].
-Winds occurring within this region tend to show increasing patterns away from the coast whereas, winds occurring in the northern region exhibits relatively less seasonal variation [@shannon1985; @pitcher).
+Winds occurring within this region tend to show increasing patterns away from the coast whereas, winds occurring in the northern region exhibits relatively less seasonal variation [@shannon1985; @pitcher].
 Upwelling occurring along the South Coast of the Cape Province within South Africa, specifically along the coastline is generated by local wind.
@@
 Evidently, Benguela Niños are expressed as regions of abnormal, persistent high sea surface temperatures.
-Many researchers often associate the equatorial interannual variability pattern in the Atlantic to the ENSO phenomenon found within the Pacific region (Sutton et al.
-2000; Zebiak 1993).
+Many researchers often associate the equatorial interannual variability pattern in the Atlantic to the ENSO phenomenon found within the Pacific region [@sutton; @zebiak].
 Benguela Niños have been recorded to occur in 1934, 1949, 1963, 1984 and 1995 [@shannon; @gammelsrod].
@@
 Futher studies on extreme events occurred within the Florida Reef Tract, where anomolously cold events resulted in the mass mortality of coral reefs and fish.
-Acropora cervicornis, was found to be particularly susceptible to low temperatures and experienced extensive mortality (Shinn 1966; Davis 1982).
+Acropora cervicornis, was found to be particularly susceptible to low temperatures and experienced extensive mortality [@shinn; @davis].
 While these phenomenons are not as common as warming events, they have a dramtic and long-lasting effect on coral communities.
-However, studies in January 2010 reveal that the Florida Reef Tract experienced a cold-water anomaly, marine cold spells (MCSs) with temperatures below 16 ◦ C recorded for up to 14 days (Lirman et al.
-2011).
+However, studies in January 2010 reveal that the Florida Reef Tract experienced a cold-water anomaly, marine cold spells (MCSs) with temperatures below 16 ◦ C recorded for up to 14 days [@lirman].
 Species population distributon are often set by cold distribution boundaries.
@@
 
-The warming of the global climate as a result of anthropogenic activity, has negatively influenced both terrestrial and marine ecosystems, and ultimately inhibits humanity and ecological functioning (McCauley et al., 2015).
-This climate variability, manifested as increases in frequency and duration of extreme events is regarded as a great threat (Thompsonet al., 2013).
+The warming of the global climate as a result of anthropogenic activity, has negatively influenced both terrestrial and marine ecosystems, and ultimately inhibits humanity and ecological functioning [@mcCauley].
+This climate variability, manifested as increases in frequency and duration of extreme events is regarded as a great threat [@thompsonet].
 Very few studies have addressed the occurrence of cold temperature spells or cold events within the Benguela region.
@@
 
+Missing refs: 
 Picaut, J., 1985: Major dynamics affecting the eastern tropical At-

### 2019-01-25 Methods_tidy.Rmd
updated

@@
 
-ggplot(data = SACTN_3yr, aes(x = date, y = insitu_temp, colour = site)) +
+insitu_temp_plot <- ggplot(data = SACTN_3yr, aes(x = date, y = insitu_temp, colour = site)) +
 geom_line(aes(group = site)) +
@@
 theme(axis.text.x = element_text(angle = 45)) +
-theme_linedraw() +
+#theme_linedraw() +
 theme(legend.position = "top")
+
+# ggsave(plot = insitu_temp_plot, filename = "Figures/insitu_temp.pdf")
 ```
@@
 MUR_plot <- match_plot(df = insitu_MUR)
-ggplot(data = insitu_MUR, aes(x = date, y = temp, colour = site)) +
+mur_temp_plot <- ggplot(data = insitu_MUR, aes(x = date, y = temp, colour = site)) +
 geom_line(aes(group = site)) +
@@
 theme(axis.text.x = element_text(angle = 45)) +
-theme_linedraw() +
+#theme_linedraw() +
 theme(legend.position = "top")
+
+# ggsave(plot = mur_temp_plot, filename = "Figures/mur_temp_plot.pdf")
+
 ```
@@
 # load("~/Documents/MastersProject/Data/OISST_sites.RData")
-
+load("~/Documents/Masters_2019/MastersProject/Data/OISST_sites.RData")
 OISST_sites <- OISST_sites %>% 
@@
 
-ggplot(data = insitu_OISST, aes(x = date, y = temp, colour = site)) +
+OISST_temp_plot <- ggplot(data = insitu_OISST, aes(x = date, y = temp, colour = site)) +
 geom_line(aes(group = site)) +
@@
 theme(axis.text.x = element_text(angle = 45)) +
-theme_linedraw() +
+# theme_linedraw() +
 theme(legend.position = "top")
+# ggsave(plot = OISST_temp_plot, filename = "Figures/OISST_temp_plot.pdf")
 ```
@@
 
-ggplot(data = insitu_G1SST, aes(x = date, y = temp, colour = site)) +
+G1SST_temp_plot <- ggplot(data = insitu_G1SST, aes(x = date, y = temp, colour = site)) +
 geom_line(aes(group = site)) +
@@
 theme(axis.text.x = element_text(angle = 45)) +
-theme_linedraw() +
+# theme_linedraw() +
 theme(legend.position = "top")
+# ggsave(plot = G1SST_temp_plot, filename = "Figures/G1SST_temp_plot.pdf")
 ```

### 2019-02-12 Write_up/Writing/Introduction_outline.Rmd
minor_updates

@@
 Temperature recordings are still ongoing at many of these sites, but unfortunately, the durations and extents of these are uneve per site.
-With the longest time series in the dataset being that of Gordons Bay, recorded by SAWS.
 Data collected for this region started on 13 September 1972 and concluded on 26 January 2017, with recordings still continuing daily.
@@
 The data used within this study and comprehensive script used for data analyses, and production of figures can be found at https://github.com/AmierohAbrahams/MASTERSPROJECT.
+
+## Satellite SST data
+
+This study made use of four satellite-derived Sea surface temperatures (SST) datasets and the SACTN in situ coastal SST have been observed by satellite instruments since 1981 and are regarded as one of the longest satellite-based records of any earth climate variable (REFERENCE).
+The National Oceanic and Atmospheric Administration (NOAA) ¼° daily Optimum Interpolation Sea Surface Temperature (OISST) is an analysis of different platforms (satellites, ships and buoys) on a global grid.
+This gridded daily SST product assimilates both remotely sensed and in situ sources of data to create a level-4 gap free product (Banzon et al., 2016).
+A spatially complete sea suface temperature map is produced by interpolation to fill in gaps within the dataset.
+The method used to fill these gaps includes the adjustment of satellite and ship observations which compensate for the bias differences (REFERENCE).
+There are two types of daily OISST, named after the relevant satellite sensors.
+These are the Advanced Very High Resolution Radiometer (AVHRR) and Advanced Microwave Scanning Radiometer on the Earth Observing System (AMSR-E).
+Currently the AVHRR has the longest record starting from 1981 to the present.
+Infrared instruments, like AVHRR, can make observations at relatively high resolutions.
+Microwave instruments, like AMSR-E, can measure SSTs in most weather conditions but not adjacent to land.
+The AVHRR-only Optimally-Interpolated Sea Surface.
+Temperature (OISST) was used to determine SST within the study region.
+The AVHRR datasets have been provided global SSTs for more than four decades (Reynolds and Smith, 1994; Pearce et al 2006).
+
+The Multi-scale Ultra-high Resolution (MUR) Sea Surface Temperature Analysis, the second dataset, is produced using satellite instruments with datasets spanning 1 June 2002 to present times (PO.DAAC, https://podaac.jpl.nasa.gov/).
+MUR provides SST data at a spatial resolution of 0.01° in longitude-latitude coordinates, roughly at 1 km intervals.
+This dataset is currently among the highest resolution SST datasets available.
+The goal in creating the MUR SST is to develop a coherent and consistent daily map of SST at the highest spatial resolution possible.
+The MUR sea surface temperature dataset combines data from the Advanced Very High Resolution Radiometer (AVHRR), Moderate Imaging Spectroradiometer (MODIS) Terra and Aqua, and Advanced Microwave Spectroradiometer-EOS (AMSR-E) instruments in an optimal way to produce 1km global maps representing the sea surface temperature.
+
+The Moderate-resolution Imaging Spectroradiometer (MODIS) is a scientific instrument (radiometer) on board the NASA Terra and Aqua satellite platforms with a spatial resolution in 0.041° longitude-latitude coordinates.
+MODIS was launched in 1991 and 2002 to study global dynamics of the earths atmosphere, ice and ocean.
+MODIS captures data in 36 spectral bands ranging in wavelength from 0.4 um to 14.4 um and at varying spatial resolutions (2 bands at 250 m, 5 bands at 500 m and 29 bands at 1 km), providing complete global coverage of the Earth every one to two days.
+MODIS Terra Global level 3 mapped thermal SST products consists of sea surface temperature data derived from the 11 and 12um thermal IR infrared (IR) Bands.
+
+The CMC dataset constitutes the fourth dataset and is a version 3.0 Group for High Resolution Sea Surface Temperature (GHRSST) Level 4 dataset with a 10 km resolution constructed by the Canadian Meteorological Center (CMC).
+The CMC dataset combines infrared satellite SST at numerous points in the time series from the AVHRR, the European Meteorological Operational-A (METOP-A) and Operational-B (METOP-B) platforms, and microwave SST data from the Advanced Microwave Scanning Radiometer 2 in conjunction with in situ observations of SST from ships and buoys from the ICOADS program.
+
+This Global 1 km SST (G1SST) analysis, with a 0.01° resolution, uses satellite data from sensors that include the Advanced Very High Resolution Radiometer (AVHRR), the Advanced Along Track Scanning Radiometer (AATSR), the Spinning Enhanced Visible and Infrared Imager (SEVIRI), the Advanced Microwave Scanning Radiometer-EOS (AMSRE), the Tropical Rainfall Measuring Mission Microwave Imager (TMI), the Moderate Resolution Imaging Spectroradiometer (MODIS), the Geostationary Operational Environmental Satellite (GOES) Imager, the Multi-Functional Transport Satellite 1R (MTSAT-1R) radiometer, and in situ data from drifting and moored buoys.
+Thid dataset produces a product that combines satellite data with the model output.
 
@@
 The differing responses of species as a result of climate change are important to understand as these changes affect, among many things, the trophic dynamics and fisheries yield of several regional ecosystems (Murawski 1993).
-
 Understanding the relationship between global warming and climate change as a result of anthropogenic activity is a vast and challenging process.
@@
 Terrestrial ecosystems are more familiar to the general populace; whereas marine ecosystems are associated with high levels of primary production [@Chavez] and are highly at risk as a result of climate change [@osterblom].
-This has been proven to decrease primary production thus affecting the trophic levels and leading to trophic turbulence and destruction of marine fauna [@lewandowska; @dell, @mcCauley].
+Climate change has been proven to decrease primary production thus affecting the trophic levels and leading to trophic turbulence and destruction of marine fauna [@lewandowska; @dell, @mcCauley].
 
@@
 It is possible that equatorial Kelvin waves may play a role in the evolution of Benguela Niños.
+--- Need to make changes to this paragraph
 
@@
 Due to the lack of in situ collected data, cold spells are mainly examined between the period 2014-2017 for four sites along the South African coastline falling within the Benguela upwelling cell.
-Here we aim to assess whether the intensity and duration of these extreme events may show any changes in frequency and intensity on a seasonal basis.
+Here I aim to assess whether the intensity and duration of these extreme events may show any changes in frequency and intensity on a seasonal basis.
 In this research project I make use of various satellite derived observations of sea surface temperature (SST) and wind stress to further investigate the Benguela upwelling system along the South African coastline, whilst considering cold events and warm events occurring within this region

### 2019-02-12 wavelets.Rmd
updated

---
title: "Wavelets"
author: "Amieroh Abrahams"
date: "09 February 2019"
output: html_document
---

Working with remotely-sensed SST and *in-situ* collected coastal seawater temperature, in the analyses bellow I aim to show how the intensity and duration of upwelling events varies. It is important to remember that each of the datasets used within this study obtained SSTs at different resolutions. Here i was only required to plot a time series of 1year.

# Libraries

First I need to find, install and load various packages. These packages will be available on CRAN and can be accessed and installed in the usual way.

```{r prelim_opts, echo=FALSE}
knitr::opts_chunk$set(
  comment = "R>",
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(plyr)
library(lubridate)
library(ggpubr)
library(zoo)
library(lubridate)
library(FNN)
library(forecast)
library(astrochron)
library(WaveletComp)
library(data.table)
```

### Load site list SACTN data
Now to get to the data. The first step involves the loading of the site list. The statistical properties of the seawater temperature representing the South African coastline, such as the mean, minimum and maximum temperatures. These values vary among coastal sections due to the influence of the cold Benguala and warm Agulhas currents. Here we will only focus on the temperature data found along the west coast (wc) (i.e. sites influenced by the Benguela current, EBUS).

```{r}
load("Data/site_list_v4.2.RData")
load("Data/SACTN_daily_v4.2.RData")
```

## Load in the rest of the datasets

* See the Dectect_MCS.Rmd script for more info. The datasets being loaded here are only for temperatures collected between the years 1992-2016. However if a longer time series is required changes may  be made to the Dectect_MCS.Rmd script (line 136)

```{r}
load("~/Documents/Masters_2019/MastersProject/Data/overlap_ts.RData")
load("~/Documents/Masters_2019/MastersProject/Data/insitu_MUR.RData")
load("~/Documents/Masters_2019/MastersProject/Data/insitu_OISST.RData")
load("~/Documents/Masters_2019/MastersProject/Data/insitu_G1SST.RData")
load("~/Documents/Masters_2019/MastersProject/Data/insitu_K10.RData")
```

# Time series of one year

Here I only look at upwelling duration over a period of 1 year. from 2014 - 2015

```{r}
one_year <- function(df){
  upwelling<- df %>% 
  filter(year(date) %in% seq(2014, 2015)) %>% 
  drop_na()
}

SACTN_1yr_US1 <- one_year(df = overlap_ts) 
MUR_1yr_US1 <- one_year(df = insitu_MUR) 
OISST_1yr_US1 <- one_year(df = insitu_OISST) 
G1SST_1yr_US1 <- one_year(df = insitu_G1SST) 
K10_1yr_US1 <- one_year(df = insitu_K10) 
```

# Wavelet analyses

```{r}
temp.d <- function(df){
temp.d <- df %>% 
  dplyr::group_by(site, date) %>% 
  dplyr::summarise(temp = mean(temp, na.rm = T)) %>%
  dplyr::mutate(no = seq(1:n())) %>%
  dplyr::ungroup() %>% 
  dplyr::select(site, no, temp, date)
}

o
temp.d <- temp.d(df = overlap_ts) 
# temp.d <- temp.d(df = insitu_MUR) 
# temp.d <- temp.d(df = insitu_OISST) 
# temp.d <- temp.d(df = insitu_G1SST) 
# temp.d <- temp.d(df = insitu_K10)

temp.d <- temp.d %>% 
  select(site, no, temp, date)

prewhite_fun <- function(x) {
  df <- x[, 2:3]
  out <- prewhiteAR(df, order = 3, method = "mle", aic = TRUE,
             genplot = FALSE, verbose = FALSE)
  colnames(out) <- c("no", "temp")
  return(out)
  }
PN_prewhite <- as.tibble(ddply(temp.d, .(site), prewhite_fun))

ggplot(PN_prewhite, aes(x = no, y = temp)) +
  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ site, nrow = 4) +
  theme_bw()

wl.fun <- function(x) {
  analyze.wavelet(x, "temp", loess.span = 0, dt = 1,
                   dj = 1/50, lowerPeriod = 2, make.pval = TRUE, n.sim = 50, 
                   method = "white.noise", verbose = FALSE)
}

PN_wave <- dlply(PN_prewhite, .(site), wl.fun)

for (i in 1:length(PN_wave)) {
  attributes(PN_wave[[i]]) <- c(attributes(PN_wave[[i]]), ref = names(PN_wave)[i])
}
# x <- PN_wave[[3]]
plot_fun <- function(x, plot_name = attributes(x)$ref) {
  png(filename = paste0("Figures",plot_name,"_wavelet.png"),
      width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
  wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
           timelab = "Days", verbose = FALSE, useRaster = TRUE,
           periodlab = "Period", lwd = 1, graphics.reset = FALSE,
           main = plot_name)
  dev.off()
}
ldply(PN_wave, plot_fun)
```

### 2019-02-12 Detect_MCSs.Rmd
MCS

---
title: "Long_time_series"
author: "Amieroh Abrahams"
date: "04 February 2019"
output: html_document
---

# Time series

The first few lines of this script only looks at a time series of 10 years or greater. However, from line 355 a longer time series is used. This time series has a length of greater than 30 years and so only a few sites are used. With a longer time series I am able to detect for extreme events such as Marine Cold Spells (MCSs).

This script is very repetitive. Wavelet.Rmd creates a function to do this.

# Libraries

First I need to find, install and load various packages. These packages will be available on CRAN and can be accessed and installed in the usual way.

```{r prelim_opts, echo=FALSE}
knitr::opts_chunk$set(
  comment = "R>",
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(plyr)
library(lubridate)
library(ggpubr)
library(zoo)
library(lubridate)
library(FNN)
library(forecast)
library(astrochron)
library(WaveletComp)
library(data.table)
library(heatwaveR)
```

### Load site list SACTN data
Now to get to the data. The first step involves the loading of the site list. The statistical properties of the seawater temperature representing the South African coastline, such as the mean, minimum and maximum temperatures. These values vary among coastal sections due to the influence of the cold Benguala and warm Agulhas currents. Here we will only focus on the temperature data found along the west coast (wc) (i.e. sites influenced by the Benguela current, EBUS). The SACTN dataset comprise of 129 *in situ* coastal seawater temperatures derived from daily measurements over up to 40 years. The SACTN temperature dataset was compiled by measuring coastal temperatures at 129 sites along the coast of South Africa, daily from 1972 until 2017.

```{r load_files1, include=TRUE}
load("Data/site_list_v4.2.RData")
load("Data/SACTN_daily_v4.2.RData")
```

## Selecting the sites and creating new *in situ* datasets
Now we select only the sites occuring along the west coast. Thereafter, I include all the sites which had a time series of 10 years or greater.

```{r}

# # Upwelling Cell 1:
# site_list_sub <- site_list %>%
#   filter(coast == "wc") %>%
#   filter(length > 3650)
# 
# # upwelling_sites <- site_list_sub[-6,] # Here we had two rows of data for Paternoster; i removed the DEA row: All data here are now collected by SAWS
# upwelling_sites <- upwelling_sites[-10:-14,]
# 
# # # # Here I find a time series with the data with a relativley long time series
# UC_1 <- upwelling_sites[c(-2, -3, -5, -6, -8),]
# save(UC_1, file = "Data/UC_1.RData")
# SACTN_daily_temps <- SACTN_daily_v4.2 %>%
#   left_join(site_list[,c(4,13)], by = "index") %>%
#   filter(index %in% upwelling_sites$index)
# save(SACTN_daily_temps, file = "Data/SACTN_daily_temps.RData")
# # 
# SACTN_US1 <- SACTN_daily_temps %>%
#   separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>%
#   dplyr::rename(insitu_temp = temp)
# save(SACTN_US1, file = "Data/SACTN_US1.RData")

# Upwelling Cell 2:
# site_list_sub_UC <- site_list %>%
#   filter(coast == "wc") %>%
#   filter(length > 3650)
# 
# upwelling_sites_UC <- site_list_sub_UC[-1:-8,]
# UC_2 <- upwelling_sites_UC[c(-2,-5),] # New sites falling within the upwelling cell before yzerfontein
# save(UC_2, file = "Data/UC_2.RData")
# 
# 
# SACTN_daily_temps_UC <- SACTN_daily_v4.2 %>%
#   left_join(site_list[,c(4,13)], by = "index") %>%
#   filter(index %in% upwelling_sites_UC$index)
# save(SACTN_daily_temps, file = "Data/SACTN_daily_temps.RData")
# 
# SACTN_US2 <- SACTN_daily_temps_UC %>%
#   separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>%
#   dplyr::rename(insitu_temp = temp)
# 
# save(SACTN_US2, file = "Data/SACTN_US2.RData")

####################################################################################

combined_US <- rbind(UC_1,UC_2)
#save(combined_US, file = "Data/combined_US.RData")
SACTN_temps_US <- SACTN_daily_v4.2 %>%
  left_join(site_list[,c(4,13)], by = "index") %>%
  filter(index %in% combined_US$index)

SACTN_temps_US <- SACTN_temps_US %>%
  separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>%
  dplyr::rename(insitu_temp = temp)

# save(SACTN_temps_US, file = "Data/SACTN_temps_US.RData")
```

I now create two datasets, `SACTN_split_US1` and `SACTN_split_US2`. These two datasets represents different groups of sites within the upwelling center based on the location of each of the sites. `SACTN_split_US1`: Port Nolloth, Lamberts Bay, Saldanna and Yzerfontein. `SACTN_split_US2`: Dassen Island, Sea Point, Hout Bay and Kommetjie. Here I also add monthly, yearly and seasonal column. This allows us to compare temperature variation on a seasonal basis. Research studies done suggests that the intensity and duration of upwelling varies on a seasonal basis.

```{r}
SACTN_monthly <- function(df) {
  SACTN_split_monthly <- df %>%
  mutate(month = month(date, abbr = T, label = T),
         year = year(date)) %>%
  group_by(index, month, year) %>%
    mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",       
                           ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
                                ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
                                       ifelse(month %in% c("Oct", "Nov", "Dec"), "Spring","Error")))))
  return(SACTN_split_monthly)
}

# SACTN_split_US1 <- SACTN_monthly(df = SACTN_US1)
# SACTN_split_US2 <- SACTN_monthly(df = SACTN_US2)
# save(SACTN_split_US1, file = "SACTN_split_US1.RData")
# save(SACTN_split_US2, file = "SACTN_split_US2.RData")
```

Find where different sites have overlapping time series within this `SACTN_temps_US` dataset. Finding a long time series.

```{r}
load("~/Documents/Masters_2019/MastersProject/Data/combined_US.RData")
load("~/Documents/Masters_2019/MastersProject/Data/SACTN_temps_US.RData")
# Find a period where each of the sites within that overlap.

overlap_ts <- SACTN_temps_US %>% 
  filter(year(date) %in% seq(1992, 2016)) %>% # The length of the time may vary depending on the question
  drop_na() # Removing NA values within the dataset
save(overlap_ts, file = "Data/overlap_ts.RData")

temp_plot <- function(df){
  plot <- ggplot(data = df, aes(x = date, y = insitu_temp, colour = site)) +
    geom_line(aes(group = site)) +
    labs(x = "", y = "Temperature (°C)") +
    theme(axis.text.x = element_text(angle = 45)) +
    theme(legend.position = "top")
}

SACTN_temps_US_plot <- temp_plot(df = overlap_ts)
```

In this thesis, different remotely-sensed SST datasets are used. Hence, these datasets represent different resolutions at which data was obtained. We will make use of five remotely-sensed SST datasets, these include; the  MUR, AVHRR OISST, K10 and G1SST data. **MODIS SST data need to be downloaded and used too**

## MUR dataset
```{r}
MUR_Lamberts_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Lamberts Bay_SST_timeseries_5nearest.csv")
MUR_Port_Nolloth <- read_csv("Data/MUR_nearest5pixels/MUR_Port Nolloth_SST_timeseries_5nearest.csv")
MUR_Saldanha_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Saldanha Bay_SST_timeseries_5nearest.csv")
MUR_Yzerfontein <- read_csv("Data/MUR_nearest5pixels/MUR_Yzerfontein_SST_timeseries_5nearest.csv")
MUR_Dassen_Island <- read_csv("Data/MUR_nearest5pixels/MUR_Dassen Island_SST_timeseries_5nearest.csv")
MUR_Hout_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Hout Bay_SST_timeseries_5nearest.csv")
MUR_Kommetjie <- read_csv("Data/MUR_nearest5pixels/MUR_Kommetjie_SST_timeseries_5nearest.csv")
MUR_Sea_Point <- read_csv("Data/MUR_nearest5pixels/MUR_Sea Point_SST_timeseries_5nearest.csv")
MUR_Oudekraal <- read_csv("Data/MUR_nearest5pixels/MUR_Oudekraal_SST_timeseries_5nearest.csv")

MUR_SST <- rbind(MUR_Lamberts_Bay,MUR_Yzerfontein, MUR_Port_Nolloth,MUR_Saldanha_Bay, MUR_Dassen_Island, MUR_Hout_Bay, MUR_Kommetjie, MUR_Sea_Point, MUR_Oudekraal) %>%
  dplyr::rename(site = station)

MUR_SST$date <- (ymd(MUR_SST$date))
# save(MUR_SST, file = "Data/MUR_SST.RData")

load("~/Documents/Masters_2019/MastersProject/Data/MUR_SST.RData")
MUR_SST <- MUR_SST %>% 
  drop_na()
```

Here I write a function to match the remotely-sensed SST to the *in-situ* collected SST

```{r}
MUR_SST <- MUR_SST %>% 
  dplyr::rename(temp = nearest1) # nearest1- 1 pixel as applied to the OISST data

match_func <- function(df){
  match <- overlap_ts  %>%  
  left_join(df, by = c("site", "date")) %>% 
  na.trim()
  return(match)
}
insitu_MUR <- match_func(df = MUR_SST) %>%
  drop_na()
# save(insitu_MUR, file = "Data/insitu_MUR.RData")

match_plot <- function(df){
  plot1 <- df %>% 
ggplot(aes(x = date, y = temp)) +
  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ site, nrow = 2) +
  theme_bw()
  return(plot1)
}
MUR_plot <- match_plot(df = insitu_MUR)

temp_plot <- function(df){
  temp_plot <- df %>% 
  ggplot(aes(x = date, y = temp, colour = site)) +
    geom_line(aes(group = site)) +
    labs(x = "", y = "Temperature (°C)") +
    theme(axis.text.x = element_text(angle = 45)) +
    theme(legend.position = "top")
}
MUR_plot <- temp_plot(df = insitu_MUR)
```

## OISST dataset
```{r}
# # AVHRR <- read_csv("data/csvavhrr-only-v2-19810901-20180630.csv")
# 
# OISSTDir <- "~/Documents/Masters_2019/Data_SST"
# OISST <- fread(paste0(OISSTDir, "/csvavhrr-only-v2-19810901-20180630.csv"),
#             col.names = c("lon", "lat", "temp", "date"))
# 
# #OISST <- as.tibble(fread("~/Documents/Masters_2019/MastersProject/Data/csvavhrr-only-v2-19810901-20180630.csv"))
# # Quick visualisation of the data
# # To explore the data I visualise the min temperatures along the South african coastline.
# 
# OISST %>%
#   filter(date == min(date)) %>%
#   ggplot(aes(x = lon, y = lat)) +
#   geom_raster(aes(fill = temp))
# ggsave(plot = OISST, filename = "Figures/OISST")
```

## Here I find the nearest SST pixels

Now we apply the FNN (Fast Nearest Neighbor) package to determine the nearesr SST pixel to the insitu collected sites.

```{r}
# unique_pixel <- OISST %>%
#   select(lon, lat) %>%
#   unique()
# 
# # Select nearest 1 pixels (k = 1)
# # here we use knnx to find the closes 1 pixels to the insitu sites
# match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
#                           query = as.matrix(combined_US[,5:6]), k = 1)
# 
# # Select SST pixels nearest to insitu sites
# pixel_match <- unique_pixel[match_index,] %>%
#   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
#   mutate(site = combined_US$site)
# 
# ## Subsetting the OISST data to match the upwelling sites within the in situ collected temperature data
# OISST_match <- OISST %>%
#   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
#   filter(combi %in% pixel_match$combi)
# 
# ## Visualisation
# ggplot(data = OISST_match, aes(x = date, y = temp_OISST)) +
#   geom_line() +
#   facet_wrap(~combi)
# ggsave(plot = OISST_match, filename = "Figures/combined_OISST_match")
# 
# OISST_match %>%
#   filter(date == max(date)) %>%
#   ggplot(aes(x = lon, y = lat)) +
#   geom_raster(aes(fill = temp))
```

```{r}
# OISST_sites <- OISST_match %>%
#   left_join(pixel_match, by = c("combi", "lon", "lat")) %>% 
#   dplyr::rename(temp_OISST =temp)
# 
# # Check that the pixel that is shared is duplicated
# # unique(combined_OISST_sites$site)
# # Visualise time series
# ggplot(data = OISST_sites, aes(x = date, y = temp)) +
#   geom_line() +
#   facet_wrap(~site)

# save(OISST_sites, file = "Data/OISST_sites.RData")
load("~/Documents/Masters_2019/MastersProject/Data/OISST_sites.RData")
OISST_sites <- OISST_sites %>% 
  dplyr::rename(temp = temp_OISST) %>%
  dplyr::mutate(date = as.Date(date)) %>% 
  drop_na()
```

## Matching the Insitu data with the OISST SST data
```{r}
insitu_OISST <- match_func(df = OISST_sites)
# save(insitu_OISST, file = "Data/insitu_OISST.RData")
OISST_plot <- match_plot(df = insitu_OISST)
OISST_temp_plot <- temp_plot(df = insitu_OISST)
```

## G1SST dataset
```{r}
Lamberts_Bay <- read_csv("Data/G1SST_sub/Lamberts Bay_SST_timeseries_5nearest.csv")
Port_Nolloth <- read_csv("Data/G1SST_sub/Port Nolloth_SST_timeseries_5nearest.csv")
Saldanha_Bay <- read_csv("Data/G1SST_sub/Saldanha Bay_SST_timeseries_5nearest.csv")
Yzerfontein <- read_csv("Data/G1SST_sub/Yzerfontein_SST_timeseries_5nearest.csv")
Dassen_Island <- read_csv("Data/G1SST_sub/Dassen Island_SST_timeseries_5nearest.csv")
Hout_Bay_SST <- read_csv("Data/G1SST_sub/Hout Bay_SST_timeseries_5nearest.csv")
Kommetjie_SST <- read_csv("Data/G1SST_sub/Kommetjie_SST_timeseries_5nearest.csv")
Oudekraal_SST <- read_csv("Data/G1SST_sub/Oudekraal_SST_timeseries_5nearest.csv")
Sea_Point_SST<- read_csv("Data/G1SST_sub/Sea Point_SST_timeseries_5nearest.csv")

G1SSTsub_SST <- rbind(Lamberts_Bay,Port_Nolloth, Saldanha_Bay, Yzerfontein, Dassen_Island, Hout_Bay_SST, Kommetjie_SST, Oudekraal_SST, Sea_Point_SST) %>%
  dplyr::rename(site = station)

G1SSTsub_SST$date <- (ymd(G1SSTsub_SST$date))
save(G1SSTsub_SST, file = "Data/G1SSTsub_SST.RData")

# load("Data/G1SSTsub_SST.RData")

G1SSTsub_SST <- G1SSTsub_SST %>% 
  drop_na() %>% 
  dplyr::rename(temp = nearest1)

insitu_G1SST <- match_func(df = G1SSTsub_SST) 
# save(insitu_G1SST, file = "Data/insitu_G1SST.RData")
G1SST_plot <- match_plot(df = insitu_G1SST )
G1SST_temp_plot <- temp_plot(df = insitu_G1SST)
```

# K10
```{r}
Lamberts_Bay <- read_csv("Data/K10/K10_Lamberts Bay_SST_timeseries_5nearest.csv")
Port_Nolloth <- read_csv("Data/K10/K10_Port Nolloth_SST_timeseries_5nearest.csv")
Saldanha_Bay <- read_csv("Data/K10/K10_Saldanha Bay_SST_timeseries_5nearest.csv")
Yzerfontein <- read_csv("Data/K10/K10_Yzerfontein_SST_timeseries_5nearest.csv")
Dassen_Island <- read_csv("Data/K10/K10_Dassen Island_SST_timeseries_5nearest.csv")
Hout_Bay <- read_csv("Data/K10/K10_Hout Bay_SST_timeseries_5nearest.csv")
Kommetjie <- read_csv("Data/K10/K10_Kommetjie_SST_timeseries_5nearest.csv")
Oudekraal <- read_csv("Data/K10/K10_Oudekraal_SST_timeseries_5nearest.csv")
Sea_Point <- read_csv("Data/K10/K10_Sea Point_SST_timeseries_5nearest.csv")

K10_SST <- rbind(Lamberts_Bay, Yzerfontein, Port_Nolloth, Saldanha_Bay, Dassen_Island, Hout_Bay, Kommetjie, Oudekraal, Sea_Point) %>%
  dplyr::rename(site = station)
K10_SST$date <- (ymd(K10_SST$date))
# save(K10_SST, file = "Data/K10_SST.RData")

load("~/Documents/Masters_2019/MastersProject/Data/K10_SST.RData")
K10_SST <- K10_SST %>% 
  drop_na()
K10_SST <- K10_SST %>% 
  dplyr::rename(temp = nearest1)

insitu_K10 <- match_func(df = K10_SST) 
# save(insitu_K10, file = "Data/insitu_K10.RData")
K10_plot <- match_plot(df = insitu_K10)
k10_temp_plot <- temp_plot(df = insitu_K10)
```

# Long term time series (>30years)

Working with remotely-sensed SST and *in-situ* collected coastal seawater temperature, in the analyses bellow, I find a long term time series (1996-2016) of seawater temperature.

# Anatomy of marine cold spells within the upwelling center
Make use of the make_whole_fast function within the heatwaveR package. This function constructs a continuous, uninterrupted time series of temperatures (faster).

## Detecting cold spells based on the original datasets

```{r}
site_list_sub <- site_list %>% 
  filter(coast == "wc") %>% 
  filter(length > 10950) #time series of greater than 30years 
# This yield 5sites: Porrt Nolloth, Lamberts Bay, Saldanha Bay, Sea Point and Hout Bay

# # The upwelling cell was then divided into two groups:
# # Port Nolloth, Lamberts Bay and Saldanha Bay was grouped together
# # Sea Point and Hout Bay were grouped together
# US_O1 <- site_list_sub[c(1:3),] 
# US_02 <- site_list_sub[c(4:5),] 
# 
# SACTN_temps_US_O1 <- SACTN_daily_v4.2 %>% 
#   left_join(site_list[,c(4,13)], by = "index") %>% 
#   filter(index %in% US_O1$index)
# # save(SACTN_temps_US_O1, file = "Data/SACTN_temps_US_O1.RData")
# 
# SACTN_temps_US_02 <- SACTN_daily_v4.2 %>% 
#   left_join(site_list[,c(4,13)], by = "index") %>% 
#   filter(index %in% US_02$index)
# # save(SACTN_temps_US_02, file = "Data/SACTN_temps_US_02.RData")

SACTN_temps_US <- SACTN_daily_v4.2 %>% 
  left_join(site_list[,c(4,13)], by = "index") %>% 
  filter(index %in% site_list_sub$index)
# save(SACTN_temps_US, file = "Data/SACTN_temps_US.RData")

SACTN_split <- SACTN_temps_US %>% 
  separate(index, into = c("site", "src"), sep = "/", remove = FALSE)

SACTN_PN <- SACTN_split %>%
  filter(site == "Port Nolloth") %>%
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()

SACTN_LB <- SACTN_split %>%
  filter(site == "Lamberts Bay") %>%
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()

SACTN_SB <- SACTN_split %>%
  filter(site == "Saldanha Bay") %>%
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()

SACTN_SP <- SACTN_split %>%
  filter(site == "Sea Point") %>%
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()

SACTN_HB <- SACTN_split %>%
  filter(site == "Hout Bay") %>%
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()
```

#SACTN
```{r}
ts <- ts2clm(SACTN_PN, climatologyPeriod = c("1974-01-01", "2016-12-31"),
             pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology # Show a portion of the climatology
MCS_events <- res$event[1:5,1:10] # Show some of the cold-spells
# MCS <- MCS[14408:15503,] # Get the years from 2013 - 2015

MCS_SACTN_PN <- event_line(res, spread = 100, metric = "intensity_cumulative",
           start_date = "1974-01-01", end_date = "1977-12-31")

#ggsave(plot = MCS1974, filename = "MCS1974.png")
```

# MUR
```{r}
load("~/Documents/Masters_2019/MastersProject/Data/MUR_SST.RData")
MUR_SST_PN <- MUR_SST %>%
  filter(site == "Port Nolloth") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(MUR_SST_PN, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_MUR_PN <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2011-12-31", end_date = "2014-10-30") +
  ggtitle("MUR_PN") +
  theme(legend = "Bottom")

MUR_SST_LB <- MUR_SST %>% 
  filter(site == "Lamberts Bay") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(MUR_SST_LB, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_MUR_LB <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2011-12-31", end_date = "2014-10-30") +
  ggtitle("MUR_LB")

MUR_SST_YZ <- MUR_SST %>%
  filter(site == "Yzerfontein") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(MUR_SST_YZ, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_MUR_YZ <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2011-12-31", end_date = "2014-10-30") +
  ggtitle("MUR_YZ")

MUR_SST_SB <- MUR_SST %>%
  filter(site == "Saldanha Bay") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(MUR_SST_SB, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_MUR_SB <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2011-12-31", end_date = "2014-10-30") +
  ggtitle("MUR_SB")

MUR_SST_DI <- MUR_SST %>%
  filter(site == "Dassen Island") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(MUR_SST_DI, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_MUR_DI <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2011-12-31", end_date = "2014-10-30") +
  ggtitle("MUR_DI")

MUR_SST_HB <- MUR_SST %>%
  filter(site == "Hout Bay") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(MUR_SST_HB, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_MUR_HB <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2011-12-31", end_date = "2014-10-30") +
  ggtitle("MUR_HB")

MUR_SST_SP <- MUR_SST %>%
  filter(site == "Sea Point") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(MUR_SST_SP, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_MUR_SP <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2011-12-31", end_date = "2014-10-30") +
  ggtitle("MUR_SP")

MUR_SST_O <- MUR_SST %>%
  filter(site == "Oudekraal") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(MUR_SST_O, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_MUR_O <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2011-12-31", end_date = "2014-10-30") +
  ggtitle("MUR_O")

MUR_SST_Kom <- MUR_SST %>%
  filter(site == "Kommetjie") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(MUR_SST_Kom, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_MUR_Kom <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2011-12-31", end_date = "2014-10-30") +
  ggtitle("MUR_Kom")

combined_MUR <- ggarrange(MCS_MUR_PN, MCS_MUR_LB, MCS_MUR_O, MCS_MUR_HB, MCS_MUR_SP, MCS_MUR_Kom, MCS_MUR_DI, MCS_MUR_SB, MCS_MUR_YZ, ncol = 3, nrow = 3)
```

# G1SST
```{r}
load("Data/G1SSTsub_SST.RData")

G1SSTsub_SST_PN <- G1SSTsub_SST %>% 
  filter(site == "Port Nolloth") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(G1SSTsub_SST_PN, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_g1sst_PN <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "1991-09-01", end_date = "1992-12-30") +
  ggtitle("G1SST_PN")

G1SSTsub_SST_LB <- G1SSTsub_SST %>% 
  filter(site == "Lamberts Bay") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(G1SSTsub_SST_LB, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_g1sst_LB <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "1991-09-01", end_date = "1992-12-30") +
  ggtitle("G1SST_LB")

G1SSTsub_SST_SB <- G1SSTsub_SST %>% 
  filter(site == "Saldanha Bay") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(G1SSTsub_SST_SB, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_g1sst_SB <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "1991-09-01", end_date = "1992-12-30") +
  ggtitle("G1SST_SB")

G1SSTsub_SST_YZ <- G1SSTsub_SST %>% 
  filter(site == "Yzerfontein") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(G1SSTsub_SST_YZ, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_g1sst_YZ <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "1991-09-01", end_date = "1992-12-30") +
  ggtitle("G1SST_YZ")

G1SSTsub_SST_HB <- G1SSTsub_SST %>% 
  filter(site == "Hout Bay") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(G1SSTsub_SST_HB, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_g1sst_HB <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "1991-09-01", end_date = "1992-12-30") +
  ggtitle("G1SST_HB")

G1SSTsub_SST_SP <- G1SSTsub_SST %>% 
  filter(site == "Sea Point") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(G1SSTsub_SST_SP, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_g1sst_SP <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "1991-09-01", end_date = "1992-12-30") +
  ggtitle("G1SST_SP")

G1SSTsub_SST_O <- G1SSTsub_SST %>% 
  filter(site == "Oudekraal") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(G1SSTsub_SST_O, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_g1sst_O <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "1991-09-01", end_date = "1992-12-30") +
  ggtitle("G1SST_O")

G1SSTsub_SST_Kom <- G1SSTsub_SST %>% 
  filter(site == "Kommetjie") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(G1SSTsub_SST_Kom, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_g1sst_Kom <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "1991-09-01", end_date = "1992-12-30") +
  ggtitle("G1SST_Kom")

G1SSTsub_SST_DI <- G1SSTsub_SST %>% 
  filter(site == "Dassen Island") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(G1SSTsub_SST_DI, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_g1sst_DI <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "1991-09-01", end_date = "1992-12-30") +
  ggtitle("G1SST_DI")

G1SSTsub_SST_SB <- G1SSTsub_SST %>% 
  filter(site == "Saldanha Bay") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(G1SSTsub_SST_SB, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_g1sst_SB <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "1991-09-01", end_date = "1992-12-30") +
  ggtitle("G1SST_SB")

G1SSTsub_SST_HB <- G1SSTsub_SST %>% 
  filter(site == "Hout Bay") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(G1SSTsub_SST_HB, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_g1sst_HB <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "1991-09-01", end_date = "1992-12-30") +
  ggtitle("G1SST_HB")

combined_G1SST <- ggarrange(MCS_g1sst_PN, MCS_g1sst_LB, MCS_g1sst_O, MCS_g1sst_SP, MCS_g1sst_Kom, MCS_g1sst_DI, MCS_g1sst_YZ, MCS_g1sst_SB, MCS_g1sst_HB, ncol = 3, nrow = 3)
```

# OISST 
```{r}
load("~/Documents/Masters_2019/MastersProject/Data/OISST_sites.RData")
OISST_sites <- OISST_sites %>% 
  dplyr::rename(temp = temp_OISST) %>%
  dplyr::mutate(date = as.Date(date)) %>% 
  drop_na()

OISST_PN <- OISST_sites %>% 
  filter(site == "Port Nolloth") %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(OISST_PN, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_OISST_PN <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2016-09-01", end_date = "2018-06-30") +
  ggtitle("OISST_PN")

OISST_LB <- OISST_sites %>% 
  filter(site == "Lamberts Bay") %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(OISST_LB, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_OISST_LB <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2016-09-01", end_date = "2018-06-30") +
  ggtitle("OISST_LB")

OISST_SB <- OISST_sites %>% 
  filter(site == "Saldanha Bay") %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(OISST_SB, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_OISST_SB <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2016-09-01", end_date = "2018-06-30") +
  ggtitle("OISST_SB")

OISST_YZ <- OISST_sites %>% 
  filter(site == "Yzerfontein") %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(OISST_YZ, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_OISST_YZ <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2016-09-01", end_date = "2018-06-30") +
  ggtitle("OISST_YZ")

OISST_DI <- OISST_sites %>% 
  filter(site == "Dassen Island") %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(OISST_DI, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_OISST_DI <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2016-09-01", end_date = "2018-06-30") +
  ggtitle("OISST_DI")

OISST_SP <- OISST_sites %>% 
  filter(site == "Sea Point") %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(OISST_SP, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_OISST_SP <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2016-09-01", end_date = "2018-06-30") +
  ggtitle("OISST_SP")

OISST_Kom <- OISST_sites %>% 
  filter(site == "Kommetjie") %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(OISST_Kom, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_OISST_Kom <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2016-09-01", end_date = "2018-06-30") +
  ggtitle("OISST_Kom")

OISST_O <- OISST_sites %>% 
  filter(site == "Oudekraal") %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(OISST_O, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_OISST_O <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2016-09-01", end_date = "2018-06-30") +
  ggtitle("OISST_O")

OISST_HB <- OISST_sites %>% 
  filter(site == "Hout Bay") %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(OISST_HB, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
MCS_OISST_HB <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2016-09-01", end_date = "2018-06-30") +
  ggtitle("OISST_HB")

combined_OISST <- ggarrange(MCS_OISST_PN, MCS_OISST_LB, MCS_OISST_SB, MCS_OISST_YZ, MCS_OISST_O, MCS_OISST_HB, MCS_OISST_Kom,MCS_OISST_SP, MCS_OISST_DI)
```

# K10
Time series is less than 10 years and so the detection of MCS is not effective. 
heatwaveR package may be applied to datasets that have a climatological period of >10years
```{r}
# load("~/Documents/Masters_2019/MastersProject/Data/K10_SST.RData")
# 
# K10_PN<- K10_SST %>% 
#   filter(site == "Port Nolloth") %>% 
#   dplyr::rename(temp = nearest1) %>% 
#   select(date, temp) %>% 
#   dplyr::rename(t = date)
# 
# ts <- ts2clm(K10_PN, climatologyPeriod = c("2008-04-02", "2015-12-31"), pctile = 10)
# res <- detect_event(ts, coldSpells = TRUE)
# MCS <- res$climatology
# MCS_K10_PN <- event_line(res, spread = 100, metric = "intensity_cumulative",
#            start_date = "2013-04-02", end_date = "2015-12-31") +
#   ggtitle("K10_PN")
# 
# K10_LB <- K10_SST %>% 
#   filter(site == "Lamberts Bay") %>% 
#   dplyr::rename(temp = nearest1) %>% 
#   select(date, temp) %>% 
#   dplyr::rename(t = date)
# 
# ts <- ts2clm(K10_LB, climatologyPeriod = c("2008-04-02", "2015-12-31"), pctile = 10)
# res <- detect_event(ts, coldSpells = TRUE)
# MCS <- res$climatology
# MCS_K10_LB <- event_line(res, spread = 100, metric = "intensity_cumulative",
#            start_date = "2013-04-02", end_date = "2015-12-31") +
#   ggtitle("K10_LB")
# 
# K10_SB <- K10_SST %>% 
#   filter(site == "Saldanha Bay") %>% 
#   dplyr::rename(temp = nearest1) %>% 
#   select(date, temp) %>% 
#   dplyr::rename(t = date)
# 
# ts <- ts2clm(K10_SB, climatologyPeriod = c("2008-04-02", "2015-12-31"), pctile = 10)
# res <- detect_event(ts, coldSpells = TRUE)
# MCS <- res$climatology
# MCS_K10_SB <- event_line(res, spread = 100, metric = "intensity_cumulative",
#            start_date = "2013-04-02", end_date = "2015-12-31") +
#   ggtitle("K10_SB")
# 
# K10_YZ <- K10_SST %>% 
#   filter(site == "Yzerfontein") %>% 
#   dplyr::rename(temp = nearest1) %>% 
#   select(date, temp) %>% 
#   dplyr::rename(t = date)
# 
# ts <- ts2clm(K10_YZ, climatologyPeriod = c("2008-04-02", "2015-12-31"), pctile = 10)
# res <- detect_event(ts, coldSpells = TRUE)
# MCS <- res$climatology
# MCS_K10_YZ <- event_line(res, spread = 100, metric = "intensity_cumulative",
#            start_date = "2013-04-02", end_date = "2015-12-31") +
#   ggtitle("K10_YZ")
# 
# K10_SP <- K10_SST %>% 
#   filter(site == "Sea Point") %>% 
#   dplyr::rename(temp = nearest1) %>% 
#   select(date, temp) %>% 
#   dplyr::rename(t = date)
# 
# ts <- ts2clm(K10_SP, climatologyPeriod = c("2008-04-02", "2015-12-31"), pctile = 10)
# res <- detect_event(ts, coldSpells = TRUE)
# MCS <- res$climatology
# MCS_K10_SP <- event_line(res, spread = 100, metric = "intensity_cumulative",
#            start_date = "2013-04-02", end_date = "2015-12-31") +
#   ggtitle("K10_SP")
# 
# K10_HB <- K10_SST %>% 
#   filter(site == "Hout Bay") %>% 
#   dplyr::rename(temp = nearest1) %>% 
#   select(date, temp) %>% 
#   dplyr::rename(t = date)
# 
# ts <- ts2clm(K10_HB, climatologyPeriod = c("2008-04-02", "2015-12-31"), pctile = 10)
# res <- detect_event(ts, coldSpells = TRUE)
# MCS <- res$climatology
# MCS_K10_HB <- event_line(res, spread = 100, metric = "intensity_cumulative",
#            start_date = "2013-04-02", end_date = "2015-12-31") +
#   ggtitle("K10_HB")
# 
# K10_Kom <- K10_SST %>% 
#   filter(site == "Kommetjie") %>% 
#   dplyr::rename(temp = nearest1) %>% 
#   select(date, temp) %>% 
#   dplyr::rename(t = date)
# 
# ts <- ts2clm(K10_Kom, climatologyPeriod = c("2008-04-02", "2015-12-31"), pctile = 10)
# res <- detect_event(ts, coldSpells = TRUE)
# MCS <- res$climatology
# MCS_K10_Kom <- event_line(res, spread = 100, metric = "intensity_cumulative",
#            start_date = "2013-04-02", end_date = "2015-12-31") +
#   ggtitle("K10_Kom")
# 
# K10_O <- K10_SST %>% 
#   filter(site == "Oudekraal") %>% 
#   dplyr::rename(temp = nearest1) %>% 
#   select(date, temp) %>% 
#   dplyr::rename(t = date)
# 
# ts <- ts2clm(K10_O, climatologyPeriod = c("2008-04-02", "2015-12-31"), pctile = 10)
# res <- detect_event(ts, coldSpells = TRUE)
# MCS <- res$climatology
# MCS_K10_O <- event_line(res, spread = 100, metric = "intensity_cumulative",
#            start_date = "2013-04-02", end_date = "2015-12-31") +
#   ggtitle("K10_O")
# 
# K10_DI <- K10_SST %>% 
#   filter(site == "Dassen Island") %>% 
#   dplyr::rename(temp = nearest1) %>% 
#   select(date, temp) %>% 
#   dplyr::rename(t = date)
# 
# ts <- ts2clm(K10_DI, climatologyPeriod = c("2008-04-02", "2015-12-31"), pctile = 10)
# res <- detect_event(ts, coldSpells = TRUE)
# MCS <- res$climatology
# MCS_K10_DI <- event_line(res, spread = 100, metric = "intensity_cumulative",
#            start_date = "2013-04-02", end_date = "2015-12-31") +
#   ggtitle("K10_DI")
# 
# 
# combined_K10 <- ggarrange(MCS_K10_PN, MCS_K10_LB, MCS_K10_DI, MCS_K10_SB, MCS_K10_SP, MCS_K10_O, MCS_K10_Kom, MCS_K10_HB, MCS_K10_YZ, ncol = 3, nrow = 3)
```

### 2019-02-13 1.netCDF2csv.R
netCDF

@@ -1,0 +1,122 @@
+# netCDF2csv.R
+
+# NOTES ON USING THIS SCRIPT ----------------------------------------------
+
+# 1. SST data processed by this script can be retrieved from:
+# https://podaac.jpl.nasa.gov/ using the python script `subset_dataset.py`.
+# 2. Subsetting and the selection of the time steps to be done via the python
+# script `subset_dataset.py`.
+# 3. This R script requires the already subsetted netCDFs to reside inside a
+# directory whose path is specified by `in.dir`, below.
+# 4. The .csv files produced will be placed inside of the directory named by
+# `csv.dir` (make sure this directory already exists).
+# 5. The dates that will be included with the final .csv file will be extracted
+# directly from the names of the daily netCDF files; please, therefore, make
+# sure to never change them by manually editing them.
+# 6. The base name of the new .csv file will be partly based on the name of the
+# input netCDF files, with the start and end dates appended at the end. These
+# things are hard coded into the script below.
+# 7. I am sure I have missed some things, or that some things may break somehow;
+# please let me know if this happens and I shall fix it.
+# 8. This file may take a while to run (10s of minutes to hours, depending on
+# the amount of data processed); please be patient while it does its thing.
+
+# Author: AJ Smit
+# Date: 27 April 2018
+# e-mail: ajsmit@uwc.ac.za
+
+
+# CAUTION -----------------------------------------------------------------
+
+# This function will append data to the end of an existing file that had been
+# previously produced by this script. This will result in duplicate data. If you
+# need to rerun the script for some reason, please make sure to delete the file
+# created as the result of the previous run from the `csv.dir`.
+
+
+# LOAD LIBRARIES ----------------------------------------------------------
+
+library(ncdf4) # library for processing netCDFs
+library(plyr) # for `llply()`
+library(data.table) # for fast .csv write function, `fwrite()`
+library(tidyverse) # misc. data processing conveniences
[…]

### 2019-02-18 Tester_code/cold_spells.Rmd
updated

---
title: "cold_spells"
author: "Amieroh Abrahams"
date: "27 January 2019"
output: html_document
---

# Libraries

First I need to find, install and load various packages. These packages will be available on CRAN and can be accessed and installed in the usual way.

```{r prelim_opts, echo=FALSE}
knitr::opts_chunk$set(
  comment = "R>",
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(plyr)
library(lubridate)
library(ggpubr)
library(zoo)
library(lubridate)
library(FNN)
library(data.table)
library(heatwaveR)
```

### Load site list SACTN data

Now to get to the data. The first step involves the loading of the site list. The statistical properties of the seawater temperature representing the South African coastline, such as the mean, minimum and maximum temperatures. These values vary among coastal sections due to the influence of the cold Benguala and warm Agulhas currents. Here we will only focus on the temperature data found along the west coast (wc) (i.e. sites influenced by the Benguela current, EBUS). The SACTN dataset comprise of 129 *in situ* coastal seawater temperatures derived from daily measurements over up to 40 years. The SACTN temperature dataset was compiled by measuring coastal temperatures at 129 sites along the coast of South Africa, daily from 1972 until 2017.

```{r load_files1, include=TRUE}
load("Data/site_list_v4.2.RData")
load("Data/SACTN_daily_v4.2.RData")
```

Now we select only the sites occuring along the west coast. Thereafter, I include all the sites which had a time series of greater than 10 years.

```{r}
site_list_sub <- site_list %>% 
  filter(coast == "wc")

SACTN_daily_wc <- site_list_sub %>% 
  filter(length > 3650)

upwelling_sites <- SACTN_daily_wc[c(-6, -13),]

SACTN_temps <- SACTN_daily_v4.2 %>% 
  left_join(site_list[,c(4,13)], by = "index") %>% 
  filter(index %in% upwelling_sites$index)
# save(SACTN_daily_temps, file = "SACTN_daily_temps.RData")

SACTN_split <- SACTN_temps %>% 
  separate(index, into = c("site", "src"), sep = "/", remove = FALSE)

SACTN_split_PN <- SACTN_split %>%
  filter(site == "Port Nolloth") %>% 
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()
```

Detecting MCS within the *in situ* and remotely SST datasets

```{r}
ts <- ts2clm(SACTN_split_PN, climatologyPeriod = c("1974-01-01", "2016-12-31"),
             pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology # Show a portion of the climatology
MCS_events <- res$event[1:5,1:10] # Show some of the cold-spells
# MCS <- MCS[14408:15503,] # Get the years from 2013 - 2015
```

PLotting

```{r}
MCS1974 <- event_line(res, spread = 100, metric = "intensity_cumulative",
           start_date = "1974-01-01", end_date = "1977-12-31")

#ggsave(plot = MCS1974, filename = "MCS1974.png")
```

```{r}
ts <- ts2clm(SACTN_split_PN, climatologyPeriod = c("1974-01-01", "2016-12-31"),
             pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology # Show a portion of the climatology
res$event[1:5,1:10] # Show some of the cold-spells
# MCS <- MCS[14408:15503,] # Get the years from 2013 - 2015

library(ggplot2)

# ts <- ts2clm(SACTN_split_PN, climatologyPeriod = c("1974-01-01", "2016-12-31"),
#              pctile = 10)
# res <- detect_event(ts)
# event_line(res, spread = 100, metric = "intensity_cumulative",
#            start_date = "2013-12-31", end_date = "2014-12-31")
# 
# event_line(res, spread = 100, start_date = "2013-12-31",
#            end_date = "2014-12-31", category = TRUE)

MCS_SACTN_PN <- ggplot(MCS, aes(x = t, y = temp)) +
geom_flame(aes(y2 = thresh), colour = "blue2", fill = "skyblue1") +
geom_text(aes(x = as.Date("2011-02-01"), y = 20,
              label = "PN_MCS_SACTN")) +
  xlab("Date") + ylab(expression(paste("Temperature [", degree, "C]"))) +
  ggtitle("SACTN")
# 
# ggsave(plot = MCS_SACTN_PN, filename = "Figures/event_detect/PN/MCS_SACTN_PN.pNG")
# ts <- ts2clm(SACTN_split_PN, climatologyPeriod = c("1974-01-01", "2016-12-31"), pctile = 10)
# res <- detect_event(ts, coldSpells = TRUE)
# MCS <- res$event
# 
# loli_SACTN_PN <- ggplot(MCS, aes(x = date_peak, y = duration)) +
#   geom_lolli(aes(colour = intensity_cumulative)) +
#   scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") +
#   xlab("Date") + ylab("Event duration [days]")
# ggsave(plot = loli_SACTN_PN, filename = "Figures/event_detect/PN/loli_SACTN_PN.pNG")

# ggplot(MCS, aes(x = event_no, y = intensity_max)) +
#   geom_lolli(shape = 5, aes(colour = rate_onset), linetype = "dotted") +
#   scale_color_distiller(palette = "RdYlGn", name = "Rate \nonset") +
#   xlab("Event number") + ylab("Max intensity [degree C]")

```

# Repeat this on all of the datasets but only for PN
# MUR
```{r}
load("~/Documents/Masters_2019/MastersProject/Data/MUR_SST.RData")
MUR_SST <- MUR_SST %>%
  filter(site == "Port Nolloth") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(MUR_SST, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
# MCS <- MCS[14408:15503,]

library(ggplot2)

MCS_MUR_PN <- ggplot(MCS, aes(x = t, y = temp)) +
geom_flame(aes(y2 = thresh), colour = "blue2", fill = "skyblue1") +
geom_text(aes(x = as.Date("2011-02-01"), y = 20,
              label = "PN_MCS")) +
  xlab("Date") + ylab(expression(paste("Temperature [", degree, "C]")))
# ggsave(plot = MCS_MUR_PN, filename = "Figures/MCS.pNG")

# Here i get the years from 2013 to 2016
MCS <- MCS[3868:4961,]
MCS_MUR_PN <- ggplot(MCS, aes(x = t, y = temp)) +
geom_flame(aes(y2 = thresh), colour = "blue2", fill = "skyblue1") +
geom_text(aes(x = as.Date("2011-02-01"), y = 20,
              label = "PN_MCS")) +
  xlab("Date") + ylab(expression(paste("Temperature [", degree, "C]"))) +
  ggtitle("MUR")
# ggsave(plot = MCS_MUR_PN, filename = "Figures/event_detect/PN/MCS_MUR_PN.pNG")

ts <- ts2clm(MUR_SST, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$event

loli_MUR_PN <- ggplot(MCS, aes(x = date_peak, y = duration)) +
  geom_lolli(aes(colour = intensity_cumulative)) +
  scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") +
  xlab("Date") + ylab("Event duration [days]")
# ggsave(plot = loli_MUR_PN, filename = "Figures/event_detect/PN/loli_MUR_PN.pNG")
```

#G1SST

```{r}
load("Data/G1SSTsub_SST.RData")

G1SSTsub_SST <- G1SSTsub_SST %>% 
  filter(site == "Port Nolloth") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(G1SSTsub_SST, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
# MCS <- MCS[14408:15503,]

library(ggplot2)

MCS_G1SST_PN <- ggplot(MCS, aes(x = t, y = temp)) +
geom_flame(aes(y2 = thresh), colour = "blue2", fill = "skyblue1") +
geom_text(aes(x = as.Date("2011-02-01"), y = 20,
              label = "PN_MCS")) +
  xlab("Date") + ylab(expression(paste("Temperature [", degree, "C]")))
# ggsave(plot = MCS_MUR_PN, filename = "Figures/MCS.pNG")

# Here i get the years from 2013 to 2016
MCS <- MCS[7797:8888,]
MCS_G1SST_PN <- ggplot(MCS, aes(x = t, y = temp)) +
geom_flame(aes(y2 = thresh), colour = "blue2", fill = "skyblue1") +
geom_text(aes(x = as.Date("2011-02-01"), y = 20,
              label = "PN_MCS")) +
  xlab("Date") + ylab(expression(paste("Temperature [", degree, "C]"))) +
  ggtitle("G1SST")
ggsave(plot = MCS_G1SST_PN, filename = "Figures/event_detect/PN/MCS_G1SST_PN.pNG")

ts <- ts2clm(G1SSTsub_SST, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$event

loli_G1SST_PN <- ggplot(MCS, aes(x = date_peak, y = duration)) +
  geom_lolli(aes(colour = intensity_cumulative)) +
  scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") +
  xlab("Date") + ylab("Event duration [days]")

# ggsave(plot = loli_G1SST_PN, filename = "Figures/event_detect/PN/loli_G1SST_PN.pNG")
```

OISST

```{r}
load("~/Documents/Masters_2019/MastersProject/Data/OISST_sites.RData")
OISST_sites <- OISST_sites %>% 
  dplyr::rename(temp = temp_OISST) %>%
  dplyr::mutate(date = as.Date(date)) %>% 
  group_by(date) %>% 
  drop_na() %>% 
  filter(site == "Port Nolloth") %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(OISST_sites, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
# MCS <- MCS[14408:15503,]

library(ggplot2)

MCS_OISST_PN <- ggplot(MCS, aes(x = t, y = temp)) +
geom_flame(aes(y2 = thresh), colour = "blue2", fill = "skyblue1") +
geom_text(aes(x = as.Date("2011-02-01"), y = 20,
              label = "PN_MCS")) +
  xlab("Date") + ylab(expression(paste("Temperature [", degree, "C]")))
# ggsave(plot = MCS_MUR_PN, filename = "Figures/MCS.pNG")

# Here i get the years from 2013 to 2016
MCS <- MCS[11446:12541,]
MCS_OISST_PN <- ggplot(MCS, aes(x = t, y = temp)) +
geom_flame(aes(y2 = thresh), colour = "blue2", fill = "skyblue1") +
geom_text(aes(x = as.Date("2011-02-01"), y = 20,
              label = "PN_MCS")) +
  xlab("Date") + ylab(expression(paste("Temperature [", degree, "C]"))) +
  ggtitle("OISST")
ggsave(plot = MCS_OISST_PN, filename = "Figures/event_detect/PN/MCS_OISST_PN.pNG")

ts <- ts2clm(G1SSTsub_SST, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$event

loli_G1SST_PN <- ggplot(MCS, aes(x = date_peak, y = duration)) +
  geom_lolli(aes(colour = intensity_cumulative)) +
  scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") +
  xlab("Date") + ylab("Event duration [days]")

#ggsave(plot = loli_G1SST_PN, filename = "Figures/event_detect/PN/loli_G1SST_PN.pNG")
```

```{r}
combined_flame <- ggarrange(MCS_SACTN_PN,MCS_OISST_PN,MCS_G1SST_PN, MCS_MUR_PN )
ggsave(plot = combined_flame, filename = "Figures/event_detect/PN/combined_flame.pNG")
```

### 2019-02-18 wavelets.Rmd
fixedBits

@@
 
-Working with remotely-sensed SST and *in-situ* collected coastal seawater temperature, in the analyses bellow I aim to show how the intensity and duration of upwelling events varies.
+Working with remotely-sensed SST and *in-situ* collected coastal seawater temperature, in the analyses bellow I aim to show how the intensity and duration of upwelling events varies per sites along the west coast of South Africa.
 It is important to remember that each of the datasets used within this study obtained SSTs at different resolutions.
-Here i was only required to plot a time series of 1year.
+Here i was only required to plot a time series of 2 years.
 
@@
 ```{r}
-one_year <- function(df){
+filtered_years <- function(df){
   upwelling<- df %>% 
@@
 
-SACTN_1yr_US1 <- one_year(df = overlap_ts) 
-MUR_1yr_US1 <- one_year(df = insitu_MUR) 
-OISST_1yr_US1 <- one_year(df = insitu_OISST) 
-G1SST_1yr_US1 <- one_year(df = insitu_G1SST) 
-K10_1yr_US1 <- one_year(df = insitu_K10) 
+SACTN_fyears <- filtered_years(df = overlap_ts) 
+MUR_fyears <- filtered_years(df = insitu_MUR) 
+OISST_fyears <- filtered_years(df = insitu_OISST) 
+G1SST_fyears <- filtered_years(df = insitu_G1SST) 
+K10_fyears <- filtered_years(df = insitu_K10) 
 ```
@@
 ```{r}
+
+# When doing a wavelet analyses on the SACTN dataset the "insitu_temp" column needed to be renamed to "temp" for the rest of the code to run.
+The rest of this code is kept constant for each of the datasets (Remotely-sensed SST and In-situ seawater temperature)
+# SACTN_fyears <- SACTN_fyears %>% 
+#   dplyr::rename(temp = insitu_temp)
+
 temp.d <- function(df){
@@
 
-o
-temp.d <- temp.d(df = overlap_ts) 
-# temp.d <- temp.d(df = insitu_MUR) 
-# temp.d <- temp.d(df = insitu_OISST) 
-# temp.d <- temp.d(df = insitu_G1SST) 
-# temp.d <- temp.d(df = insitu_K10)
+# temp.d <- temp.d(df = MUR_fyears) 
+# temp.d <- temp.d(df = OISST_fyears)
+# temp.d <- temp.d(df = G1SST_fyears) 
+temp.d <- temp.d(df = K10_fyears)
 
@@
 plot_fun <- function(x, plot_name = attributes(x)$ref) {
-  png(filename = paste0("Figures",plot_name,"_wavelet.png"),
+  png(filename = paste0("Figures",plot_name,"_wavelets.png"),
       width = 800, height = 600, units = "px", pointsize = 12, bg = "white")

### 2019-02-19 sites_metadata.Rmd
study_site

---
title: "sites_metadata"
author: "Amieroh Abrahams"
date: "19 February 2019"
output: html_document
---

```{r}
library(tidyverse)
library(viridis)
library(plyr)
library(zoo)
library(lubridate)
library(maptools)
library(sp)
library(geosphere)
library(marmap)
library(PBSmapping)
library(scales)
library(grid)
library(gridExtra)
library(mapproj)
```

```{r}
load("~/Documents/Masters_2019/MastersProject/Data/site_list_v4.2.RData")

# # Upwelling Cell 1:
# site_list_sub <- site_list %>%
#   filter(coast == "wc") %>%
#   filter(length > 3650)
#
# # upwelling_sites <- site_list_sub[-6,] # Here we had two rows of data for Paternoster; i removed the DEA row: All data here are now collected by SAWS
# upwelling_sites <- upwelling_sites[-10:-14,]
#
# # # # Here I find a time series with the data with a relativley long time series
# UC_1 <- upwelling_sites[c(-2, -3, -5, -6, -8),]
# save(UC_1, file = "Data/UC_1.RData")
# SACTN_daily_temps <- SACTN_daily_v4.2 %>%
#   left_join(site_list[,c(4,13)], by = "index") %>%
#   filter(index %in% upwelling_sites$index)
# save(SACTN_daily_temps, file = "Data/SACTN_daily_temps.RData")
# #
# SACTN_US1 <- SACTN_daily_temps %>%
#   separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>%
#   dplyr::rename(insitu_temp = temp)
# save(SACTN_US1, file = "Data/SACTN_US1.RData")

# Upwelling Cell 2:
# site_list_sub_UC <- site_list %>%
#   filter(coast == "wc") %>%
#   filter(length > 3650)
#
# upwelling_sites_UC <- site_list_sub_UC[-1:-8,]
# UC_2 <- upwelling_sites_UC[c(-2,-5),] # New sites falling within the upwelling cell before yzerfontein
# save(UC_2, file = "Data/UC_2.RData")
#
#
# SACTN_daily_temps_UC <- SACTN_daily_v4.2 %>%
#   left_join(site_list[,c(4,13)], by = "index") %>%
#   filter(index %in% upwelling_sites_UC$index)
# save(SACTN_daily_temps, file = "Data/SACTN_daily_temps.RData")
#
# SACTN_US2 <- SACTN_daily_temps_UC %>%
#   separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>%
#   dplyr::rename(insitu_temp = temp)
#
# save(SACTN_US2, file = "Data/SACTN_US2.RData")

# combined_US <- rbind(UC_1,UC_2)
# save(combined_US, file = "Data/combined_US.RData")
load("~/Documents/Masters_2019/MastersProject/Data/combined_US.RData")
metaData2 <- combined_US
```

## Calculate mean meta data for each coast and add to metaData2
# Dictate coastal groupings

wc <- c("Hout Bay", "Kommetjie", "Port Nolloth", "Sea Point", "Lamberts Bay", "Saldanha Bay", "Yzerfontein", "Dassen Island", "Oudekraal")

```{r}
west <- metaData2[metaData2$site %in% wc,]
west$coast <- "west"
west2 <- data.frame(site = "west coast", coast = "west", lon = round(mean(west$lon), 4),
                    lat = round(mean(west$lat), 4), depth = round(mean(west$depth), 1),
                   type = NA, 'start date' = min(west$`start date`),'end date' = max(west$`end date`), 
                   length = round(mean(west$length), 0), 'temp days' = round(mean(west$'temp days'), 0), 
                   'NA days' = round(mean(west$'NA days'), 0), 'NA%' = round(mean(west$`NA%`),1), 
                   mean = round(mean(west$mean),1), sd = round(mean(west$sd),1), 
                   min = round(mean(west$min),1), max = round(mean(west$max),1))

names(west2)[c(7:8,10:12)] <- c("start date", "end date", "temp days", "NA days", "NA%")
metaData2 <- data.frame(ID = seq(1, length(metaData2$site),1), metaData2)
metaData2$lon <- round(metaData2$lon,2); metaData2$lat <- round(metaData2$lat,2)
metaData2$depth <- NULL; metaData2$NA.days <- NULL

save(metaData2, file = "Data/metaData2.Rdata")
```

# Creating a quick visualisation

```{r}
load("Data/south_africa_coast.RData")
load("Data/sa_provinces_new.RData")

# Setting the plotting parameters
sa_lats <- c(-35.5, -26); sa_lons <- c(14, 34)

sa_site <- ggplot() + 
  geom_polygon(data = south_africa_coast, aes(x = long, y = lat, group = group), 
               show_guide = FALSE, fill = "grey80") +
  geom_path(data = sa_provinces_new, aes(x = long, y = lat, group = group), colour = "grey50") +
  geom_point(data = metaData2, aes(x = lon, y = lat, colour = metaData2$coast, 
                                   shape = metaData2$type), alpha = 0.8, size = 2.2) +
  guides(colour = guide_legend(title = "coast"),
         shape = guide_legend(title = "type"))

ggsave("Figures/map01.pdf", width = 7.5, height = 4.5, pointsize = 10)
```

### 2019-02-19 studysite_plot.Rmd
study_site

---
title: "Plotting"
author: "Amieroh Abrahams"
date: "18 February 2019"
output: html_document
---

```{r}
library(tidyverse)
library(viridis)
library(plyr)
library(zoo)
library(lubridate)
library(maptools)
library(sp)
library(geosphere)
library(marmap)
library(PBSmapping)
library(scales)
library(grid)
library(gridExtra)
library(mapproj)
library(fossil)
```

# Loading in the data

```{r}
load("Data/metaData2.Rdata")
```

```{r}
# Transects
site_transects <- data.frame()
for(i in 1:length(metaData2$site)){
 site <- metaData2[i,]
 site_transect <- shore.normal.transect(site, 2)
 site_transects <- rbind(site_transects, site_transect)
}

# # Manually correct some transects
site_transects$heading[2:4] <- 290
site_transects$heading[5:6] <- 178
# site_transects$heading[13:14] <- 177.9855
save(site_transects, file = "Data/site_transects.RData")
#write.csv(site_transects, file = "data/site_transects.csv", row.names = F)
load("Data/site_transects.RData")

# Pixel points
site_pixels <- data.frame()
for(i in 1:length(metaData2$site)){
  site <- site_transects[i,]
  site_pixel <- transect.pixel.isobath(site, 25000, -200)
  site_pixels <- rbind(site_pixels, site_pixel)
}
save(site_pixels, file = "Data/site_pixels.RData")
# write.csv(site_pixels, file = "data/site_pixels.csv", row.names = F)
load("Data/site_pixels.RData")

# Bounding box
  # Only one is made in order to know how large the the geom_point() squares should be made to match
test <- data.frame(xmin = destPoint(p = site_pixels[1,2:3], b = 270, d = 12500)[1],
                   xmax = destPoint(p = site_pixels[1,2:3], b = 90, d = 12500)[1],
                   ymin = destPoint(p = site_pixels[1,2:3], b = 180, d = 12500)[2],
                   ymax = destPoint(p = site_pixels[1,2:3], b = 0, d = 12500)[2])

#############################################################################
## Create 200m isobath polygon for plotting

# SA coastline
load("Data/south_africa_coast.RData")
names(south_africa_coast)[1] <- "lon"
south_africa_coast$site <- "SA"

# Manually divide up coastline
wc <- south_africa_coast[291:410,]
sc <- south_africa_coast[132:291,]
ec <- south_africa_coast[23:132,]

# Function for calculating bounding boxes
boundingBoxIso <- function(dat, distance = 400000, isobath = -200){
  #df <- data.frame()
  for(i in rev(1:length(dat$lon))){
    distances <- seq(from = 0, to = distance, by = 1000)
    heading <- shore.normal.transect(dat[i,])
    coords <- data.frame(lon = heading$lon, lat = heading$lat)
    distances2 <- as.data.frame(destPoint(p = coords, b = heading$heading, d = distances))
    sitesIdx <- knnx.index(sa_bathy[,1:2], as.matrix(distances2), k = 1)
    bathy2 <- sa_bathy[sitesIdx,]
    bathy2 <- bathy2[complete.cases(bathy2[,3]),]
    bathy2 <- bathy2[bathy2$depth >= isobath,]
    bathy2 <- bathy2[length(bathy2$depth),]
    if(nrow(bathy2) < 1){
      dat2 <- data.frame(dat[i, ])
    }else{
      dat2 <- cbind(bathy2[,1:2], dat[i,3:8])
    }
    dat <- rbind(dat, dat2)
  }
  return(dat)
}

wcBox <- boundingBoxIso(wc)
wcBox <- wcBox[-c(140:143),] # Remove island
scBox <- boundingBoxIso(sc)
ecBox <- boundingBoxIso(ec)

# isobath200 <- sa_bathy[sa_bathy$depth >= -200,]
# isobath200$coast <- NA
# isobath200$coast[1:1500] <- "west"
# isobath200$coast[1501:3200] <- "south"
# isobath200$coast[3201:4448] <- "east"

#############################################################################
## Create figure showing all the stuffs

# Define plotting parameters
sa_lats <- c(-37, -27); sa_lons <- c(14, 34)

# Plotting with pixel boxes
fig1 <- ggplot() +#coord_equal() + 
  geom_raster(data = sa_bathy, aes(x = lon, y = lat, fill = depth)) +
  #geom_raster(data = isobath200, aes(x = lon, y = lat, fill = coast)) +
  stat_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth, alpha = ..level..), 
               colour = "black", size = 0.2, binwidth = 200, na.rm = TRUE, show_guide = FALSE) +
  geom_polygon(data = south_africa_coast, aes(x = lon, y = lat, group = group), 
               size = 1, colour = "black", fill = "white") +
  geom_path(data = sa_provinces_new, aes(x = long, y = lat, group = group)) +
  geom_polygon(data = wcBox, aes(x = lon, y = lat, group = group), alpha = 0.20, 
               colour = "bisque1", fill = "bisque1") +
  geom_point(data = metaData2, aes(x = lon, y = lat, colour = "white"), alpha = 0.8, size = 1.6) +
  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "white", shape = 0, alpha = 0.8, size = 2.1) +
  geom_rect(data = test, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), 
            alpha = 0.4, colour = "red", size = 0.1, linetype = 1) +
  scale_alpha_continuous(breaks = c(-200, -1000, -2000, -3000, -4000, -5000),
                         guide_legend(title = "depth (m)")) +
  scale_fill_gradient(low = "dodgerblue4", high = "dodgerblue", na.value = "steelblue4", 
                      breaks = c(-1000, -2000, -3000, -4000, -5000),
                      guide_legend(title = "depth (m)")) +
  labs(title = NULL, x = NULL, y = NULL) +
  guides(colour = guide_legend(override.aes = list(size = 3))) +
  #guide_legend(ncol = 2) +
  theme(legend.key = element_rect(colour = NA, size = 0.2),
        legend.key.height = unit(0.4, "cm"),
        legend.background = element_blank()) +
        #legend.justification = c(1,0), legend.position = c(0.5, 0.4)) +
  coord_cartesian(xlim = sa_lons, ylim = sa_lats)
```

### 2019-02-19 1.netCDF2csv.R

@@ -1,122 +1,0 @@
-# netCDF2csv.R
-
-# NOTES ON USING THIS SCRIPT ----------------------------------------------
-
-# 1. SST data processed by this script can be retrieved from:
-# https://podaac.jpl.nasa.gov/ using the python script `subset_dataset.py`.
-# 2. Subsetting and the selection of the time steps to be done via the python
-# script `subset_dataset.py`.
-# 3. This R script requires the already subsetted netCDFs to reside inside a
-# directory whose path is specified by `in.dir`, below.
-# 4. The .csv files produced will be placed inside of the directory named by
-# `csv.dir` (make sure this directory already exists).
-# 5. The dates that will be included with the final .csv file will be extracted
-# directly from the names of the daily netCDF files; please, therefore, make
-# sure to never change them by manually editing them.
-# 6. The base name of the new .csv file will be partly based on the name of the
-# input netCDF files, with the start and end dates appended at the end. These
-# things are hard coded into the script below.
-# 7. I am sure I have missed some things, or that some things may break somehow;
-# please let me know if this happens and I shall fix it.
-# 8. This file may take a while to run (10s of minutes to hours, depending on
-# the amount of data processed); please be patient while it does its thing.
-
-# Author: AJ Smit
-# Date: 27 April 2018
-# e-mail: ajsmit@uwc.ac.za
-
-
-# CAUTION -----------------------------------------------------------------
-
-# This function will append data to the end of an existing file that had been
-# previously produced by this script. This will result in duplicate data. If you
-# need to rerun the script for some reason, please make sure to delete the file
-# created as the result of the previous run from the `csv.dir`.
-
-
-# LOAD LIBRARIES ----------------------------------------------------------
-
-library(ncdf4) # library for processing netCDFs
-library(plyr) # for `llply()`
-library(data.table) # for fast .csv write function, `fwrite()`
-library(tidyverse) # misc. data processing conveniences
[…]

### 2019-02-19 studysite_plot.Rmd
fixed_plot

@@
                colour = "bisque1", fill = "bisque1") +
-  geom_point(data = metaData2, aes(x = lon, y = lat, colour = "white"), alpha = 0.8, size = 1.6) +
+  geom_point(data = metaData2, aes(x = lon, y = lat), alpha = 0.8, size = 1.6) +
   geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "white", shape = 0, alpha = 0.8, size = 2.1) +

### 2019-02-19 studysite_plot.Rmd
Updated

@@
   geom_polygon(data = south_africa_coast, aes(x = lon, y = lat, group = group), 
-               size = 1, colour = "black", fill = "white") +
-  geom_path(data = sa_provinces_new, aes(x = long, y = lat, group = group)) +
+               size = 0.5, colour = "black", fill = "white") +
+  #geom_path(data = sa_provinces_new, aes(x = long, y = lat, group = group)) +
   geom_polygon(data = wcBox, aes(x = lon, y = lat, group = group), alpha = 0.20, 

### 2019-02-19 functions/wind.rose.R
functions

@@ -1,0 +1,136 @@
+# WindRose.R
+require(ggplot2)
+require(RColorBrewer)
+
+plot.windrose <- function(data,
+                          spd,
+                          dir,
+                          spdres = 2,
+                          dirres = 30,
+                          spdmin = 2,
+                          spdmax = 22,
+                          spdseq = NULL,
+                          palette = "YlGnBu",
+                          countmax = NA,
+                          debug = 0){
+  
+  
+  # Look to see what data was passed in to the function
+  if (is.numeric(spd) & is.numeric(dir)) {
+    # assume that we've been given vectors of the speed and direction vectors
+    data <- data.frame(spd = spd,
+                       dir = dir)
+    spd = "spd"
+    dir = "dir"
+  } else if (exists("data")) {
+    # Assume that we've been given a data frame, and the name of the speed 
+    # and direction columns. This is the format we want for later use.    
+  }  
+  
+  # Tidy up input data ----
+  n.in <- NROW(data)
+  dnu <- (is.na(data[[spd]]) | is.na(data[[dir]]))
+  data[[spd]][dnu] <- NA
+  data[[dir]][dnu] <- NA
+  
+  # figure out the wind speed bins ----
+  if (missing(spdseq)) {
+    spdseq <- seq(spdmin,spdmax,spdres)
+  } else {
+    if (debug > 0) {
+      cat("Using custom speed bins \n")
+    }
+  }
+  # get some information about the number of bins, etc.
+  n.spd.seq <- length(spdseq)
+  n.colors.in.range <- n.spd.seq - 1
+  
+  # create the color map
+  spd.colors <- colorRampPalette(brewer.pal(min(max(3,
+                                                    n.colors.in.range),
+                                                min(9,
+                                                    n.colors.in.range)),                                               
+                                            palette))(n.colors.in.range)
+  
+  if (max(data[[spd]],na.rm = TRUE) > spdmax) {    
+    spd.breaks <- c(spdseq,
+                    max(data[[spd]],na.rm = TRUE))
+    spd.labels <- c(paste0(c(spdseq[1:n.spd.seq - 1]),
+                          ':',
+                          c(spdseq[2:n.spd.seq])),
+                    paste0(spdmax,
+                          ":",
[…]

### 2019-02-23 Wind_analyses.Rmd
some bits

---
title: "Wind_analyses"
author: "Amieroh Abrahams"
date: "18 February 2019"
output: html_document
---

# Script Info
The aim of this script is to plot wind rose diagrams representing the wind directions for the sites along the west coast of South Africa. The data used to plot the wind rose diagrams will correspond with the data used to plot the wavelet analyses. Using this we will be able to make deductions as to weather or not the intensity and duration of upwelling varied as a result of wind direction.

The wind data used in this analyses were obtained from the SAWS as well as the NCEP wind data. *Other wind datasets still to be used*

# Libraries

First I need to find, install and load various packages. These packages will be available on CRAN and can be accessed and installed in the usual way.

```{r prelim_opts, echo=FALSE}
knitr::opts_chunk$set(
  comment = "R>",
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(ggpubr)
library(lubridate)
library(stringr)
library(circular)
library(broom)
library(purrr)
library(stlplus)
source("functions/wind.rose.R")
```

## Wind and wave data obtained from SAWS

Creating a function which allows me to load the wind data:

```{r define_fun8, include=TRUE}
load_wave <- function(wave_files) {
  site_name <- sapply(strsplit(as.character(wave_files), "/"), "[[", 3)
  site_info <- sapply(strsplit(as.character(wave_files), "/"), "[[", 4)
  site_info <- sapply(strsplit(site_info, ".txt"), "[[", 1)
  site_info <- strsplit(site_info, "_")
  site_depth <- sapply(site_info, "[[", 1)
  site_num <- sapply(site_info, "[[", 2)

  wave <- read_table(wave_files, col_types = c("dddddd"),
                     col_names = c("date", "hs", "tp", "dir", "dirw", "spw")) %>%
    filter(tp != -999) %>%
    mutate(date = as.POSIXct(as.character(date), "%Y%m%d%H%M", tz = "Africa/Johannesburg")) %>%
    mutate(site = site_name,
           num = site_num,
           depth = site_depth) %>%
    select(site, num, depth, everything()) %>% 
    na.omit()
  return(wave)
}
```

Loadng the wind data for some of the sites along the wc of South Africa

```{r}
sea_point <- map_dfr(dir("Data/wave_data/Sea Point", 
                            full.names = TRUE, pattern = "*.txt"), load_wave)
lamberts_bay <- map_dfr(dir("Data/wave_data/Lamberts Bay", 
                          full.names = TRUE, pattern = "*.txt"), load_wave)
port_nolloth <- map_dfr(dir("Data/wave_data/Port Nolloth", 
                            full.names = TRUE, pattern = "*.txt"), load_wave)
saldanha_bay <- map_dfr(dir("Data/wave_data/Saldanha Bay",
                        full.names = TRUE, pattern = "*.txt"), load_wave)

wave_data <- rbind(sea_point,lamberts_bay,port_nolloth,saldanha_bay)
rm(sea_point,lamberts_bay,port_nolloth,saldanha_bay)
```

Converting the 3-hour resolution wave data into daily data. https://cran.r-project.org/web/packages/circular/circular.pdf (Pg 130). The circular function creates circular objects around the wind and wave direction. The wave and wind direction was collected every three hours and by using the circular function we calcuated the daily mean wave and wind direction. There is a numerically large gap between 360 and 2 where as in degrees its not as large. The circular mean function returns the mean direction of a vector of circular data. Here I also split the daily wave data into depths of 7m and 15m respectivley. For this analyses I will only be working with 7m analyses.

```{r}
wave_daily_dir <- wave_data %>%
  mutate(date = as.Date(date)) %>%
  group_by(site, num, depth, date) %>%
  summarise(dir_circ = mean.circular(circular(dir, units = "degrees")),
            dirw_circ = mean.circular(circular(dirw, units = "degrees"))) 
wave_daily <- wave_data %>%
  mutate(date = as.Date(date)) %>%
  group_by(site, num, depth, date) %>%
  summarise_all(funs(mean = mean, sd = sd), na.rm = T) %>%
  ungroup() %>% 
  left_join(wave_daily_dir)
# save(wave_daily, file = "Data/wave_daily.RData")

load("~/Documents/Masters_2019/MastersProject/Data/wave_daily.RData")
wave_data_7 <- wave_daily %>%
  filter(depth == "7m")
# wave_date_15 <- wave_data %>%
#   filter(depth == "15m")
```

```{r}
load("~/Documents/Masters_2019/MastersProject/Data/temp_2014_2015/G1SST_fyears.RData")
load("~/Documents/Masters_2019/MastersProject/Data/temp_2014_2015/OISST_fyears.RData")
load("~/Documents/Masters_2019/MastersProject/Data/temp_2014_2015/MUR_fyears.RData")
load("~/Documents/Masters_2019/MastersProject/Data/temp_2014_2015/SACTN_fyears.RData")
load("~/Documents/Masters_2019/MastersProject/Data/temp_2014_2015/K10_fyears.RData")
```

Here i match the wave data with the datasets

```{r}
match_wave_7 <- function(df) {
  func1 <- df %>% 
    left_join(wave_data_7, by = c("site", "date")) %>% 
    na.trim() %>% 
    group_by(site)
    return(func1)
}

MUR_wind <- match_wave_7(df = MUR_fyears)
SACTN_wind <-  match_wave_7(df = SACTN_fyears)
OISST_wind <-  match_wave_7(df = OISST_fyears)
G1SST_wind <-  match_wave_7(df = G1SST_fyears)
# K10_wind <-  match_wave_7(df = overlap_ts)
```

# Wind rose diagrams plotting the wind action

```{r}
source("functions/wind.rose.R")

wave_daily_renamed <- MUR_wind %>% 
  dplyr::rename(spd = spw_mean) %>%
  dplyr::rename(dir = dirw_mean)

p.wr2 <- plot.windrose(data = wave_daily_renamed,
              spd = "spd",
              dir = "dir")

p.wr3 <- p.wr2 + facet_wrap(~ site) +
  theme(strip.text.x = element_text(size = 20))
p.wr3

```

## Wavelet analyses on the wind data corresponding to the wavelet data showing temperature

```{r}

# When doing a wavelet analyses on the SACTN dataset the "insitu_temp" column needed to be renamed to "temp" for the rest of the code to run. The rest of this code is kept constant for each of the datasets (Remotely-sensed SST and In-situ seawater temperature)
# SACTN_fyears <- SACTN_fyears %>% 
#   dplyr::rename(temp = insitu_temp)

temp.d <- function(df){
temp.d <- df %>% 
  dplyr::group_by(site, date) %>% 
  dplyr::summarise(dirw_circ = mean(dirw_circ, na.rm = T)) %>%
  dplyr::mutate(no = seq(1:n())) %>%
  dplyr::ungroup() %>% 
  dplyr::select(site, no, dirw_circ, date)
}

temp.d <- temp.d(df = MUR_wind)
# temp.d <- temp.d(df = OISST_fyears)
# temp.d <- temp.d(df = G1SST_fyears) 
# temp.d <- temp.d(df = K10_fyears)

temp.d <- temp.d %>% 
  select(site, no, dirw_circ, date)

prewhite_fun <- function(x) {
  df <- x[, 2:3]
  out <- prewhiteAR(df, order = 3, method = "mle", aic = TRUE,
             genplot = FALSE, verbose = FALSE)
  colnames(out) <- c("no", "temp")
  return(out)
  }
PN_prewhite <- as.tibble(ddply(temp.d, .(site), prewhite_fun))

ggplot(PN_prewhite, aes(x = no, y = dirw)) +
  geom_hline(aes(yintercept = mean(dirw)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ site, nrow = 4) +
  theme_bw()

wl.fun <- function(x) {
  analyze.wavelet(x, "temp", loess.span = 0, dt = 1,
                   dj = 1/50, lowerPeriod = 2, make.pval = TRUE, n.sim = 50, 
                   method = "white.noise", verbose = FALSE)
}

PN_wave <- dlply(PN_prewhite, .(site), wl.fun)

for (i in 1:length(PN_wave)) {
  attributes(PN_wave[[i]]) <- c(attributes(PN_wave[[i]]), ref = names(PN_wave)[i])
}
# x <- PN_wave[[3]]
plot_fun <- function(x, plot_name = attributes(x)$ref) {
  png(filename = paste0("Figures",plot_name,"_wind.png"),
      width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
  wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
           timelab = "Days", verbose = FALSE, useRaster = TRUE,
           periodlab = "Period", lwd = 1, graphics.reset = FALSE,
           main = plot_name)
  dev.off()
}
ldply(PN_wave, plot_fun)
```

### 2019-02-23 Write_up/Writing/Introduction_outline.Rmd
some bits

@@
 
-## Satellite SST data
+## Satellite data
 
-This study made use of four satellite-derived Sea surface temperatures (SST) datasets and the SACTN in situ coastal SST have been observed by satellite instruments since 1981 and are regarded as one of the longest satellite-based records of any earth climate variable (REFERENCE).
+This study made use of four satellite-derived Sea Surface Temperatures (SST) datasets and the SACTN in situ coastal SST have been observed by satellite instruments since 1981 and are regarded as one of the longest satellite-based records of any earth climate variable (REFERENCE).
 The National Oceanic and Atmospheric Administration (NOAA) ¼° daily Optimum Interpolation Sea Surface Temperature (OISST) is an analysis of different platforms (satellites, ships and buoys) on a global grid.
@@
 
+```{r}
+wc <- site_list %>% 
+  filter(coast == "wc")
+save(wc, file = "wc.csv")
+```
+
 Missing refs: 

### 2019-02-23 studysite_plot.Rmd
fixed
some bits

--- 2019-02-19 studysite_plot.Rmd
@@
   coord_cartesian(xlim = sa_lons, ylim = sa_lats)
+#ggsave("Figures/fig1.pdf", width = 7.5, height = 3.5, pointsize = 10)
 ```

### 2019-02-23 wavelets.Rmd
some bits

@@
 K10_fyears <- filtered_years(df = insitu_K10)
+
+save(SACTN_fyears, file = "Data/temp_2014_2015/SACTN_fyears.RData")
+save(MUR_fyears, file = "Data/temp_2014_2015/MUR_fyears.RData")
+save(OISST_fyears, file = "Data/temp_2014_2015/OISST_fyears.RData")
+save(G1SST_fyears, file = "Data/temp_2014_2015/G1SST_fyears.RData")
+save(K10_fyears, file = "Data/temp_2014_2015/K10_fyears.RData")
 ```
@@
 
+# temp.d <- temp.d(df = SACTN_fyears) 
 # temp.d <- temp.d(df = MUR_fyears) 

### 2019-02-23 transcet_func-Rob.R

@@ -1,0 +1,71 @@
+library(FNN)
+library(fossil)
+library(rgeos)
+library(maptools)
+library(geosphere)
+source("function/earthdist.R")
+
+# loading South African Coastline data
+load("~/Documents/Masters_2019/MastersProject/Data/africa_coast.RData")
+
+# Download mid-res bathymetry data
+# sa_lat <- c(-38, -24.5); sa_lon <- c(11.5, 35.5)
+# sa_bathy <- as.xyz(getNOAA.bathy(lon1 = sa_lon[1], lon2 = sa_lon[2], lat1 = sa_lat[1], lat2 = sa_lat[2], resolution = 4))
+# colnames(sa_bathy) <- c("lon", "lat", "depth")
+# sa_bathy <- sa_bathy[sa_bathy$depth <= 0,]
+# save(sa_bathy, file = "data/bathy/sa_bathy.RData")
+load("Data/bathy/sa_bathy.RData")
+plot(as.bathy(sa_bathy), image = T)
+
+# This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
+shore.normal.transect <- function(site, width = 2){
+  # Find the site on the coastline and it's nearest neighbour points
+  coords <- data.frame(lon = site$lon, lat = site$lat)
+  coords2 <- knnx.index(africa_coast[,1:2], as.matrix(coords), k = 1)
+  coords3 <- data.frame(site = site$site, africa_coast[c(coords2-width, coords2+width),]) # This line controls how wide the spread of points used to determine the shore-normal transect is. A wider spread gives a more even, less sensitive transect
+  coords3 <- coords3[2:1,1:3]
+  # Define the shore normal transect bearing
+  heading <- earth.bear(coords3[1,2], coords3[1,3], coords3[2,2], coords3[2,3]) + 90
+  if(heading >= 360){
+    heading <- heading-360
+  } else {
+    heading <- heading
+  }
+  heading2 <- data.frame(site = site$site, lon = site$lon, lat = site$lat, heading)
+  return(heading2)
+}
+
+
+# This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
+# It then extracts a lat/ lon point every X kilometres until reaching a specified isobath
+transect.pixel.isobath <- function(site, distance = 25000, isobath = -200){
+  # Extract coordinates
[…]

### 2019-02-23 Write_up/Writing/Chapter1_Intro.Rmd
upadted

--- 2019-02-23 Write_up/Writing/Introduction_outline.Rmd
@@
 ---
-title: "Introduction_outline"
+title: "Dynamical properties of the Benguela Current as seen through a cold-spell algorithm."
 author: "Amieroh Abrahams"

### 2019-02-24 Write_up/Writing/Chapter1_Intro.Rmd
updated

@@
 
-Chapter 1: General Introduction
+# Chapter 1: Introduction
 
@@
 Regime shifts are often defined as a shift from one stable ecosystem state to another.
-This process affects several species within a short period of time [@cury; @jarre; @blamey).
+This process affects several species within a short period of time [@cury; @jarre; @blamey].
 Regime shifts are thought to be driven by abiotic processes such as climate change [@polovina; @blamey].
@@
 Nevertheless, satellite data products are extremely useful when investigating the ocean at broad scale and can provide important SST data.
-<Will add more on the different types of data and the varying resolutions>
 
@@
 
-This study made use of four satellite-derived Sea Surface Temperatures (SST) datasets and the SACTN in situ coastal SST have been observed by satellite instruments since 1981 and are regarded as one of the longest satellite-based records of any earth climate variable (REFERENCE).
-The National Oceanic and Atmospheric Administration (NOAA) ¼° daily Optimum Interpolation Sea Surface Temperature (OISST) is an analysis of different platforms (satellites, ships and buoys) on a global grid.
-This gridded daily SST product assimilates both remotely sensed and in situ sources of data to create a level-4 gap free product (Banzon et al., 2016).
+Complete datasets are often difficult to
+
+This study made use of various satellite-derived Sea Surface Temperatures (SST) datasets.
+One of these datasets are the National Oceanic and Atmospheric Administration (NOAA) ¼° daily Optimum Interpolation Sea Surface Temperature (OISST).
+This dataset is an analysis of seawater temperature by different platforms (satellites, ships and buoys) on a global grid.
+The gridded daily SST product assimilates both remotely sensed and *in situ* sources of data to create a level-4 gap free product (Banzon et al., 2016).
 A spatially complete sea suface temperature map is produced by interpolation to fill in gaps within the dataset.
-The method used to fill these gaps includes the adjustment of satellite and ship observations which compensate for the bias differences (REFERENCE).
+The method used to fill these gaps includes the adjustment of satellite and ship observations which compensate for the bias differences (NOAA.GOV, https://www.ncdc.noaa.gov/oisst).
 There are two types of daily OISST, named after the relevant satellite sensors.
@@
 Temperature (OISST) was used to determine SST within the study region.
-The AVHRR datasets have been provided global SSTs for more than four decades (Reynolds and Smith, 1994; Pearce et al 2006).
+The AVHRR datasets have been provided global SSTs for more than four decades [@reynolds; @Pearce; @corlett].
 
@@
 MODIS was launched in 1991 and 2002 to study global dynamics of the earths atmosphere, ice and ocean.
-MODIS captures data in 36 spectral bands ranging in wavelength from 0.4 um to 14.4 um and at varying spatial resolutions (2 bands at 250 m, 5 bands at 500 m and 29 bands at 1 km), providing complete global coverage of the Earth every one to two days.
-MODIS Terra Global level 3 mapped thermal SST products consists of sea surface temperature data derived from the 11 and 12um thermal IR infrared (IR) Bands.
+MODIS captures data in 36 spectral bands ranging in wavelength from 0.4 um to 14.4 um and at varying spatial resolutions (2 bands at 250 m, 5 bands at 500 m and 29 bands at 1 km), providing complete global coverage of the Earth every one to two days [@hosoda; @gentemann, @dong].
+MODIS Terra Global level 3 mapped thermal SST products consists of sea surface temperature data derived from the 11 and 12um thermal IR infrared (IR) Bands (NASA.GOV, http://oceancolor.gsfc.nasa.gov).
 
@@
 
-This Global 1 km SST (G1SST) analysis, with a 0.01° resolution, uses satellite data from sensors that include the Advanced Very High Resolution Radiometer (AVHRR), the Advanced Along Track Scanning Radiometer (AATSR), the Spinning Enhanced Visible and Infrared Imager (SEVIRI), the Advanced Microwave Scanning Radiometer-EOS (AMSRE), the Tropical Rainfall Measuring Mission Microwave Imager (TMI), the Moderate Resolution Imaging Spectroradiometer (MODIS), the Geostationary Operational Environmental Satellite (GOES) Imager, the Multi-Functional Transport Satellite 1R (MTSAT-1R) radiometer, and in situ data from drifting and moored buoys.
-Thid dataset produces a product that combines satellite data with the model output.
+The fifth dataset, the Global 1 km SST (G1SST) analysis, representing a 0.01° resolution, uses satellite data from sensors that include the Advanced Very High Resolution Radiometer (AVHRR), the Advanced Along Track Scanning Radiometer (AATSR), the Spinning Enhanced Visible and Infrared Imager (SEVIRI), the Advanced Microwave Scanning Radiometer-EOS (AMSRE), the Tropical Rainfall Measuring Mission Microwave Imager (TMI), the Moderate Resolution Imaging Spectroradiometer (MODIS), the Geostationary Operational Environmental Satellite (GOES) Imager, the Multi-Functional Transport Satellite 1R (MTSAT-1R) radiometer, and in situ data from drifting and moored buoys [@dash}.
+Thid dataset produces a product that combines satellite data with the model output (PO.DAAC, https://podaac.jpl.nasa.gov/).
 
@@
 Increased ocean stratification as a result of surface warming may reduce the ventilation of deep-water masses.
-As the deep water is being up welled to the surface, decreases in oxygen and pH levels are predicted along with an increase in nutrient levels [@rykaczewski].
+As the deep water is being upwelled to the surface, decreases in oxygen and pH levels are predicted along with an increase in nutrient levels [@rykaczewski].
 Generally, regional ocean climate oscillations like the Benguela Niño for the Benguela Current System impose considerable variability on temperatures and complicate attempts to assess long-term trends [@demarcq].
@@
 Research by Mead et al (2013) found that coastal ecosystems are at high risk as a result of climate change, later, Whitfield et al (2016) confirmed that coastal bio-diversity have been observed to be affected.
-Given the importance of primary productivity, it was determined that the Benguela upwelling systerm within the South African region should become the focus of this thesis.
+Given the importance of primary productivity, it was determined that the Benguela upwelling system within the South African region should become the focus of this thesis.
 Here the increase in global temperatures are one of the main focus points along with the chlorophyll concentration.
@@
 This increase will undoubtedly have an impact on the world’s oceans and the marine life within, both directly and indirectly.
-The oceanic process of coastal upwelling is one aspect that is likely to be affected as changes in global temperature will affect temperature and pressure gradients, and in turn affect alongshore winds which are responsible for driving this process within EBUS regions.
-Because upwelling is important for productivity within coastal systems, a decrease in this process will negatively impact on marine fauna and flora, and on human activities such as fishing.
-To assess changes in upwelling events, here I aim to use SST, wind, and chlorophyll data to monitor the durations and intensities of upwelling events within the Benguela EBUS and identify wether this resulted in an extreme anomalous cold event.
+The oceanic process of coastal upwelling is one aspect that is likely to be affected as changes in global temperature will affect temperature and pressure gradients, and in turn affect alongshore winds which are responsible for driving the upwelling process within EBUS regions.
+Because upwelling is important for productivity within coastal systems, a decrease in this process will negatively impact marine fauna and flora, and human activities such as fishing.
+To assess changes in upwelling events, here I aim to use SST, wind, and chlorophyll data to monitor the durations and intensities of upwelling events within the Benguela EBUS and to identify wether this resulted in an extreme anomalous cold event.
 Atmospheric-ocean coupling is known to be affected by climate change, likely changing wind patterns and leading to upwelling favourable conditions (García-Reyes et al., 2015).
@@
 
-This study focusses on the Benguela EBUS.
-The Benguela EBUS is unique in that it is bounded at both the equator ward and pole ward ends by warm water regimes.
-These currents are forced locally by the wind stress field off South-West Africa [@nelson ; @fennel].
-This upwelling system stretches along the south-western coast of Africa from southern Angola to Cape Agulhas [@cole, 1999).
+The area of study for this thesis will focus on the west coast of South Africa, which is an eastern boundry upwelling system.
+The Benguela EBUS is unique in that it is bounded at both the equatorward and poleward ends by warm water regimes.
+These currents are forced locally by the wind stress field off southwest Africa [@nelson ; @fennel].
+This upwelling system stretches along the south-western coast of Africa from southern Angola to Cape Agulhas [@cole].
 Water is coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S [@Lutjeharms].
@@
 The seasonal temperature pattern within the northern Benguela is somewhat similar to that in the eastern equatorial Atlantic and the regular southward intrusion of warm water into the northern Benguela during the summer season.
-During these warming periods the equator ward wind stress over the northern Benguela region was stronger than normal; yet upwelling was suppressed by the southward movement of warm saline waters from the Angolan region.
+During these warming periods the equatorward wind stress over the northern Benguela region was stronger than normal; yet upwelling was suppressed by the southward movement of warm saline waters from the Angolan region.
 
@@
 Winds occurring within this region tend to show increasing patterns away from the coast whereas, winds occurring in the northern region exhibits relatively less seasonal variation [@shannon1985; @pitcher].
-Upwelling occurring along the South Coast of the Cape Province within South Africa, specifically along the coastline is generated by local wind.
+Upwelling occurring along the south coast of the Cape Province within South Africa, specifically along the coastline is generated by local wind.
 The wind increases somewhat away from the coast.
@@
 It is possible that equatorial Kelvin waves may play a role in the evolution of Benguela Niños.
---- Need to make changes to this paragraph
+***Need to make changes to this paragraph, random but feels important***
 
@@
 This investigation examines recent trends and variation in the frequency and duration of cold spells occurring in this region.
-Due to the lack of in situ collected data, cold spells are mainly examined between the period 2014-2017 for four sites along the South African coastline falling within the Benguela upwelling cell.
 Here I aim to assess whether the intensity and duration of these extreme events may show any changes in frequency and intensity on a seasonal basis.
-In this research project I make use of various satellite derived observations of sea surface temperature (SST) and wind stress to further investigate the Benguela upwelling system along the South African coastline, whilst considering cold events and warm events occurring within this region
+In this research project I make use of various satellite derived observations of sea surface temperature (SST) and wind stress to further investigate the Benguela upwelling system along the South African coastline, whilst considering cold events and warm events occurring within this region.
 
-```{r}
-wc <- site_list %>% 
-  filter(coast == "wc")
-save(wc, file = "wc.csv")
-```
+## Study1
+
+## Study 2
 
@@
 lantic and Pacific Oceans. CalCOFI Report, Vol 26.
+
+# Chapter 2: Upwelling
+
+## Introduction
+
+## Methods

### 2019-02-25 Write_up/Writing/dissertation.Rmd
Updated_commit

--- 2019-02-24 Write_up/Writing/Chapter1_Intro.Rmd
@@
 
-Complete datasets are often difficult to
+Reliable datasets are often difficult to maintain and obtain.
+In the field of biological science research, it is often abiotic data products that are impaired thus leading to the use of satellite derived Sst products for the study of specific fields such as coastal and marine ecology as well as oceanography.
+These datasets however, may be inaccurate along the coast and local in situ collected data should be used when possible (Smit et al., 2013).
+Remotely sensed data products are obtained by several different organisations worldwide.
+Satellites launched into space detect the infra-red light emitted by the ocean which is then taken as a proxy from SST.
+Given the extensive distance at which these satellites record data, a certain level of coarseness is displayed within the dataset.
+This is demonstrated as the size of the pixels one sees in the final product.
+
+Satellite SST products undergo a range of quality control and validation processes (e.g.
+Reynolds and Smith, 1994; Brown et al., 1999; Martin et al., 2012), making the data reliable and more useful.
+As previously mentioned, satellites may show a warm or cold bias along the coastal region but they are still useful for broader scale investigations (Smit et al., 2013, Smale and Wernberg, 2009; Castillo and Lima, 2010; Dufois and Rouault).
+Remotely-sensed SST products are obtained at different resolutions, and as a result do not show the same level of bias against in situ collected seawater temperature.
+Research done by Smit et al (2013) showed that MODIS represented the least amount of bias.
 
@@
 
-## Problem statement:
-
-With global warming continuing at a rapid pace, the amount of greenhouse gases within the atmosphere are going to increase in the future.
-This increase will undoubtedly have an impact on the world’s oceans and the marine life within, both directly and indirectly.
-The oceanic process of coastal upwelling is one aspect that is likely to be affected as changes in global temperature will affect temperature and pressure gradients, and in turn affect alongshore winds which are responsible for driving the upwelling process within EBUS regions.
-Because upwelling is important for productivity within coastal systems, a decrease in this process will negatively impact marine fauna and flora, and human activities such as fishing.
-To assess changes in upwelling events, here I aim to use SST, wind, and chlorophyll data to monitor the durations and intensities of upwelling events within the Benguela EBUS and to identify wether this resulted in an extreme anomalous cold event.
-Atmospheric-ocean coupling is known to be affected by climate change, likely changing wind patterns and leading to upwelling favourable conditions (García-Reyes et al., 2015).
-Here we test if an increase in upwelling may lead to a detection of more MCSs simply because these cold temperatures caused by aseasonal upwelling will reflect in historic climatologies.
-
 ## Benguela :
@@
 
+### Event detection
+
+As specified earlier, the extreme cold events occurring along the west coast of South Africa, which a central to this thesis will be detected, and correlated to upwelling events with the algorithm developed by Hobday et al.
+(2016).
+The events that recorde in situ data from SACTN would be compared against those detected in satellite derived SST.
+For this analyses the R package ‘RmarineHeatWaves’ and ‘heatwaveR’ were used to calculate the extreme events.
+In order to obtain the basis for these calculations, the temperature metrics must be aggregated into mean values based on the year in which they occurred.
+With these annual values it is possible to calculate long-term trends in extreme events.
+
+## Analyses
+
+The analyses conducted in this thesis varied according to the investigations performed.
+The one common thread uniting all of the analyses was that they were conducted in R-programming language (R Core Team, 2017).
+This provided reproducibility for all of the analyses and visualisations used.
+
+## Problem statement:
+
+With global warming continuing at a rapid pace, the amount of greenhouse gases within the atmosphere are going to increase in the future.
+This increase will undoubtedly have an impact on the world’s oceans and the marine life within, both directly and indirectly.
+The oceanic process of coastal upwelling is one aspect that is likely to be affected as changes in global temperature will affect temperature and pressure gradients, and in turn affect alongshore winds which are responsible for driving the upwelling process within EBUS regions.
+Because upwelling is important for productivity within coastal systems, a decrease in this process will negatively impact marine fauna and flora, and human activities such as fishing.
+To assess changes in upwelling events, here I aim to use SST, wind, and chlorophyll data to monitor the durations and intensities of upwelling events within the Benguela EBUS and to identify wether this resulted in an extreme anomalous cold event.
+Atmospheric-ocean coupling is known to be affected by climate change, likely changing wind patterns and leading to upwelling favourable conditions (García-Reyes et al., 2015).
+Here we test if an increase in upwelling may lead to a detection of more MCSs simply because these cold temperatures caused by aseasonal upwelling will reflect in historic climatologies.
+
 ## Aims and approach:
@@
 
+The primary aim of the first study was to make use of various remotely sensed SST data, in-situ collected coastal seawater temperature data and chlorophyll-a data.
+It was also useful to examine how the intensity and duration of upwelling varies per dataset as these datasets are collected at different resolutions, with the assumption that high resolution data may reveal a more accurate detection of the duration and intensity of upwelling events.
+Wind data used in these analyses were predicted to show clear patterns between weak and favourable upwelling winds and its impact on the intensity and duration of upwelling.
+Chlorophyll a data are then used to examine whether or not the intensity and duration of upwelling may lead to an increase or a decrease in the chlorophyll a concentration at the various sites.
+It was hypothesised that an increase in the intensity and duration of upwelling will show an increase in the abundance of chlorophyll a concentration.
+The testing of the hypothesis would be made possible with a series of analyses that allows one to select specific time periods from a range of datasets.
+
 ## Study 2
+
+The primary aim of the second study was to determine the occurrence of MCSs along the west coast of Southern Africa between the nearshore (<400m from the shore) and offshore.
+The most appropriate dataset for the measurement of the nearshore temperatures within South Africa is the South African Coastal Temperature Network (SACTN) dataset (Smit et al.
+2013).
+With the knowledge obtained from study 1, it was now possible to investigate whether a more intense upwelling event may lead to an increase in  the intensity and duration of MCSs.
+It is decided that an investigation seeking to answer this, taken from historical records of the occerence of every thermal  event recorded along the west coast of South Africa for both in situ and remotely sensed seawater temperatures.
+The SACTN dataset and a total of five SST datasets could be used to investigate the occurrence of marine cold-spells (MCSs) with the methodology introduced by Hobday et al.
+2016.
+This research also assists in the understanding of  local oceanographic patterns along the west coast of South Africa and the impact anthropogenically forced climate change may have for the recent changes in patterns and long-term trends.
+
+## Research assumptions
+
+The central assumption of this project was that remotely-sensed SST data and in situ seawater temperature data  may be compared.
+Remotely sensed SST data and in situ seawater temperature data were obtained at different resolutions and so slight changes may be found within the different datasets.
+An additional assumption within this research focused on the effect of upwelling on MCSs.
+Very little research has been done on the driving factors leading to MCSs and which processes may result in an increase in the intensity and duration of these events.
+For the purpose of this research  the influence of upwelling on MCSs was the primary focus.
 
@@
 
-# Chapter 2: Upwelling
+### Chapter 2 Upwelling
 
-## Introduction
+##Introduction
+
+Upwelling events are known to have a large influence on the abundance, diversity and distribution of marine organisms and critically affect marine ecosystem processes.
+The changes in the intensity and duration of upwelling events have proven to cause large changes in ecosystem functioning [@wang].
+Globally, there are four major eastern boundary upwelling systems (EBUS).
+Each of these systems are characterized as vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans [@bakun ; @pauly].
+Eastern boundary upwelling regions only cover 1% of the worlds ocean surface, but are among the most productive regions of the world.
+This is because these regions are influenced by upwelling events.
+Upwelling is defined as a process whereby cold, nutrient rich waters are pushed to the surface as a result of alongshore equatorial winds.
+The South African coastline, being approximately 3000km in length, is unique in that it is bordered by the Agulhas current along the east coast and the Benguela current, an eastern boundary upwelling system, along the west coast.
+These regions display noticeable differences in seawater temperatures in comparison to each other, primarily due to the influences of the neighbouring ocean currents (Bolton et al.
+2004; Mead et al.
+2013; Schlegel and Smit 2016).
+The Benguela Current is driven by the anticyclone high pressures systems and eastward moving cyclones which determine the boundaries of the Benguela upwelling region (Shannon 2006; Hutchings et al.
+2009).
+The atmospheric temperatures along the coastline are largely influenced by the Benguela Current [@van], however, upwelling systems may be forced by various atmospheric conditions [@Lutjeharms; @roberts; @hutchings].
+Regionally, the Benguela Current assists in transporting cold water northwards from the Southern Ocean to the coast (Lüning 1990; Lutjeharms, Cooper, and Roberts 2000; Hutchings et al.
+2009; Schlegel et al.
+2017).
+
+Over the past three decades, climate change as a result of anthropogenic activities has established global concern and is shown to have negatively affected marine and terrestrial ecosystems.
+The consequences of these changes are also shown to have direct effects on the behaviour, survival and functioning of marine organisms, thus ultimately affecting species trophic levels and integration [@doney].
+According to the Bakun hypothesis, an increase in greenhouse gases, will result in an increase in day time warming and night time cooling, this ultimately causes an increase in temperature gradients and this further results in a strong pressure gradient.
+Pressure gradients are known to affect wind patterns and wind patterns subsequently control the intensity and duration of upwelling events  [@hsieh; @mote; @lima].
+
+Several research studies provide significant evidence suggesting an increase in ocean surface temperatures.
+These temperatures are agreed to have increased as a result of climate change  [@gruber].
+This warming is shown to slowly move down into the water column thus increasing thermal stratification.
+This thermal stratification restricts the depth at which water is upwelled and thus limits the amount of nutrients brought up to the surface with a decrease in oxygen and pH levels [@chhak; @jacox2011; @jocox2016].
+This process is thus strongly linked to primary productivity.
+Climate change impacting ocean processes has major ecological consequences (Bradley et al.
+1999; Walther et al.
+2002).
+For example, several regions show an increase in sea surface temperatures.
+This rise in temperatures largely influence species survival, reproduction and abundance.
+These changes are known to yield several economic consequences (Murawski 1993), thus, understanding the impact and cause of climate change is a challenging phenomenon.
+Similarly, climate change also largely influences coastal species such as corals, muscles and oysters.
+Given the importance of primary production, it was determined that the Benguela upwelling system, the eastern boundary upwelling system in South Africa, should be the focus of this paper.
+
+The Benguela system is bounded by the equatorward and poleward ends by warm water regimes.
+These currents are shown to be forced by local wind stress fields off southwest Africa[@nelson ; @fennel].
+This EBUS streches along the south-western coast of Africa from southern Angola to Cape Agulhas [@cole].
+Water is coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S [@Lutjeharms].
+The upwelling period in the benguela north of 25°C  persist throughout the entire year but peaks during August.
+The Benguela upwelling region is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region within the Namibian region [@shannon; @cole].
+Several research studies have shown extreme warming and cooling periods occurring in the northern and southern benguela systems [@stander; @walker; @shannon1984; @boyd; @brainard].
+Research studies done by Walter (1937) also show anomously warm periods which may have resulted as a lack of upwelling or poor upwelling events.
+During the periods of warming the equatorward wind stress over the northern region of the benguela system was stronger than usual; yet upwelling was suppressed by the southward movement of warm saline waters from the Angolan region.
+In the southern benguela region, continental shelf bathymetry and strong upwelling favourable winds are known to result in more intense upwelling events.
+Upwelling occuring along the south coastline of the of Southern Africa are particularly generated by local winds.
+
+Global warming is increasing at an extremely high rate.
+This will result in an increase in greenhouse gases which in turn will impact the ocean and terrestrial life.
+The process of upwelling is one such factor that is predicted to change with an increase in global warming.
+As prreviously mentioned, changes in temperature results in changes in pressure gradients, these pressure gradients control wind patterns which ultimately controls the intensity and duration of upwelling.
+To assess the changes within the upwelling events, various datasets would be used.
+These are satellite sea surface temperatures (SST), in situ collected seawater temperatures, wind data as well as satellite obtained chlorophyll-a data.
 
 ## Methods
+
+In order to compare abiotic variables such as wind action and direction along the South African coastline, large historical datasets for temperature and wind were analysed and accessed.
+
+## Wind data
+
+Wind action was an important variable in this study as it was hypothesised to exhibit a direct
+influence on seawater temperatures (Sinnett and Feddersen 2014) and a direct influence on the intensity and duration of upwelling events; consequently, wind action was investigated for its impact on upwelling at specific sites along the west coast of South Africa.
+Wind data was obtained from the South African Weather Service (SAWS), and were provided at three-hour resolutions.
+Specific wind characteristics were measured, namely, wind direction (dirw) and wind speed (spw).
+A resolution of 200m was modelled at both 7 and 15m isobaths.
+
+## Temperature data sources
+
+This research study lies within the borders of South Africa along the west coast in particular.
+The locations of each of the points used within this study can be seen in Fig.1.
+The South African Coastal Temperature Network was the primary dataset used to specify the various sites along the west coast of South Africa.
+This dataset consists of coastal seawater temperatures obtained from 129 sites along the entire South African coastline which is approximately 3000km long and is distinct in that it is bordered by the Benguela and Agulhas Currents (e.g.,Roberts 2005; Hutchings et al.
+2009).
+These currents are known to largely influence the seawater temperature and marine life along the South African coastline.
+The SACTN has collected SST data along this unique coastline from as early as 1972 till present date, with contributions made from several organisations and govermental departments.
+As a result, temperature within this datasets were obtained by using different methods and instruments, such as hand held thermometers or Underwater temperature reorders( UTRs).
+The precision of the instruments used vary accordingly, with hand-held thermometers recording temperature at a precision of 0.5°C as opposed to the underwater temperature recorders which have an instrument precision of 0.001°C.
+The duration and extent of recordings per sites were uneven with the longest time series in the dataset being that of Gordons Bay which started on 13 September 1972 and concluded on 26 January 2017.
+During the 1970s, a total of 11 time series began recording.
+A further 53 entries were added during the 1980s, 34 entries were added during the 1990s, and 18 entries were added during the 2000s.
+Recordings are still ongoing at many of these sites.
+
+This study also made use of a range of remotely sensed temperature datasets.
+Satellites provide a system of large spatial and temporal data retrieval.
+Remotely sensed SST products are produced by a number of organisations across the globe.
+The AVHRR-only Optimally-Interpolated Sea Surface Temperature (OISST) was used to determine SST within the study region.
+The AVHRR datasets have been provided global SSTs for more than four decades (Reynolds and Smith, 1994; Pearce et al 2006).
+OISST is a global 1/4° gridded daily SST product that assimilates both remotely sensed and in situ sources of data to create a level-4 gap free product (Banzon et al., 2016).
+The Multi-scale Ultra-high Resolution (MUR) Sea Surface Temperature Analysis, the second dataset, is produced using satellite instruments with datasets spanning 1 June 2002 to present times (PO.DAAC, https://podaac.jpl.nasa.gov/).
+MUR provides SST data at a spatial resolution of 0.01° in longitude-latitude coordinates and is currently among the highest resolution SST datasets available.
+The CMC dataset constitutes the third dataset and is a version 3.0 Group for High Resolution Sea Surface Temperature (GHRSST) Level 4 dataset with a 10km resolution constructed by the Canadian Meteorological Center (CMC).
+The CMC dataset combines infrared satellite SST at numerous points in the time series from the AVHRR, the European Meteorological Operational-A (METOP-A) and Operational-B (METOP-B) platforms, and microwave SST data from the Advanced Microwave Scanning Radiometer 2 in conjunction with in situ observations of SST from ships and buoys from the ICOADS program.
+
+## Chlorophyll-a data
+
+In order to observe the patterns in the variation of the chlorophyll-a concentration with the intensity and duration of upwelling events,  MODIS-Aqua Chlorophyll a concentration, OCI Algorithm, 8-day composite were used.
+The National Aeronautics and Space Administration’s (NASA) Moderate Resolution imaging Spectroradiometer (MODIS) Aqua satellite is an example of a satellite currently collecting the reflected wavelength energy from the earth.
+The MODIS Aqua satellite records wavelength data down to the 1% photosynthetically available depth (0.1 PAR depth) of the water column, a vital component in the estimation of chlorophyll (NASA.GOV, http://oceancolor.gsfc.nasa.gov).
+Because chlorophyll-a concentration retrieval is only possible over cloud-free areas the daily data offered may have large data gaps due to cloud cover.
+The quality of the chlorophyll-a concentration  estimates depends strongly on the quality of the cloud mask.
+The NASA Aqua satellite contains the MODIS sensor that is used for detecting chlorophyll-a concentrations (ocean color) in the world's oceans, among other applications (NASA.GOV, http://oceancolor.gsfc.nasa.gov).
+The Level 3 standard mapped image (SMI) chlorophyll-a dataset has a monthly temporal resolution and 4.6 km spatial resolution.
+The SMI dataset is an image representation of binned MODIS data.
+
+## Site selection
+
+In order to compare the impact of wind action on the intensity and duration of upwelling, the appropriate sites along the west coast of South Africa (i.e.
+sites influenced by the Benguela current) were selected.
+Since temperature data were not evenly recorded for each of the sites, Firstly the full dataset was narrowed to only those sites that could be adequately compared.
+To do this, the filter () function in R was used.
+This function selects all the sites with a time series of greater than 10 years.
+Once the sites were minimised, one years worth of data was selected from the SACTN data between the years 2014 – 2015.
+This yield a total of  seven sites along the west coast of South Africa.
+Hereafter, each of the remotely sensed SST datasets obtained were matched with the SACTN dataset.
+This was done for two reasons, to find the exact same points at which temperature was collected for the SACTN sites and secondly, to confirm that the dates and time that temperature were collected correspond accordingly.

### 2019-02-26 Chlorophyll_a .Rmd
chloro

---
title: "Chlorophylla"
author: "Amieroh Abrahams"
date: "18 February 2019"
output: html_document
---

The aim of this script is to extract the chlorophyll a netCDF files and convert it to CSV. The chlorophylla data was obtained from MODIS Aqua. The aim of working with this chlorophyll a data is to analyse how it changes/varies with upwelling.

```{r}
library(ncdf4)
library(data.table)
library(tidyverse)
library(reshape2)
library(plyr)
library(lubridate)
library(stringr)
library(doMC); doMC::registerDoMC(cores = 4)
library(ggpubr)
library(fasttime)
library(FNN)

MODIS_chlor.dir <- "/home/amieroh/Documents/Data/Datasets/Chlorophyll_a"
MODIS_chlor.csv.dir <- "/home/amieroh/Documents/Data/Datasets"

region <- "BC" # Benguela Current
coords <- bbox[, region]
coords <- c(-35, -20, 10, 20) # this is the BC

ncList <- list.files(path = MODIS_chlor.dir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
strt.date <- str_sub(basename(ncList[1]), start = 2, end = 8)
end.date <- str_sub(basename(ncList[length(ncList)]), start = 2, end = 8)
nc.init <- nc_open(ncList[1])
LatIdx <- which(nc.init$dim$lat$vals > coords[1] & nc.init$dim$lat$vals < coords[2])
LonIdx <- which(nc.init$dim$lon$vals > coords[3] & nc.init$dim$lon$vals < coords[4])
nc_close(nc.init)

ncFun <- function(nc.file = nc.files, csv.dir = csv.dir) {
nc <- nc_open(nc.file)
  instrument <- ncatt_get(nc, 0, "instrument")$value
  platform <- ncatt_get(nc, 0, "platform")$value
  product_name <- ncatt_get(nc, 0, "product_name")$value
  fNameStem <- substr(product_name, 17, 38)
  timeStamp <- substr(product_name, 2, 8)
  origin <- paste0(substr(timeStamp, 1, 4), "-01-01")
  date <- as.Date(as.numeric(substr(timeStamp, 5, 7)), origin)
  chl <- round(ncvar_get(nc,
                   varid = "chlor_a",
                   start = c(LonIdx[1], LatIdx[1]),
                   count = c(length(LonIdx), length(LatIdx))),
               3)
  dimnames(chl) <- list(lon = nc$dim$lon$vals[LonIdx],
                        lat =  nc$dim$lat$vals[LatIdx])
  nc_close(nc)
  chl <-
    as.data.table(melt(chl, value.name = "chl"), row.names = NULL) %>%
    mutate(t = date) %>%
    na.omit()
  fwrite(chl,
         file = paste(csv.dir, "/", region, "-", instrument, ".",platform, ".",
                      fNameStem, "-", strt.date, "-", end.date, ".csv", sep = ""),
         append = TRUE, col.names = FALSE)
  rm(chl)
}

llply(ncList, ncFun, csv.dir = MODIS_chlor.csv.dir, .parallel = TRUE)
```

## Setting up the chlorophylla dataset

```{r}
MODIS_Chloro <- "~/Documents/Masters_2019/MastersProject/Data"
MODIS_Chloro <- fread(paste0(MODIS_Chloro, "/BC-MODIS.Aqua.L3m_8D_CHL_chlor_a_9km-2002185-2018345.csv"),
            col.names = c("lon", "lat", "chloro", "date"))

# save(MODIS_Chloro, file = "Data/MODIS_Chloro.RData")
load("Data/MODIS_Chloro.RData")
load("Data/combined_US.RData")

daily_chloro_data <- MODIS_Chloro %>%
  dplyr::mutate(date = as.Date(date)) %>% 
  dplyr::group_by(lon, lat, date) %>% 
  mutate(month = month(date),
         day = day(date),
         year = year(date)) %>% 
  dplyr::summarise(chloro = mean(chloro, na.rm = TRUE))

# fastDate <- function(x, tz = NULL) {
#   as.Date(fastPOSIXct(x, tz = tz))
# }
# 
# daily2monthly <- function(dailyData) {
#   monthlyData <- dailyData %>%
#     dplyr::mutate(time = fastDate(time)) %>%
#     dplyr::mutate(month = floor_date(time, unit = "month")) %>%
#     dplyr::group_by(lon, lat, month) %>%
#     dplyr::summarise(chlor_a = mean(chlor_a, na.rm = TRUE)) %>%
#     dplyr::mutate(year = year(month)) %>%
#     dplyr::group_by(lon, lat) %>%
#     dplyr::mutate(num = seq(1:length(chlor_a))) %>%
#     dplyr::ungroup()
#   return(monthlyData)
# }
#  daily2monthly(MODIS_Chloro)

# Pixels and matching
# match_index <- knnx.index(data = as.matrix(daily_chloro[,4:5]),
#                           query = as.matrix(combined_US[,5:6]), k = 1)
# 
# # # Select SST pixels nearest to insitu sites
# pixel_match <- daily_chloro[match_index,] %>%
#   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
#   unite(col = day, year, month, sep = "-", remove = F) %>% 
#   mutate(site = combined_US$site)
# 
# daily_chloro_match <- daily_chloro %>%
#   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
#   unite(col = day, year, month, sep = "-", remove = F) %>% 
#   filter(combi %in% pixel_match$combi)
# 
# daily_chloro_match_sites <- daily_chloro_match %>%
#   left_join(pixel_match, by = c("combi", "lon", "lat")) %>%
#   group_by(day, site) %>% 
#   dplyr::summarise(chloro = mean(chloro, na.rm = TRUE))
```

### 2019-02-27 Chlorophyll_a .Rmd
Updated

@@
 The chlorophylla data was obtained from MODIS Aqua.
-The aim of working with this chlorophyll a data is to analyse how it changes/varies with upwelling.
+The aim of working with this chlorophyll a data is to examine how the chlorophyll a  concentration varies with upwelling.
 
@@
 
+Chlorophyll-a data are 8day composites.
+In order to compare the chlorophyll concentration with upwelling events, take 8days, take the average.
+This is one value.
+Repeat this so what i have a time series with chlorophyll-a data every 8days.
+
+Plot this using a line graph?... Upwelling evetns were plot using wavelet analyses
+
 ```{r}
-MODIS_Chloro <- "~/Documents/Masters_2019/MastersProject/Data"
-MODIS_Chloro <- fread(paste0(MODIS_Chloro, "/BC-MODIS.Aqua.L3m_8D_CHL_chlor_a_9km-2002185-2018345.csv"),
-            col.names = c("lon", "lat", "chloro", "date"))
+# MODIS_Chloro <- "~/Documents/Masters_2019/MastersProject/Data"
+# MODIS_Chloro <- fread(paste0(MODIS_Chloro, "/BC-MODIS.Aqua.L3m_8D_CHL_chlor_a_9km-2002185-2018345.csv"),
+#             col.names = c("lon", "lat", "chloro", "date"))
 
@@
   dplyr::mutate(date = as.Date(date)) %>% 
-  dplyr::group_by(lon, lat, date) %>% 
   mutate(month = month(date),
@@
          year = year(date)) %>% 
+  dplyr::group_by(lon, lat, day, month, year) %>% 
   dplyr::summarise(chloro = mean(chloro, na.rm = TRUE))
 
-# fastDate <- function(x, tz = NULL) {
-#   as.Date(fastPOSIXct(x, tz = tz))
-# }
-# 
-# daily2monthly <- function(dailyData) {
-#   monthlyData <- dailyData %>%
-#     dplyr::mutate(time = fastDate(time)) %>%
-#     dplyr::mutate(month = floor_date(time, unit = "month")) %>%
-#     dplyr::group_by(lon, lat, month) %>%
-#     dplyr::summarise(chlor_a = mean(chlor_a, na.rm = TRUE)) %>%
-#     dplyr::mutate(year = year(month)) %>%
-#     dplyr::group_by(lon, lat) %>%
-#     dplyr::mutate(num = seq(1:length(chlor_a))) %>%
-#     dplyr::ungroup()
-#   return(monthlyData)
-# }
-#  daily2monthly(MODIS_Chloro)
+# Pixels and matching
+match_index <- knnx.index(data = as.matrix(daily_chloro_data[,1:2]),
+                          query = as.matrix(combined_US[,5:6]), k = 1)
 
-# Pixels and matching
-# match_index <- knnx.index(data = as.matrix(daily_chloro[,4:5]),
-#                           query = as.matrix(combined_US[,5:6]), k = 1)
-# 
 # # # Select SST pixels nearest to insitu sites
-# pixel_match <- daily_chloro[match_index,] %>%
-#   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
-#   unite(col = day, year, month, sep = "-", remove = F) %>% 
-#   mutate(site = combined_US$site)
-# 
-# daily_chloro_match <- daily_chloro %>%
-#   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
-#   unite(col = day, year, month, sep = "-", remove = F) %>% 
-#   filter(combi %in% pixel_match$combi)
-# 
-# daily_chloro_match_sites <- daily_chloro_match %>%
-#   left_join(pixel_match, by = c("combi", "lon", "lat")) %>%
-#   group_by(day, site) %>% 
-#   dplyr::summarise(chloro = mean(chloro, na.rm = TRUE))
+pixel_match <- daily_chloro_data[match_index,] %>%
+  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
+  mutate(site = combined_US$site)
+
+daily_chloro_match <- daily_chloro_data %>%
+  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
+  filter(combi %in% pixel_match$combi)
+
+daily_chloro_match_sites <- daily_chloro_match %>%
+  left_join(pixel_match, by = c("combi", "lon", "lat", "month", "day", "year")) 
 ```

### 2019-02-27 Chlorophyll_a .Rmd
A rough fix to create daily cholorphyl-a data from the 8 day data

@@
 The aim of this script is to extract the chlorophyll a netCDF files and convert it to CSV.
-The chlorophylla data was obtained from MODIS Aqua.
-The aim of working with this chlorophyll a data is to examine how the chlorophyll a  concentration varies with upwelling.
+The chlorophylla data were obtained from MODIS Aqua.
+The aim of working with these chlorophyll a data is to examine how the chlorophyll a concentration varies with upwelling.
 
@@
 library(reshape2)
-library(plyr)
+library(plyr) # Never load plyr when also loading the tidyverse. It causes a lot of conflicts.
 library(lubridate)
@@
 library(FNN)
+library(heatwaveR)
 
@@
 
-Chlorophyll-a data are 8day composites.
-In order to compare the chlorophyll concentration with upwelling events, take 8days, take the average.
+Chlorophyll-a data are 8 day composites.
+In order to compare the chlorophyll concentration with upwelling events, take 8 days, take the average.
 This is one value.
-Repeat this so what i have a time series with chlorophyll-a data every 8days.
+Repeat this so what I have a time series with chlorophyll-a data every 8 days.
 
@@
 
-daily_chloro_data <- MODIS_Chloro %>%
-  dplyr::mutate(date = as.Date(date)) %>% 
-  mutate(month = month(date),
-         day = day(date),
-         year = year(date)) %>% 
-  dplyr::group_by(lon, lat, day, month, year) %>% 
-  dplyr::summarise(chloro = mean(chloro, na.rm = TRUE))
-
 # Pixels and matching
@@
 
-daily_chloro_match <- daily_chloro_data %>%
+daily_chloro_match <- MODIS_Chloro %>%
   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
@@
 
-daily_chloro_match_sites <- daily_chloro_match %>%
-  left_join(pixel_match, by = c("combi", "lon", "lat", "month", "day", "year")) 
+# With the chloro pixels filter, it is now faster to create the daily data
+daily_chloro_data <- daily_chloro_match %>%
+  dplyr::mutate(date = as.Date(date)) %>% 
+  dplyr::group_by(combi, lon, lat, date) %>% 
+  dplyr::summarise(chloro = mean(chloro, na.rm = TRUE)) %>% 
+  dplyr::group_by(combi, lon, lat) %>% 
+  tidyr::nest() %>% 
+  dplyr::mutate(clims = purrr::map(data, ts2clm, x = date, y = chloro,
+                       climatologyPeriod = c("2002-07-13", "2018-12-04"), # This doesn't mean anything, but the function needs it to work
+                       maxPadLength = 8)) %>% # Adjust maxPadLength = X for longer interpolation
+  dplyr::select(-data) %>% 
+  tidyr::unnest() %>% 
+  select(-doy, -seas, -thresh)
+
+ggplot(daily_chloro_data, aes(x = date, y = chloro)) +
+  geom_line() +
+  facet_wrap(~combi, ncol = 1)
+
+# daily_chloro_match_sites <- daily_chloro_data %>%
+#   left_join(pixel_match, by = c("combi", "lon", "lat", "month", "day", "year")) 
 ```

### 2019-02-28 Chlorophyll_a .Rmd
updated

@@
 
-# Pixels and matching
-match_index <- knnx.index(data = as.matrix(daily_chloro_data[,1:2]),
+unique_pixel <- MODIS_Chloro %>%
+  select(lon, lat) %>%
+  unique()
+
+# # Select nearest 1 pixels (k = 1)
+# # here we use knnx to find the closes 1 pixels to the insitu sites
+match_index <- knnx.index(data = as.matrix(MODIS_Chloro[,1:2]),
                           query = as.matrix(combined_US[,5:6]), k = 1)
@@
 # # # Select SST pixels nearest to insitu sites
-pixel_match <- daily_chloro_data[match_index,] %>%
+pixel_match <- unique_pixel[match_index,] %>%
   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
@@
 ```
+
+```{r}
+daily_chloro_data <- MODIS_Chloro %>%
+  dplyr::mutate(date = as.Date(date)) %>% 
+  mutate(month = month(date),
+         day = day(date),
+         year = year(date)) %>% 
+  dplyr::group_by(lon, lat, day, month, year) %>% 
+  dplyr::summarise(chloro = mean(chloro, na.rm = TRUE))
+
+match_index <- knnx.index(data = as.matrix(daily_chloro_data[,1:2]),
+                          query = as.matrix(combined_US[,5:6]), k = 1)
+
+pixel_match <- daily_chloro_data[match_index,] %>%
+  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
+  mutate(site = combined_US$site)
+
+daily_chloro_match <- MODIS_Chloro %>%
+  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
+  filter(combi %in% pixel_match$combi)
+
+# With the chloro pixels filter, it is now faster to create the daily data
+daily_chloro_data <- daily_chloro_match %>%
+  dplyr::mutate(date = as.Date(date)) %>% 
+  dplyr::group_by(combi, lon, lat, date) %>% 
+  dplyr::summarise(chloro = mean(chloro, na.rm = TRUE)) %>% 
+  dplyr::group_by(combi, lon, lat) %>% 
+  tidyr::nest() %>% 
+  dplyr::mutate(clims = purrr::map(data, ts2clm, x = date, y = chloro,
+                       climatologyPeriod = c("2002-07-13", "2018-12-04"), # This doesn't mean anything, but the function needs it to work
+                       maxPadLength = 8)) %>% # Adjust maxPadLength = X for longer interpolation
+  dplyr::select(-data) %>% 
+  tidyr::unnest() %>% 
+  select(-doy, -seas, -thresh)
+```

### 2019-02-28 Chlorophyll_a .Rmd
....

--- 2019-02-27 Chlorophyll_a .Rmd
@@
 
-# Pixels and matching
+unique_pixel <- MODIS_Chloro %>%
+  select(lon, lat) %>%
+  unique()
+
+daily_chloro_data <- MODIS_Chloro %>%
+  dplyr::mutate(date = as.Date(date)) %>% 
+  mutate(month = month(date),
+         day = day(date),
+         year = year(date)) %>% 
+  dplyr::group_by(lon, lat, day, month, year) %>% 
+  dplyr::summarise(chloro = mean(chloro, na.rm = TRUE))
+
 match_index <- knnx.index(data = as.matrix(daily_chloro_data[,1:2]),
@@
 
-# # # Select SST pixels nearest to insitu sites
 pixel_match <- daily_chloro_data[match_index,] %>%

### 2019-02-28 Chlorophyll_a .Rmd
Update
fixed....maybe..

--- 2019-02-27 Chlorophyll_a .Rmd
@@
 
-# Pixels and matching
-match_index <- knnx.index(data = as.matrix(daily_chloro_data[,1:2]),
+match_index <- knnx.index(data = as.matrix(MODIS_Chloro[,1:2]),
                           query = as.matrix(combined_US[,5:6]), k = 1)
 
-# # # Select SST pixels nearest to insitu sites
-pixel_match <- daily_chloro_data[match_index,] %>%
+pixel_match <- MODIS_Chloro[match_index,] %>%
   unite(col = combi, lon, lat, sep = "/", remove = F) %>%

### 2019-02-28 Chlorophyll_a .Rmd
fixed

@@
 
+#Plotting daily data
 ggplot(daily_chloro_data, aes(x = date, y = chloro)) +
+  geom_line() +
+  facet_wrap(~combi, ncol = 1)
+
+# Plotting only years 2014-2015
+daily_chloro_data_tester <- daily_chloro_data %>% 
+  filter(year(date) %in% seq(2014, 2015))
+
+ggplot(daily_chloro_data_tester, aes(x = date, y = chloro)) +
   geom_line() +

### 2019-03-01 Detect_MCSs.Rmd
fixed
study_site
updated

@@
 
-# Time series
+# Script Info (Time series)
 
 The first few lines of this script only looks at a time series of 10 years or greater.
-However, from line 355 a longer time series is used.
+However, from line 360 a longer time series is used.
 This time series has a length of greater than 30 years and so only a few sites are used.
@@
 ```{r}
-# # AVHRR <- read_csv("data/csvavhrr-only-v2-19810901-20180630.csv")
-# 
+# AVHRR <- read_csv("data/csvavhrr-only-v2-19810901-20180630.csv")
+
 # OISSTDir <- "~/Documents/Masters_2019/Data_SST"
@@
 #             col.names = c("lon", "lat", "temp", "date"))
-# 
-# #OISST <- as.tibble(fread("~/Documents/Masters_2019/MastersProject/Data/csvavhrr-only-v2-19810901-20180630.csv"))
+
+OISST <- as.tibble(fread("Data/csvavhrr-only-v2-19810901-20180630.csv"))
 # # Quick visualisation of the data

### 2019-03-02 Chlorophyll_a .Rmd
updated
progress

--- 2019-02-28 Chlorophyll_a .Rmd
@@
   dplyr::summarise(chloro = mean(chloro, na.rm = TRUE)) %>% 
-  dplyr::group_by(combi, lon, lat) %>% 
-  tidyr::nest() %>% 
-  dplyr::mutate(clims = purrr::map(data, ts2clm, x = date, y = chloro,
-                       climatologyPeriod = c("2002-07-13", "2018-12-04"), # This doesn't mean anything, but the function needs it to work
-                       maxPadLength = 8)) %>% # Adjust maxPadLength = X for longer interpolation
-  dplyr::select(-data) %>% 
-  tidyr::unnest() %>% 
-  select(-doy, -seas, -thresh)
+  dplyr::group_by(combi, lon, lat) #%>% 
+  # tidyr::nest() %>% 
+  # dplyr::mutate(clims = purrr::map(data, ts2clm, x = date, y = chloro,
+  #                      climatologyPeriod = c("2002-07-13", "2018-12-04"), # This doesn't mean anything, but the function needs it to work
+  #                      maxPadLength = 8)) %>% # Adjust maxPadLength = X for longer interpolation
+  # dplyr::select(-data) %>% 
+  # tidyr::unnest() %>% 
+  # select(-doy, -seas, -thresh)
+s
+#Plotting daily data
+# ggplot(daily_chloro_data, aes(x = date, y = chloro)) +
+#   geom_line() +
+#   facet_wrap(~combi, ncol = 1)
 
-ggplot(daily_chloro_data, aes(x = date, y = chloro)) +
+# Plotting only years 2014-2015
+daily_chloro_data_tester <- daily_chloro_data %>% 
+  filter(year(date) %in% seq(2014, 2015))
+
+rough_plot <- ggplot(daily_chloro_data_tester, aes(x = date, y = chloro)) +
   geom_line() +

### 2019-03-02 wavelets.Rmd
progress

@@
 ```{r}
-load("~/Documents/Masters_2019/MastersProject/Data/overlap_ts.RData")
-load("~/Documents/Masters_2019/MastersProject/Data/insitu_MUR.RData")
-load("~/Documents/Masters_2019/MastersProject/Data/insitu_OISST.RData")
-load("~/Documents/Masters_2019/MastersProject/Data/insitu_G1SST.RData")
-load("~/Documents/Masters_2019/MastersProject/Data/insitu_K10.RData")
+load("Data/overlap_ts.RData")
+load("Data/insitu_MUR.RData")
+load("Data/insitu_OISST.RData")
+load("Data/insitu_G1SST.RData")
+load("Data/insitu_K10.RData")
 ```
@@
 
-save(SACTN_fyears, file = "Data/temp_2014_2015/SACTN_fyears.RData")
-save(MUR_fyears, file = "Data/temp_2014_2015/MUR_fyears.RData")
-save(OISST_fyears, file = "Data/temp_2014_2015/OISST_fyears.RData")
-save(G1SST_fyears, file = "Data/temp_2014_2015/G1SST_fyears.RData")
-save(K10_fyears, file = "Data/temp_2014_2015/K10_fyears.RData")
+# save(SACTN_fyears, file = "Data/temp_2014_2015/SACTN_fyears.RData")
+# save(MUR_fyears, file = "Data/temp_2014_2015/MUR_fyears.RData")
+# save(OISST_fyears, file = "Data/temp_2014_2015/OISST_fyears.RData")
+# save(G1SST_fyears, file = "Data/temp_2014_2015/G1SST_fyears.RData")
+# save(K10_fyears, file = "Data/temp_2014_2015/K10_fyears.RData")
 ```
@@
 The rest of this code is kept constant for each of the datasets (Remotely-sensed SST and In-situ seawater temperature)
-# SACTN_fyears <- SACTN_fyears %>% 
-#   dplyr::rename(temp = insitu_temp)
+SACTN_fyears <- SACTN_fyears %>%
+  dplyr::rename(temp = insitu_temp)
 
@@
 
-# temp.d <- temp.d(df = SACTN_fyears) 
+temp.d <- temp.d(df = SACTN_fyears) 
 # temp.d <- temp.d(df = MUR_fyears) 
@@
 # temp.d <- temp.d(df = G1SST_fyears) 
-temp.d <- temp.d(df = K10_fyears)
+# temp.d <- temp.d(df = K10_fyears)
 
@@
 plot_fun <- function(x, plot_name = attributes(x)$ref) {
-  png(filename = paste0("Figures",plot_name,"_wavelets.png"),
+  png(filename = paste0("Figures/new",plot_name,"_wavelets.png"),
       width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
@@
 ```
+
+## Examining seasonal upwelling
+
+Creating a seasonal column.
+Allowing me to compare the intensity/duration of upwelling on a seasonal basis.
+This should to be repeated for each of the sites within each of the datasets (SACTN, MUR, OISST, G1SST and K10).
+The code bellow needs to be edited.
+Here I only focused on Port Nolloth within the SACTN dataset.
+
+```{r}
+load("Data/temp_2014_2015/G1SST_fyears.RData")
+load("Data/temp_2014_2015/K10_fyears.RData")
+load("Data/temp_2014_2015/MUR_fyears.RData")
+load("Data/temp_2014_2015/OISST_fyears.RData")
+load("Data/temp_2014_2015/SACTN_fyears.RData")
+
+seasonal_col <- function(df){
+  temp_data <- df %>% 
+    mutate(month = month(date, abbr = T, label = T),
+           year = year(date)) %>% 
+    group_by(site, month, year) %>% 
+     mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",        
+                           ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
+                                ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
+                                       ifelse(month %in% c("Oct", "Nov", "Dec"), "Spring","Error")))))
+  return(temp_data)
+}
+
+SACTN_seasonal <- seasonal_col(df = SACTN_fyears)
+MUR_seasonal <- seasonal_col(df = MUR_fyears)
+OISST_seasonal <- seasonal_col(df = OISST_fyears)
+G1SST_seasonal <- seasonal_col(df = G1SST_fyears)
+K10_seasonal <- seasonal_col(df = K10_fyears)
+
+# save(SACTN_seasonal, file = "Data/temp_2014_2015/SACTN_seasonal.RData")
+# save(MUR_seasonal, file = "Data/temp_2014_2015/MUR_seasonal.RData")
+# save(OISST_seasonal, file = "Data/temp_2014_2015/OISST_seasonal.RData")
+# save(G1SST_seasonal, file = "Data/temp_2014_2015/G1SST_seasonal.RData")
+# save(K10_seasonal, file = "Data/temp_2014_2015/K10_seasonal.RData")
+
+load("Data/temp_2014_2015/SACTN_seasonal.RData")
+load("Data/temp_2014_2015/MUR_seasonal.RData")
+load("Data/temp_2014_2015/OISST_seasonal.RData")
+load("Data/temp_2014_2015/G1SST_seasonal.RData")
+load("Data/temp_2014_2015/K10_seasonal.RData")
+```
+
+A seasonal plotting representing the sea surface temperature for 2014 and 2015
+
+```{r}
+PN_14 <- SACTN_seasonal %>% 
+  filter(site == "Port Nolloth") %>% 
+  filter(year == 2014 )
+
+PN_15 <- SACTN_seasonal %>% 
+  filter(year == 2015 )
+  # filter(site == "Port Nolloth") %>% 
+  # filter(year == 2015 )
+
+combined_seasonal <- function(df){
+  plot <- df %>% 
+    ggplot(aes(x = date, y = insitu_temp)) +
+    geom_line() +
+    facet_wrap(~season, ncol = 2, nrow = 2) +
+    theme_bw()
+}
+
+combined_seasonal_2014 <- combined_seasonal(df = PN_14)
+combined_seasonal_2015 <- combined_seasonal(df = PN_15)
+combined_PortNolloth <- ggarrange(combined_seasonal_2014, combined_seasonal_2015)
+
+# ggsave(plot = combined_PortNolloth, filename = "Figures/combined_PortNolloth.png")
+```
+
+Creating a wavelet analyses for the year 2015. This code creates a plot per season.
+
+```{r}
+seasonal_PN_15 <- PN_15 %>% 
+  dplyr::rename(temp = insitu_temp)
+
+temp.d <- function(df){
+temp.d <- df %>% 
+  dplyr::group_by(site, date, season) %>% 
+  dplyr::summarise(temp = mean(temp, na.rm = T)) %>%
+  dplyr::mutate(no = seq(1:n())) %>%
+  dplyr::ungroup() %>% 
+  dplyr::select(season, site, no, temp)
+}
+
+temp.d <- temp.d(df = seasonal_PN_15)
+
+temp.d <- temp.d %>% 
+  select(season, site, no, temp)
+
+prewhite_fun <- function(x) {
+  df <- x[, 3:4]
+  out <- prewhiteAR(df, order = 3, method = "mle", aic = TRUE,
+             genplot = FALSE, verbose = FALSE)
+  colnames(out) <- c("no", "temp")
+  return(out)
+  }
+PN_prewhite <- as.tibble(ddply(temp.d, .(site, season), prewhite_fun))
+
+temp_seasonal <- ggplot(temp.d, aes(x = date, y = temp)) +
+  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
+  geom_line() +
+  facet_wrap(~ season, nrow = 4) +
+  theme_bw()
+
+wl.fun <- function(x) {
+  analyze.wavelet(x, "temp", loess.span = 0, dt = 1,
+                   dj = 1/50, lowerPeriod = 2, make.pval = TRUE, n.sim = 50, 
+                   method = "white.noise", verbose = FALSE)
+}
+
+PN_wave <- dlply(temp.d, .(site, season), wl.fun)
+
+for (i in 1:length(PN_wave)) {
+  attributes(PN_wave[[i]]) <- c(attributes(PN_wave[[i]]), ref = names(PN_wave)[i])
+}
+# x <- PN_wave[[3]]
+plot_fun <- function(x, plot_name = attributes(x)$ref) {
+  png(filename = paste0("Figures/new",plot_name,"_wavelets.png"),
+      width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
+  wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
+           timelab = "Days", verbose = FALSE, useRaster = TRUE,
+           periodlab = "Period", lwd = 1, graphics.reset = FALSE,
+           main = plot_name)
+  dev.off()
+}
+ldply(PN_wave, plot_fun)
+```

### 2019-03-03 Write_up/Writing/dissertation.Rmd
updated

@@
 
-## Eastern boundary upwelling systems:
+## Eastern boundary upwelling systems
 
 Globally, there are four major eastern boundary upwelling systems (EBUS).
-These are the Benguela Current off south-western Africa, the Canary Current off north-western Africa including its northern extension off the Iberian Peninsula of south-western Europe, the Peru-Humboldt Current off western South America, and the California Current off the western continental USA and north-western Mexico [@bakun ; @pauly].
+Upwelling systems include the Benguela Current off south-western Africa, the Canary Current off north-western Africa together with its northern extension off the Iberian Peninsula of south-western Europe, the Peru-Humboldt Current off western South America, and the California Current off the western continental USA and north-western Mexico [@bakun ; @pauly].
 Each of these systems are characterized as vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans.
-EBUS cover only 1% of the world’s ocean surface, but despite these are among the most productive regions of the world [@pauly].
-This is because these systems are subject to coastal upwelling, which is defined as the process whereby cold, nutrient rich, high CO2 concentrated, low pH, and low oxygen waters are pushed to the surface as a result of alongshore winds interacting with the earth’s rotation [@huyer].
-As a result of upwelling, diverse marine fauna and flora are able to thrive in these areas because of the abundance of nutrients and favourable temperatures.
+EBUS cover only 1% of the world’s ocean surface, but despite this these are among the most productive regions of the world [@pauly].
+This is due to the  systems exposure to coastal upwelling, which is defined as the process whereby cold, nutrient rich, high concentrated CO2, low pH, and low oxygenated waters are pushed to the surface as a result of alongshore winds interacting with the earth’s rotation [@huyer].
+As a result of upwelling, diverse marine fauna and flora are able to thrive in these areas due to the abundance of nutrients and favourable temperatures.
 
@@
 These winds are caused by cross-shore atmospheric pressure gradients, and these gradients occur predominantly during heating periods.
-Bakun then proposed that if there is an increase in greenhouse gases, it will cause more heating during the day and more cooling during the night which cause the temperature gradient to increase and would result in a strong pressure gradient, and as we know the pressure gradient will affect the wind and therefore will affect upwelling.
-Several people have tried to model this using different datasets of SST and wind pattern data at various resolution and initially found that increases in gases, such as CO^2 causes a reduction in trade winds, and therefore a reduction in upwelling.
-However, these models were based on coarse data and when higher resolutions are used, such as for the Benguela EBUS, the models have shown that upwelling will intensify in accordance to the upwelling intensification hypotheses [@hsieh; @mote; @lima].
+Bakun proposed that if there was an increase in greenhouse gases, it would cause more heating during the day and more cooling during the night causing the temperature gradient to increase resulting in a strong pressure gradient, Ultimately effecting wind patterns and therefore upwelling.
+Several people have tried to model this using different datasets of SST and wind pattern data at various resolutions and initially found that increases in gases, such as CO^2 causes a reduction in trade winds, and therefore a reduction in upwelling.
+These models however, were based on coarse data and when higher resolutions were used, such as for the Benguela EBUS, the models have shown that upwelling will intensify in accordance to the upwelling intensification hypotheses [@hsieh; @mote; @lima].
 
-### Ecology and primary production within upwelling zones:
+### Ecology and primary production within upwelling zones
 
-The duration and intensity of coastal upwelling are known to have a critical role in the phenology of key marine ecosystem processes.
-Similarly, changes in the characteristics of upwelling are shown to have cause large disturbances to ecosystems at several trophic levels [@wang].
+The duration and intensity of coastal upwelling are known to play a critical role in the phenology of key marine ecosystem processes.
+Similarly, changes in the characteristics of upwelling are shown to cause large disturbances to ecosystems at several trophic levels [@wang].
 Cold, nutrient rich, low pH, and low oxygen upwelled waters fuels primary production as a consequence of the development of phytoplankton blooms [@ryther; @Cushing; @mcGillicuddy; @oschlies].
@@
 Mesoscale and sub-mesoscale oceanic movements are key in managing the exchange and transportation of organic resources and marine organisms from the coastal zone to the open ocean.
-Most of the world’s major fishing industries are dependent on upwelling, because this process brings nutrient rich waters into the euphotic zone, where it forms the foundation of a localized food web that ecologically differs from surrounding, non-upwelled regions [@bosman; @barshis].
+Most of the world’s major fishing industries are dependent on upwelling, as this process brings nutrient rich waters into the euphotic zone, where it forms the foundation of a localized food web that ecologically differs from surrounding, non-upwelled regions [@bosman; @barshis].
 Whilst nutrients play a primary role in the survival of these systems, the cool temperatures of these waters are critical as well.
-There is an increasingly large body of research that has provided evidence showing that sea surface temperatures (SST) have and are continuing to increase.
-However, some regions still show a lower or non-significant increase, and in some cases cooling trends are evident (REFERENCES).
+There is an increasingly large body of research that has provided evidence showing that sea surface temperatures (SST) have and are continuing to increase, however some regions still show a lower or non-significant increase, and in some cases cooling trends are evident (REFERENCES).
 Upwelling occurring along the Cape south coast is proven to have less of an economic importance in contrast to the west coast.
-Inshore regions and embayments that are protected from wind action may be subjected to intense episodes of primary production [@graham].
+Inshore regions and embayments that are protected from wind action are likely to be subjected to intense episodes of primary production [@graham].
 Vander Wouder et al.
@@
 The southern Benguela upwelling system supports a wide variety of commercially-exploited fish species and exhibits substantial variability in both oceanographic and biological components.
-Over the past 3 three decades, a number of species occurring within this region have undergone shifts in their distributions and Roy et al.
-(2007) linked these shifts to cooling waters along the inner shelf of the Agulhas bank on the south coast, east of Cape Agulhas [@blamey].
+Over the past  three decades, a number of species occurring within this region have undergone shifts in their distributions and Roy et al.
+(2007) associated these shifts to cooling waters along the inner shelf of the Agulhas bank on the south coast, east of Cape Agulhas [@blamey].
 This coast is associated with high levels of harmful toxic algal blooms, which negatively impact commercial and recreational interests along the coastal region.
@@
 
-### Atmosphere:
+### Atmosphere
 
-Throughout region surrounding the Benguela EBUS, the atmosphere is dominated by anticyclonic high pressure cells with quasi-stationary positions.
+Throughout regions surrounding the Benguela EBUS, the atmosphere is dominated by anticyclonic high-pressure cells with quasi-stationary positions.
 The South Atlantic Ocean High is situated along the west, drawing cool, dry air onto the west of the subcontinent [@van].
-Solar heating during summer may result in the development of low pressure cells, known as heat lows.
-These cells are absent during winter time [@tyson].
-When the anticyclones shift north, during winter months, the cold westerlies have a substantial impact on the weather of the southern tip of the South African subcontinent [@van].
-The atmospheric temperatures along the coastline are largely influenced by the Benguela Current [@van].
-However, upwelling systems may be forced by various atmospheric conditions [@Lutjeharms; @roberts; @hutchings].
+Solar heating during summer may result in the development of low-pressure cells, known as heat lows, which are absent during the winter [@tyson].
+As the anticyclones shift north during winter months, the cold westerlies have a substantial impact on the weather of the southern tip of the South African subcontinent [@van].
+The atmospheric temperatures along the coastline are largely influenced by the Benguela Current [@van], however upwelling systems may be forced by various atmospheric conditions [@Lutjeharms; @roberts; @hutchings].
 
-### Sea surface temperature and data collection:
+### Sea surface temperature and data collection
 
@@
 Regardless of the objectives, whether it be studies based on climate change, extreme events or ecosystem functioning, it is important that the research is conducted through accurate, reliable datasets.
-Unfortunately, not all data are created equally and under the same conditions and as such not all datasets are usable for a particular task.
-
-Remotely-sensed satellite SST products are produced and maintained by a number of organisations worldwide.
-In order to obtain these SST data, advanced technology is launched into space, in order to obtain infra-red light being emitted by the ocean.
-This irradiance is then taken as a proxy for SST.
-Due to the distance at which these SST are being collected, certain coarseness within the datasets are unavoidable.
-Satellite SST products undergo a range of quality control processes that are well researched and tested [@reynolds], making their use more reliable.
-However, Smit et al.
-(2013) have shown that SST data are not particularly reliable for coastal regions and should be used cautiously as they can be prone to both warm and cold bias.
-Nevertheless, satellite data products are extremely useful when investigating the ocean at broad scale and can provide important SST data.
+Unfortunately, not all data are created equally and under the same conditions and as such, not all datasets are appropriate for a particular task.
 
@@
 The South African Coastal Temperature Network (SACTN).
-The availability of the in situ collected coastal temperature data products for several decades and provides a reliable source of accurate coastal seawater temperature data [@smit].
+The availability of the in situ collected coastal temperature data products for several decades provides a reliable source of accurate coastal seawater temperature data [@smit].
 SACTN has collected coastal seawater temperature along the South African coastline from as early as 1972, with contributions from a total of seven different organisations and governmental departments.
-At present, this dataset contained temperature of 129 sites along the three distinct coasts of the South African coastline.
-Temperature recordings are still ongoing at many of these sites, but unfortunately, the durations and extents of these are uneve per site.
+At present, this dataset contains temperatures of 129 sites along the three distinct coasts of the South African coastline.
+Temperature recordings are still ongoing at many of these sites, but unfortunately, the durations and extents of these are uneven per site.
 Data collected for this region started on 13 September 1972 and concluded on 26 January 2017, with recordings still continuing daily.
@@
 The uneven and inconsistent nature of the SATCN dataset can be attributed to several factors.
-For instance, these may be as a result of variances in the equipment used by the various organisations and governmental bodies who recorded the data.
-One problematic outcome of this sampling methodology is that due to specific work of the South African Weather Service (SAWS) and Kwazulu Natal Sharks Board (KZNSB), two thirds of the SACTN data that were sampled were obtained with hand-held thermometers and records at a data precision of 0.500 ◦ C.
-However, the current global standard is to make use of underwater temperature recorders (UTRs), which have a data precision of 0.001◦ C [@jarraud].
-Because of these inconsistencies, this dataset required verification from an independent source before its use within research.
+For instance, as a result of variances in the equipment used by the various organisations and governmental bodies who recorded the data.
+One problematic outcome of this sampling methodology was that, due to specific work of the South African Weather Service (SAWS) and Kwa-Zulu Natal Sharks Board (KZNSB), two thirds of the SACTN data that were sampled were obtained with hand-held thermometers and recorded at a data precision of 0.500 ◦ C, however, the current global standard is to make use of underwater temperature recorders (UTRs), which have a data precision of 0.001◦ C [@jarraud].
+Due of these inconsistencies, this dataset required verification from an independent source before its use in research.
 Smit et al.
 (2013) thoroughly compared the temperature data within the SACTN dataset to those of several remotely-sensed SST products and found bias results.
-Using these in situ collected temperature data along with remotely sensed SST data requires a clear understanding in the characteristics of the contributing data sources in order to understand how accurate, reliable and useful these long term measurements are for use in climate change studies.
+Using these in situ collected temperature data along with remotely sensed SST data requires a clear understanding of the characteristics of the contributing data sources in order to understand how accurate, reliable and useful these long-term measurements are for use in climate change studies.
 
-It must also be acknowledged that most of the time series within the SACTN dataset do not have meta-data records that meet the international climate change research requirements [@aguilar].
-Specifically uncertainty lies in the drift of the instrument, who was sampling and where the exact sampling locations were collected throughout the collection period.
-We are however certain that all of the thermometer time series were collected by thermometers and that the UTR time series were collected by UTRs.
+It should also be acknowledged that most of the time series within the SACTN dataset do not have meta-data records that meet the international climate change research requirements [@aguilar].
+Specifically, uncertainty lies within the drift of the instrument, who was sampling and where the exact sampling locations were collected throughout the collection period.
+There is certainty that all of the thermometer time series were collected by thermometers and that the UTR time series were collected by UTRs in this study.
 Given these issues associated with the mete-data, recent work has been started to develop a nation wide standard.
@@
 Reynolds and Smith, 1994; Brown et al., 1999; Martin et al., 2012), making the data reliable and more useful.
-As previously mentioned, satellites may show a warm or cold bias along the coastal region but they are still useful for broader scale investigations (Smit et al., 2013, Smale and Wernberg, 2009; Castillo and Lima, 2010; Dufois and Rouault).
-Remotely-sensed SST products are obtained at different resolutions, and as a result do not show the same level of bias against in situ collected seawater temperature.
+As previously mentioned, satellites may show a warm or cold bias along the coastal region but are still useful for broader scale investigations (Smit et al., 2013, Smale and Wernberg, 2009; Castillo and Lima, 2010; Dufois and Rouault).
+Remotely-sensed SST products are obtained at different resolutions, and as a result do not represent the same level of bias against in situ collected seawater temperature.
 Research done by Smit et al (2013) showed that MODIS represented the least amount of bias.
 
-This study made use of various satellite-derived Sea Surface Temperatures (SST) datasets.
-One of these datasets are the National Oceanic and Atmospheric Administration (NOAA) ¼° daily Optimum Interpolation Sea Surface Temperature (OISST).
-This dataset is an analysis of seawater temperature by different platforms (satellites, ships and buoys) on a global grid.
+This study made use of various satellite-derived Sea Surface Temperature (SST) datasets.
+One of datasets used was the National Oceanic and Atmospheric Administration (NOAA) ¼° daily Optimum Interpolation Sea Surface Temperature (OISST).
+This dataset is an analysis of seawater temperature collected by different platforms (satellites, ships and buoys) on a global grid.
 The gridded daily SST product assimilates both remotely sensed and *in situ* sources of data to create a level-4 gap free product (Banzon et al., 2016).
-A spatially complete sea suface temperature map is produced by interpolation to fill in gaps within the dataset.
-The method used to fill these gaps includes the adjustment of satellite and ship observations which compensate for the bias differences (NOAA.GOV, https://www.ncdc.noaa.gov/oisst).
-There are two types of daily OISST, named after the relevant satellite sensors.
-These are the Advanced Very High Resolution Radiometer (AVHRR) and Advanced Microwave Scanning Radiometer on the Earth Observing System (AMSR-E).
-Currently the AVHRR has the longest record starting from 1981 to the present.
+A spatially complete sea surface temperature map is produced by interpolation to fill in gaps within the dataset.
+The method used to fill these gaps include the adjustment of satellite and ship observations which compensate for bias differences (NOAA.GOV, https://www.ncdc.noaa.gov/oisst).
+There are two types of daily OISST, named after the relevant satellite sensors, these are the Advanced Very High Resolution Radiometer (AVHRR) and Advanced Microwave Scanning Radiometer on the Earth Observing System (AMSR-E).
+Currently the AVHRR has the longest record starting from 1981 to present day.
 Infrared instruments, like AVHRR, can make observations at relatively high resolutions.
-Microwave instruments, like AMSR-E, can measure SSTs in most weather conditions but not adjacent to land.
-The AVHRR-only Optimally-Interpolated Sea Surface.
-Temperature (OISST) was used to determine SST within the study region.
-The AVHRR datasets have been provided global SSTs for more than four decades [@reynolds; @Pearce; @corlett].
+Microwave instruments, such as AMSR-E, can measure SSTs in most weather conditions but not adjacent to land.
+The AVHRR-only Optimally-Interpolated Sea Surface Temperature (OISST) was used to determine SST within the study region.
+The AVHRR datasets have provided global SSTs for more than four decades [@reynolds; @Pearce; @corlett].
 
-The Multi-scale Ultra-high Resolution (MUR) Sea Surface Temperature Analysis, the second dataset, is produced using satellite instruments with datasets spanning 1 June 2002 to present times (PO.DAAC, https://podaac.jpl.nasa.gov/).
+The Multi-scale Ultra-high Resolution (MUR) Sea Surface Temperature Analysis, used as the second dataset, was produced using satellite instruments with datasets spanning from1 June 2002 to present times (PO.DAAC, https://podaac.jpl.nasa.gov/).
 MUR provides SST data at a spatial resolution of 0.01° in longitude-latitude coordinates, roughly at 1 km intervals.
-This dataset is currently among the highest resolution SST datasets available.
+This dataset is currently among the highest resolution SST dataset available.
 The goal in creating the MUR SST is to develop a coherent and consistent daily map of SST at the highest spatial resolution possible.
-The MUR sea surface temperature dataset combines data from the Advanced Very High Resolution Radiometer (AVHRR), Moderate Imaging Spectroradiometer (MODIS) Terra and Aqua, and Advanced Microwave Spectroradiometer-EOS (AMSR-E) instruments in an optimal way to produce 1km global maps representing the sea surface temperature.
+The MUR sea surface temperature dataset combines data from the Advanced Very High Resolution Radiometer (AVHRR), Moderate Imaging Spectroradiometer (MODIS) Terra and Aqua, and Advanced Microwave Spectroradiometer-EOS (AMSR-E) instruments in an optimal way in order to produce 1km global maps representing sea surface temperature.
 
 The Moderate-resolution Imaging Spectroradiometer (MODIS) is a scientific instrument (radiometer) on board the NASA Terra and Aqua satellite platforms with a spatial resolution in 0.041° longitude-latitude coordinates.
-MODIS was launched in 1991 and 2002 to study global dynamics of the earths atmosphere, ice and ocean.
+MODIS was launched in 1991 and 2002 to study global dynamics of the earth’s atmosphere, ice and ocean.
 MODIS captures data in 36 spectral bands ranging in wavelength from 0.4 um to 14.4 um and at varying spatial resolutions (2 bands at 250 m, 5 bands at 500 m and 29 bands at 1 km), providing complete global coverage of the Earth every one to two days [@hosoda; @gentemann, @dong].
@@
 
-The CMC dataset constitutes the fourth dataset and is a version 3.0 Group for High Resolution Sea Surface Temperature (GHRSST) Level 4 dataset with a 10 km resolution constructed by the Canadian Meteorological Center (CMC).
+The CMC dataset comprises the fourth dataset and is a version 3.0 Group for High Resolution Sea Surface Temperature (GHRSST) Level 4 dataset with a 10 km resolution constructed by the Canadian Meteorological Center (CMC).
 The CMC dataset combines infrared satellite SST at numerous points in the time series from the AVHRR, the European Meteorological Operational-A (METOP-A) and Operational-B (METOP-B) platforms, and microwave SST data from the Advanced Microwave Scanning Radiometer 2 in conjunction with in situ observations of SST from ships and buoys from the ICOADS program.
 
-The fifth dataset, the Global 1 km SST (G1SST) analysis, representing a 0.01° resolution, uses satellite data from sensors that include the Advanced Very High Resolution Radiometer (AVHRR), the Advanced Along Track Scanning Radiometer (AATSR), the Spinning Enhanced Visible and Infrared Imager (SEVIRI), the Advanced Microwave Scanning Radiometer-EOS (AMSRE), the Tropical Rainfall Measuring Mission Microwave Imager (TMI), the Moderate Resolution Imaging Spectroradiometer (MODIS), the Geostationary Operational Environmental Satellite (GOES) Imager, the Multi-Functional Transport Satellite 1R (MTSAT-1R) radiometer, and in situ data from drifting and moored buoys [@dash}.
-Thid dataset produces a product that combines satellite data with the model output (PO.DAAC, https://podaac.jpl.nasa.gov/).
+The fifth dataset, the Global 1 km SST (G1SST) analysis, representing a 0.01° resolution, uses satellite data from sensors that include the Advanced Very High Resolution Radiometer (AVHRR), the Advanced Along Track Scanning Radiometer (AATSR), the Spinning Enhanced Visible and Infrared Imager (SEVIRI), the Advanced Microwave Scanning Radiometer-EOS (AMSRE), the Tropical Rainfall Measuring Mission Microwave Imager (TMI), the Moderate Resolution Imaging Spectroradiometer (MODIS), the Geostationary Operational Environmental Satellite (GOES) Imager, the Multi-Functional Transport Satellite 1R (MTSAT-1R) radiometer, and in situ data from drifting and moored buoys [@dash].
+This dataset produces a product that combines satellite data with the model output (PO.DAAC, https://podaac.jpl.nasa.gov/).
 
-### Climate change:
+### Climate change
 
-Climate change as a result of anthropogenic activities is  an established global concern of both scientific and political importance.
+Climate change as a result of anthropogenic activities is a global concern of both scientific and political importance.
 As the human population grows, the rate of global warming increases, resulting in profound effects on marine and other ecosystems.
 Climbing temperatures create a host of additional changes to marine systems, such as rising sea levels, increased ocean stratification, decreased sea-ice extent, and altered patterns of ocean circulation and precipitation [@doney].
-Furthermore, direct effects on ocean temperature changes and chemistry may alter the functioning and behaviour or marine organisms.
-In turn, these changes lead to a change in species integration and trophic pathways.
+Direct effects on ocean temperature changes and chemistry may alter the functioning and behaviour of marine organisms and in turn, these changes lead to a change in species integration and trophic pathways.
 
-Significant research studies accept the climate change and variability have great impacts on marine ecosystems [@stenseth; @harley; @hoegh].
-The Intergovernmental Panel on Climate Change (IPCC) predicts future projections with a mean increase in air temperature of 1.1°C by the 2020’s and 1.8°C by the 2040’s as a result of greenhouse gas concentrations [@solomon].
+Significant research studies accept that climate change and variability have great impacts on marine ecosystems [@stenseth; @harley; @hoegh].
+The Intergovernmental Panel on Climate Change (IPCC) predicts future projections with a mean increase in air temperature of 1.1°C by the 2020’s and 1.8°C by the 2040’s as a result of the increase of greenhouse gas concentrations [@solomon].
 A large amount of this heat will be absorbed by the world’s ocean.
-In current years, as a result of changes in air temperature, the heat content of the ocean has increased largely.
+In current years, as a result of changes in air temperature, the heat content of the ocean has increased significantly.
 Given ocean warming in association with their physical changes such as sea level rise and declines in sea ice, marine life is expected to intensify in future decades.
 Global warming represents distinct warming and cooling periods throughout the 20th and 21st century (IPCC, 2007).
-Global warming however is not evenly distributed across the global ocean, and thus several regions are warming faster or slower than then García-Reyes et al., 2015 the global average [@levitus; @santos].
-In particular, the Atlantic Ocean contributes most to the increase of the heat content [@nerem; @levitus].
+Global warming however is not evenly distributed across the global ocean, and thus several regions are warming faster or slower than García-Reyes et al., 2015 the global average [@levitus; @santos].
+In particular, the Atlantic Ocean contributes most to the increase of heat content [@nerem; @levitus].
 
 As a result of climate change, research studies provide significant evidence relating to the increase in the warming of the ocean surface.
-This warming slowly propagates down through the water column and in turn increases the thermal stratification [@gruber].
+This warming slowly propagates down through the water column and in turn increases thermal stratification [@gruber].
 Strong stratification limits the depth at which water is upwelled and hence controls the amount of nutrients brought up [@chhak; @jacox2011; @jocox2016].
-The interplay between upwelling winds and stratification determines the amount of nutrients present within the system and is therfore strongly linked to biological productivity.
+The interplay between upwelling winds and stratification determines the amount of nutrients present within the system and is therefore strongly linked to biological productivity.
 Globally, as a result of climate change, stratification of the water-column between 0 and 200 m has increased by an estimated 4% between 1971 to 2010 [@stocker].
 Increased ocean stratification as a result of surface warming may reduce the ventilation of deep-water masses.
-As the deep water is being upwelled to the surface, decreases in oxygen and pH levels are predicted along with an increase in nutrient levels [@rykaczewski].
+As the deep water is being upwelled to the surface, decreases in oxygen and pH levels are predicted with an increase in nutrient levels [@rykaczewski].
 Generally, regional ocean climate oscillations like the Benguela Niño for the Benguela Current System impose considerable variability on temperatures and complicate attempts to assess long-term trends [@demarcq].
-It is predicted that these climate oscillations may increase the variance with climate change [@timmermann; @kuzmina; @sydeman] which will further result in unusual trends.
+It is predicted that these climate oscillations may increase the variance with climate change [@timmermann; @kuzmina; @sydeman] which may result in unusual trends.
 
@@
 2002).
-For example, in the past two decades, sea surface temperature has risen in many regions.
-Seasonal rise in temperature can act as an important cue in species survival, reproduction and abundance (Cushing 1982; Cushing 1990).
-The differing responses of species as a result of climate change are important to understand as these changes affect, among many things, the trophic dynamics and fisheries yield of several regional ecosystems (Murawski 1993).
+In the past two decades, sea surface temperature has risen in many regions and this Seasonal rise in temperature can act as an important cue in species survival, reproduction and abundance (Cushing 1982; Cushing 1990).
+The different responses of species as a result of climate change are important to understand as these changes affect, the trophic dynamics and fisheries yield of several regional ecosystems among many things (Murawski 1993).
 Understanding the relationship between global warming and climate change as a result of anthropogenic activity is a vast and challenging process.
@@
 According to this concept, within each of these boundaries exists a level of safety that should not be exceeded.
-For example, the planetary boundary for biodiversity loss, has been surpassed [@steffen].
+According to this concept, the planetary boundary for biodiversity loss, has been surpassed [@steffen].
 Similarly, the safety boundary for warming of the globe by greenhouse gas emission has also been surpassed.
-The idea of planetary boundaries is essential in that it helps our understanding of the threats we cause via the continued destruction of the natural environment.
+The idea of planetary boundaries is essential in that it helps in understanding the threats humans cause via the continued destruction of the natural environment.
 Consequently, studies of anthropogenic influences on the ecosystem are generally divided into land or sea.
 Terrestrial ecosystems are more familiar to the general populace; whereas marine ecosystems are associated with high levels of primary production [@Chavez] and are highly at risk as a result of climate change [@osterblom].
-Climate change has been proven to decrease primary production thus affecting the trophic levels and leading to trophic turbulence and destruction of marine fauna [@lewandowska; @dell, @mcCauley].
+Climate change has been proven to decrease primary production thus affecting trophic levels and leading to trophic turbulence and the destruction of marine fauna [@lewandowska; @dell, @mcCauley].
 
-Climate change similarly affects coastal species such as oysters, corals and, sea grass.
-Lord (2017) also found evidence that suggests that a change in climate as a result of rising sea temperatures may increase the spread of invasive species and increase the chances of an outbreak of diseases [@altizer].
-Research by Mead et al (2013) found that coastal ecosystems are at high risk as a result of climate change, later, Whitfield et al (2016) confirmed that coastal bio-diversity have been observed to be affected.
+Climate change affects coastal species such as oysters, corals and, sea grass.
+Lord (2017) also found evidence that suggested that a change in climate as a result of rising sea temperatures may increase the spread of invasive species and increase the chances of an outbreak of diseases [@altizer].
+Research by Mead et al (2013) found that coastal ecosystems are at high risk as a result of climate change, later, Whitfield et al (2016) confirmed that coastal bio-diversity has been observed to be affected.
 Given the importance of primary productivity, it was determined that the Benguela upwelling system within the South African region should become the focus of this thesis.
-Here the increase in global temperatures are one of the main focus points along with the chlorophyll concentration.
-By making use of these two measurements, I am able to analyses how chlorophyll concentration vary with a change in temperature.
+Here the increase in global temperatures is one of the main focus points along with chlorophyll concentration.
+By making use of these two measurements, an analysis of how chlorophyll concentration varies with temperature changes could be conducted.
 
-## Benguela :
+## Benguela
 
-The area of study for this thesis will focus on the west coast of South Africa, which is an eastern boundry upwelling system.
+The area of study for this thesis will focus on the west coast of South Africa, which is an eastern boundary upwelling system.
 The Benguela EBUS is unique in that it is bounded at both the equatorward and poleward ends by warm water regimes.
@@
 Water is coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S [@Lutjeharms].
-The main upwelling period in the Benguela north of 25°S extends mainly from March to November, peaking during August [@shannon1985], but upwelling persists throughout the year.
+The main upwelling period in the Benguela, north of 25°S extends mainly from March to November, peaking during August [@shannon1985], but upwelling persists throughout the year.
 Upwelling season in the northern Benguela extends between September and March, with very little upwelling taking place during winter (June-August).
-In the central Benguela and also at Cape Frio upwelling is continuous throughout the year [@shannon1985].
-The cold Benguela waters are bound in the south by the warm Agulus retroflexion region and in the north by the southward flowing Angola current.
+In the central Benguela and also at Cape Frio, upwelling is continuous throughout the year [@shannon1985].
+The cold Benguela waters are bound in the south by the warm Agulhas retroflexion region and in the north by the southward flowing Angola current.
 The flow of the Benguela is often considered to be topographically guided [@nelson; @barange; @barange].
@@
 Upwelling in the northern Benguela is in phase with the seasonal insulation cycle, whereas in the southern Benguela it is not.
-The result is that there is a pronounced seasonal cycle in SST in the north while in the south there is no clear SST cycle.
+This results in a pronounced seasonal cycle in SST in the north while in the south there is no clear SST cycle.
 The seasonal warming occurring during the late summer season off northern and central Namibia is a regular occurrence and is associated with warmer, higher salinity water of Angolan origin into the region [@o; @shannon].
@@
 Shannon et al.
-(1986) illustrated the warm (21-23°C) water near Walvis Bay as a large warm water patch offshore west of Luderitz.
-The absence of upwelling between Walvis Bay and north of Luderitz is evident however, the upwelling centre in the Benguela system [@shannon1986], and upwelling was minimal.
+(1986) illustrated the warm (21-23°C) water near Walvis Bay as a large warm water patch offshore, west of Luderitz.
+The absence of upwelling between Walvis Bay and north of Luderitz is evident, however, the upwelling centre in the Benguela system [@shannon1986], and upwelling was minimal.
 Interestingly, in the southern Benguela region, there was no evidence of any unusual patterns in temperature distributions.
-The seasonal temperature pattern within the northern Benguela is somewhat similar to that in the eastern equatorial Atlantic and the regular southward intrusion of warm water into the northern Benguela during the summer season.
+The seasonal temperature pattern within the northern Benguela was somewhat similar to that in the eastern equatorial Atlantic and the regular southward intrusion of warm water into the northern Benguela during the summer season.
 During these warming periods the equatorward wind stress over the northern Benguela region was stronger than normal; yet upwelling was suppressed by the southward movement of warm saline waters from the Angolan region.
@@
 Northward, the Benguela Current Upwelling system is linked by the Angola Benguela Front which converge between the Benguela current and the Benguela coastal current as well as the warm saline waters of the Angola Current [@peterson].
-Research done by Yamagata and Iizuka (1995) discovered slight indication that coastal Kelvin waves, originating from equatorial Kelvin waves, carry the down welling process southward.
+Research done by Yamagata and Iizuka (1995) discovered that coastal Kelvin waves, originating from equatorial Kelvin waves, carry the down welling process southward.
 Southward, the coastal upwelling system is bounded by the Agulhas Bank.
@@
 Winds occurring within this region tend to show increasing patterns away from the coast whereas, winds occurring in the northern region exhibits relatively less seasonal variation [@shannon1985; @pitcher].
-Upwelling occurring along the south coast of the Cape Province within South Africa, specifically along the coastline is generated by local wind.
-The wind increases somewhat away from the coast.
-This facilitated the idea of wind stress curl, as discussed in Bakun and Nelson (1991) and Shannon and Nelson (1996).
+Upwelling occurring along the south coast of the Cape Province within South Africa, specifically along the coastline is generated by local wind, with wind increasing somewhat away from the coast.
+This facilitated the idea of a wind stress curl, as discussed in Bakun and Nelson (1991) and Shannon and Nelson (1996).
 
 Benguela Niños are defined as anomalous warm events occurring between the southward flowing Angola current and the Benguela upwelling system off south western Africa [@shannon1986].
-These warm events profoundly impact on fisheries and the climate of this region.
+These warm events profoundly impact fisheries and the climate of this region.
 They are proven to induce unexpected rainfall events and drastically influence fish abundance and distribution [@rouault; @boyer].
@@
 @gordon; @stott; @perkins].
-These events occur rapidly, and unexpectedly and they profoundly impact on species diversity, distributions, growth and survival where they take place.
+These events occur rapidly, and unexpectedly and have significant impacts on species diversity, distributions, growth and survival where they take place.
 These events are often undetected, but studies have found that extreme events have occurred as recently as approximately four decades ago.
@@
 Little is known about extreme events occurring within the South African region.
-Few South African researchers have specifically focussed on investigating this topic, and published literature are sparse.
-During the 1980’s however, Birkett and Cook analysed the impacts of extreme warming events on the life cycle of white muscles, Donax serra Röding.
-Futher studies on extreme events occurred within the Florida Reef Tract, where anomolously cold events resulted in the mass mortality of coral reefs and fish.
-Acropora cervicornis, was found to be particularly susceptible to low temperatures and experienced extensive mortality [@shinn; @davis].
-While these phenomenons are not as common as warming events, they have a dramtic and long-lasting effect on coral communities.
-However, studies in January 2010 reveal that the Florida Reef Tract experienced a cold-water anomaly, marine cold spells (MCSs) with temperatures below 16 ◦ C recorded for up to 14 days [@lirman].
-Species population distributon are often set by cold distribution boundaries.
+Few South African researchers have specifically focussed on investigating this topic, and published literature is sparse.
+During the 1980’s however, Birkett and Cook analysed the impacts of extreme warming events on the life cycle of white muscles, *Donax serra Röding*.
+Further studies on extreme events occurred within the Florida Reef Tract, where anomolously cold events resulted in the mass mortality of coral reefs and fish.
+*Acropora cervicornis*, was found to be particularly susceptible to low temperatures and experienced a high mortality rate [@shinn; @davis].
+While these phenomena are not as common as warming events, they have a dramatic and long-lasting effect on coral communities.
+Studies in January 2010 revealed that the Florida Reef Tract experienced a cold-water anomaly, marine cold spells (MCSs) with temperatures below 16 ◦ C recorded for up to 14 days [@lirman].
+Species population distribution are often set by cold distribution boundaries.
 More recently, Schlegel et al, (2017, 2018) examined the relationship between the nearshore and offshore occurrences of marine heatwaves and cold-spells as well as the influence of predominant atmospheric and oceanic patterns during Coastal Marine Heatwaves.
@@
 In a warming planet one would assume that the occurrence and intensity of the severity of marine cold spells (MCSs) are likely to decrease.
-This thesis aims to address this assumption.
+This thesis aims to address that assumption.
 
@@
 
-As specified earlier, the extreme cold events occurring along the west coast of South Africa, which a central to this thesis will be detected, and correlated to upwelling events with the algorithm developed by Hobday et al.
+As specified earlier, the extreme cold events occurring along the west coast of South Africa, which is central to this thesis will be detected and correlated to upwelling events with the algorithm developed by Hobday et al.
 (2016).
-The events that recorde in situ data from SACTN would be compared against those detected in satellite derived SST.
-For this analyses the R package ‘RmarineHeatWaves’ and ‘heatwaveR’ were used to calculate the extreme events.
+The events that recorded in situ data from SACTN would be compared against those detected in satellite derived SST.
+For this analyses, the R packages ‘RmarineHeatWaves’ and ‘heatwaveR’ were used to calculate the extreme events.
 In order to obtain the basis for these calculations, the temperature metrics must be aggregated into mean values based on the year in which they occurred.
@@
 
-With global warming continuing at a rapid pace, the amount of greenhouse gases within the atmosphere are going to increase in the future.
-This increase will undoubtedly have an impact on the world’s oceans and the marine life within, both directly and indirectly.
+With global warming continuing at a rapid pace, the amount of greenhouse gases within the atmosphere will subsequently  increase in the future.
+This increase will undoubtedly have an impact on the world’s oceans and the marine life within it,  both directly and indirectly.
 The oceanic process of coastal upwelling is one aspect that is likely to be affected as changes in global temperature will affect temperature and pressure gradients, and in turn affect alongshore winds which are responsible for driving the upwelling process within EBUS regions.
-Because upwelling is important for productivity within coastal systems, a decrease in this process will negatively impact marine fauna and flora, and human activities such as fishing.
-To assess changes in upwelling events, here I aim to use SST, wind, and chlorophyll data to monitor the durations and intensities of upwelling events within the Benguela EBUS and to identify wether this resulted in an extreme anomalous cold event.
+Because upwelling is important for productivity within coastal systems, a decrease in this process will negatively impact marine fauna and flora as well as human activities such as fishing.
+To assess changes in upwelling events, the aim of this investigation is to use SST, wind, and chlorophyll data to monitor the durations and intensities of upwelling events within the Benguela EBUS and to identify whether this resulted in an extreme anomalous cold event.
 Atmospheric-ocean coupling is known to be affected by climate change, likely changing wind patterns and leading to upwelling favourable conditions (García-Reyes et al., 2015).
-Here we test if an increase in upwelling may lead to a detection of more MCSs simply because these cold temperatures caused by aseasonal upwelling will reflect in historic climatologies.
+Here, it is tested whether an increase in upwelling may lead to a detection of more MCSs simply because these cold temperatures caused by aseasonal upwelling will reflect in historic climatology’s.
 
-## Aims and approach:
+## Aims and approach
 
 The warming of the global climate as a result of anthropogenic activity, has negatively influenced both terrestrial and marine ecosystems, and ultimately inhibits humanity and ecological functioning [@mcCauley].
-This climate variability, manifested as increases in frequency and duration of extreme events is regarded as a great threat [@thompsonet].
+This climate variability, manifested as increases in the frequency and duration of extreme events is regarded as a great threat [@thompsonet].
 Very few studies have addressed the occurrence of cold temperature spells or cold events within the Benguela region.
 This investigation examines recent trends and variation in the frequency and duration of cold spells occurring in this region.
-Here I aim to assess whether the intensity and duration of these extreme events may show any changes in frequency and intensity on a seasonal basis.
-In this research project I make use of various satellite derived observations of sea surface temperature (SST) and wind stress to further investigate the Benguela upwelling system along the South African coastline, whilst considering cold events and warm events occurring within this region.
+Its aims to assess whether the intensity and duration of these extreme events may show any changes in frequency and intensity on a seasonal basis.
+In this research project, various satellite derived observations of sea surface temperature (SST) and wind stress will be used to further investigate the Benguela upwelling system along the South African coastline, whilst considering cold events and warm events occurring within this region.
 
@@
 Wind data used in these analyses were predicted to show clear patterns between weak and favourable upwelling winds and its impact on the intensity and duration of upwelling.
-Chlorophyll a data are then used to examine whether or not the intensity and duration of upwelling may lead to an increase or a decrease in the chlorophyll a concentration at the various sites.
+Chlorophyll a data were then used to examine whether or not the intensity and duration of upwelling may lead to an increase or a decrease in the chlorophyll a concentration at the various sites.
 It was hypothesised that an increase in the intensity and duration of upwelling will show an increase in the abundance of chlorophyll a concentration.
@@
 The primary aim of the second study was to determine the occurrence of MCSs along the west coast of Southern Africa between the nearshore (<400m from the shore) and offshore.
-The most appropriate dataset for the measurement of the nearshore temperatures within South Africa is the South African Coastal Temperature Network (SACTN) dataset (Smit et al.
+The most appropriate dataset for the measurement of the nearshore temperatures within South Africa was the South African Coastal Temperature Network (SACTN) dataset (Smit et al.
 2013).
 With the knowledge obtained from study 1, it was now possible to investigate whether a more intense upwelling event may lead to an increase in  the intensity and duration of MCSs.
-It is decided that an investigation seeking to answer this, taken from historical records of the occerence of every thermal  event recorded along the west coast of South Africa for both in situ and remotely sensed seawater temperatures.
 The SACTN dataset and a total of five SST datasets could be used to investigate the occurrence of marine cold-spells (MCSs) with the methodology introduced by Hobday et al.
 2016.
-This research also assists in the understanding of  local oceanographic patterns along the west coast of South Africa and the impact anthropogenically forced climate change may have for the recent changes in patterns and long-term trends.
+This research could also assists in the understanding of  local oceanographic patterns along the west coast of South Africa and the impact anthropogenically forced climate change could have for the recent changes in patterns and long-term trends.
 
@@
 Very little research has been done on the driving factors leading to MCSs and which processes may result in an increase in the intensity and duration of these events.
-For the purpose of this research  the influence of upwelling on MCSs was the primary focus.
+For the purpose of this study  the influence of upwelling on MCSs was the primary focus.
 

### 2019-03-05 Chlorophyll_a .Rmd
updated

--- 2019-02-28 Chlorophyll_a .Rmd
@@
 
+## Script info
 The aim of this script is to extract the chlorophyll a netCDF files and convert it to CSV.
@@
 library(heatwaveR)
+```
 
+This function convert chlorophyll netCDF datasets to a single CSV datset
+
+```{r}
 MODIS_chlor.dir <- "/home/amieroh/Documents/Data/Datasets/Chlorophyll_a"
@@
 llply(ncList, ncFun, csv.dir = MODIS_chlor.csv.dir, .parallel = TRUE)
-```
 
-## Setting up the chlorophylla dataset
-
-Chlorophyll-a data are 8 day composites.
-In order to compare the chlorophyll concentration with upwelling events, take 8 days, take the average.
-This is one value.
-Repeat this so what I have a time series with chlorophyll-a data every 8 days.
-
-Plot this using a line graph?... Upwelling evetns were plot using wavelet analyses
-
-```{r}
 # MODIS_Chloro <- "~/Documents/Masters_2019/MastersProject/Data"
@@
 # save(MODIS_Chloro, file = "Data/MODIS_Chloro.RData")
-load("Data/MODIS_Chloro.RData")
-load("Data/combined_US.RData")
+```
 
-match_index <- knnx.index(data = as.matrix(MODIS_Chloro[,1:2]),
+## Setting up the chlorophylla dataset
+
+```{r}
+
+# Loading the data
+load("Data/MODIS_Chloro.RData") # Extracted chlorophyll data
+load("Data/combined_US.RData") # Site list of all the sites along the west coast
+
+chloro_plot <- MODIS_Chloro %>%
+  filter(date == min(date)) %>%
+  ggplot(aes(x = lon, y = lat)) +
+  geom_raster(aes(fill = chloro))
+
+unique_pixel <- MODIS_Chloro %>% 
+  select(lon, lat) %>% 
+  unique()
+
+match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
                           query = as.matrix(combined_US[,5:6]), k = 1)
 
-pixel_match <- MODIS_Chloro[match_index,] %>%
+# Matching the sites within the chlorophyll data to the sites along the west coast
+pixel_match <- unique_pixel[match_index,] %>%
   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
@@
 
-daily_chloro_match <- MODIS_Chloro %>%
+chloro_match <- MODIS_Chloro %>%
   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
@@
 
-# With the chloro pixels filter, it is now faster to create the daily data
-daily_chloro_data <- daily_chloro_match %>%
-  dplyr::mutate(date = as.Date(date)) %>% 
-  dplyr::group_by(combi, lon, lat, date) %>% 
-  dplyr::summarise(chloro = mean(chloro, na.rm = TRUE)) %>% 
-  dplyr::group_by(combi, lon, lat) %>% 
-  tidyr::nest() %>% 
-  dplyr::mutate(clims = purrr::map(data, ts2clm, x = date, y = chloro,
-                       climatologyPeriod = c("2002-07-13", "2018-12-04"), # This doesn't mean anything, but the function needs it to work
-                       maxPadLength = 8)) %>% # Adjust maxPadLength = X for longer interpolation
-  dplyr::select(-data) %>% 
-  tidyr::unnest() %>% 
-  select(-doy, -seas, -thresh)
+chloro_match %>%
+  filter(date == max(date)) %>%
+  ggplot(aes(x = lon, y = lat)) +
+  geom_raster(aes(fill = chloro))
 
-ggplot(daily_chloro_data, aes(x = date, y = chloro)) +
+combined_chloro_sites <- chloro_match %>%
+  left_join(pixel_match, by = c("combi", "lon", "lat")) %>% 
+  dplyr::mutate(date = as.Date(date))
+
+# Save the complete dataset
+# save(combined_chloro_sites, file = "Data/combined_chloro_sites.RData")
+
+Chloro_plot_complete <- ggplot(combined_chloro_sites, aes(x = date, y = chloro)) +
   geom_line() +
-  facet_wrap(~combi, ncol = 1)
+  facet_wrap(~site, ncol = 1)
+Chloro_plot_complete
+
+# Plotting only years 2014-2015
+daily_chloro_data_tester <- combined_chloro_sites %>% 
+  filter(year(date) %in% seq(2014, 2015))
+
+Chloro_ggplot(daily_chloro_data_tester, aes(x = date, y = chloro)) +
+  geom_line() +
+  facet_wrap(~site, ncol = 1)
+Chloro_plot_filtered
+
+# # With the chloro pixels filter, it is now faster to create the daily data
+# daily_chloro_data <- daily_chloro_match %>%
+#   dplyr::mutate(date = as.Date(date)) %>%
+#   dplyr::group_by(combi, lon, lat, date) %>%
+#   dplyr::summarise(chloro = mean(chloro, na.rm = TRUE)) %>%
+#   dplyr::group_by(combi, lon, lat) %>%
+#   tidyr::nest() %>%
+#   dplyr::mutate(clims = purrr::map(data, ts2clm, x = date, y = chloro,
+#                      climatologyPeriod = c("2002-07-13", "2018-12-04"), # This doesn't mean anything, but the function needs it to work
+#                      maxPadLength = 8)) %>% # Adjust maxPadLength = X for longer interpolation
+#   dplyr::select(-data) %>%
+#   tidyr::unnest() %>%
+#   select(-doy, -seas, -thresh)
+
+## Plotting daily data
+# ggplot(daily_chloro_data, aes(x = date, y = chloro)) +
+#   geom_line() +
+#   facet_wrap(~combi, ncol = 1)
 
@@
 ```
+
+Here I match the chlorophyll a data with the temperature data. This is done to explore how chlorophyll a concentration varies with temperature.
+
+```{r}
+load("Data/combined_chloro_sites.RData")
+
+## Load in the rest of the datasets
+## See the Dectect_MCS.Rmd script for more info.
+The datasets loaded here are only for temperatures collected between the years 1992-2016.
+However if a longer time series is required changes may  be made to the Dectect_MCS.Rmd script (line 136)
+
+load("Data/overlap_ts.RData")
+load("Data/insitu_MUR.RData")
+load("Data/insitu_OISST.RData")
+load("Data/insitu_G1SST.RData")
+load("Data/insitu_K10.RData")
+
+filtered_years <- function(df){
+  upwelling<- df %>% 
+  filter(year(date) %in% seq(2014, 2015)) %>% 
+  drop_na()
+}
+
+SACTN_fyears <- filtered_years(df = overlap_ts) 
+MUR_fyears <- filtered_years(df = insitu_MUR) 
+OISST_fyears <- filtered_years(df = insitu_OISST) 
+G1SST_fyears <- filtered_years(df = insitu_G1SST) 
+K10_fyears <- filtered_years(df = insitu_K10)
+
+# save(SACTN_fyears, file = "Data/temp_2014_2015/SACTN_fyears.RData")
+# save(MUR_fyears, file = "Data/temp_2014_2015/MUR_fyears.RData")
+# save(OISST_fyears, file = "Data/temp_2014_2015/OISST_fyears.RData")
+# save(G1SST_fyears, file = "Data/temp_2014_2015/G1SST_fyears.RData")
+# save(K10_fyears, file = "Data/temp_2014_2015/K10_fyears.RData")
+```
+
+The next step is to now match the insitu temperature data with the SST data
+
+```{r}
+temp_chloro <- function(df){
+  combined <- df %>% 
+  left_join(combined_chloro_sites, by = c("site", "date")) %>% 
+  select(-combi, -lat, -lon) %>% # Temp.x is the insitu collected data
+  #dplyr::rename(temp = temp.y) %>%
+  na.trim()
+  return(combined)
+}
+
+temp_chloro_SACTN <- temp_chloro(df = SACTN_fyears)
+temp_chloro_MUR <- temp_chloro(df = MUR_fyears)
+temp_chloro_OISST <- temp_chloro(df = OISST_fyears)
+temp_chloro_G1SST <- temp_chloro(df = G1SST_fyears)
+temp_chloro_G1SST <- temp_chloro(df = K10_fyears)
+```

### 2019-03-05 Wind_analyses.Rmd
updated

@@
 # Script Info
-The aim of this script is to plot wind rose diagrams representing the wind directions for the sites along the west coast of South Africa.
-The data used to plot the wind rose diagrams will correspond with the data used to plot the wavelet analyses.
-Using this we will be able to make deductions as to weather or not the intensity and duration of upwelling varied as a result of wind direction.
+The aim of this script is to plot wind rose diagrams and wavelets representing the wind for the sites which correspond with the upwelling sites along the west coast of South Africa.
+Using this we will be able to make deductions as to whether or not the intensity and duration of upwelling varied as a result of wind patterns.
+*Latest wind data on line 149*
 
-The wind data used in this analyses were obtained from the SAWS as well as the NCEP wind data. *Other wind datasets still to be used*
+The wind data used in this analyses were obtained from the SAWS  *Other wind datasets still to be used*
 
@@
 library(stlplus)
+library(zoo)
 source("functions/wind.rose.R")
@@
 
-Creating a function which allows me to load the wind data:
+Creating a function to load the wind and wave data:
 
 ```{r define_fun8, include=TRUE}
+# Building the  function (file path): data/wave_data/Sea Point/15m_359.txt
+#                          1/   2     /    3    /     4
+
 load_wave <- function(wave_files) {
@@
 }
-```
 
-Loadng the wind data for some of the sites along the wc of South Africa
-
-```{r}
 sea_point <- map_dfr(dir("Data/wave_data/Sea Point", 
@@
 ```{r}
-load("~/Documents/Masters_2019/MastersProject/Data/temp_2014_2015/G1SST_fyears.RData")
-load("~/Documents/Masters_2019/MastersProject/Data/temp_2014_2015/OISST_fyears.RData")
-load("~/Documents/Masters_2019/MastersProject/Data/temp_2014_2015/MUR_fyears.RData")
-load("~/Documents/Masters_2019/MastersProject/Data/temp_2014_2015/SACTN_fyears.RData")
-load("~/Documents/Masters_2019/MastersProject/Data/temp_2014_2015/K10_fyears.RData")
+load("Data/temp_2014_2015/G1SST_fyears.RData")
+load("Data/temp_2014_2015/OISST_fyears.RData")
+load("Data/temp_2014_2015/MUR_fyears.RData")
+load("Data/temp_2014_2015/SACTN_fyears.RData")
+load("Data/temp_2014_2015/K10_fyears.RData")
 ```
 
-Here i match the wave data with the datasets
+#############################################################
+#############################################################
+This following code is unable to plot the wind patterns (wind roses) as temperature collection occured between 2014-2015 and we do not have wind data that corresponds with these years
+
+Here I match the wind data with the SST datasets
 
@@
 ```
+#############################################################
+#############################################################
 
-## Wavelet analyses on the wind data corresponding to the wavelet data showing temperature
+## Writing a funciton to load the new wind data
 
-```{r}
+Data/wind_data.txt(SAWS)/wind1/wind1.txt
 
-# When doing a wavelet analyses on the SACTN dataset the "insitu_temp" column needed to be renamed to "temp" for the rest of the code to run.
-The rest of this code is kept constant for each of the datasets (Remotely-sensed SST and In-situ seawater temperature)
-# SACTN_fyears <- SACTN_fyears %>% 
-#   dplyr::rename(temp = insitu_temp)
+```{r define_fun8, include=TRUE}
+load_wind <- function(wind_files) {
+ site_info <- sapply(strsplit(as.character(wind_files), "/"), "[[", 4)
+ site_info <- sapply(strsplit(site_info, ".txt"), "[[", 1)
 
-temp.d <- function(df){
-temp.d <- df %>% 
-  dplyr::group_by(site, date) %>% 
-  dplyr::summarise(dirw_circ = mean(dirw_circ, na.rm = T)) %>%
-  dplyr::mutate(no = seq(1:n())) %>%
-  dplyr::ungroup() %>% 
-  dplyr::select(site, no, dirw_circ, date)
+  wave <- read_table(wind_files, col_types = c("dddddd"),
+                     col_names = c("station_number", "station_name", "date", "hour", "speed", "dir")) %>% 
+       na.omit()
+  return(wave)
 }
 
-temp.d <- temp.d(df = MUR_wind)
-# temp.d <- temp.d(df = OISST_fyears)
-# temp.d <- temp.d(df = G1SST_fyears) 
-# temp.d <- temp.d(df = K10_fyears)
-
-temp.d <- temp.d %>% 
-  select(site, no, dirw_circ, date)
-
-prewhite_fun <- function(x) {
-  df <- x[, 2:3]
-  out <- prewhiteAR(df, order = 3, method = "mle", aic = TRUE,
-             genplot = FALSE, verbose = FALSE)
-  colnames(out) <- c("no", "temp")
-  return(out)
-  }
-PN_prewhite <- as.tibble(ddply(temp.d, .(site), prewhite_fun))
-
-ggplot(PN_prewhite, aes(x = no, y = dirw)) +
-  geom_hline(aes(yintercept = mean(dirw)), colour = "salmon") +
-  geom_line() +
-  facet_wrap(~ site, nrow = 4) +
-  theme_bw()
-
-wl.fun <- function(x) {
-  analyze.wavelet(x, "temp", loess.span = 0, dt = 1,
-                   dj = 1/50, lowerPeriod = 2, make.pval = TRUE, n.sim = 50, 
-                   method = "white.noise", verbose = FALSE)
-}
-
-PN_wave <- dlply(PN_prewhite, .(site), wl.fun)
-
-for (i in 1:length(PN_wave)) {
-  attributes(PN_wave[[i]]) <- c(attributes(PN_wave[[i]]), ref = names(PN_wave)[i])
-}
-# x <- PN_wave[[3]]
-plot_fun <- function(x, plot_name = attributes(x)$ref) {
-  png(filename = paste0("Figures",plot_name,"_wind.png"),
-      width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
-  wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
-           timelab = "Days", verbose = FALSE, useRaster = TRUE,
-           periodlab = "Period", lwd = 1, graphics.reset = FALSE,
-           main = plot_name)
-  dev.off()
-}
-ldply(PN_wave, plot_fun)
+tester <- map_dfr(dir("Data/wind_data.txt(SAWS)/wind1", 
+                            full.names = TRUE, pattern = "*.txt"), load_wind)
 ```

### 2019-03-05 wavelets.Rmd
updated

@@
 
-Working with remotely-sensed SST and *in-situ* collected coastal seawater temperature, in the analyses bellow I aim to show how the intensity and duration of upwelling events varies per sites along the west coast of South Africa.
+## Script info
+
+Wavelet anlyses showing the temperature variation between sites along the west coast for the years 2014 and 2015.
+Thereafter, wavelet analyses ae created per season for each individual site along the west coast of South Africa.
+Working with remotely-sensed SST and *in-situ* collected coastal seawater temperature, in the analyses bellow I aim to show how the intensity and duration of upwelling events vary between sites along the west coast of South Africa.
 It is important to remember that each of the datasets used within this study obtained SSTs at different resolutions.
-Here i was only required to plot a time series of 2 years.
 
@@
 library(tidyverse)
-library(plyr)
-library(lubridate)
 library(ggpubr)
@@
 library(lubridate)
-library(FNN)
+library(stringr)
+library(circular)
+library(broom)
+library(ggrepel)
+library(purrr)
+library(stlplus)
 library(forecast)
@@
 library(data.table)
+# library(plyr)
 ```
@@
 * See the Dectect_MCS.Rmd script for more info.
-The datasets being loaded here are only for temperatures collected between the years 1992-2016.
+The datasets loaded here are only for temperatures collected between the years 1992-2016.
 However if a longer time series is required changes may  be made to the Dectect_MCS.Rmd script (line 136)
@@
 
-Here I only look at upwelling duration over a period of 1 year. from 2014 - 2015
+Examining the intensity and duration of upwelling for the years 2014 - 2015
 
@@
 This should to be repeated for each of the sites within each of the datasets (SACTN, MUR, OISST, G1SST and K10).
-The code bellow needs to be edited.
-Here I only focused on Port Nolloth within the SACTN dataset.
 
@@
 ```{r}
-PN_14 <- SACTN_seasonal %>% 
-  filter(site == "Port Nolloth") %>% 
-  filter(year == 2014 )
+# PN_14 <- SACTN_seasonal %>% 
+#   filter(site == "Port Nolloth") %>% 
+#   filter(year == 2014 )
 
-PN_15 <- SACTN_seasonal %>% 
+# creating a function to only select the year 2015
+SST_season_func <- function(df){
+seasonal_15 <- df %>% 
   filter(year == 2015 )
-  # filter(site == "Port Nolloth") %>% 
-  # filter(year == 2015 )
+}
 
+seasonal_2015_SACTN <- SST_season_func(df = SACTN_seasonal)
+seasonal_2015_MUR <- SST_season_func(df = MUR_seasonal)
+seasonal_2015_OISST <- SST_season_func(df = OISST_seasonal)
+seasonal_2015_G1SST <- SST_season_func(df = G1SST_seasonal)
+seasonal_2015_K10 <- SST_season_func(df = K10_seasonal)
+
+## Plotting function
 combined_seasonal <- function(df){
@@
 
-combined_seasonal_2014 <- combined_seasonal(df = PN_14)
-combined_seasonal_2015 <- combined_seasonal(df = PN_15)
-combined_PortNolloth <- ggarrange(combined_seasonal_2014, combined_seasonal_2015)
-
-# ggsave(plot = combined_PortNolloth, filename = "Figures/combined_PortNolloth.png")
+combined_seasonal_2015_SACTN <- combined_seasonal(df = seasonal_2015_SACTN)
+combined_seasonal_2015_MUR <- combined_seasonal(df = seasonal_2015_MUR)
+combined_seasonal_2015_OISST <- combined_seasonal(df = seasonal_2015_OISST)
+combined_seasonal_2015_G1SST <- combined_seasonal(df = seasonal_2015_G1SST)
+combined_seasonal_2015K10 <- combined_seasonal(df = seasonal_2015_K10)
 ```
 
-Creating a wavelet analyses for the year 2015. This code creates a plot per season.
+Creating a wavelet analyses for the year 2015. This code creates a plot per season. The outcome and plots of this code may be found in the figures folder->seasonal_wavelets (Figures/seasonal_wavelets)
 
 ```{r}
-seasonal_PN_15 <- PN_15 %>% 
-  dplyr::rename(temp = insitu_temp)
+
+# Here I rename the column "insitu-temp" to "temp", this is done in order to have it run on the same code as each of the other SST datasets
+# seasonal_15 <- seasonal_2015_SACTN %>%
+#   dplyr::rename(temp = insitu_temp)
 
@@
 
-temp.d <- temp.d(df = seasonal_PN_15)
+###### Remove the hash's to run the same code on each of these datasets####
+# temp.d <- temp.d(df = seasonal_15)  for the SACTN  seawater temp dataset
+# temp.d <- temp.d(df = seasonal_2015_MUR) #for the MUR SST dataset
+# temp.d <- temp.d(df = seasonal_2015_OISST) #for the OISST SST dataset
+# temp.d <- temp.d(df = seasonal_2015_G1SST) #for the G1SST SST dataset
+# temp.d <- temp.d(df = seasonal_2015_K10) #for the K10 SST dataset
 
@@
   select(season, site, no, temp)
+
+temp_seasonal <- ggplot(temp.d, aes(x = date, y = temp)) +
+  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
+  geom_line() +
+  facet_wrap(~ season, nrow = 4) +
+  theme_bw()
 
@@
 
-temp_seasonal <- ggplot(temp.d, aes(x = date, y = temp)) +
-  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
-  geom_line() +
-  facet_wrap(~ season, nrow = 4) +
-  theme_bw()
-
 wl.fun <- function(x) {
@@
 
-PN_wave <- dlply(temp.d, .(site, season), wl.fun)
+sites_wave <- dlply(temp.d, .(site, season), wl.fun)
 
-for (i in 1:length(PN_wave)) {
-  attributes(PN_wave[[i]]) <- c(attributes(PN_wave[[i]]), ref = names(PN_wave)[i])
+for (i in 1:length(sites_wave)) {
+  attributes(sites_wave[[i]]) <- c(attributes(sites_wave[[i]]), ref = names(sites_wave)[i])
 }
-# x <- PN_wave[[3]]
+# x <- sites_wave[[3]]
 plot_fun <- function(x, plot_name = attributes(x)$ref) {
-  png(filename = paste0("Figures/new",plot_name,"_wavelets.png"),
+  png(filename = paste0("Figures/seasonal",plot_name,"_wavelets.png"),
       width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
@@
 }
-ldply(PN_wave, plot_fun)
+ldply(sites_wave, plot_fun)
 ```
+
+```{r}
+temp.d <- temp.d %>% 
+  select(season, site, no, temp)
+
+temp_seasonal <- ggplot(temp.d, aes(x = date, y = temp)) +
+  geom_line(aes(group = site)) +
+  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
+  facet_wrap(~ season, nrow = 4) +
+  theme_bw()
+
+```

### 2019-03-06 Wind_analyses.Rmd
Wind

@@
 
-Data/wind_data.txt(SAWS)/wind1/wind1.txt
+```{r define_fun8, include=TRUE}
+# load_wind <- function(wind_files) {
+#  site_info <- sapply(strsplit(as.character(wind_files), "/"), "[[", 4)
+#  site_info <- sapply(strsplit(site_info, ".txt"), "[[", 1)
+#   
+#   wave <- read_table(wind_files, col_types = c("dddddd"),
+#                      col_names = c("station_number", "station_name", "date", "hour", "speed", "dir")) %>% 
+#        na.omit()
+#   return(wave)
+# }
+# 
+# tester <- map_dfr(dir("Data/wind_data.txt(SAWS)/wind1", 
+#                             full.names = TRUE, pattern = "*.txt"), load_wind)
 
-```{r define_fun8, include=TRUE}
-load_wind <- function(wind_files) {
- site_info <- sapply(strsplit(as.character(wind_files), "/"), "[[", 4)
- site_info <- sapply(strsplit(site_info, ".txt"), "[[", 1)
+wind_1 <- read.delim("/home/amieroh/Documents/Masters_2019/MastersProject/Data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings="") 
+colnames(wind_1) <- c("station_number", "station_name", "date", "hour", "sub", "speed", "dir")
 
-  wave <- read_table(wind_files, col_types = c("dddddd"),
-                     col_names = c("station_number", "station_name", "date", "hour", "speed", "dir")) %>% 
-       na.omit()
-  return(wave)
+wind_2 <- read.delim("/home/amieroh/Documents/Masters_2019/MastersProject/Data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings="")
+colnames(wind_2) <- c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir")
+
+wind_3 <- read.delim("/home/amieroh/Documents/Masters_2019/MastersProject/Data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings="")
+colnames(wind_3) <- c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir")
+
+# Slecting the important columns for each of the datasets
+wind_fix <- function(df){
+wind <- df%>% 
+  select(station_name, date, hour, dir, speed)
 }
 
-tester <- map_dfr(dir("Data/wind_data.txt(SAWS)/wind1", 
-                            full.names = TRUE, pattern = "*.txt"), load_wind)
+wind_1 <- wind_fix(df = wind_1)
+wind_2 <- wind_fix(df = wind_2)
+wind_3 <- wind_fix(df = wind_3)
+
+## Renaming the sites within the wind datasets to match the name of the sites at which seawater temperature was collected
+## The wind data was obtained from the SAWS and the wind stations used were the closes stations to which temperature was collected
+
+renaming_sites_1 <- function(df) {
+  sites <- df %>%
+    mutate(temp_sites = ifelse(station_name %in% c("CAPE COLUMBINE"), "St Helena Bay",        
+                           ifelse(station_name %in% c("KOINGNAAS"), "Hondeklipbaai",
+                                ifelse(station_name %in% c("PORT NOLLOTH"), "Port Nolloth",
+                                       ifelse(station_name %in% c("LAMBERTSBAAI NORTIER"), "Lamberts Bay",
+                                              ifelse(station_name %in% c("LANGEBAANWEG AWS"), "Saldanha Bay
+","Error"))))))
+  return(sites)
+}
+
+wind_sitesmatched_1 <-  renaming_sites_1(df = wind_1)
+wind_sitesmatched_1 <- wind_sitesmatched_1[-c(121352, 121353, 379892, 379893, 609324, 609325, 843506, 843507, 1014585), ]
+
+renaming_sites_2 <- function(df) {
+  sites <- df %>%
+    mutate(temp_sites = ifelse(station_name %in% c("ROBBENEILAND"), "Koeberg Basin",        
+                           ifelse(station_name %in% c("GEELBEK"), "Yzerfontein",
+                                ifelse(station_name %in% c("DASSEN ISLAND"), "Dassen Island",
+                                       ifelse(station_name %in% c("ATLANTIS"), "Koeberg Basin","Error")))))
+  return(sites)
+}
+
+wind_sitesmatched_2 <-  renaming_sites_2(df = wind_2)
+wind_sitesmatched_2 <- wind_sitesmatched_2[-c(228372, 228373, 313441, 313442, 364881, 364882, 557392), ]
+
+renaming_sites_3 <- function(df) {
+  sites <- df %>%
+    mutate(temp_sites = ifelse(station_name %in% c("CAPE TOWN SLANGKOP"), "Kommetjie",  
+                           ifelse(station_name %in% c("CAPE TOWN - ROYAL YACHT CLUB"), "Sea Point",
+                           ifelse(station_name %in% c("CAPE TOWN TABLE BAY"), "Sea Point","Error"))))
+  return(sites)
+}
+
+wind_sitesmatched_3 <-  renaming_sites_3(df = wind_3)
+wind_sitesmatched_3 <- wind_sitesmatched_3[-c(119317, 119318, 196551, 196552, 346759), ]
+
+## CAPE TOWN SLANGKOP may be used for Kommetjie and for Houtbay
+wind_3_HoutBay <- wind_3 %>% 
+  filter(station_name == "CAPE TOWN SLANGKOP") %>% 
+  mutate(temp_sites = ifelse(station_name %in% c("CAPE TOWN SLANGKOP"), "Hout Bay","Error"))
+
+wind_data <- rbind(wind_3_HoutBay,wind_sitesmatched_3,wind_sitesmatched_2,wind_sitesmatched_1)
+wind_data <- wind_data %>% 
+  dplyr::rename(site = temp_sites) %>% 
+  select(site, date, hour, dir, speed)
 ```
+
+Things to do*******
+
+- Match temp and wind data
+- Create plots shwoing predominant wind direction during an upwelling evnt

### 2019-03-08 Wind_analyses.Rmd
updated

@@
   dplyr::rename(site = temp_sites) %>% 
+  mutate(date = as.Date(date)) %>% 
+  mutate(hour = as.numeric(hour),
+         dir = as.numeric(dir),
+         speed = as.numeric(speed)) %>% 
   select(site, date, hour, dir, speed)
+
+# save(wind_data, file = "Data/wind_data.RData")
+
+# The wind data
+
+wind_daily <- wind_data %>%
+  mutate(date = as.Date(date)) %>%
+  group_by(site, date) %>%
+  summarise(dir_circ = mean.circular(circular(dir, units = "degrees")),
+            speed_circ = mean.circular(circular(speed, units = "degrees")))
+# save(wind_daily, file = "Data/wind_daily.RData")
+
+load("Data/wind_daily.RData") # Loading the daily wind data for each of the sites
 ```
 
-Things to do*******
+## Matching remotely sensed SST's with the wind data
 
-- Match temp and wind data
+Loading the SST datasets: 
+```{r}
+
+## Load in the sea surface temperatue datasets
+## See the Dectect_MCS.Rmd script for more info.
+The datasets loaded here are only for temperatures collected between the years 1992-2016.
+However if a longer time series is required changes may  be made to the Dectect_MCS.Rmd script (line 136)
+
+load("Data/overlap_ts.RData")
+load("Data/insitu_MUR.RData")
+load("Data/insitu_OISST.RData")
+load("Data/insitu_G1SST.RData")
+load("Data/insitu_K10.RData")
+
+# Select only the years 2014 and 2015
+filtered_years <- function(df){
+  upwelling<- df %>% 
+  filter(year(date) %in% seq(2014, 2015)) %>% 
+  drop_na()
+}
+
+SACTN_fyears <- filtered_years(df = overlap_ts) 
+MUR_fyears <- filtered_years(df = insitu_MUR) 
+OISST_fyears <- filtered_years(df = insitu_OISST) 
+G1SST_fyears <- filtered_years(df = insitu_G1SST) 
+K10_fyears <- filtered_years(df = insitu_K10)
+
+# save(SACTN_fyears, file = "Data/temp_2014_2015/SACTN_fyears.RData")
+# save(MUR_fyears, file = "Data/temp_2014_2015/MUR_fyears.RData")
+# save(OISST_fyears, file = "Data/temp_2014_2015/OISST_fyears.RData")
+# save(G1SST_fyears, file = "Data/temp_2014_2015/G1SST_fyears.RData")
+# save(K10_fyears, file = "Data/temp_2014_2015/K10_fyears.RData")
+
+temp_wind <- function(df){
+  combined <- df %>% 
+  left_join(wind_daily, by = c("site", "date")) %>% 
+  #select(-combi) %>% 
+  na.trim()
+  return(combined)
+}
+
+temp_wind_MUR <- temp_wind(df = MUR_fyears)
+temp_wind_SACTN <- temp_chloro(df = SACTN_fyears)
+temp_wind_OISST <- temp_chloro(df = OISST_fyears)
+temp_wind_G1SST <- temp_chloro(df = G1SST_fyears)
+temp_wind_K10 <- temp_chloro(df = K10_fyears)
+```
+
 - Create plots shwoing predominant wind direction during an upwelling evnt
+- Create wavelet analyses

### 2019-03-09 Wind_analyses.Rmd
updated

@@
 source("functions/wind.rose.R")
+library(stlplus)
+library(forecast)
+library(astrochron)
+library(WaveletComp)
+library(data.table)
+# library(plyr)
 ```
 
+#############################################################
+Ignore: run from line 154
+#############################################################
 ## Wind and wave data obtained from SAWS
@@
 
-#############################################################
-#############################################################
 This following code is unable to plot the wind patterns (wind roses) as temperature collection occured between 2014-2015 and we do not have wind data that corresponds with these years
@@
   summarise(dir_circ = mean.circular(circular(dir, units = "degrees")),
-            speed_circ = mean.circular(circular(speed, units = "degrees")))
+            mean_speed = mean(speed))
+
 # save(wind_daily, file = "Data/wind_daily.RData")
@@
 temp_wind_MUR <- temp_wind(df = MUR_fyears)
-temp_wind_SACTN <- temp_chloro(df = SACTN_fyears)
-temp_wind_OISST <- temp_chloro(df = OISST_fyears)
-temp_wind_G1SST <- temp_chloro(df = G1SST_fyears)
-temp_wind_K10 <- temp_chloro(df = K10_fyears)
+temp_wind_SACTN <- temp_wind(df = SACTN_fyears)
+temp_wind_OISST <- temp_wind(df = OISST_fyears)
+temp_wind_G1SST <- temp_wind(df = G1SST_fyears)
+temp_wind_K10 <- temp_wind(df = K10_fyears)
 ```
 
-- Create plots shwoing predominant wind direction during an upwelling evnt
-- Create wavelet analyses
+## Visualising the wind data
+
+```{r}
+
+source("functions/wind.rose.R")
+
+renaming_func <- function(df){ 
+  renaming <- df %>% 
+  dplyr::rename(spd = mean_speed) %>%
+  dplyr::rename(dir = dir_circ)
+}
+
+temp_wind_MUR <- renaming_func(df = temp_wind_MUR)
+temp_wind_SACTN <- renaming_func(df = temp_wind_SACTN)
+temp_wind_OISST <- renaming_func(df = temp_wind_OISST)
+temp_wind_G1SST <- renaming_func(df = temp_wind_G1SST)
+temp_wind_K10 <- renaming_func(df = temp_wind_K10)
+
+## Plotting function
+
+p.wr2 <- plot.windrose(data = temp_wind_SACTN,
+              spd = "spd",
+              dir = "dir")
+
+p.wr3 <- p.wr2 + facet_wrap(~ site) +
+  theme(strip.text.x = element_text(size = 20))
+p.wr3
+```
+
+# Wavelet analyses
+
+```{r}
+
+library(plyr)
+
+temp_wind_SACTN <-  temp_wind_SACTN %>% 
+  drop_na()
+
+temp.d <- function(df){
+temp.d <- df %>% 
+  dplyr::mutate(no = seq(1:n())) %>%
+  dplyr::ungroup() %>% 
+  dplyr::select(site, no, dir_circ, date)
+}
+
+temp.d <- temp.d(df = temp_wind_SACTN) 
+# temp_wind_MUR <- temp_wind(df = MUR_fyears)
+# temp_wind_SACTN <- temp_wind(df = SACTN_fyears)
+# temp_wind_OISST <- temp_wind(df = OISST_fyears)
+# temp_wind_G1SST <- temp_wind(df = G1SST_fyears)
+# temp_wind_K10 <- temp_wind(df = K10_fyears)
+
+temp.d <- temp.d %>% 
+  select(site, no, dir_circ, date)
+
+prewhite_fun <- function(x) {
+  df <- x[, 2:3]
+  out <- prewhiteAR(df, order = 3, method = "mle", aic = TRUE,
+             genplot = FALSE, verbose = FALSE)
+  colnames(out) <- c("no", "dir_circ")
+  return(out)
+  }
+PN_prewhite <- as.tibble(ddply(temp.d, .(site), prewhite_fun))
+
+ggplot(PN_prewhite, aes(x = no, y = dir_circ)) +
+  geom_hline(aes(yintercept = mean(dir_circ)), colour = "salmon") +
+  geom_line() +
+  facet_wrap(~ site, nrow = 4) +
+  theme_bw()
+
+wl.fun <- function(x) {
+  analyze.wavelet(x, "dir_circ", loess.span = 0, dt = 1,
+                   dj = 1/50, lowerPeriod = 2, make.pval = TRUE, n.sim = 50, 
+                   method = "white.noise", verbose = FALSE)
+}
+
+PN_wave <- dlply(PN_prewhite, .(site), wl.fun)
+
+for (i in 1:length(PN_wave)) {
+  attributes(PN_wave[[i]]) <- c(attributes(PN_wave[[i]]), ref = names(PN_wave)[i])
+}
+# x <- PN_wave[[3]]
+plot_fun <- function(x, plot_name = attributes(x)$ref) {
+  png(filename = paste0("Figures/new",plot_name,"_wind.png"),
+      width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
+  wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
+           timelab = "Days", verbose = FALSE, useRaster = TRUE,
+           periodlab = "Period", lwd = 1, graphics.reset = FALSE,
+           main = plot_name)
+  dev.off()
+}
+ldply(PN_wave, plot_fun)
+```

### 2019-03-09 Wind_analyses.Rmd
fixed_path

@@
 
-wind_1 <- read.delim("/home/amieroh/Documents/Masters_2019/MastersProject/Data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings="") 
+wind_1 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings="") 
 colnames(wind_1) <- c("station_number", "station_name", "date", "hour", "sub", "speed", "dir")
 
-wind_2 <- read.delim("/home/amieroh/Documents/Masters_2019/MastersProject/Data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings="")
+wind_2 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings="")
 colnames(wind_2) <- c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir")
 
-wind_3 <- read.delim("/home/amieroh/Documents/Masters_2019/MastersProject/Data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings="")
+wind_3 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings="")
 colnames(wind_3) <- c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir")

### 2019-03-15 Wind_analyses.Rmd
progress..

@@
 
-## Load in the sea surface temperatue datasets
+## Load the sea surface temperatue datasets
 ## See the Dectect_MCS.Rmd script for more info.
@@
 
-load("Data/overlap_ts.RData")
+load("Data/overlap_ts.RData") # overlap: Sites within the SACTN dataset overlap, these sites 
 load("Data/insitu_MUR.RData")
@@
 
-temp_wind_MUR <- temp_wind(df = MUR_fyears)
+temp_wind_MUR <- temp_wind(df = MUR_fyears) %>% 
+  drop_na()
 temp_wind_SACTN <- temp_wind(df = SACTN_fyears)
@@
 temp_wind_K10 <- temp_wind(df = K10_fyears)
+
+speed_plot <- temp_wind_MUR %>% 
+  select(lat, lon, mean_speed) %>% 
+  group_by(lat, lon) %>% 
+  summarise(max_speed = max(mean_speed)) %>% 
+  summarise(lat = round(lat, na.rm = T),3)#%>% 
+  # dplyr::rename(Z = "mean_speed") %>% 
+  # dplyr::rename(X = "lat") %>% 
+  # dplyr::rename(Y = "lon")
+
+ maxw <- cbind(speed_plot$lon, speed_plot$lat, speed_plot$max_speed)  
+head(maxw)
+
+rmax <- rasterFromXYZ(maxw)  
+ acol <- colorRampPalette(c("white", "blue", "darkblue"))  
+ plot(rmax, col=acol(1000), main= "Maximum wind speed reported",  
+     xlab="Longitude", ylab="Lattitude")  + lines(getMap(resolution = "high"), lwd=2)
+
+ library(rworldmap)  
+ lines(getMap(resolution = "high"), lwd=2)  
 ```
@@
 source("functions/wind.rose.R")
-
-renaming_func <- function(df){ 
-  renaming <- df %>% 
-  dplyr::rename(spd = mean_speed) %>%
-  dplyr::rename(dir = dir_circ)
-}
-
-temp_wind_MUR <- renaming_func(df = temp_wind_MUR)
-temp_wind_SACTN <- renaming_func(df = temp_wind_SACTN)
-temp_wind_OISST <- renaming_func(df = temp_wind_OISST)
-temp_wind_G1SST <- renaming_func(df = temp_wind_G1SST)
-temp_wind_K10 <- renaming_func(df = temp_wind_K10)
-
-## Plotting function
-
-p.wr2 <- plot.windrose(data = temp_wind_SACTN,
-              spd = "spd",
-              dir = "dir")
-
-p.wr3 <- p.wr2 + facet_wrap(~ site) +
-  theme(strip.text.x = element_text(size = 20))
-p.wr3
+# 
+# renaming_func <- function(df){
+#   renaming <- df %>%
+#   dplyr::rename(spd = mean_speed) %>%
+#   dplyr::rename(dir = dir_circ)
+# }
+# 
+# temp_wind_MUR <- renaming_func(df = temp_wind_MUR)
+# # temp_wind_SACTN <- renaming_func(df = temp_wind_SACTN)
+# # temp_wind_OISST <- renaming_func(df = temp_wind_OISST)
+# # temp_wind_G1SST <- renaming_func(df = temp_wind_G1SST)
+# # temp_wind_K10 <- renaming_func(df = temp_wind_K10)
+# 
+# ## Plotting function
+# 
+# p.wr2 <- plot.windrose(data = temp_wind_MUR,
+#               spd = "spd",
+#               dir = "dir")
+# 
+# p.wr3 <- p.wr2 + facet_wrap(~ site) +
+#   theme(strip.text.x = element_text(size = 20))
+# p.wr3
 ```
@@
 
+For the wavelet analyses
+
 ```{r}
@@
 library(plyr)
-
-temp_wind_SACTN <-  temp_wind_SACTN %>% 
-  drop_na()
 
@@
 
-temp.d <- temp.d(df = temp_wind_SACTN) 
-# temp_wind_MUR <- temp_wind(df = MUR_fyears)
-# temp_wind_SACTN <- temp_wind(df = SACTN_fyears)
-# temp_wind_OISST <- temp_wind(df = OISST_fyears)
-# temp_wind_G1SST <- temp_wind(df = G1SST_fyears)
-# temp_wind_K10 <- temp_wind(df = K10_fyears)
+#temp.d <- temp.d(df = temp_wind_SACTN)
+# temp_wind_MUR <- temp_wind_MUR %>% 
+#   drop_na()
+# temp.d <- temp.d(df = temp_wind_MUR)
+# temp_wind_OISST <- temp_wind_OISST %>% 
+#   drop_na()
+# temp.d <- temp.d(df = temp_wind_OISST)
+# temp_wind_G1SST <- temp_wind_G1SST %>% 
+#   drop_na()
+# temp.d <- temp.d(df = temp_wind_G1SST)
+# temp.d <- temp.d(df = temp_wind_K10)
 
@@
 ```
+
+## Wind speed
+
+```{r}
+temp_wind_MUR <- temp_wind_MUR %>% 
+    mutate(dir_circ = as.numeric(dir_circ))
+
+SACTN_wind_dir <- ggplot(temp_wind_SACTN, aes(x = date)) +
+  geom_path(aes(y = mean_speed)) +
+  geom_path(aes(y = insitu_temp), colour = "orange") +
+  facet_wrap(~ site, ncol = 1)  +
+  # scale_y_continuous(breaks = c(10, 20, 30 ,40, 50), labels = c("10", "20", "30", "40", "50")) +
+  theme_bw()
+```

### 2019-03-15 wavelets.Rmd
progress..

@@
 The datasets loaded here are only for temperatures collected between the years 1992-2016.
-However if a longer time series is required changes may  be made to the Dectect_MCS.Rmd script (line 136)
+However if a longer time series is required changes may  be made to the Dectect_MCS.Rmd script (line 136) .
+MCS using a long term time series to detect any extreme events occuring within the region
 
@@
 
-SACTN_seasonal <- seasonal_col(df = SACTN_fyears)
-MUR_seasonal <- seasonal_col(df = MUR_fyears)
-OISST_seasonal <- seasonal_col(df = OISST_fyears)
-G1SST_seasonal <- seasonal_col(df = G1SST_fyears)
-K10_seasonal <- seasonal_col(df = K10_fyears)
+# SACTN_seasonal <- seasonal_col(df = SACTN_fyears)
+# MUR_seasonal <- seasonal_col(df = MUR_fyears)
+# OISST_seasonal <- seasonal_col(df = OISST_fyears)
+# G1SST_seasonal <- seasonal_col(df = G1SST_fyears)
+# K10_seasonal <- seasonal_col(df = K10_fyears)
 
@@
 }
+seasonal <-  function(df)
 

### 2019-03-20 Write_up/Writing/dissertation.Rmd
Geostrophic_currents

@@
 Reliable datasets are often difficult to maintain and obtain.
-In the field of biological science research, it is often abiotic data products that are impaired thus leading to the use of satellite derived Sst products for the study of specific fields such as coastal and marine ecology as well as oceanography.
+In the field of biological science research, it is often abiotic data products that are impaired thus leading to the use of satellite derived SST products for the study of specific fields such as coastal and marine ecology as well as oceanography.
 These datasets however, may be inaccurate along the coast and local in situ collected data should be used when possible (Smit et al., 2013).
@@
 The Benguela system is bounded by the equatorward and poleward ends by warm water regimes.
-These currents are shown to be forced by local wind stress fields off southwest Africa[@nelson ; @fennel].
+These currents are shown to be forced by local wind stress fields off southwest Africa[@nelson; @fennel].
 This EBUS streches along the south-western coast of Africa from southern Angola to Cape Agulhas [@cole].

### 2019-03-20 geostrophic_currents.Rmd
Geostrophic currents

---
title: "Geostrophic currents"
author: "Amieroh Abrahams"
date: "20 March 2019"
output: html_document
---

Geostrophic current:
- Oceanic current in which the pressure gradient force is balanced by the coriolis effect
- Direction of geostrophic flows is parallel to the isobars
- Use AVISO data
- u and v values

## Extracting the AVISO data

```{r}
library(ncdf4) 
library(data.table) 
library(tidyverse) 
library(reshape2) 
library(plyr) 
library(lubridate)
library(stringr)

nc.dir <- "/home/amieroh/Documents/spatial/test/netCDF"
csv.dir <- "/home/amieroh/Documents/spatial/test/csv"

#          1         2         3         4         5         6        
# 1234567890123456789012345678901234567890123456789012345678901234567890
# dataset-duacs-rep-global-merged-allsat-phy-l4-v3_19930101-19951231.nc

ncread <- function(nc.dir = nc.dir, csv.dir = dir){
  nc.files <- list.files(path = nc.dir, pattern = "*.nc", full.names = T, include.dirs = T)
  strt.date <- str_sub(basename(nc.files[1]), start = 51, end = 66)
  end.date <- str_sub(basename(nc.files[length(nc.files)]), start = 51, end = 66)
  nc.file <- nc.files[1]
  ncFun <- function(nc.file = nc.files, csv.dir = csv.dir) {
    nc <- nc_open(nc.file)
    name.stem <- substr(basename(nc.file), 1, 48) 
    date.stamp <- substr(basename(nc.file), 51, 66)
    ugos <- ncvar_get(nc, varid = "ugos") %>%
      round(4)
    dimnames(ugos) <- list(lon = nc$dim$lon$vals,
                           lat = nc$dim$lat$vals,
                           time = nc$dim$time$vals)
    sla <-
      as.tibble(melt(ugos, value.name = "ugos"), row.names = NULL) %>%
      dplyr::mutate(vgos = as.vector(ncvar_get(nc, varid = "vgos")),
                    adt = as.vector(ncvar_get(nc, varid = "adt")),
                    sla = as.vector(ncvar_get(nc, varid = "sla"))) %>%
      dplyr::mutate(time = as.Date(time, origin = "1950-01-01 00:00:00")) %>%
      na.omit()
    nc_close(nc)
    fwrite(sla,
           file = paste0(csv.dir, name.stem, "-", strt.date, "-", end.date, ".csv"),
           append = TRUE, col.names = FALSE)
    rm(sla)
  }
  llply(nc.files, ncFun, csv.dir = csv.dir, .parallel = FALSE)
}

ncread(nc.dir, csv.dir)
```

## Loading the AVISO data

```{r}

AVISODir <- "~/Documents/Masters_2019/Data_SST"
AVISO <- fread(paste0(AVISODir, "/csvdataset-duacs-rep-global-merged-allsat-phy-l4-v3-9930101-19951231-0160101-20170515.csv"),
            col.names = c("lon", "lat", "time", "u", "v", "sla", "adt"))

AVISO <- AVISO %>%
  dplyr::group_by(lon, lat) %>%
  dplyr::summarise(u = mean(u, na.rm = TRUE),
                   v = mean(v, na.rm = TRUE),
                   mean.adt = mean(adt, na.rm = TRUE),
                   mean.sla = mean(sla, na.rm = TRUE)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(velocity = sqrt((v^2) + (u^2)),
                eke = 0.5 * (u^2 + v^2),
                arrow_size = ((abs(u * v) / max(abs(u * v))) + 0.3) / 6)

save(AVISO, file = "Data/AVISO.RData")
```

### 2019-04-01 studysite_plot.Rmd
updated

@@
 library(fossil)
+
+source("functions/earthdist.R") # lies in the folder labeled funtions
+source("functions/scale.bar.func.R")
+source("functions/shoreNormalTransectFunc.R")
 ```

### 2019-04-02 Detect_MCSs.Rmd
updated

@@
 
-This script is very repetitive. Wavelet.Rmd creates a function to do this.
+** This script is very repetitive. Wavelet.Rmd creates a function to do this **
 
@@
 
-combined_US <- rbind(UC_1,UC_2)
+# combined_US <- rbind(UC_1,UC_2)
 #save(combined_US, file = "Data/combined_US.RData")
@@
 
-Here I write a function to match the remotely-sensed SST to the *in-situ* collected SST
+This function matches the remotely-sensed SST to the *in-situ* collected SST
 
@@
 
-#SACTN
+#SACTN.....
 ```{r}
@@
 
-MCS_SACTN_PN <- event_line(res, spread = 100, metric = "intensity_cumulative",
-           start_date = "1974-01-01", end_date = "1977-12-31")
+MCS_SACTN_PN <- event_line(res, spread = 500, metric = "intensity_cumulative",
+           start_date = "1974-01-01", end_date = "2015-12-31")
 
@@
            start_date = "2011-12-31", end_date = "2014-10-30") +
-  ggtitle("MUR_PN") +
-  theme(legend = "Bottom")
+  ggtitle("MUR_PN")
 
@@
 
-MUR_SST_DI <- MUR_SST %>%
-  filter(site == "Dassen Island") %>% 
-  dplyr::rename(temp = nearest1) %>% 
-  select(date, temp) %>% 
-  dplyr::rename(t = date)
-
-ts <- ts2clm(MUR_SST_DI, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
-MCS <- res$climatology
-MCS_MUR_DI <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "2011-12-31", end_date = "2014-10-30") +
-  ggtitle("MUR_DI")
-
-MUR_SST_HB <- MUR_SST %>%
-  filter(site == "Hout Bay") %>% 
-  dplyr::rename(temp = nearest1) %>% 
-  select(date, temp) %>% 
-  dplyr::rename(t = date)
-
-ts <- ts2clm(MUR_SST_HB, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
-MCS <- res$climatology
-MCS_MUR_HB <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "2011-12-31", end_date = "2014-10-30") +
-  ggtitle("MUR_HB")
-
 MUR_SST_SP <- MUR_SST %>%
@@
 
-combined_MUR <- ggarrange(MCS_MUR_PN, MCS_MUR_LB, MCS_MUR_O, MCS_MUR_HB, MCS_MUR_SP, MCS_MUR_Kom, MCS_MUR_DI, MCS_MUR_SB, MCS_MUR_YZ, ncol = 3, nrow = 3)
+combined_MUR <- ggarrange(MCS_MUR_PN, MCS_MUR_LB, MCS_MUR_O, MCS_MUR_SP, MCS_MUR_KoM, MCS_MUR_SB, MCS_MUR_YZ, ncol = 2, nrow = 4)
 ```
@@
 # G1SST
+
 ```{r}
@@
 MCS_g1sst_PN <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "1991-09-01", end_date = "1992-12-30") +
+           start_date = "2011-12-31", end_date = "2014-10-30") +
   ggtitle("G1SST_PN")
@@
 MCS_g1sst_LB <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "1991-09-01", end_date = "1992-12-30") +
+           start_date = "2011-12-31", end_date = "2014-10-30") +
   ggtitle("G1SST_LB")
@@
 MCS_g1sst_SB <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "1991-09-01", end_date = "1992-12-30") +
+           start_date = "2011-12-31", end_date = "2014-10-30") +
   ggtitle("G1SST_SB")
@@
 MCS_g1sst_YZ <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "1991-09-01", end_date = "1992-12-30") +
+           start_date = "2011-12-31", end_date = "2014-10-30") +
   ggtitle("G1SST_YZ")
-
-G1SSTsub_SST_HB <- G1SSTsub_SST %>% 
-  filter(site == "Hout Bay") %>% 
-  dplyr::rename(temp = nearest1) %>% 
-  select(date, temp) %>% 
-  dplyr::rename(t = date)
-
-ts <- ts2clm(G1SSTsub_SST_HB, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
-MCS <- res$climatology
-MCS_g1sst_HB <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "1991-09-01", end_date = "1992-12-30") +
-  ggtitle("G1SST_HB")
 
@@
 MCS_g1sst_SP <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "1991-09-01", end_date = "1992-12-30") +
+           start_date = "2011-12-31", end_date = "2014-10-30") +
   ggtitle("G1SST_SP")
@@
 MCS_g1sst_O <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "1991-09-01", end_date = "1992-12-30") +
+           start_date = "2011-12-31", end_date = "2014-10-30") +
   ggtitle("G1SST_O")
@@
 MCS_g1sst_Kom <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "1991-09-01", end_date = "1992-12-30") +
+           start_date = "2011-12-31", end_date = "2014-10-30") +
   ggtitle("G1SST_Kom")
-
-G1SSTsub_SST_DI <- G1SSTsub_SST %>% 
-  filter(site == "Dassen Island") %>% 
-  dplyr::rename(temp = nearest1) %>% 
-  select(date, temp) %>% 
-  dplyr::rename(t = date)
-
-ts <- ts2clm(G1SSTsub_SST_DI, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
-MCS <- res$climatology
-MCS_g1sst_DI <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "1991-09-01", end_date = "1992-12-30") +
-  ggtitle("G1SST_DI")
 
@@
 MCS_g1sst_SB <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "1991-09-01", end_date = "1992-12-30") +
+           start_date = "2011-12-31", end_date = "2014-10-30") +
   ggtitle("G1SST_SB")
 
-G1SSTsub_SST_HB <- G1SSTsub_SST %>% 
-  filter(site == "Hout Bay") %>% 
-  dplyr::rename(temp = nearest1) %>% 
-  select(date, temp) %>% 
-  dplyr::rename(t = date)
-
-ts <- ts2clm(G1SSTsub_SST_HB, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
-MCS <- res$climatology
-MCS_g1sst_HB <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "1991-09-01", end_date = "1992-12-30") +
-  ggtitle("G1SST_HB")
-
-combined_G1SST <- ggarrange(MCS_g1sst_PN, MCS_g1sst_LB, MCS_g1sst_O, MCS_g1sst_SP, MCS_g1sst_Kom, MCS_g1sst_DI, MCS_g1sst_YZ, MCS_g1sst_SB, MCS_g1sst_HB, ncol = 3, nrow = 3)
+combined_G1SST <- ggarrange(MCS_g1sst_PN, MCS_g1sst_LB, MCS_g1sst_O, MCS_g1sst_SP, MCS_g1sst_Kom, MCS_g1sst_YZ, MCS_g1sst_SB, ncol = 2, nrow = 4)
 ```
@@
 MCS <- res$climatology
-MCS_OISST_PN <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "2016-09-01", end_date = "2018-06-30") +
+MCS_OISST_PN <- event_line(res, spread = 900, metric = "intensity_cumulative",
+           start_date = "2010-12-31", end_date = "2014-12-31") +
   ggtitle("OISST_PN")
@@
 MCS <- res$climatology
-MCS_OISST_LB <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "2016-09-01", end_date = "2018-06-30") +
+MCS_OISST_LB <- event_line(res, spread = 900, metric = "intensity_cumulative",
+           start_date = "2010-12-31", end_date = "2014-12-31") +
   ggtitle("OISST_LB")
@@
 MCS_OISST_SB <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "2016-09-01", end_date = "2018-06-30") +
+           start_date = "2010-12-31", end_date = "2014-12-31") +
   ggtitle("OISST_SB")
@@
 MCS_OISST_YZ <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "2016-09-01", end_date = "2018-06-30") +
+           start_date = "2010-12-31", end_date = "2014-12-31") +
   ggtitle("OISST_YZ")
-
-OISST_DI <- OISST_sites %>% 
-  filter(site == "Dassen Island") %>% 
-  select(date, temp) %>% 
-  dplyr::rename(t = date)
-
-ts <- ts2clm(OISST_DI, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
-MCS <- res$climatology
-MCS_OISST_DI <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "2016-09-01", end_date = "2018-06-30") +
-  ggtitle("OISST_DI")
 
@@
 MCS_OISST_SP <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "2016-09-01", end_date = "2018-06-30") +
+           start_date = "2010-12-31", end_date = "2014-12-31") +
   ggtitle("OISST_SP")
@@
 MCS_OISST_Kom <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "2016-09-01", end_date = "2018-06-30") +
+           start_date = "2010-12-31", end_date = "2014-12-31") +
   ggtitle("OISST_Kom")
@@
 MCS <- res$climatology
-MCS_OISST_O <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "2016-09-01", end_date = "2018-06-30") +
+MCS_OISST_O <- event_line(res, spread = 900, metric = "intensity_cumulative",
+           start_date = "2011-12-31", end_date = "2014-12-31") +
   ggtitle("OISST_O")
 
-OISST_HB <- OISST_sites %>% 
-  filter(site == "Hout Bay") %>% 
-  select(date, temp) %>% 
-  dplyr::rename(t = date)
-
-ts <- ts2clm(OISST_HB, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
-MCS <- res$climatology
-MCS_OISST_HB <- event_line(res, spread = 500, metric = "intensity_cumulative",
-           start_date = "2016-09-01", end_date = "2018-06-30") +
-  ggtitle("OISST_HB")
-
-combined_OISST <- ggarrange(MCS_OISST_PN, MCS_OISST_LB, MCS_OISST_SB, MCS_OISST_YZ, MCS_OISST_O, MCS_OISST_HB, MCS_OISST_Kom,MCS_OISST_SP, MCS_OISST_DI)
+combined_OISST <- ggarrange(MCS_OISST_PN, MCS_OISST_LB, MCS_OISST_SB, MCS_OISST_YZ, MCS_OISST_O, MCS_OISST_Kom,MCS_OISST_SP, ncol = 2, nrow = 4)
 ```
@@
 heatwaveR package may be applied to datasets that have a climatological period of >10years
+
 ```{r}

### 2019-04-03 Wind_analyses.Rmd
Still to do some minor fixing

@@
 wind_data <- wind_data %>% 
-  dplyr::rename(site = temp_sites) %>% 
-  mutate(date = as.Date(date)) %>% 
-  mutate(hour = as.numeric(hour),
-         dir = as.numeric(dir),
-         speed = as.numeric(speed)) %>% 
-  select(site, date, hour, dir, speed)
-
-# save(wind_data, file = "Data/wind_data.RData")
+  drop_na()
+save(wind_data, file = "Data/wind_data.RData")
 
@@
 
-wind_daily <- wind_data %>%
+wind_data <- wind_data %>% 
+  as.numeric(dir)
+
+wind_data <- wind_data %>%
   mutate(date = as.Date(date)) %>%
-  group_by(site, date) %>%
-  summarise(dir_circ = mean.circular(circular(dir, units = "degrees")),
-            mean_speed = mean(speed))
+  group_by(temp_sites, date) %>%
+  summarise(mean_dir = mean(as.numeric(dir),
+            mean_speed = mean(as.numeric(speed))))
 
-# save(wind_daily, file = "Data/wind_daily.RData")
+save(wind_data, file = "Data/wind_daily.RData")
 
-load("Data/wind_daily.RData") # Loading the daily wind data for each of the sites
+# load("Data/wind_daily.RData") # Loading the daily wind data for each of the sites
 ```
@@
   upwelling<- df %>% 
-  filter(year(date) %in% seq(2014, 2015)) %>% 
+  filter(year(date) %in% seq(2012, 2015)) %>% 
   drop_na()
@@
   #select(-combi) %>% 
-  na.trim()
+  na.trim() %>% 
+  group_by(site)
   return(combined)
@@
 temp_wind_K10 <- temp_wind(df = K10_fyears)
+```
 
-speed_plot <- temp_wind_MUR %>% 
-  select(lat, lon, mean_speed) %>% 
-  group_by(lat, lon) %>% 
-  summarise(max_speed = max(mean_speed)) %>% 
-  summarise(lat = round(lat, na.rm = T),3)#%>% 
-  # dplyr::rename(Z = "mean_speed") %>% 
-  # dplyr::rename(X = "lat") %>% 
-  # dplyr::rename(Y = "lon")
+```{r}
+histogram_1 <- ggplot(data = temp_wind_MUR, aes(x = dir_circ)) +
+geom_histogram() +
+facet_wrap(~site, scales = "free_x", ncol = 2)
+histogram_1
+```
 
- maxw <- cbind(speed_plot$lon, speed_plot$lat, speed_plot$max_speed)  
-head(maxw)
+```{r}
 
-rmax <- rasterFromXYZ(maxw)  
- acol <- colorRampPalette(c("white", "blue", "darkblue"))  
- plot(rmax, col=acol(1000), main= "Maximum wind speed reported",  
-     xlab="Longitude", ylab="Lattitude")  + lines(getMap(resolution = "high"), lwd=2)
+wind_daily <- wind_daily %>% 
+  drop_na()
+wind_daily_renamed <- wind_daily %>% 
+  dplyr::rename(spd = mean_speed) %>%
+  dplyr::rename(dir = dir_circ)
 
- library(rworldmap)  
- lines(getMap(resolution = "high"), lwd=2)  
+p.wr2 <- plot.windrose(data = wind_daily_renamed,
+              spd = "spd",
+              dir = "dir")
 ```

### 2019-04-03 geostrophic_currents.Rmd
updated: Require BC data

@@
 library(stringr)
+library(doMC); doMC::registerDoMC(cores = 1)
 
@@
 ncread(nc.dir, csv.dir)
+
+# coords <- c(-35, -20, 10, 20) # this is the BC
 ```
@@
 AVISO <- AVISO %>%
-  dplyr::group_by(lon, lat) %>%
+  dplyr::group_by(lon, lat, date) %>%
   dplyr::summarise(u = mean(u, na.rm = TRUE),
@@
   dplyr::mutate(velocity = sqrt((v^2) + (u^2)),
-                eke = 0.5 * (u^2 + v^2),
-                arrow_size = ((abs(u * v) / max(abs(u * v))) + 0.3) / 6)
+                eke = 0.5 * (u^2 + v^2))
 
-save(AVISO, file = "Data/AVISO.RData")
+AVISO_test1 <- AVISO%>%
+  dplyr::mutate(lon = round(lon, 0),
+                lat = round(lat, 0)) %>%
+  dplyr::group_by(lon, lat) %>%
+  dplyr::summarise(u = mean(u, na.rm = TRUE),
+                   v = mean(v, na.rm = TRUE),
+                   velocity = mean(velocity, na.rm = TRUE)) %>%
+  dplyr::ungroup()
+
+plot1 <- ggplot(AVISO_test1, aes(x = lon, y = lat)) +
+    geom_raster(aes(fill = velocity)) +
+    scale_fill_gradientn(colours = rev(rainbow(7, end = 4/6)),
+                         space = "Lab", limits = c(0, 1.45),
+                         guide = guide_colorbar(title = expression(Velocity~(m~s^{-1})),
+                                                position = "bottom",
+                                                direction = "horizontal",
+                                                barheight = unit(2, units = "mm"),
+                                                barwidth = unit(50, units = "mm"),
+                                                draw.ulim = F,
+                                                title.position = 'top',
+                                                title.hjust = 0.5,
+                                                label.hjust = 0.5)) +
+  geom_segment(data = vec,
+               aes(xend = lon + u * current_uv_scalar,
+                   yend = lat + v * current_uv_scalar, alpha = velocity),
+               arrow = arrow(angle = 20, length = unit(0.1, "cm"), type = "open"), size = 0.35, show.legend = FALSE) +
+  scale_alpha_continuous(range = c(0, 1.0))
+
 ```

### 2019-04-08 Paper1_results.Rmd
house_keeping

---
title: "Results_paper1"
author: "Amieroh Abrahams"
date: "07 April 2019"
output: html_document
---

This scripts comprise of the analyses needed for the first section of my masters project.

# Libraries

First I need to find, install and load various packages. These packages will be available on CRAN and can be accessed and installed in the usual way.

```{r prelim_opts, echo=FALSE}
knitr::opts_chunk$set(
  comment = "R>",
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(plyr)
library(lubridate)
library(ggpubr)
library(zoo)
library(lubridate)
library(FNN)
library(forecast)
library(astrochron)
library(WaveletComp)
library(data.table)
library(heatwaveR)
```

## Load site list SACTN data

Now to get to the data. The first step involves the loading of the site list. The statistical properties of the seawater temperature representing the South African coastline, such as the mean, minimum and maximum temperatures. These values vary among coastal sections due to the influence of the cold Benguala and warm Agulhas currents. Here we will only focus on the temperature data found along the west coast (wc) (i.e. sites influenced by the Benguela current, EBUS). The SACTN dataset comprise of 129 *in situ* coastal seawater temperatures derived from daily measurements over up to 40 years. The SACTN temperature dataset was compiled by measuring coastal temperatures at 129 sites along the coast of South Africa, daily from 1972 until 2017.

```{r load_files1, include=TRUE}
load("Data/site_list_v4.2.RData")
load("Data/SACTN_daily_v4.2.RData")
```

## Selecting the sites and creating new *in situ* datasets

Now we select only the sites occuring along the west coast. The purpose of this research task is to assess whether or not the intensity of upwelling varies at va range of distances from the coastline. The data used here is obtained from different satellites and thus possess different resolutions. Here I attempt to find an overlapping time series for all of the sites along the west coast. The sites here vary between upwelling centres and further away from the upwelling centre

```{r}
site_list_sub <- site_list %>%
  filter(coast == "wc") %>%
  filter(length > 3650) 
site_list_sub <- site_list_sub[c(-2, -3, -5, -6, -7,-9, -12, -13, -14),] # Here I keep all the sites with temperature obtained up to 2017
save(site_list_sub, file = "Data_P1/site_list_sub")

SACTN_US <- SACTN_daily_v4.2 %>%
  left_join(site_list[,c(4,13)], by = "index") %>%
  filter(index %in% site_list_sub$index) %>%
  separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>%
  dplyr::rename(insitu_temp = temp)

# The next step is to identify the period where all sites within this dataset overlap
SACTN_overlap <- SACTN_US %>% 
  filter(year(date) %in% seq(1992, 2015)) %>% # The length of the time may vary depending on the question
  drop_na() # Removing NA values within the dataset
# save(SACTN_overlap, file = "Data_P1/SACTN_overlap.RData")

# Visualising the data
temp_plot <- function(df){
  plot <- ggplot(data = df, aes(x = date, y = insitu_temp, colour = site)) +
    geom_line(aes(group = site)) +
    labs(x = "", y = "Temperature (°C)") +
    theme(axis.text.x = element_text(angle = 45)) +
    theme(legend.position = "top")
}

SACTN_plot <- temp_plot(df = SACTN_overlap)
```

In this thesis, different remotely-sensed SST datasets are used. Hence, these datasets represent different resolutions at which data was obtained. We will make use of five remotely-sensed SST datasets, these include; the  MUR, AVHRR OISST, K10 and G1SST data. **MODIS SST data need to be downloaded and used too**

## MUR dataset

```{r}
MUR_Lamberts_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Lamberts Bay_SST_timeseries_5nearest.csv")
MUR_Port_Nolloth <- read_csv("Data/MUR_nearest5pixels/MUR_Port Nolloth_SST_timeseries_5nearest.csv")
MUR_Saldanha_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Saldanha Bay_SST_timeseries_5nearest.csv")
MUR_Yzerfontein <- read_csv("Data/MUR_nearest5pixels/MUR_Yzerfontein_SST_timeseries_5nearest.csv")
MUR_Kommetjie <- read_csv("Data/MUR_nearest5pixels/MUR_Kommetjie_SST_timeseries_5nearest.csv")
MUR_Sea_Point <- read_csv("Data/MUR_nearest5pixels/MUR_Sea Point_SST_timeseries_5nearest.csv")

MUR_SST <- rbind(MUR_Lamberts_Bay,MUR_Yzerfontein, MUR_Port_Nolloth,MUR_Saldanha_Bay, MUR_Kommetjie, MUR_Sea_Point) %>%
  dplyr::rename(site = station)

MUR_SST$date <- (ymd(MUR_SST$date))
MUR_SST <- MUR_SST %>% 
  dplyr::rename(temp = nearest1) %>% 
  drop_na()

# save(MUR_SST, file = "Data_P1/MUR_SST.RData")
```

This function matches the remotely-sensed SST to the *in-situ* collected SST

```{r}
match_func <- function(df){
  match <- SACTN_overlap  %>%  
  left_join(df, by = c("site", "date")) %>% 
  na.trim()
  return(match)
}

insitu_MUR <- match_func(df = MUR_SST) %>%
  drop_na()
# save(insitu_MUR, file = "Data_P1/insitu_MUR.RData")

# Basic visualisations
match_plot <- function(df){
  plot1 <- df %>% 
ggplot(aes(x = date, y = temp)) +
  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ site, nrow = 2) +
  theme_bw()
  return(plot1)
}

MUR_plot <- match_plot(df = insitu_MUR)

temp_plot <- function(df){
  temp_plot <- df %>% 
  ggplot(aes(x = date, y = temp, colour = site)) +
    geom_line(aes(group = site)) +
    labs(x = "", y = "Temperature (°C)") +
    theme(axis.text.x = element_text(angle = 45)) +
    theme(legend.position = "top")
}
MUR_plot <- temp_plot(df = insitu_MUR)
```

## G1SST dataset

```{r}
Lamberts_Bay <- read_csv("Data/G1SST_sub/Lamberts Bay_SST_timeseries_5nearest.csv")
Port_Nolloth <- read_csv("Data/G1SST_sub/Port Nolloth_SST_timeseries_5nearest.csv")
Saldanha_Bay <- read_csv("Data/G1SST_sub/Saldanha Bay_SST_timeseries_5nearest.csv")
Yzerfontein <- read_csv("Data/G1SST_sub/Yzerfontein_SST_timeseries_5nearest.csv")
Kommetjie_SST <- read_csv("Data/G1SST_sub/Kommetjie_SST_timeseries_5nearest.csv")
Sea_Point_SST<- read_csv("Data/G1SST_sub/Sea Point_SST_timeseries_5nearest.csv")

G1SSTsub_SST <- rbind(Lamberts_Bay, Port_Nolloth, Saldanha_Bay, Yzerfontein, Kommetjie_SST, Sea_Point_SST) %>%
  dplyr::rename(site = station)

G1SSTsub_SST$date <- (ymd(G1SSTsub_SST$date))
# save(G1SSTsub_SST, file = "Data_P1/G1SSTsub_SST.RData")

G1SSTsub_SST <- G1SSTsub_SST %>% 
  drop_na() %>% 
  dplyr::rename(temp = nearest1)

insitu_G1SST <- match_func(df = G1SSTsub_SST) 
save(insitu_G1SST, file = "Data_P1/insitu_G1SST.RData")
G1SST_plot <- match_plot(df = insitu_G1SST )
G1SST_temp_plot <- temp_plot(df = insitu_G1SST)
```

# K10

```{r}
Lamberts_Bay <- read_csv("Data/K10/K10_Lamberts Bay_SST_timeseries_5nearest.csv")
Port_Nolloth <- read_csv("Data/K10/K10_Port Nolloth_SST_timeseries_5nearest.csv")
Saldanha_Bay <- read_csv("Data/K10/K10_Saldanha Bay_SST_timeseries_5nearest.csv")
Yzerfontein <- read_csv("Data/K10/K10_Yzerfontein_SST_timeseries_5nearest.csv")
Kommetjie <- read_csv("Data/K10/K10_Kommetjie_SST_timeseries_5nearest.csv")
Sea_Point <- read_csv("Data/K10/K10_Sea Point_SST_timeseries_5nearest.csv")

K10_SST <- rbind(Lamberts_Bay, Yzerfontein, Port_Nolloth, Saldanha_Bay, Kommetjie, Sea_Point) %>%
  dplyr::rename(site = station)
K10_SST$date <- (ymd(K10_SST$date))
save(K10_SST, file = "Data_P1/K10_SST.RData")

K10_SST <- K10_SST %>% 
  drop_na()
K10_SST <- K10_SST %>% 
  dplyr::rename(temp = nearest1)

insitu_K10 <- match_func(df = K10_SST) 
# save(insitu_K10, file = "Data_P1/insitu_K10.RData")

K10_plot <- match_plot(df = insitu_K10)
k10_temp_plot <- temp_plot(df = insitu_K10)
```

## OISST dataset

```{r}

OISSTDir <- "~/Documents/Masters_2019/Data_SST"
OISST <- fread(paste0(OISSTDir, "/csvavhrr-only-v2-19810901-20180630.csv"),
            col.names = c("lon", "lat", "temp", "date"))

# Visualise the data
# To explore the data I visualise the min temperatures along the South african coastline.

OISST %>%
  filter(date == min(date)) %>%
  ggplot(aes(x = lon, y = lat)) +
  geom_raster(aes(fill = temp))
```

## Here I find the nearest SST pixels

Now we apply the FNN (Fast Nearest Neighbor) package to determine the nearesr SST pixel to the insitu collected sites.

```{r}

load("Data_P1/SACTN_overlap.RData") # Loading the SACTN_overlap dataset which allows me to match it to the OISST data

unique_pixel <- OISST %>%
  select(lon, lat) %>%
  unique()

# Select nearest 1 pixels (k = 1)
# Here I use knnx to find the closes 1 pixels to the insitu sites
match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
                      query = as.matrix(SACTN_overlap[,5:6]), k = 1)

# Select SST pixels nearest to insitu sites
pixel_match <- unique_pixel[match_index,] %>%
  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
  mutate(site = SACTN_overlap$site)
# save(pixel_match, file = "Data_P1/pixel_match.RData")
load("Data_P1/pixel_match.RData")

# Subsetting the OISST data to match the upwelling sites within the in situ collected temperature data
OISST_match <- OISST %>%
  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
  filter(combi %in% pixel_match$combi)
# save(OISST_match, file = "Data_P1/OISST_match.RData")

load("Data_P1/OISST_match.RData")
OISST_sites <- OISST_match %>%
  left_join(pixel_match, by = c("combi", "lon", "lat")) %>%
  dplyr::rename(temp_OISST =temp)

# Visualise time series
ggplot(data = OISST_sites, aes(x = date, y = temp)) +
  geom_line() +
  facet_wrap(~site)

# save(OISST_sites, file = "Data/OISST_sites.RData")
load("~/Documents/Masters_2019/MastersProject/Data/OISST_sites.RData")
OISST_sites <- OISST_sites %>% 
  dplyr::rename(temp = temp_OISST) %>%
  dplyr::mutate(date = as.Date(date)) %>% 
  drop_na()
```

## Matching the Insitu data with the OISST SST data
```{r}
insitu_OISST <- match_func(df = OISST_sites)
# save(insitu_OISST, file = "Data/insitu_OISST.RData")
OISST_plot <- match_plot(df = insitu_OISST)
OISST_temp_plot <- temp_plot(df = insitu_OISST)
```

### 2019-04-08 Exploring/Tester_code/Methods.Rmd
housekeeping

--- 2019-01-24 Methods.Rmd
@@
 
-save(SACTN_3yr, file = "Data/SACTN_3yr.RData")
+# save(SACTN_3yr, file = "Data/SACTN_3yr.RData")
 ```
@@
 OISSTDir <- "~/Documents/Masters_2019/Data_SST"
-OISST <- fread(paste0(OISSTDir, "/csvavhrr-only-v2-19810901-20180630.csv"),
-            col.names = c("lon", "lat", "temp", "date"))
+# OISST <- fread(paste0(OISSTDir, "/csvavhrr-only-v2-19810901-20180630.csv"),
+#             col.names = c("lon", "lat", "temp", "date"))
 

### 2019-04-08 Exploring/Tester_code/Methods_tidy.Rmd
housekeeping

--- 2019-01-25 Methods_tidy.Rmd
@@
 ### Load site list SACTN data
-
 Now to get to the data.
@@
 
-Now we select only the sites occuring along the west coast.
-Thereafter, I include all the sites which had a time series of 10 years or greater.
-This yield a 2 rows of data for Paternoster.
-I removed the row of data obtained by the DEA, thus ending up with all of the data being obtained from SAWS in this dataset.
-I then exclude the sites that do not fall within the upwelling cell.
-Once those sites were removed, we have a total of 4 sites.
-These are: Port Nolloth, Lamberts Bay, Saldanna and Yzerfontein.
+## Selecting the sites and creating new *in situ* datasets
+Now we select only the sites occuring along the west coast. Thereafter, I include all the sites which had a time series of 10 years or greater.
 
 ```{r}
-site_list_sub <- site_list %>% 
-  filter(coast == "wc")
+# site_list_sub <- site_list %>%
+#   filter(coast == "wc")
+# 
+# SACTN_daily_wc <- site_list_sub %>%
+#   filter(length >= 3650)
+# 
+# upwelling_sites <- SACTN_daily_wc[-6,] # Here we had two rows of data for Paternoster; i removed the DEA row: All data here are now collected by SAWS
+# upwelling_sites <- upwelling_sites[-10:-14,] # Here i remove the sites that does not fall within the upwelling cell
+# 
+# # Here I find a time series with the data occuring in the most recent years
+# UC_1 <- upwelling_sites[c(-2, -3, -5, -6, -8),]
+# save(UC_1, file = "Data/UC_1.RData")
+# 
+# SACTN_daily_temps <- SACTN_daily_v4.2 %>%
+#   left_join(site_list[,c(4,13)], by = "index") %>%
+#   filter(index %in% upwelling_sites$index)
+# # save(SACTN_daily_temps, file = "SACTN_daily_temps.RData")
+# 
+# SACTN_US1 <- SACTN_daily_temps %>%
+#   separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>%
+#   dplyr::rename(insitu_temp = temp)
+# save(SACTN_US1, file = "Data/SACTN_US1.RData")
 
-SACTN_daily_wc <- site_list_sub %>% 
-  filter(length >= 3650)
+# site_list_sub_UC <- site_list %>%
+#   filter(coast == "wc")
+# 
+# SACTN_daily_wc <- site_list_sub_UC %>%
+#   filter(length >= 3650)
+# 
+# upwelling_sites_UC <- SACTN_daily_wc[-1:-8,]
+# UC_2 <- upwelling_sites_UC[c(-2,-5),] # New sites falling within the upwelling cell before yzerfontein
+# save(UC_2, file = "Data/UC_2.RData")
 
-upwelling_sites <- SACTN_daily_wc[-6,] # Here we had two rows of data for Paternoster; i removed the DEA row: All data here are now collected by SAWS 
-upwelling_sites <- upwelling_sites[-10:-14,] # Here i remove the sites that does not fall within the upwelling cell
-
-# Here I find a time series with the data occuring in the most recent years
-upwelling_sites <- upwelling_sites[c(-2, -3, -5, -6, -8),] 
-# save(upwelling_sites, file = "Data/upwelling_sites.RData")
-
-SACTN_daily_temps <- SACTN_daily_v4.2 %>% 
-  left_join(site_list[,c(4,13)], by = "index") %>% 
-  filter(index %in% upwelling_sites$index)
-# save(SACTN_daily_temps, file = "SACTN_daily_temps.RData")
-
-SACTN_daily_temps_indexsplit <- SACTN_daily_temps %>% 
-  separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>% 
-  dplyr::rename(insitu_temp = temp)
-
-# save(SACTN_daily_temps_indexsplit, file = "Data/SACTN_daily_temps_indexsplit.RData")
+# SACTN_daily_temps_UC <- SACTN_daily_v4.2 %>%
+#   left_join(site_list[,c(4,13)], by = "index") %>%
+#   filter(index %in% upwelling_sites_UC$index)
+# # save(SACTN_daily_temps, file = "SACTN_daily_temps.RData")
+# 
+# SACTN_US2 <- SACTN_daily_temps_UC %>%
+#   separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>%
+#   dplyr::rename(insitu_temp = temp)
+# 
+# save(SACTN_US2, file = "Data/SACTN_US2.RData")
 ```
@@
 
-# SACTN_split <- SACTN_monthly(df = SACTN_daily_temps_indexsplit)
-# save(SACTN_split, file = "SACTN_split.RData")
+# SACTN_split_US1 <- SACTN_monthly(df = SACTN_US1)
+# SACTN_split_US2 <- SACTN_monthly(df = SACTN_US2)
+# save(SACTN_split_US1, file = "SACTN_split_US1.RData")
+# save(SACTN_split_US2, file = "SACTN_split_US2.RData")
 ```
 
-Find where different sites have overlapping time series within this `SACTN_split` dataset.
+Find where different sites have overlapping time series within this `SACTN_split` dataset. Finding a long time series
 
 ```{r}
-load("~/Documents/Masters_2019/MastersProject/Data/SACTN_split.RData")
+load("~/Documents/Masters_2019/MastersProject/Data/SACTN_split_US1.RData")
+load("~/Documents/Masters_2019/MastersProject/Data/SACTN_split_US2.RData")
 
-SACTN_3yr <- SACTN_split %>% 
-  filter(year(date) %in% seq(2014, 2015)) %>% 
+# Find a period where each of the sites within SACTN_US1 and SACTN_US2 that overlap. See dataset UC_1.RData and UC_2.RData
+
+UC1 <- SACTN_split_US1 %>% 
+  filter(year(date) %in% seq(1992, 2016)) %>% # see datsets UC_1.RData
   drop_na() # Removing NA values within the dataset
 
-save(SACTN_3yr, file = "Data/SACTN_3yr.RData")
+# Find a period where each of the sites within SACTN_US1 and SACTN_US2 that overlap
+UC2 <- SACTN_split_US2 %>% 
+  filter(year(date) %in% seq(1992, 2016)) %>% # see datsets UC_2.RData
+  drop_na() # Removing NA values within the dataset
 
-insitu_temp_plot <- ggplot(data = SACTN_3yr, aes(x = date, y = insitu_temp, colour = site)) +
+UC1 <- temp_plot(df = UC1)
+UC2 <- temp_plot(df = UC2)
+
+temp_plot <- function(df){
+  plot <- ggplot(data = df, aes(x = date, y = insitu_temp, colour = site)) +
     geom_line(aes(group = site)) +
@@
     theme(axis.text.x = element_text(angle = 45)) +
-#theme_linedraw() +
     theme(legend.position = "top")
+}
 
-# ggsave(plot = insitu_temp_plot, filename = "Figures/insitu_temp.pdf")
+UC1 <- temp_plot(df = UC1)
+UC2 <- temp_plot(df = UC2)
 ```
@@
 ## MUR dataset
-
 ```{r}
@@
 MUR_SST$date <- (ymd(MUR_SST$date))
-# save(MUR_SST, file = "Data/MUR_SST.RData")
+#save(MUR_SST, file = "Data/MUR_SST.RData")
 
@@
 match_func <- function(df){
-  match <- SACTN_3yr  %>%  
+  match <- SACTN_split_US1  %>%  
   left_join(df, by = c("site", "date")) %>% 
@@
 }
-
 insitu_MUR <- match_func(df = MUR_SST) %>%
@@
 }
+MUR_plot <- match_plot(df = insitu_MUR)
 
-MUR_plot <- match_plot(df = insitu_MUR)
-mur_temp_plot <- ggplot(data = insitu_MUR, aes(x = date, y = temp, colour = site)) +
+temp_plot <- function(df){
+  temp_plot <- df %>% 
+  ggplot(aes(x = date, y = temp, colour = site)) +
     geom_line(aes(group = site)) +
@@
     theme(axis.text.x = element_text(angle = 45)) +
-#theme_linedraw() +
     theme(legend.position = "top")
-
-# ggsave(plot = mur_temp_plot, filename = "Figures/mur_temp_plot.pdf")
-
+}
+MUR_plot <- temp_plot(df = insitu_MUR)
 ```
@@
 ## OISST dataset
-
 ```{r}
@@
 ## Matching the Insitu data with the OISST SST data
-
 ```{r}
@@
 OISST_plot <- match_plot(df = insitu_OISST)
-
-OISST_temp_plot <- ggplot(data = insitu_OISST, aes(x = date, y = temp, colour = site)) +
-geom_line(aes(group = site)) +
-labs(x = "", y = "Temperature (°C)") +
-theme(axis.text.x = element_text(angle = 45)) +
-# theme_linedraw() +
-theme(legend.position = "top")
-# ggsave(plot = OISST_temp_plot, filename = "Figures/OISST_temp_plot.pdf")
+OISST_temp_plot <- temp_plot(df = insitu_OISST)
 ```
@@
 ## G1SST dataset
+```{r}
+# Lamberts_Bay <- read_csv("Data/G1SST_sub/Lamberts Bay_SST_timeseries_5nearest.csv")
+# Port_Nolloth <- read_csv("Data/G1SST_sub/Port Nolloth_SST_timeseries_5nearest.csv")
+# Saldanha_Bay <- read_csv("Data/G1SST_sub/Saldanha Bay_SST_timeseries_5nearest.csv")
+# Yzerfontein <- read_csv("Data/G1SST_sub/Yzerfontein_SST_timeseries_5nearest.csv")
+# 
+# G1SSTsub_SST <- rbind(Lamberts_Bay,Port_Nolloth, Saldanha_Bay, Yzerfontein) %>%
+#   dplyr::rename(site = station)
+# 
+# G1SSTsub_SST$date <- (ymd(G1SSTsub_SST$date))
+# # save(G1SSTsub_SST, file = "Data/G1SSTsub_SST.RData")
 
-```{r}
-Lamberts_Bay <- read_csv("Data/G1SST_sub/Lamberts Bay_SST_timeseries_5nearest.csv")
-Port_Nolloth <- read_csv("Data/G1SST_sub/Port Nolloth_SST_timeseries_5nearest.csv")
-Saldanha_Bay <- read_csv("Data/G1SST_sub/Saldanha Bay_SST_timeseries_5nearest.csv")
-Yzerfontein <- read_csv("Data/G1SST_sub/Yzerfontein_SST_timeseries_5nearest.csv")
-
-G1SSTsub_SST <- rbind(Lamberts_Bay,Port_Nolloth, Saldanha_Bay, Yzerfontein) %>%
-  dplyr::rename(site = station)
-
-G1SSTsub_SST$date <- (ymd(G1SSTsub_SST$date))
-
-# save(G1SSTsub_SST, file = "Data/G1SSTsub_SST.RData")
 load("Data/G1SSTsub_SST.RData")
@@
 G1SSTsub_SST <- G1SSTsub_SST %>% 
-  drop_na()
-
-G1SSTsub_SST <- G1SSTsub_SST %>% 
+  drop_na() %>% 
   dplyr::rename(temp = nearest1)
 
-insitu_G1SST <- match_func(df = G1SSTsub_SST) %>%
+insitu_G1SST <- match_func(df = G1SSTsub_SST) 
+G1SST_plot <- match_plot(df = insitu_G1SST )
+G1SST_temp_plot <- temp_plot(df = insitu_G1SST)
+```
+
+# K10
+```{r}
+# Lamberts_Bay <- read_csv("Data/K10/K10_Lamberts Bay_SST_timeseries_5nearest.csv")
+# Port_Nolloth <- read_csv("Data/K10/K10_Port Nolloth_SST_timeseries_5nearest.csv")
+# Saldanha_Bay <- read_csv("Data/K10/K10_Saldanha Bay_SST_timeseries_5nearest.csv")
+# Yzerfontein <- read_csv("Data/K10/K10_Yzerfontein_SST_timeseries_5nearest.csv")
+# 
+# K10_SST <- rbind(Lamberts_Bay, Yzerfontein, Port_Nolloth, Saldanha_Bay) %>%
+#   dplyr::rename(site = station)
+# K10_SST$date <- (ymd(K10_SST$date))
+# save(K10_SST, file = "Data/K10_SST.RData")
+
+load("~/Documents/Masters_2019/MastersProject/Data/K10_SST.RData")
+K10_SST <- K10_SST %>% 
   drop_na()
-G1SST_plot <- match_plot(df = insitu_G1SST )
+K10_SST <- K10_SST %>% 
+  dplyr::rename(temp = nearest1)
 
-G1SST_temp_plot <- ggplot(data = insitu_G1SST, aes(x = date, y = temp, colour = site)) +
-geom_line(aes(group = site)) +
-labs(x = "", y = "Temperature (°C)") +
-theme(axis.text.x = element_text(angle = 45)) +
-# theme_linedraw() +
-theme(legend.position = "top")
-# ggsave(plot = G1SST_temp_plot, filename = "Figures/G1SST_temp_plot.pdf")
+insitu_K10 <- match_func(df = K10_SST) 
+K10_plot <- match_plot(df = insitu_K10)
+k10_temp_plot <- temp_plot(df = insitu_K10)
 ```
+
+# K10
+
+```{r}
+# OISST
+# load("Data/OISST_sites.RData")
+
+temp.d <- insitu_K10_SST %>% 
+  dplyr::group_by(site, date) %>% 
+  dplyr::summarise(temp = mean(temp, na.rm = T)) %>%
+  dplyr::mutate(no = seq(1:n())) %>%
+  dplyr::ungroup() %>% 
+  dplyr::select(site, no, temp, date)
+
+auto.arima(dplyr::filter(insitu_K10_SST, site == "Port Nolloth")[, "temp"], max.p = 3, max.q = 3, stationary = FALSE, 
+           seasonal = FALSE)
+
+PN <- temp.d %>% 
+  dplyr::filter(site == "Port Nolloth") %>% 
+  dplyr::select(temp, -date, -site) %>% 
+  dplyr::mutate(no = seq(1:n())) %>% 
+  dplyr::select(no, temp)
+
+temp.d <- temp.d %>% 
+  select(site, no, temp, date)
+
+prewhite_fun <- function(x) {
+  df <- x[, 2:3]
+  out <- prewhiteAR(df, order = 3, method = "mle", aic = TRUE,
+             genplot = FALSE, verbose = FALSE)
+  colnames(out) <- c("no", "temp")
+  return(out)
+  }
+PN_prewhite <- as.tibble(ddply(temp.d, .(site), prewhite_fun))
+
+ggplot(PN_prewhite, aes(x = no, y = temp)) +
+  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
+  geom_line() +
+  facet_wrap(~ site, nrow = 4) +
+  theme_bw()
+
+wl.fun <- function(x) {
+  analyze.wavelet(x, "temp", loess.span = 0, dt = 1,
+                   dj = 1/50, lowerPeriod = 2, make.pval = TRUE, n.sim = 50, 
+                   method = "white.noise", verbose = FALSE)
+}
+
+PN_wave <- dlply(PN_prewhite, .(site), wl.fun)
+
+for (i in 1:length(PN_wave)) {
+  attributes(PN_wave[[i]]) <- c(attributes(PN_wave[[i]]), ref = names(PN_wave)[i])
+}
+# x <- PN_wave[[3]]
+plot_fun <- function(x, plot_name = attributes(x)$ref) {
+  png(filename = paste0("Figures/K10/",plot_name,"_wavelet.png"),
+      width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
+  wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
+           timelab = "Days", verbose = FALSE, useRaster = TRUE,
+           periodlab = "Period", lwd = 1, graphics.reset = FALSE,
+           main = plot_name)
+  dev.off()
+}
+ldply(PN_wave, plot_fun)
+```
+
+# SACTN
 
@@
 
+# OISST
+
 ```{r}
@@
 ```
+
+# MUR
 
@@
 
+# G1SST
+
 ```{r}

### 2019-04-08 Exploring/Tester_code/methods_UC.Rmd
housekeeping

---
title: "methods_UC"
author: "Amieroh Abrahams"
date: "25 January 2019"
output: html_document
---

# Libraries

First I need to find, install and load various packages. These packages will be available on CRAN and can be accessed and installed in the usual way.

```{r prelim_opts, echo=FALSE}
knitr::opts_chunk$set(
  comment = "R>",
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(plyr)
library(lubridate)
library(ggpubr)
library(zoo)
library(lubridate)
library(FNN)
library(forecast)
library(astrochron)
library(WaveletComp)
library(data.table)
```

### Load site list SACTN data

Now to get to the data. The first step involves the loading of the site list. The statistical properties of the seawater temperature representing the South African coastline, such as the mean, minimum and maximum temperatures. These values vary among coastal sections due to the influence of the cold Benguala and warm Agulhas currents. Here we will only focus on the temperature data found along the west coast (wc) (i.e. sites influenced by the Benguela current, EBUS). The SACTN dataset comprise of 129 *in situ* coastal seawater temperatures derived from daily measurements over up to 40 years. The SACTN temperature dataset was compiled by measuring coastal temperatures at 129 sites along the coast of South Africa, daily from 1972 until 2017.

```{r load_files1, include=TRUE}
load("Data/site_list_v4.2.RData")
load("Data/SACTN_daily_v4.2.RData")
```

Now we select only the sites occuring along the west coast. Thereafter, I include all the sites which had a time series of 10years or greater. This yield a 2 rows of data for Paternoster. I removed the row of data obtained by the DEA, thus ending up with all of the data being obtained from SAWS in this dataset. I then exclude the sites that do not fall within the upwelling cell. Once those sites were removed, we have a total of 4 sites. These are: Port Nolloth, Lamberts Bay, Saldanna and Yzerfontein.

```{r}
site_list_sub_UC <- site_list %>% 
  filter(coast == "wc")

SACTN_daily_wc <- site_list_sub_UC %>% 
  filter(length >= 3650)

upwelling_sites_UC <- SACTN_daily_wc[-1:-8,] 
upwelling_sites_UC <- upwelling_sites_UC[c(-2,-5),] # New sites falling within the upwelling cell before yzerfontein

SACTN_daily_temps_UC <- SACTN_daily_v4.2 %>% 
  left_join(site_list[,c(4,13)], by = "index") %>% 
  filter(index %in% upwelling_sites_UC$index)
# save(SACTN_daily_temps, file = "SACTN_daily_temps.RData")

SACTN_daily_temps_indexsplit_UC <- SACTN_daily_temps_UC %>% 
  separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>% 
  dplyr::rename(insitu_temp = temp)

save(SACTN_daily_temps_indexsplit_UC, file = "Data/SACTN_daily_temps_indexsplit_UC.RData")
```

```{r}
SACTN_monthly <- function(df) {
  SACTN_split_monthly <- df %>%
  mutate(month = month(date, abbr = T, label = T),
         year = year(date)) %>%
  group_by(index, month, year) %>%
    mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",       
                           ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
                                ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
                                       ifelse(month %in% c("Oct", "Nov", "Dec"), "Spring","Error")))))
  return(SACTN_split_monthly)
}

# SACTN_split_UC <- SACTN_monthly(df = SACTN_daily_temps_indexsplit_UC)
load("~/Documents/Masters_2019/MastersProject/Data/SACTN_split_UC.RData")
SACTN_3yr_UC <- SACTN_split_UC %>% 
  filter(year(date) %in% seq(2014, 2015)) %>% 
  drop_na() # Removing NA values within the dataset

#save(SACTN_3yr_UC, file = "Data/SACTN_3yr_UC.RData")

insitu_UC_temp <- ggplot(data = SACTN_3yr_UC, aes(x = date, y = insitu_temp, colour = site)) +
geom_line(aes(group = site)) +
labs(x = "", y = "Temperature (°C)") +
theme(axis.text.x = element_text(angle = 45)) +
# theme_linedraw() +
theme(legend.position = "top")
ggsave(plot = insitu_UC_temp, filename = "Figures/insitu_UC_temp.pdf")
```

```{r}
## MUR dataset
MUR_Kommetjie_SST <- read_csv("Data/MUR_UC/MUR_Kommetjie_SST_timeseries_5nearest.csv")
MUR_Oudekraal_SST <- read_csv("Data/MUR_UC/MUR_Oudekraal_SST_timeseries_5nearest.csv")
MUR_Sea_Point_SST <- read_csv("Data/MUR_UC/MUR_Sea Point_SST_timeseries_5nearest.csv")

MUR_SST_UC <- rbind( MUR_Kommetjie_SST,MUR_Sea_Point_SST,MUR_Oudekraal_SST) %>%
  dplyr::rename(site = station)

MUR_SST_UC$date <- (ymd(MUR_SST_UC$date))
save(MUR_SST_UC, file = "Data/MUR_SST_UC.RData")

load("~/Documents/Masters_2019/MastersProject/Data/MUR_SST_UC.RData")
MUR_SST_UC <- MUR_SST_UC %>% 
  drop_na()
```

Here I write a function to match the remotely-sensed SST to the *in-situ* collected SST

```{r}
MUR_SST_UC <- MUR_SST_UC %>% 
  dplyr::rename(temp = nearest1) # nearest1- 1 pixel as applied to the OISST data

match_func <- function(df){
  match <- SACTN_3yr_UC  %>%  
  left_join(df, by = c("site", "date")) %>% 
  na.trim()
  return(match)
}

insitu_MUR_UC <- match_func(df = MUR_SST_UC) %>%
  drop_na()

match_plot <- function(df){
  plot1 <- df %>% 
ggplot(aes(x = date, y = temp)) +
  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ site, nrow = 2) +
  theme_bw()
  return(plot1)
}

MUR_plot <- match_plot(df = MUR_SST_UC)
mur_temp_plot_UC <- ggplot(data = insitu_MUR_UC, aes(x = date, y = temp, colour = site)) +
geom_line(aes(group = site)) +
labs(x = "", y = "Temperature (°C)") +
theme(axis.text.x = element_text(angle = 45)) +
#theme_linedraw() +
theme(legend.position = "top")

ggsave(plot = mur_temp_plot_UC, filename = "Figures/mur_temp_plot_UC.pdf")
```

#G1SST_sub_UC

```{r}
Kommetjie_SST <- read_csv("Data/G1SST_sub_UC/Kommetjie_SST_timeseries_5nearest.csv")
Sea_Point_SST <- read_csv("Data/G1SST_sub_UC/Sea Point_SST_timeseries_5nearest.csv")
Oudekraal_SST<- read_csv("Data/G1SST_sub_UC/Oudekraal_SST_timeseries_5nearest.csv")

G1SST_SST_UC <- rbind(Kommetjie_SST,Sea_Point_SST,Oudekraal_SST) %>%
  dplyr::rename(site = station)

G1SST_SST_UC$date <- (ymd(G1SST_SST_UC$date))
save(G1SST_SST_UC, file = "Data/G1SST_SST_UC.RData")

load("~/Documents/Masters_2019/MastersProject/Data/G1SST_SST_UC.RData")
G1SST_SST_UC <- G1SST_SST_UC %>% 
  drop_na()
```

```{r}
G1SST_SST_UC <- G1SST_SST_UC %>% 
  dplyr::rename(temp = nearest1) # nearest1- 1 pixel as applied to the OISST data

match_func <- function(df){
  match <- SACTN_3yr_UC  %>%  
  left_join(df, by = c("site", "date")) %>% 
  na.trim()
  return(match)
}

insitu_G1SST_SST_UC <- match_func(df = G1SST_SST_UC) %>%
  drop_na()

match_plot <- function(df){
  plot1 <- df %>% 
ggplot(aes(x = date, y = temp)) +
  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ site, nrow = 2) +
  theme_bw()
  return(plot1)
}

G1SST_SST_UC <- match_plot(df = G1SST_SST_UC)
G1SST_temp_plot_UC <- ggplot(data = insitu_G1SST_SST_UC, aes(x = date, y = temp, colour = site)) +
geom_line(aes(group = site)) +
labs(x = "", y = "Temperature (°C)") +
theme(axis.text.x = element_text(angle = 45)) +
#theme_linedraw() +
theme(legend.position = "top")

# ggsave(plot = G1SST_temp_plot_UC, filename = "Figures/G1SST_temp_plot_UC.pdf")
```

## OISST
```{r}
# AVHRR <- read_csv("data/csvavhrr-only-v2-19810901-20180630.csv")

OISSTDir <- "~/Documents/Masters_2019/Data_SST"
OISST <- fread(paste0(OISSTDir, "/csvavhrr-only-v2-19810901-20180630.csv"),
            col.names = c("lon", "lat", "temp", "date"))

#OISST <- as.tibble(fread("~/Documents/Masters_2019/MastersProject/Data/csvavhrr-only-v2-19810901-20180630.csv"))
# Quick visualisation of the data
# To explore the data I visualise the min temperatures along the South african coastline.
```

## Here I find the nearest SST pixels

Now we apply the FNN (Fast Nearest Neighbor) package to determine the nearesr SST pixel to the insitu collected sites.

```{r}
unique_pixel <- OISST %>%
  select(lon, lat) %>%
  unique()

# Select nearest 1 pixels (k = 1)
# here we use knnx to find the closes 1 pixels to the insitu sites
match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
                          query = as.matrix(upwelling_sites_UC[,5:6]), k = 1)

# Select SST pixels nearest to insitu sites
pixel_match <- unique_pixel[match_index,] %>%
  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
  mutate(site = upwelling_sites_UC$site)

## Subsetting the OISST data to match the upwelling sites within the in situ collected temperature data
OISST_match <- OISST %>%
  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
  filter(combi %in% pixel_match$combi)

OISST_match %>%
  filter(date == max(date)) %>%
  ggplot(aes(x = lon, y = lat)) +
  geom_raster(aes(fill = temp))
```

```{r}
OISST_sites_UC <- OISST_match %>%
  left_join(pixel_match, by = c("combi", "lon", "lat")) %>% 
  dplyr::rename(temp_OISST =temp)

# Check that the pixel that is shared is duplicated
# unique(combined_OISST_sites$site)
# # Visualise time series
# ggplot(data = OISST_sites, aes(x = date, y = temp)) +
#   geom_line() +
#   facet_wrap(~site)

# save(OISST_sites_UC, file = "Data/OISST_sites_UC.RData")
# load("~/Documents/MastersProject/Data/OISST_sites.RData")
load("~/Documents/Masters_2019/MastersProject/Data/OISST_sites_UC.RData")
OISST_sites_UC <- OISST_sites_UC %>% 
  dplyr::rename(temp = temp_OISST) %>%
  dplyr::mutate(date = as.Date(date)) %>% 
  drop_na()
```

## Matching the Insitu data with the OISST SST data

```{r}
insitu_OISST_UC <- match_func(df = OISST_sites_UC)
OISST_plot_UC <- match_plot(df = insitu_OISST_UC)

OISST_temp_plot_UC <- ggplot(data = insitu_OISST_UC, aes(x = date, y = temp, colour = site)) +
geom_line(aes(group = site)) +
labs(x = "", y = "Temperature (°C)") +
theme(axis.text.x = element_text(angle = 45)) +
# theme_linedraw() +
theme(legend.position = "top")
# ggsave(plot = OISST_temp_plot_UC, filename = "Figures/OISST_temp_plot_UC.pdf")
```

## Wavelet analyses

```{r}
load("Data/SACTN_3yr_UC.RData") 
temp.d <- SACTN_3yr_UC %>% 
  dplyr::group_by(site, date) %>% 
  dplyr::summarise(temp = mean(insitu_temp, na.rm = T)) %>%
  dplyr::mutate(no = seq(1:n())) %>%
  dplyr::ungroup() %>% 
  dplyr::select(site, no, temp, date)

SP <- temp.d %>% 
  dplyr::filter(site == "Sea Point") %>% 
  dplyr::select(temp, -date, -site) %>% 
  dplyr::mutate(no = seq(1:n())) %>% 
  dplyr::select(no, temp)

temp.d <- temp.d %>% 
  select(site, no, temp, date)
prewhite_fun <- function(x) {
  df <- x[, 2:3]
  out <- prewhiteAR(df, order = 3, method = "mle", aic = TRUE,
             genplot = FALSE, verbose = FALSE)
  colnames(out) <- c("no", "temp")
  return(out)
  }
PN_prewhite <- as.tibble(ddply(temp.d, .(site), prewhite_fun))
ggplot(PN_prewhite, aes(x = no, y = temp)) +
  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ site, nrow = 4) +
  theme_bw()
wl.fun <- function(x) {
  analyze.wavelet(x, "temp", loess.span = 0, dt = 1,
                   dj = 1/50, lowerPeriod = 2, make.pval = TRUE, n.sim = 50, 
                   method = "white.noise", verbose = FALSE)
}
SP_wave <- dlply(PN_prewhite, .(site), wl.fun)
# RWS: The order of these sites was incorrect
# Rather get the names from the pre-existing data
# site.names <- list("Port Nolloth", "Lamberts Bay", "Saldanha Bay", "Yzerfontein")
for (i in 1:length(SP_wave)) {
  attributes(SP_wave[[i]]) <- c(attributes(SP_wave[[i]]), ref = names(SP_wave)[i])
}
# x <- PN_wave[[3]]
plot_fun <- function(x, plot_name = attributes(x)$ref) {
  png(filename = paste0("Figures/Insitu/",plot_name,"_wavelet.png"),
      width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
  wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
           timelab = "Days", verbose = FALSE, useRaster = TRUE,
           periodlab = "Period", lwd = 1, graphics.reset = FALSE,
           main = plot_name)
  dev.off()
}
ldply(SP_wave, plot_fun)
# ... apply the function
#op <- par(no.readonly = TRUE)
# par(mfrow = c(2, 2),
#     mar = c(2.5, 3.5, 2.5, 0.5),
#     mgp = c(2.5, 1, 0),
#     oma = c(0, 0, 1, 0))
#par(op)
# Save each of the graph seperate- graph PN, LB, saldanha Bay and Yzerfontein
# l_ply(PN_wave, plot_fun)
```

```{r}
# OISST
# load("Data/OISST_sites.RData")

temp.d <- insitu_OISST_UC %>% 
  dplyr::group_by(site, date) %>% 
  dplyr::summarise(temp = mean(temp, na.rm = T)) %>%
  dplyr::mutate(no = seq(1:n())) %>%
  dplyr::ungroup() %>% 
  dplyr::select(site, no, temp, date)

SP <- temp.d %>% 
  dplyr::filter(site == "Sea Point") %>% 
  dplyr::select(temp, -date, -site) %>% 
  dplyr::mutate(no = seq(1:n())) %>% 
  dplyr::select(no, temp)

temp.d <- temp.d %>% 
  select(site, no, temp, date)
prewhite_fun <- function(x) {
  df <- x[, 2:3]
  out <- prewhiteAR(df, order = 3, method = "mle", aic = TRUE,
             genplot = FALSE, verbose = FALSE)
  colnames(out) <- c("no", "temp")
  return(out)
  }
PN_prewhite <- as.tibble(ddply(temp.d, .(site), prewhite_fun))
ggplot(PN_prewhite, aes(x = no, y = temp)) +
  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ site, nrow = 4) +
  theme_bw()
wl.fun <- function(x) {
  analyze.wavelet(x, "temp", loess.span = 0, dt = 1,
                   dj = 1/50, lowerPeriod = 2, make.pval = TRUE, n.sim = 50, 
                   method = "white.noise", verbose = FALSE)
}
SP_wave <- dlply(PN_prewhite, .(site), wl.fun)
# RWS: The order of these sites was incorrect
# Rather get the names from the pre-existing data
# site.names <- list("Port Nolloth", "Lamberts Bay", "Saldanha Bay", "Yzerfontein")
for (i in 1:length(SP_wave)) {
  attributes(SP_wave[[i]]) <- c(attributes(SP_wave[[i]]), ref = names(SP_wave)[i])
}
# x <- PN_wave[[3]]
plot_fun <- function(x, plot_name = attributes(x)$ref) {
  png(filename = paste0("Figures/OISST/",plot_name,"_wavelet.png"),
      width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
  wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
           timelab = "Days", verbose = FALSE, useRaster = TRUE,
           periodlab = "Period", lwd = 1, graphics.reset = FALSE,
           main = plot_name)
  dev.off()
}
ldply(SP_wave, plot_fun)
```

```{r}

# MUR
# load("Data/insitu_MUR.RData")

temp.d <- insitu_MUR_UC %>% 
  dplyr::group_by(site, date) %>% 
  dplyr::summarise(temp = mean(temp, na.rm = T)) %>%
  dplyr::mutate(no = seq(1:n())) %>%
  dplyr::ungroup() %>% 
  dplyr::select(site, no, temp, date)

SP <- temp.d %>% 
  dplyr::filter(site == "Sea Point") %>% 
  dplyr::select(temp, -date, -site) %>% 
  dplyr::mutate(no = seq(1:n())) %>% 
  dplyr::select(no, temp)

temp.d <- temp.d %>% 
  select(site, no, temp, date)
prewhite_fun <- function(x) {
  df <- x[, 2:3]
  out <- prewhiteAR(df, order = 3, method = "mle", aic = TRUE,
             genplot = FALSE, verbose = FALSE)
  colnames(out) <- c("no", "temp")
  return(out)
  }
PN_prewhite <- as.tibble(ddply(temp.d, .(site), prewhite_fun))
ggplot(PN_prewhite, aes(x = no, y = temp)) +
  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ site, nrow = 4) +
  theme_bw()
wl.fun <- function(x) {
  analyze.wavelet(x, "temp", loess.span = 0, dt = 1,
                   dj = 1/50, lowerPeriod = 2, make.pval = TRUE, n.sim = 50, 
                   method = "white.noise", verbose = FALSE)
}
SP_wave <- dlply(PN_prewhite, .(site), wl.fun)
# RWS: The order of these sites was incorrect
# Rather get the names from the pre-existing data
# site.names <- list("Port Nolloth", "Lamberts Bay", "Saldanha Bay", "Yzerfontein")
for (i in 1:length(SP_wave)) {
  attributes(SP_wave[[i]]) <- c(attributes(SP_wave[[i]]), ref = names(SP_wave)[i])
}
# x <- PN_wave[[3]]
plot_fun <- function(x, plot_name = attributes(x)$ref) {
  png(filename = paste0("Figures/MUR/",plot_name,"_wavelet.png"),
      width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
  wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
           timelab = "Days", verbose = FALSE, useRaster = TRUE,
           periodlab = "Period", lwd = 1, graphics.reset = FALSE,
           main = plot_name)
  dev.off()
}
ldply(SP_wave, plot_fun)
```

```{r}
# load("Data/insitu_G1SST.RData")

temp.d <- insitu_G1SST_SST_UC %>% 
  dplyr::group_by(site, date) %>% 
  dplyr::summarise(temp = mean(temp, na.rm = T)) %>%
  dplyr::mutate(no = seq(1:n())) %>%
  dplyr::ungroup() %>% 
  dplyr::select(site, no, temp, date)

SP <- temp.d %>% 
  dplyr::filter(site == "Sea Point") %>% 
  dplyr::select(temp, -date, -site) %>% 
  dplyr::mutate(no = seq(1:n())) %>% 
  dplyr::select(no, temp)

temp.d <- temp.d %>% 
  select(site, no, temp, date)
prewhite_fun <- function(x) {
  df <- x[, 2:3]
  out <- prewhiteAR(df, order = 3, method = "mle", aic = TRUE,
             genplot = FALSE, verbose = FALSE)
  colnames(out) <- c("no", "temp")
  return(out)
  }
PN_prewhite <- as.tibble(ddply(temp.d, .(site), prewhite_fun))
ggplot(PN_prewhite, aes(x = no, y = temp)) +
  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
  geom_line() +
  facet_wrap(~ site, nrow = 4) +
  theme_bw()
wl.fun <- function(x) {
  analyze.wavelet(x, "temp", loess.span = 0, dt = 1,
                   dj = 1/50, lowerPeriod = 2, make.pval = TRUE, n.sim = 50, 
                   method = "white.noise", verbose = FALSE)
}
SP_wave <- dlply(PN_prewhite, .(site), wl.fun)
# RWS: The order of these sites was incorrect
# Rather get the names from the pre-existing data
# site.names <- list("Port Nolloth", "Lamberts Bay", "Saldanha Bay", "Yzerfontein")
for (i in 1:length(SP_wave)) {
  attributes(SP_wave[[i]]) <- c(attributes(SP_wave[[i]]), ref = names(SP_wave)[i])
}
# x <- PN_wave[[3]]
plot_fun <- function(x, plot_name = attributes(x)$ref) {
  png(filename = paste0("Figures/G1SST/",plot_name,"_wavelet.png"),
      width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
  wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
           timelab = "Days", verbose = FALSE, useRaster = TRUE,
           periodlab = "Period", lwd = 1, graphics.reset = FALSE,
           main = plot_name)
  dev.off()
}
ldply(SP_wave, plot_fun)
```

### 2019-04-08 Exploring/Detect_MHWs.Rmd
housekeeping

--- 2019-04-02 Detect_MCSs.Rmd
@@
 ---
-title: "Long_time_series"
+title: "Detect_MHWs"
 author: "Amieroh Abrahams"
-date: "04 February 2019"
+date: "02 April 2019"
 output: html_document
 ---
-
-# Script Info (Time series)
-
-The first few lines of this script only looks at a time series of 10 years or greater.
-However, from line 360 a longer time series is used.
-This time series has a length of greater than 30 years and so only a few sites are used.
-With a longer time series I am able to detect for extreme events such as Marine Cold Spells (MCSs).
-
-** This script is very repetitive. Wavelet.Rmd creates a function to do this **
 
@@
 
-## Selecting the sites and creating new *in situ* datasets
-Now we select only the sites occuring along the west coast. Thereafter, I include all the sites which had a time series of 10 years or greater.
-
-```{r}
-
-# # Upwelling Cell 1:
-# site_list_sub <- site_list %>%
-#   filter(coast == "wc") %>%
-#   filter(length > 3650)
-# 
-# # upwelling_sites <- site_list_sub[-6,] # Here we had two rows of data for Paternoster; i removed the DEA row: All data here are now collected by SAWS
-# upwelling_sites <- upwelling_sites[-10:-14,]
-# 
-# # # # Here I find a time series with the data with a relativley long time series
-# UC_1 <- upwelling_sites[c(-2, -3, -5, -6, -8),]
-# save(UC_1, file = "Data/UC_1.RData")
-# SACTN_daily_temps <- SACTN_daily_v4.2 %>%
-#   left_join(site_list[,c(4,13)], by = "index") %>%
-#   filter(index %in% upwelling_sites$index)
-# save(SACTN_daily_temps, file = "Data/SACTN_daily_temps.RData")
-# # 
-# SACTN_US1 <- SACTN_daily_temps %>%
-#   separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>%
-#   dplyr::rename(insitu_temp = temp)
-# save(SACTN_US1, file = "Data/SACTN_US1.RData")
-
-# Upwelling Cell 2:
-# site_list_sub_UC <- site_list %>%
-#   filter(coast == "wc") %>%
-#   filter(length > 3650)
-# 
-# upwelling_sites_UC <- site_list_sub_UC[-1:-8,]
-# UC_2 <- upwelling_sites_UC[c(-2,-5),] # New sites falling within the upwelling cell before yzerfontein
-# save(UC_2, file = "Data/UC_2.RData")
-# 
-# 
-# SACTN_daily_temps_UC <- SACTN_daily_v4.2 %>%
-#   left_join(site_list[,c(4,13)], by = "index") %>%
-#   filter(index %in% upwelling_sites_UC$index)
-# save(SACTN_daily_temps, file = "Data/SACTN_daily_temps.RData")
-# 
-# SACTN_US2 <- SACTN_daily_temps_UC %>%
-#   separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>%
-#   dplyr::rename(insitu_temp = temp)
-# 
-# save(SACTN_US2, file = "Data/SACTN_US2.RData")
-
-####################################################################################
-
-# combined_US <- rbind(UC_1,UC_2)
-#save(combined_US, file = "Data/combined_US.RData")
-SACTN_temps_US <- SACTN_daily_v4.2 %>%
-  left_join(site_list[,c(4,13)], by = "index") %>%
-  filter(index %in% combined_US$index)
-
-SACTN_temps_US <- SACTN_temps_US %>%
-  separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>%
-  dplyr::rename(insitu_temp = temp)
-
-# save(SACTN_temps_US, file = "Data/SACTN_temps_US.RData")
-```
-
-I now create two datasets, `SACTN_split_US1` and `SACTN_split_US2`.
-These two datasets represents different groups of sites within the upwelling center based on the location of each of the sites.
-`SACTN_split_US1`: Port Nolloth, Lamberts Bay, Saldanna and Yzerfontein.
-`SACTN_split_US2`: Dassen Island, Sea Point, Hout Bay and Kommetjie.
-Here I also add monthly, yearly and seasonal column.
-This allows us to compare temperature variation on a seasonal basis.
-Research studies done suggests that the intensity and duration of upwelling varies on a seasonal basis.
-
-```{r}
-SACTN_monthly <- function(df) {
-  SACTN_split_monthly <- df %>%
-  mutate(month = month(date, abbr = T, label = T),
-         year = year(date)) %>%
-  group_by(index, month, year) %>%
-    mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",       
-                           ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
-                                ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
-                                       ifelse(month %in% c("Oct", "Nov", "Dec"), "Spring","Error")))))
-  return(SACTN_split_monthly)
-}
-
-# SACTN_split_US1 <- SACTN_monthly(df = SACTN_US1)
-# SACTN_split_US2 <- SACTN_monthly(df = SACTN_US2)
-# save(SACTN_split_US1, file = "SACTN_split_US1.RData")
-# save(SACTN_split_US2, file = "SACTN_split_US2.RData")
-```
-
-Find where different sites have overlapping time series within this `SACTN_temps_US` dataset. Finding a long time series.
-
-```{r}
-load("~/Documents/Masters_2019/MastersProject/Data/combined_US.RData")
-load("~/Documents/Masters_2019/MastersProject/Data/SACTN_temps_US.RData")
-# Find a period where each of the sites within that overlap.
-
-overlap_ts <- SACTN_temps_US %>% 
-  filter(year(date) %in% seq(1992, 2016)) %>% # The length of the time may vary depending on the question
-  drop_na() # Removing NA values within the dataset
-save(overlap_ts, file = "Data/overlap_ts.RData")
-
-temp_plot <- function(df){
-  plot <- ggplot(data = df, aes(x = date, y = insitu_temp, colour = site)) +
-    geom_line(aes(group = site)) +
-    labs(x = "", y = "Temperature (°C)") +
-    theme(axis.text.x = element_text(angle = 45)) +
-    theme(legend.position = "top")
-}
-
-SACTN_temps_US_plot <- temp_plot(df = overlap_ts)
-```
-
-In this thesis, different remotely-sensed SST datasets are used.
-Hence, these datasets represent different resolutions at which data was obtained.
-We will make use of five remotely-sensed SST datasets, these include; the  MUR, AVHRR OISST, K10 and G1SST data.
-**MODIS SST data need to be downloaded and used too**
-
-## MUR dataset
-```{r}
-MUR_Lamberts_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Lamberts Bay_SST_timeseries_5nearest.csv")
-MUR_Port_Nolloth <- read_csv("Data/MUR_nearest5pixels/MUR_Port Nolloth_SST_timeseries_5nearest.csv")
-MUR_Saldanha_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Saldanha Bay_SST_timeseries_5nearest.csv")
-MUR_Yzerfontein <- read_csv("Data/MUR_nearest5pixels/MUR_Yzerfontein_SST_timeseries_5nearest.csv")
-MUR_Dassen_Island <- read_csv("Data/MUR_nearest5pixels/MUR_Dassen Island_SST_timeseries_5nearest.csv")
-MUR_Hout_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Hout Bay_SST_timeseries_5nearest.csv")
-MUR_Kommetjie <- read_csv("Data/MUR_nearest5pixels/MUR_Kommetjie_SST_timeseries_5nearest.csv")
-MUR_Sea_Point <- read_csv("Data/MUR_nearest5pixels/MUR_Sea Point_SST_timeseries_5nearest.csv")
-MUR_Oudekraal <- read_csv("Data/MUR_nearest5pixels/MUR_Oudekraal_SST_timeseries_5nearest.csv")
-
-MUR_SST <- rbind(MUR_Lamberts_Bay,MUR_Yzerfontein, MUR_Port_Nolloth,MUR_Saldanha_Bay, MUR_Dassen_Island, MUR_Hout_Bay, MUR_Kommetjie, MUR_Sea_Point, MUR_Oudekraal) %>%
-  dplyr::rename(site = station)
-
-MUR_SST$date <- (ymd(MUR_SST$date))
-# save(MUR_SST, file = "Data/MUR_SST.RData")
-
-load("~/Documents/Masters_2019/MastersProject/Data/MUR_SST.RData")
-MUR_SST <- MUR_SST %>% 
-  drop_na()
-```
-
-This function matches the remotely-sensed SST to the *in-situ* collected SST
-
-```{r}
-MUR_SST <- MUR_SST %>% 
-  dplyr::rename(temp = nearest1) # nearest1- 1 pixel as applied to the OISST data
-
-match_func <- function(df){
-  match <- overlap_ts  %>%  
-  left_join(df, by = c("site", "date")) %>% 
-  na.trim()
-  return(match)
-}
-insitu_MUR <- match_func(df = MUR_SST) %>%
-  drop_na()
-# save(insitu_MUR, file = "Data/insitu_MUR.RData")
-
-match_plot <- function(df){
-  plot1 <- df %>% 
-ggplot(aes(x = date, y = temp)) +
-  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
-  geom_line() +
-  facet_wrap(~ site, nrow = 2) +
-  theme_bw()
-  return(plot1)
-}
-MUR_plot <- match_plot(df = insitu_MUR)
-
-temp_plot <- function(df){
-  temp_plot <- df %>% 
-  ggplot(aes(x = date, y = temp, colour = site)) +
-    geom_line(aes(group = site)) +
-    labs(x = "", y = "Temperature (°C)") +
-    theme(axis.text.x = element_text(angle = 45)) +
-    theme(legend.position = "top")
-}
-MUR_plot <- temp_plot(df = insitu_MUR)
-```
-
-## OISST dataset
-```{r}
-# AVHRR <- read_csv("data/csvavhrr-only-v2-19810901-20180630.csv")
-
-# OISSTDir <- "~/Documents/Masters_2019/Data_SST"
-# OISST <- fread(paste0(OISSTDir, "/csvavhrr-only-v2-19810901-20180630.csv"),
-#             col.names = c("lon", "lat", "temp", "date"))
-
-OISST <- as.tibble(fread("Data/csvavhrr-only-v2-19810901-20180630.csv"))
-# # Quick visualisation of the data
-# # To explore the data I visualise the min temperatures along the South african coastline.
-# 
-# OISST %>%
-#   filter(date == min(date)) %>%
-#   ggplot(aes(x = lon, y = lat)) +
-#   geom_raster(aes(fill = temp))
-# ggsave(plot = OISST, filename = "Figures/OISST")
-```
-
-## Here I find the nearest SST pixels
-
-Now we apply the FNN (Fast Nearest Neighbor) package to determine the nearesr SST pixel to the insitu collected sites.
-
-```{r}
-# unique_pixel <- OISST %>%
-#   select(lon, lat) %>%
-#   unique()
-# 
-# # Select nearest 1 pixels (k = 1)
-# # here we use knnx to find the closes 1 pixels to the insitu sites
-# match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
-#                           query = as.matrix(combined_US[,5:6]), k = 1)
-# 
-# # Select SST pixels nearest to insitu sites
-# pixel_match <- unique_pixel[match_index,] %>%
-#   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
-#   mutate(site = combined_US$site)
-# 
-# ## Subsetting the OISST data to match the upwelling sites within the in situ collected temperature data
-# OISST_match <- OISST %>%
-#   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
-#   filter(combi %in% pixel_match$combi)
-# 
-# ## Visualisation
-# ggplot(data = OISST_match, aes(x = date, y = temp_OISST)) +
-#   geom_line() +
-#   facet_wrap(~combi)
-# ggsave(plot = OISST_match, filename = "Figures/combined_OISST_match")
-# 
-# OISST_match %>%
-#   filter(date == max(date)) %>%
-#   ggplot(aes(x = lon, y = lat)) +
-#   geom_raster(aes(fill = temp))
-```
-
-```{r}
-# OISST_sites <- OISST_match %>%
-#   left_join(pixel_match, by = c("combi", "lon", "lat")) %>% 
-#   dplyr::rename(temp_OISST =temp)
-# 
-# # Check that the pixel that is shared is duplicated
-# # unique(combined_OISST_sites$site)
-# # Visualise time series
-# ggplot(data = OISST_sites, aes(x = date, y = temp)) +
-#   geom_line() +
-#   facet_wrap(~site)
-
-# save(OISST_sites, file = "Data/OISST_sites.RData")
-load("~/Documents/Masters_2019/MastersProject/Data/OISST_sites.RData")
-OISST_sites <- OISST_sites %>% 
-  dplyr::rename(temp = temp_OISST) %>%
-  dplyr::mutate(date = as.Date(date)) %>% 
-  drop_na()
-```
-
-## Matching the Insitu data with the OISST SST data
-```{r}
-insitu_OISST <- match_func(df = OISST_sites)
-# save(insitu_OISST, file = "Data/insitu_OISST.RData")
-OISST_plot <- match_plot(df = insitu_OISST)
-OISST_temp_plot <- temp_plot(df = insitu_OISST)
-```
-
-## G1SST dataset
-```{r}
-Lamberts_Bay <- read_csv("Data/G1SST_sub/Lamberts Bay_SST_timeseries_5nearest.csv")
-Port_Nolloth <- read_csv("Data/G1SST_sub/Port Nolloth_SST_timeseries_5nearest.csv")
-Saldanha_Bay <- read_csv("Data/G1SST_sub/Saldanha Bay_SST_timeseries_5nearest.csv")
-Yzerfontein <- read_csv("Data/G1SST_sub/Yzerfontein_SST_timeseries_5nearest.csv")
-Dassen_Island <- read_csv("Data/G1SST_sub/Dassen Island_SST_timeseries_5nearest.csv")
-Hout_Bay_SST <- read_csv("Data/G1SST_sub/Hout Bay_SST_timeseries_5nearest.csv")
-Kommetjie_SST <- read_csv("Data/G1SST_sub/Kommetjie_SST_timeseries_5nearest.csv")
-Oudekraal_SST <- read_csv("Data/G1SST_sub/Oudekraal_SST_timeseries_5nearest.csv")
-Sea_Point_SST<- read_csv("Data/G1SST_sub/Sea Point_SST_timeseries_5nearest.csv")
-
-G1SSTsub_SST <- rbind(Lamberts_Bay,Port_Nolloth, Saldanha_Bay, Yzerfontein, Dassen_Island, Hout_Bay_SST, Kommetjie_SST, Oudekraal_SST, Sea_Point_SST) %>%
-  dplyr::rename(site = station)
-
-G1SSTsub_SST$date <- (ymd(G1SSTsub_SST$date))
-save(G1SSTsub_SST, file = "Data/G1SSTsub_SST.RData")
-
-# load("Data/G1SSTsub_SST.RData")
-
-G1SSTsub_SST <- G1SSTsub_SST %>% 
-  drop_na() %>% 
-  dplyr::rename(temp = nearest1)
-
-insitu_G1SST <- match_func(df = G1SSTsub_SST) 
-# save(insitu_G1SST, file = "Data/insitu_G1SST.RData")
-G1SST_plot <- match_plot(df = insitu_G1SST )
-G1SST_temp_plot <- temp_plot(df = insitu_G1SST)
-```
-
-# K10
-```{r}
-Lamberts_Bay <- read_csv("Data/K10/K10_Lamberts Bay_SST_timeseries_5nearest.csv")
-Port_Nolloth <- read_csv("Data/K10/K10_Port Nolloth_SST_timeseries_5nearest.csv")
-Saldanha_Bay <- read_csv("Data/K10/K10_Saldanha Bay_SST_timeseries_5nearest.csv")
-Yzerfontein <- read_csv("Data/K10/K10_Yzerfontein_SST_timeseries_5nearest.csv")
-Dassen_Island <- read_csv("Data/K10/K10_Dassen Island_SST_timeseries_5nearest.csv")
-Hout_Bay <- read_csv("Data/K10/K10_Hout Bay_SST_timeseries_5nearest.csv")
-Kommetjie <- read_csv("Data/K10/K10_Kommetjie_SST_timeseries_5nearest.csv")
-Oudekraal <- read_csv("Data/K10/K10_Oudekraal_SST_timeseries_5nearest.csv")
-Sea_Point <- read_csv("Data/K10/K10_Sea Point_SST_timeseries_5nearest.csv")
-
-K10_SST <- rbind(Lamberts_Bay, Yzerfontein, Port_Nolloth, Saldanha_Bay, Dassen_Island, Hout_Bay, Kommetjie, Oudekraal, Sea_Point) %>%
-  dplyr::rename(site = station)
-K10_SST$date <- (ymd(K10_SST$date))
-# save(K10_SST, file = "Data/K10_SST.RData")
-
-load("~/Documents/Masters_2019/MastersProject/Data/K10_SST.RData")
-K10_SST <- K10_SST %>% 
-  drop_na()
-K10_SST <- K10_SST %>% 
-  dplyr::rename(temp = nearest1)
-
-insitu_K10 <- match_func(df = K10_SST) 
-# save(insitu_K10, file = "Data/insitu_K10.RData")
-K10_plot <- match_plot(df = insitu_K10)
-k10_temp_plot <- temp_plot(df = insitu_K10)
-```
-
 # Long term time series (>30years)
@@
 
-# Anatomy of marine cold spells within the upwelling center
+# Anatomy of marine heatwaves within the upwelling center
 Make use of the make_whole_fast function within the heatwaveR package. This function constructs a continuous, uninterrupted time series of temperatures (faster).
@@
 
-ts <- ts2clm(MUR_SST_PN, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
+ts <- ts2clm(MUR_SST_PN, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 90)
+res <- detect_event(ts)
 MCS <- res$climatology
-MCS_MUR_PN <- event_line(res, spread = 500, metric = "intensity_cumulative",
+MHW_MUR_PN <- event_line(res, spread = 500, metric = "intensity_cumulative",
            start_date = "2011-12-31", end_date = "2014-10-30") +
@@
 
-ts <- ts2clm(MUR_SST_LB, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
+ts <- ts2clm(MUR_SST_LB, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 90)
+res <- detect_event(ts)
 MCS <- res$climatology
-MCS_MUR_LB <- event_line(res, spread = 500, metric = "intensity_cumulative",
+MHW_MUR_LB <- event_line(res, spread = 500, metric = "intensity_cumulative",
            start_date = "2011-12-31", end_date = "2014-10-30") +
@@
 
-ts <- ts2clm(MUR_SST_YZ, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
+ts <- ts2clm(MUR_SST_YZ, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 90)
+res <- detect_event(ts)
 MCS <- res$climatology
-MCS_MUR_YZ <- event_line(res, spread = 500, metric = "intensity_cumulative",
+MHW_MUR_YZ <- event_line(res, spread = 500, metric = "intensity_cumulative",
            start_date = "2011-12-31", end_date = "2014-10-30") +
@@
 
-ts <- ts2clm(MUR_SST_SB, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
+ts <- ts2clm(MUR_SST_SB, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 90)
+res <- detect_event(ts)
 MCS <- res$climatology
-MCS_MUR_SB <- event_line(res, spread = 500, metric = "intensity_cumulative",
+MHW_MUR_SB <- event_line(res, spread = 500, metric = "intensity_cumulative",
            start_date = "2011-12-31", end_date = "2014-10-30") +
@@
 
-ts <- ts2clm(MUR_SST_SP, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
+ts <- ts2clm(MUR_SST_SP, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 90)
+res <- detect_event(ts)
 MCS <- res$climatology
-MCS_MUR_SP <- event_line(res, spread = 500, metric = "intensity_cumulative",
+MHW_MUR_SP <- event_line(res, spread = 500, metric = "intensity_cumulative",
            start_date = "2011-12-31", end_date = "2014-10-30") +
@@
 
-ts <- ts2clm(MUR_SST_O, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
+ts <- ts2clm(MUR_SST_O, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 90)
+res <- detect_event(ts)
 MCS <- res$climatology
-MCS_MUR_O <- event_line(res, spread = 500, metric = "intensity_cumulative",
+MHW_MUR_O <- event_line(res, spread = 500, metric = "intensity_cumulative",
            start_date = "2011-12-31", end_date = "2014-10-30") +
@@
 
-ts <- ts2clm(MUR_SST_Kom, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
+ts <- ts2clm(MUR_SST_Kom, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 90)
+res <- detect_event(ts)
 MCS <- res$climatology
-MCS_MUR_Kom <- event_line(res, spread = 500, metric = "intensity_cumulative",
+MHW_MUR_Kom <- event_line(res, spread = 500, metric = "intensity_cumulative",
            start_date = "2011-12-31", end_date = "2014-10-30") +
@@
 
-combined_MUR <- ggarrange(MCS_MUR_PN, MCS_MUR_LB, MCS_MUR_O, MCS_MUR_SP, MCS_MUR_KoM, MCS_MUR_SB, MCS_MUR_YZ, ncol = 2, nrow = 4)
+combined_MUR <- ggarrange(MHW_MUR_PN, MHW_MUR_LB, MHW_MUR_O, MHW_MUR_SP,MHW_MUR_Kom, MHW_MUR_SB, MHW_MUR_YZ, ncol = 2, nrow = 4)
 ```
@@
 
-ts <- ts2clm(G1SSTsub_SST_PN, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
+ts <- ts2clm(G1SSTsub_SST_PN, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 90)
+res <- detect_event(ts)
 MCS <- res$climatology
-MCS_g1sst_PN <- event_line(res, spread = 500, metric = "intensity_cumulative",
+MHW_g1sst_PN <- event_line(res, spread = 500, metric = "intensity_cumulative",
            start_date = "2011-12-31", end_date = "2014-10-30") +
@@
 
-ts <- ts2clm(G1SSTsub_SST_LB, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
+ts <- ts2clm(G1SSTsub_SST_LB, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 90)
+res <- detect_event(ts)
 MCS <- res$climatology
-MCS_g1sst_LB <- event_line(res, spread = 500, metric = "intensity_cumulative",
+MHW_g1sst_LB <- event_line(res, spread = 500, metric = "intensity_cumulative",
            start_date = "2011-12-31", end_date = "2014-10-30") +
@@
 
-ts <- ts2clm(G1SSTsub_SST_SB, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
+ts <- ts2clm(G1SSTsub_SST_SB, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 90)
+res <- detect_event(ts)
 MCS <- res$climatology
-MCS_g1sst_SB <- event_line(res, spread = 500, metric = "intensity_cumulative",
+MHW_g1sst_SB <- event_line(res, spread = 500, metric = "intensity_cumulative",
            start_date = "2011-12-31", end_date = "2014-10-30") +
@@
 
-ts <- ts2clm(G1SSTsub_SST_YZ, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
+ts <- ts2clm(G1SSTsub_SST_YZ, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 90)
+res <- detect_event(ts)
 MCS <- res$climatology
-MCS_g1sst_YZ <- event_line(res, spread = 500, metric = "intensity_cumulative",
+MHW_g1sst_YZ <- event_line(res, spread = 500, metric = "intensity_cumulative",
            start_date = "2011-12-31", end_date = "2014-10-30") +
@@
 
-ts <- ts2clm(G1SSTsub_SST_SP, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
+ts <- ts2clm(G1SSTsub_SST_SP, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 90)
+res <- detect_event(ts)
 MCS <- res$climatology
-MCS_g1sst_SP <- event_line(res, spread = 500, metric = "intensity_cumulative",
+MHW_g1sst_SP <- event_line(res, spread = 500, metric = "intensity_cumulative",
            start_date = "2011-12-31", end_date = "2014-10-30") +
@@
 
-ts <- ts2clm(G1SSTsub_SST_O, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
+ts <- ts2clm(G1SSTsub_SST_O, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 90)
+res <- detect_event(ts)
 MCS <- res$climatology
-MCS_g1sst_O <- event_line(res, spread = 500, metric = "intensity_cumulative",
+MHW_g1sst_O <- event_line(res, spread = 500, metric = "intensity_cumulative",
            start_date = "2011-12-31", end_date = "2014-10-30") +
@@
 
-ts <- ts2clm(G1SSTsub_SST_Kom, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
+ts <- ts2clm(G1SSTsub_SST_Kom, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 90)
+res <- detect_event(ts)
 MCS <- res$climatology
-MCS_g1sst_Kom <- event_line(res, spread = 500, metric = "intensity_cumulative",
+MHW_g1sst_Kom <- event_line(res, spread = 500, metric = "intensity_cumulative",
            start_date = "2011-12-31", end_date = "2014-10-30") +
@@
 
-combined_G1SST <- ggarrange(MCS_g1sst_PN, MCS_g1sst_LB, MCS_g1sst_O, MCS_g1sst_SP, MCS_g1sst_Kom, MCS_g1sst_YZ, MCS_g1sst_SB, ncol = 2, nrow = 4)
+combined_G1SST <- ggarrange(MHW_g1sst_PN, MHW_g1sst_LB, MHW_g1sst_O, MHW_g1sst_SP, MHW_g1sst_Kom, MHW_g1sst_YZ, MHW_g1sst_SB, ncol = 2, nrow = 4)
 ```
@@
 
-ts <- ts2clm(OISST_PN, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
-MCS <- res$climatology
-MCS_OISST_PN <- event_line(res, spread = 900, metric = "intensity_cumulative",
+ts <- ts2clm(OISST_PN, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 90)
+res <- detect_event(ts)
+MHW_OISST_PN <- event_line(res, spread = 900, metric = "intensity_cumulative",
            start_date = "2010-12-31", end_date = "2014-12-31") +
@@
 
-ts <- ts2clm(OISST_LB, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
-MCS <- res$climatology
-MCS_OISST_LB <- event_line(res, spread = 900, metric = "intensity_cumulative",
+ts <- ts2clm(OISST_LB, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 90)
+res <- detect_event(ts)
+MHW_OISST_LB <- event_line(res, spread = 900, metric = "intensity_cumulative",
            start_date = "2010-12-31", end_date = "2014-12-31") +
@@
 
-ts <- ts2clm(OISST_SB, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
-MCS <- res$climatology
-MCS_OISST_SB <- event_line(res, spread = 500, metric = "intensity_cumulative",
+ts <- ts2clm(OISST_SB, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 90)
+res <- detect_event(ts)
+MHW_OISST_SB <- event_line(res, spread = 500, metric = "intensity_cumulative",
            start_date = "2010-12-31", end_date = "2014-12-31") +
@@
 
-ts <- ts2clm(OISST_YZ, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
-MCS <- res$climatology
-MCS_OISST_YZ <- event_line(res, spread = 500, metric = "intensity_cumulative",
+ts <- ts2clm(OISST_YZ, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 90)
+res <- detect_event(ts)
+MHW_OISST_YZ <- event_line(res, spread = 500, metric = "intensity_cumulative",
            start_date = "2010-12-31", end_date = "2014-12-31") +
@@
 
-ts <- ts2clm(OISST_SP, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
-MCS <- res$climatology
-MCS_OISST_SP <- event_line(res, spread = 500, metric = "intensity_cumulative",
+ts <- ts2clm(OISST_SP, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 90)
+res <- detect_event(ts)
+MHW_OISST_SP <- event_line(res, spread = 500, metric = "intensity_cumulative",
            start_date = "2010-12-31", end_date = "2014-12-31") +
@@
 
-ts <- ts2clm(OISST_Kom, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
-MCS <- res$climatology
-MCS_OISST_Kom <- event_line(res, spread = 500, metric = "intensity_cumulative",
+ts <- ts2clm(OISST_Kom, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 90)
+res <- detect_event(ts)
+MHW_OISST_Kom <- event_line(res, spread = 500, metric = "intensity_cumulative",
            start_date = "2010-12-31", end_date = "2014-12-31") +
@@
 
-ts <- ts2clm(OISST_O, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 10)
-res <- detect_event(ts, coldSpells = TRUE)
-MCS <- res$climatology
-MCS_OISST_O <- event_line(res, spread = 900, metric = "intensity_cumulative",
+ts <- ts2clm(OISST_O, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 90)
+res <- detect_event(ts)
+MHW_OISST_O <- event_line(res, spread = 900, metric = "intensity_cumulative",
            start_date = "2011-12-31", end_date = "2014-12-31") +
@@
 
-combined_OISST <- ggarrange(MCS_OISST_PN, MCS_OISST_LB, MCS_OISST_SB, MCS_OISST_YZ, MCS_OISST_O, MCS_OISST_Kom,MCS_OISST_SP, ncol = 2, nrow = 4)
+combined_OISST <- ggarrange(MHW_OISST_PN, MHW_OISST_LB, MHW_OISST_SB, MHW_OISST_YZ, MHW_OISST_O, MHW_OISST_Kom,MHW_OISST_SP, ncol = 2, nrow = 4)
 ```
-
-# K10
-Time series is less than 10 years and so the detection of MCS is not effective. 
-heatwaveR package may be applied to datasets that have a climatological period of >10years
-
-```{r}
-# load("~/Documents/Masters_2019/MastersProject/Data/K10_SST.RData")
-# 
-# K10_PN<- K10_SST %>% 
-#   filter(site == "Port Nolloth") %>% 
-#   dplyr::rename(temp = nearest1) %>% 
-#   select(date, temp) %>% 
-#   dplyr::rename(t = date)
-# 
-# ts <- ts2clm(K10_PN, climatologyPeriod = c("2008-04-02", "2015-12-31"), pctile = 10)
-# res <- detect_event(ts, coldSpells = TRUE)
-# MCS <- res$climatology
-# MCS_K10_PN <- event_line(res, spread = 100, metric = "intensity_cumulative",
-#            start_date = "2013-04-02", end_date = "2015-12-31") +
-#   ggtitle("K10_PN")
-# 
-# K10_LB <- K10_SST %>% 
-#   filter(site == "Lamberts Bay") %>% 
-#   dplyr::rename(temp = nearest1) %>% 
-#   select(date, temp) %>% 
-#   dplyr::rename(t = date)
-# 
-# ts <- ts2clm(K10_LB, climatologyPeriod = c("2008-04-02", "2015-12-31"), pctile = 10)
-# res <- detect_event(ts, coldSpells = TRUE)
-# MCS <- res$climatology
-# MCS_K10_LB <- event_line(res, spread = 100, metric = "intensity_cumulative",
-#            start_date = "2013-04-02", end_date = "2015-12-31") +
-#   ggtitle("K10_LB")
-# 
-# K10_SB <- K10_SST %>% 
-#   filter(site == "Saldanha Bay") %>% 
-#   dplyr::rename(temp = nearest1) %>% 
-#   select(date, temp) %>% 
-#   dplyr::rename(t = date)
-# 
-# ts <- ts2clm(K10_SB, climatologyPeriod = c("2008-04-02", "2015-12-31"), pctile = 10)
-# res <- detect_event(ts, coldSpells = TRUE)
-# MCS <- res$climatology
-# MCS_K10_SB <- event_line(res, spread = 100, metric = "intensity_cumulative",
-#            start_date = "2013-04-02", end_date = "2015-12-31") +
-#   ggtitle("K10_SB")
-# 
-# K10_YZ <- K10_SST %>% 
-#   filter(site == "Yzerfontein") %>% 
-#   dplyr::rename(temp = nearest1) %>% 
-#   select(date, temp) %>% 
-#   dplyr::rename(t = date)
-# 
-# ts <- ts2clm(K10_YZ, climatologyPeriod = c("2008-04-02", "2015-12-31"), pctile = 10)
-# res <- detect_event(ts, coldSpells = TRUE)
-# MCS <- res$climatology
-# MCS_K10_YZ <- event_line(res, spread = 100, metric = "intensity_cumulative",
-#            start_date = "2013-04-02", end_date = "2015-12-31") +
-#   ggtitle("K10_YZ")
-# 
-# K10_SP <- K10_SST %>% 
-#   filter(site == "Sea Point") %>% 
-#   dplyr::rename(temp = nearest1) %>% 
-#   select(date, temp) %>% 
-#   dplyr::rename(t = date)
-# 
-# ts <- ts2clm(K10_SP, climatologyPeriod = c("2008-04-02", "2015-12-31"), pctile = 10)
-# res <- detect_event(ts, coldSpells = TRUE)
-# MCS <- res$climatology
-# MCS_K10_SP <- event_line(res, spread = 100, metric = "intensity_cumulative",
-#            start_date = "2013-04-02", end_date = "2015-12-31") +
-#   ggtitle("K10_SP")
-# 
-# K10_HB <- K10_SST %>% 
-#   filter(site == "Hout Bay") %>% 
-#   dplyr::rename(temp = nearest1) %>% 
-#   select(date, temp) %>% 
-#   dplyr::rename(t = date)
-# 
-# ts <- ts2clm(K10_HB, climatologyPeriod = c("2008-04-02", "2015-12-31"), pctile = 10)
-# res <- detect_event(ts, coldSpells = TRUE)
-# MCS <- res$climatology
-# MCS_K10_HB <- event_line(res, spread = 100, metric = "intensity_cumulative",
-#            start_date = "2013-04-02", end_date = "2015-12-31") +
-#   ggtitle("K10_HB")
-# 
-# K10_Kom <- K10_SST %>% 
-#   filter(site == "Kommetjie") %>% 
-#   dplyr::rename(temp = nearest1) %>% 
-#   select(date, temp) %>% 
-#   dplyr::rename(t = date)
-# 
-# ts <- ts2clm(K10_Kom, climatologyPeriod = c("2008-04-02", "2015-12-31"), pctile = 10)
-# res <- detect_event(ts, coldSpells = TRUE)
-# MCS <- res$climatology
-# MCS_K10_Kom <- event_line(res, spread = 100, metric = "intensity_cumulative",
-#            start_date = "2013-04-02", end_date = "2015-12-31") +
-#   ggtitle("K10_Kom")
-# 
-# K10_O <- K10_SST %>% 
-#   filter(site == "Oudekraal") %>% 
-#   dplyr::rename(temp = nearest1) %>% 
-#   select(date, temp) %>% 
-#   dplyr::rename(t = date)
-# 
-# ts <- ts2clm(K10_O, climatologyPeriod = c("2008-04-02", "2015-12-31"), pctile = 10)
-# res <- detect_event(ts, coldSpells = TRUE)
-# MCS <- res$climatology
-# MCS_K10_O <- event_line(res, spread = 100, metric = "intensity_cumulative",
-#            start_date = "2013-04-02", end_date = "2015-12-31") +
-#   ggtitle("K10_O")
-# 
-# K10_DI <- K10_SST %>% 
-#   filter(site == "Dassen Island") %>% 
-#   dplyr::rename(temp = nearest1) %>% 
-#   select(date, temp) %>% 
-#   dplyr::rename(t = date)
-# 
-# ts <- ts2clm(K10_DI, climatologyPeriod = c("2008-04-02", "2015-12-31"), pctile = 10)
-# res <- detect_event(ts, coldSpells = TRUE)
-# MCS <- res$climatology
-# MCS_K10_DI <- event_line(res, spread = 100, metric = "intensity_cumulative",
-#            start_date = "2013-04-02", end_date = "2015-12-31") +
-#   ggtitle("K10_DI")
-# 
-# 
-# combined_K10 <- ggarrange(MCS_K10_PN, MCS_K10_LB, MCS_K10_DI, MCS_K10_SB, MCS_K10_SP, MCS_K10_O, MCS_K10_Kom, MCS_K10_HB, MCS_K10_YZ, ncol = 3, nrow = 3)
-```

### 2019-04-08 Exploring/wavelets.Rmd
housekeeping

--- 2019-03-15 wavelets.Rmd
@@
 The rest of this code is kept constant for each of the datasets (Remotely-sensed SST and In-situ seawater temperature)
-SACTN_fyears <- SACTN_fyears %>%
-  dplyr::rename(temp = insitu_temp)
+# SACTN_fyears <- SACTN_fyears %>%
+#   dplyr::rename(temp = insitu_temp)
 
@@
 
-temp.d <- temp.d(df = SACTN_fyears) 
+# temp.d <- temp.d(df = SACTN_fyears) 
 # temp.d <- temp.d(df = MUR_fyears) 
@@
 # temp.d <- temp.d(df = G1SST_fyears) 
-# temp.d <- temp.d(df = K10_fyears)
+temp.d <- temp.d(df = K10_fyears)
 

### 2019-04-09 Exploring/Detect_MCSs.Rmd
fixed_bits

--- 2019-04-02 Detect_MCSs.Rmd
@@
 
-# OISSTDir <- "~/Documents/Masters_2019/Data_SST"
-# OISST <- fread(paste0(OISSTDir, "/csvavhrr-only-v2-19810901-20180630.csv"),
-#             col.names = c("lon", "lat", "temp", "date"))
+OISSTDir <- "~/Documents/Masters_2019/Data_SST"
+OISST <- fread(paste0(OISSTDir, "/csvavhrr-only-v2-19810901-20180630.csv"),
+            col.names = c("lon", "lat", "temp", "date"))
 
-OISST <- as.tibble(fread("Data/csvavhrr-only-v2-19810901-20180630.csv"))
+# OISST <- as.tibble(fread("Data/csvavhrr-only-v2-19810901-20180630.csv"))
 # # Quick visualisation of the data
@@
 ```{r}
-# unique_pixel <- OISST %>%
-#   select(lon, lat) %>%
-#   unique()
+unique_pixel <- OISST %>%
+  select(lon, lat) %>%
+  unique()
 # 
@@
 # # here we use knnx to find the closes 1 pixels to the insitu sites
-# match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
-#                           query = as.matrix(combined_US[,5:6]), k = 1)
+match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
+                          query = as.matrix(combined_US[,5:6]), k = 1)
 # 

### 2019-04-09 Paper1_results.Rmd
fixed_bits

@@
 ```{r load_files1, include=TRUE}
-load("Data/site_list_v4.2.RData")
-load("Data/SACTN_daily_v4.2.RData")
+load("Data_P1/site_list_v4.2.RData")
+load("Data_P1/SACTN_daily_v4.2.RData")
 ```
@@
 site_list_sub <- site_list_sub[c(-2, -3, -5, -6, -7,-9, -12, -13, -14),] # Here I keep all the sites with temperature obtained up to 2017
-save(site_list_sub, file = "Data_P1/site_list_sub")
+save(site_list_sub, file = "Data_P1/site_list_sub.Rdata")
 
@@
 
-In this thesis, different remotely-sensed SST datasets are used.
-Hence, these datasets represent different resolutions at which data was obtained.
-We will make use of five remotely-sensed SST datasets, these include; the  MUR, AVHRR OISST, K10 and G1SST data.
-**MODIS SST data need to be downloaded and used too**
+In this thesis, different remotely-sensed SST datasets are used. Hence, these datasets represent different resolutions at which data was obtained.
 
@@
 
-insitu_MUR <- match_func(df = MUR_SST) %>%
-  drop_na()
-# save(insitu_MUR, file = "Data_P1/insitu_MUR.RData")
-
 # Basic visualisations
@@
 
-MUR_plot <- match_plot(df = insitu_MUR)
-
 temp_plot <- function(df){
@@
 }
+
+## Matching the in-situ data with the G1SST SST data
+insitu_MUR <- match_func(df = MUR_SST) %>%
+  drop_na()
+# save(insitu_MUR, file = "Data_P1/insitu_MUR.RData")
+MUR_plot <- match_plot(df = insitu_MUR)
 MUR_plot <- temp_plot(df = insitu_MUR)
@@
 
+## Matching the Insitu data with the G1SST SST data
 insitu_G1SST <- match_func(df = G1SSTsub_SST) 
-save(insitu_G1SST, file = "Data_P1/insitu_G1SST.RData")
+# save(insitu_G1SST, file = "Data_P1/insitu_G1SST.RData")
 G1SST_plot <- match_plot(df = insitu_G1SST )
@@
 
+## Matching the Insitu data with the K10 SST data
 insitu_K10 <- match_func(df = K10_SST) 
 # save(insitu_K10, file = "Data_P1/insitu_K10.RData")
-
 K10_plot <- match_plot(df = insitu_K10)
@@
 
-load("Data_P1/SACTN_overlap.RData") # Loading the SACTN_overlap dataset which allows me to match it to the OISST data
+load("Data_P1/site_list_sub.RData") # Loading the SACTN_overlap dataset which allows me to match it to the OISST data
 
@@
 match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
-                      query = as.matrix(SACTN_overlap[,5:6]), k = 1)
+                      query = as.matrix(site_list_sub[,5:6]), k = 1)
 
-# Select SST pixels nearest to insitu sites
 pixel_match <- unique_pixel[match_index,] %>%
   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
-  mutate(site = SACTN_overlap$site)
-# save(pixel_match, file = "Data_P1/pixel_match.RData")
-load("Data_P1/pixel_match.RData")
+  mutate(site = site_list_sub$site)
 
@@
 
-load("Data_P1/OISST_match.RData")
 OISST_sites <- OISST_match %>%
@@
 
-# Visualise time series
-ggplot(data = OISST_sites, aes(x = date, y = temp)) +
-  geom_line() +
-  facet_wrap(~site)
-
-# save(OISST_sites, file = "Data/OISST_sites.RData")
-load("~/Documents/Masters_2019/MastersProject/Data/OISST_sites.RData")
 OISST_sites <- OISST_sites %>% 
@@
   drop_na()
-```
+# save(OISST_sites, file = "Data_P1/OISST_sites.RData")
 
 ## Matching the Insitu data with the OISST SST data
-```{r}
 insitu_OISST <- match_func(df = OISST_sites)
-# save(insitu_OISST, file = "Data/insitu_OISST.RData")
+# save(insitu_OISST, file = "Data_P1/insitu_OISST.RData")
 OISST_plot <- match_plot(df = insitu_OISST)

### 2019-04-09 Paper1_results.Rmd
updated_script

@@
 ```
+
+## Wavelets
+
+Wavelet anlyses showing the temperature variation between sites along the west coast for the years 2014 and 2015.
+Thereafter, wavelet analyses ae created per season for each individual site along the west coast of South Africa.
+Working with remotely-sensed SST and *in-situ* collected coastal seawater temperature, in the analyses bellow I aim to show how the intensity and duration of upwelling events vary between sites along the west coast of South Africa.
+It is important to remember that each of the datasets used within this study obtained SSTs at different resolutions.
+
+### Load site list SACTN data
+Now to get to the data.
+The first step involves the loading of the site list.
+The statistical properties of the seawater temperature representing the South African coastline, such as the mean, minimum and maximum temperatures.
+These values vary among coastal sections due to the influence of the cold Benguala and warm Agulhas currents.
+Here we will only focus on the temperature data found along the west coast (wc) (i.e.
+sites influenced by the Benguela current, EBUS).
+
+```{r}
+load("Data_P1/site_list_v4.2.RData")
+load("Data_P1/SACTN_daily_v4.2.RData")
+```
+
+## Load in the rest of the datasets
+
+* See the Dectect_MCS.Rmd script for more info.
+The datasets loaded here are only for temperatures collected between the years 1992-2016.
+However if a longer time series is required changes may  be made to the Dectect_MCS.Rmd script (line 136) .
+MCS using a long term time series to detect any extreme events occuring within the region
+
+```{r}
+load("Data_P1/SACTN_overlap.RData")
+load("Data_P1/insitu_MUR.RData")
+load("Data_P1/insitu_OISST.RData")
+load("Data_P1/insitu_G1SST.RData")
+load("Data_P1/insitu_K10.RData")
+```
+
+# Time series of one year
+
+Examining the intensity and duration of upwelling for the years 2014 - 2015
+
+```{r}
+filtered_years <- function(df){
+  upwelling<- df %>% 
+  filter(year(date) %in% seq(2002, 2015)) %>% 
+  drop_na()
+}
+
+SACTN_fyears <- filtered_years(df = SACTN_overlap) 
+MUR_fyears <- filtered_years(df = insitu_MUR) 
+OISST_fyears <- filtered_years(df = insitu_OISST) 
+G1SST_fyears <- filtered_years(df = insitu_G1SST) 
+K10_fyears <- filtered_years(df = insitu_K10)
+
+# save(SACTN_fyears, file = "Data/temp_2014_2015/SACTN_fyears.RData")
+# save(MUR_fyears, file = "Data/temp_2014_2015/MUR_fyears.RData")
+# save(OISST_fyears, file = "Data/temp_2014_2015/OISST_fyears.RData")
+# save(G1SST_fyears, file = "Data/temp_2014_2015/G1SST_fyears.RData")
+# save(K10_fyears, file = "Data/temp_2014_2015/K10_fyears.RData")
+```
+
+# Wavelet analyses
+
+```{r}
+# When doing a wavelet analyses on the SACTN dataset the "insitu_temp" column needed to be renamed to "temp" for the rest of the code to run.
+The rest of this code is kept constant for each of the datasets (Remotely-sensed SST and In-situ seawater temperature)
+# SACTN_fyears <- SACTN_fyears %>%
+#   dplyr::rename(temp = insitu_temp)
+
+temp.d <- function(df){
+temp.d <- df %>% 
+  dplyr::group_by(site, date) %>% 
+  dplyr::summarise(temp = mean(temp, na.rm = T)) %>%
+  dplyr::mutate(no = seq(1:n())) %>%
+  dplyr::ungroup() %>% 
+  dplyr::select(site, no, temp, date)
+}
+
+# temp.d <- temp.d(df = SACTN_fyears) 
+# temp.d <- temp.d(df = MUR_fyears) 
+# temp.d <- temp.d(df = OISST_fyears)
+# temp.d <- temp.d(df = G1SST_fyears) 
+temp.d <- temp.d(df = K10_fyears)
+
+temp.d <- temp.d %>% 
+  select(site, no, temp, date)
+
+prewhite_fun <- function(x) {
+  df <- x[, 2:3]
+  out <- prewhiteAR(df, order = 3, method = "mle", aic = TRUE,
+             genplot = FALSE, verbose = FALSE)
+  colnames(out) <- c("no", "temp")
+  return(out)
+  }
+PN_prewhite <- as.tibble(ddply(temp.d, .(site), prewhite_fun))
+
+ggplot(PN_prewhite, aes(x = no, y = temp)) +
+  geom_hline(aes(yintercept = mean(temp)), colour = "salmon") +
+  geom_line() +
+  facet_wrap(~ site, nrow = 4) +
+  theme_bw()
+
+wl.fun <- function(x) {
+  analyze.wavelet(x, "temp", loess.span = 0, dt = 1,
+                   dj = 1/50, lowerPeriod = 2, make.pval = TRUE, n.sim = 50, 
+                   method = "white.noise", verbose = FALSE)
+}
+
+PN_wave <- dlply(PN_prewhite, .(site), wl.fun)
+
+for (i in 1:length(PN_wave)) {
+  attributes(PN_wave[[i]]) <- c(attributes(PN_wave[[i]]), ref = names(PN_wave)[i])
+}
+# x <- PN_wave[[3]]
+plot_fun <- function(x, plot_name = attributes(x)$ref) {
+  png(filename = paste0("Figures/new",plot_name,"_wavelets.png"),
+      width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
+  wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
+           timelab = "Days", verbose = FALSE, useRaster = TRUE,
+           periodlab = "Period", lwd = 1, graphics.reset = FALSE,
+           main = plot_name)
+  dev.off()
+}
+ldply(PN_wave, plot_fun)
+```

### 2019-04-12 Paper1_results.Rmd
Data_push

@@
 
-# Libraries
+# Loading libraries
 
@@
 library(heatwaveR)
+library(viridis)
+library(ggrepel)
+library(plyr)
+library(maptools)
+library(sp)
+library(geosphere)
+library(marmap)
+library(PBSmapping)
+library(scales)
+library(grid)
+library(gridExtra)
+library(fossil)
+library(mapproj)
+library(rgeos)
+
+## Functions
+source("functions/earthdist.R")
 ```
@@
 
-* See the Dectect_MCS.Rmd script for more info.
-The datasets loaded here are only for temperatures collected between the years 1992-2016.
-However if a longer time series is required changes may  be made to the Dectect_MCS.Rmd script (line 136) .
-MCS using a long term time series to detect any extreme events occuring within the region
-
 ```{r}
@@
   upwelling<- df %>% 
-  filter(year(date) %in% seq(2002, 2015)) %>% 
+  filter(year(date) %in% seq(2010, 2014)) %>% 
   drop_na()
@@
 K10_fyears <- filtered_years(df = insitu_K10) 
-
-# save(SACTN_fyears, file = "Data/temp_2014_2015/SACTN_fyears.RData")
-# save(MUR_fyears, file = "Data/temp_2014_2015/MUR_fyears.RData")
-# save(OISST_fyears, file = "Data/temp_2014_2015/OISST_fyears.RData")
-# save(G1SST_fyears, file = "Data/temp_2014_2015/G1SST_fyears.RData")
-# save(K10_fyears, file = "Data/temp_2014_2015/K10_fyears.RData")
 ```
@@
 The rest of this code is kept constant for each of the datasets (Remotely-sensed SST and In-situ seawater temperature)
-# SACTN_fyears <- SACTN_fyears %>%
-#   dplyr::rename(temp = insitu_temp)
+
+SACTN_fyears <- SACTN_fyears %>%
+  dplyr::rename(temp = insitu_temp)
 
@@
 plot_fun <- function(x, plot_name = attributes(x)$ref) {
-  png(filename = paste0("Figures/new",plot_name,"_wavelets.png"),
+  png(filename = paste0("Figures",plot_name,"_wavelets.png"),
       width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
@@
 ```
+
+############################################################################################################################################
+
+# Calculating a shore transect
+The aim of this is to find the temperature collected at different distances from the coast (distances = 15km, 25km, 35km and 50km).
+Then create wavelet analyses (or another analyses) where I compare how the intensity and duration of upwelling vary at different distances from the coastline.
+So would I have a dataset with temperature at 15km then at 25km then at 35km and then at 50km and so on...?
+
+```{r}
+load("Data_P1/site_list_sub.Rdata")
+west <- site_list_sub
+west$coast <- "west" # Chnages wc to west
+
+load("Data_P1/africa_coast.RData")
+
+## Downloading the bathy data from NOAA
+# Download mid-res bathymetry data
+# sa_lat <- c(-38, -24.5); sa_lon <- c(11.5, 35.5)
+# sa_bathy <- as.xyz(getNOAA.bathy(lon1 = sa_lon[1], lon2 = sa_lon[2], lat1 = sa_lat[1], lat2 = sa_lat[2], resolution = 4))
+# colnames(sa_bathy) <- c("lon", "lat", "depth")
+# sa_bathy <- sa_bathy[sa_bathy$depth <= 0,]
+# save(sa_bathy, file = "Data_P1/bathy/sa_bathy.RData")
+
+# Loading in the newly downloaded bathymetry data
+load("Data_P1/bathy/sa_bathy.RData")
+
+# This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
+shore.normal.transect <- function(site, width = 2){
+  # Find the site on the coastline and it's nearest neighbour points
+  coords <- data.frame(lon = site$lon, lat = site$lat)
+  coords2 <- knnx.index(africa_coast[,1:2], as.matrix(coords), k = 1)
+  coords3 <- data.frame(site = site$site, africa_coast[c(coords2-width, coords2+width),]) 
+  coords3 <- coords3[2:1,1:3]
+  # Define the shore normal transect bearing
+  heading <- earth.bear(coords3[1,2], coords3[1,3], coords3[2,2], coords3[2,3]) + 90
+  if(heading >= 360){
+    heading <- heading-360
+  } else {
+    heading <- heading
+  }
+  heading2 <- data.frame(site = site$site, lon = site$lon, lat = site$lat, heading)
+  return(heading2)
+}
+
+# Creating the transects
+site_transects <- data.frame()
+for(i in 1:length(west$site)){
+ site <- west[i,]
+ site_transect <- shore.normal.transect(site, 2)
+ site_transects <- rbind(site_transects, site_transect)
+}
+
+# Manually correcting Sea Point and Kommetjie
+site_transects$heading[5:6] <- 290 
+save(site_transects, file = "Data_P1/site_transects.RData")
+load("Data_P1/site_transects.RData")
+
+# This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
+# It then extracts a lat/ lon point every X kilometres until reaching a specified isobath
+
+transect.pixel.isobath <- function(site, distance = 25000, isobath = 999){ # Only at a distance of 25000km
+  # Extract coordinates
+  coords <- data.frame(lon = site$lon, lat = site$lat)
+  # Find lon/ lats every X metres until the chosen isobath is reached
+  pixels <- data.frame()
+  deep <- 999
+  while(deep > isobath){
+    # Find depth at distance step
+    coords2 <- as.data.frame(destPoint(p = coords, b = site$heading, d = distance))
+    sitesIdx <- knnx.index(sa_bathy[,1:2], as.matrix(coords2), k = 1)
+    bathy2 <- sa_bathy[sitesIdx,]
+    bathy2 <- bathy2[complete.cases(bathy2$depth),]
+    bathy3 <- data.frame(site = site$site, lon = bathy2$lon, lat = bathy2$lat, 
+                         heading = site$heading, depth = bathy2$depth)
+    pixels <- rbind(pixels, bathy3)
+    coords <- coords2
+    deep <- bathy2$depth
+  }
+  pixels2 <- pixels[pixels$depth > -999,] # This may remove the only row of data
+  if(nrow(pixels2) < 1){
+    pixels2 <- data.frame(site, depth = NA)
+  }else{
+    pixels2 <- pixels2
+  }
+  return(pixels2)
+}
+
+# Pixel points
+site_pixels <- data.frame()
+for(i in 1:length(west$site)){
+  site <- site_transects[i,]
+  site_pixel <- transect.pixel.isobath(site, 25000, -900)
+  site_pixels <- rbind(site_pixels, site_pixel)
+}
+
+save(site_pixels, file = "Data_P1/site_pixels.RData")
+load("Data_P1/site_pixels.RData")
+
+# Bounding box
+  # Only one is made in order to know how large the the geom_point() squares should be made to match
+test <- data.frame(xmin = destPoint(p = site_pixels[1,2:3], b = 270, d = 12500)[1],
+                   xmax = destPoint(p = site_pixels[1,2:3], b = 90, d = 12500)[1],
+                   ymin = destPoint(p = site_pixels[1,2:3], b = 180, d = 12500)[2],
+                   ymax = destPoint(p = site_pixels[1,2:3], b = 0, d = 12500)[2])
+```
+
+## Create a visualisation of the sites at the different distances from the coastline
+****Fix: Fill to show the temperature variation at Benguela coast| also fix the site names*
+
+```{r}
+
+load("Data_P1/south_africa_coast.RData")
+load("Data_P1/bathy/sa_bathy.RData")
+names(south_africa_coast)[1] <- "lon"
+
+# Manually divide up coastline
+wc <- south_africa_coast[291:410,]
+
+# Define plotting parameters
+sa_lats <- c(-37, -27); sa_lons <- c(14, 34)
+
+# Plotting with pixel boxes
+plot1 <- ggplot() +
+ geom_contour(data = bathy[bathy$depth >= -250,], aes(x = lon, y = lat, z = depth),
+               colour = "grey20", alpha = 0.7, size = 0.2, binwidth = 200, na.rm = TRUE, show.legend = FALSE) +
+  stat_contour(data = sa_bathy[sa_bathy$depth < -250,], aes(x = lon, y = lat, z = depth, alpha = ..level..),
+               colour = "grey20", size = 0.1, binwidth = 1000, na.rm = TRUE, show.legend = FALSE) +
+  geom_polygon(data = south_africa_coast, aes(x = lon, y = lat, group = group), 
+               size = 0.5, colour = "black", fill = "white") +
+   # geom_text(data = west, aes(x = lon, y = lat, label = site),
+   #          size = 1.8, colour = "black") +
+  # geom_polygon(data = wcBox, aes(x = lon, y = lat, group = group), alpha = 0.20, 
+  #              colour = "bisque1", fill = "bisque1") +
+  geom_point(data = west, aes(x = lon, y = lat), alpha = 0.8, size = 1.6) +
+  geom_text(data = west, aes(x = lon, y = lat, label = site), 
+                     size = 6, box.padding = 2, nudge_x = -0.5, nudge_y = 0.2, 
+                     segment.alpha = 0.4, force = 0.1) +
+  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "black", shape = 0, alpha = 0.8, size = 2.1) +
+  scale_alpha_continuous(breaks = c(-200, -1000, -2000, -3000, -4000, -5000),
+                         guide_legend(title = "depth (m)")) +
+  scale_fill_gradient(low = "dodgerblue4", high = "dodgerblue", na.value = "steelblue4", 
+                      breaks = c(-1000, -2000, -3000, -4000, -5000),
+                      guide_legend(title = "depth (m)")) +
+   annotate("text", label = "INDIAN\nOCEAN", x = 34.00, y = -35.0,
+             size = 6.0, angle = 0, colour = "black") +
+    annotate("text", label = "ATLANTIC\nOCEAN", x = 14.00, y = -35.0,
+             size = 6.0, angle = 0, colour = "black") +
+    annotate("text", label = "Benguela", x = 15.0, y = -31.7,
+             size = 4.0, angle = 302, colour = "black") +
+    annotate("text", label = "Agulhas", x = 32.0, y = -31.7,
+             size = 4.0, angle = 50, colour = "black") +
+  labs(title = NULL, x = NULL, y = NULL) +
+  theme(legend.key = element_rect(colour = NA, size = 0.2),
+        legend.key.height = unit(0.4, "cm"),
+        legend.background = element_blank()) +
+  coord_cartesian(xlim = sa_lons, ylim = sa_lats) 
+plot1
+```
+
+# Upwelling correlated with wind patterns
+## Loading in the Wind data
+
+```{r}
+wind_1 <- read.delim("Data_P1/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings="") 
+colnames(wind_1) <- c("station_number", "station_name", "date", "hour", "sub", "speed", "dir")
+
+wind_2 <- read.delim("Data_P1/Wind_data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings="")
+colnames(wind_2) <- c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir")
+
+wind_3 <- read.delim("Data_P1/Wind_data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings="")
+colnames(wind_3) <- c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir")
+
+# Slecting the important columns for each of the datasets
+wind_fix <- function(df){
+wind <- df%>% 
+  select(station_name, date, hour, dir, speed)
+}
+wind_1 <- wind_fix(df = wind_1)
+wind_2 <- wind_fix(df = wind_2)
+wind_3 <- wind_fix(df = wind_3)
+
+## Renaming the sites within the wind datasets to match the name of the sites at which seawater temperature was collected
+## The wind data was obtained from the SAWS and the wind stations used were the closes stations to which temperature was collected
+
+renaming_sites_1 <- function(df) {
+  sites <- df %>%
+    mutate(temp_sites = ifelse(station_name %in% c("CAPE COLUMBINE"), "St Helena Bay",        
+                           ifelse(station_name %in% c("KOINGNAAS"), "Hondeklipbaai",
+                                ifelse(station_name %in% c("PORT NOLLOTH"), "Port Nolloth",
+                                       ifelse(station_name %in% c("LAMBERTSBAAI NORTIER"), "Lamberts Bay",
+                                              ifelse(station_name %in% c("LANGEBAANWEG AWS"), "Saldanha Bay
+","Error"))))))
+  return(sites)
+}
+
+wind_sitesmatched_1 <-  renaming_sites_1(df = wind_1)
+wind_sitesmatched_1 <- wind_sitesmatched_1[-c(121352, 121353, 379892, 379893, 609324, 609325, 843506, 843507, 1014585), ]
+
+renaming_sites_2 <- function(df) {
+  sites <- df %>%
+    mutate(temp_sites = ifelse(station_name %in% c("ROBBENEILAND"), "Koeberg Basin",        
+                           ifelse(station_name %in% c("GEELBEK"), "Yzerfontein",
+                                ifelse(station_name %in% c("DASSEN ISLAND"), "Dassen Island",
+                                       ifelse(station_name %in% c("ATLANTIS"), "Koeberg Basin","Error")))))
+  return(sites)
+}
+
+wind_sitesmatched_2 <-  renaming_sites_2(df = wind_2)
+wind_sitesmatched_2 <- wind_sitesmatched_2[-c(228372, 228373, 313441, 313442, 364881, 364882, 557392), ]
+
+renaming_sites_3 <- function(df) {
+  sites <- df %>%
+    mutate(temp_sites = ifelse(station_name %in% c("CAPE TOWN SLANGKOP"), "Kommetjie",  
+                           ifelse(station_name %in% c("CAPE TOWN - ROYAL YACHT CLUB"), "Sea Point",
+                           ifelse(station_name %in% c("CAPE TOWN TABLE BAY"), "Sea Point","Error"))))
+  return(sites)
+}
+
+wind_sitesmatched_3 <-  renaming_sites_3(df = wind_3)
+wind_sitesmatched_3 <- wind_sitesmatched_3[-c(119317, 119318, 196551, 196552, 346759), ]
+
+## CAPE TOWN SLANGKOP may be used for Kommetjie and for Houtbay
+wind_3_HoutBay <- wind_3 %>% 
+  filter(station_name == "CAPE TOWN SLANGKOP") %>% 
+  mutate(temp_sites = ifelse(station_name %in% c("CAPE TOWN SLANGKOP"), "Hout Bay","Error"))
+
+wind_data <- rbind(wind_3_HoutBay,wind_sitesmatched_3,wind_sitesmatched_2,wind_sitesmatched_1)
+wind_data <- wind_data %>% 
+  na.omit()
+# save(wind_data, file = "Data_P1/wind_data.RData")
+
+# The wind data
+
+wind_data <- wind_data %>% 
+  as.numeric(dir)
+
+wind_data <- wind_data %>%
+  mutate(date = as.Date(date)) %>%
+  group_by(temp_sites, date) %>%
+  summarise(mean_dir = mean(as.numeric(dir),
+            mean_speed = mean(as.numeric(speed))))
+
+save(wind_data, file = "Data/wind_daily.RData")
+
+# load("Data/wind_daily.RData") # Loading the daily wind data for each of the sites
+```
+
+## Upwelling indeces
+
+```{r}
+
+```
+
+In particular, trends will be observed to be different at ocean and coastal locations and we hypothesize that
+those differences are due to changes in upwelling intensity.
+
+- Detemining the upwelling indeces in R
+- How does upwelling indeces vary at the different distances along the coast
+- Positive indeces display a positive upwelling.

### 2019-04-12 Paper1_results.Rmd
wind+transect

@@
 library(gridExtra)
+library(circular)
 library(fossil)
@@
 
-# K10
+## K10 dataset
 
@@
   na.omit()
+
+wind_data <- wind_data %>% 
+  group_by(date,hour,temp_sites) %>% 
+  mutate(date = as.Date(date)) 
+# write.csv(wind_data, file = "Data_P1/wind_data.csv", row.names = T)
 # save(wind_data, file = "Data_P1/wind_data.RData")
 
-# The wind data
+# LOading the wind data
+library(tidyverse)
+library(circular)
+wind_data <- read_csv("wind_data.csv", col_types = cols(X1 = col_number(), date = col_character(), dir = col_number(), hour = col_number(),  speed = col_number()))
 
-wind_data <- wind_data %>% 
-  as.numeric(dir)
+wind_daily <- wind_data %>% 
+  mutate(date = as.Date(date))%>%
+   group_by(temp_sites, date) %>%
+   summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees")),2),
+             mean_speed = round(mean(speed),2))
+save(wind_daily, file = "data/wind_daily.RData")
 
-wind_data <- wind_data %>%
-  mutate(date = as.Date(date)) %>%
-  group_by(temp_sites, date) %>%
-  summarise(mean_dir = mean(as.numeric(dir),
-            mean_speed = mean(as.numeric(speed))))
-
-save(wind_data, file = "Data/wind_daily.RData")
-
-# load("Data/wind_daily.RData") # Loading the daily wind data for each of the sites
+# Loading in the daily wind data
+load("Data_P1/wind_daily.RData")
 ```

### 2019-04-15 Write_up/Writing/dissertation.Rmd
updated_commit

@@
 All analyses were conducted in R software version 3.4.2 (http://www.r-project.org/).
-The data used within this study and comprehensive script used for data analyses, and production of figures can be found at https://github.com/AmierohAbrahams/MASTERSPROJECT.
+The data used within this study and comprehensive script used for data analyses, and production of figures can be found at https://github.com/AmierohAbrahams/MastersProject.
 
@@
 
-### Chapter 2 Upwelling
+### Chapter 2 Upwelling - Paper 1
 
-##Introduction
+## Introduction - Paper 1
 
@@
 Upwelling is defined as a process whereby cold, nutrient rich waters are pushed to the surface as a result of alongshore equatorial winds.
-The South African coastline, being approximately 3000km in length, is unique in that it is bordered by the Agulhas current along the east coast and the Benguela current, an eastern boundary upwelling system, along the west coast.
+The South African coastline, being approximately 3000km in length, is unique in that it is bordered by the Agulhas current along the east coast an d the Benguela current, an eastern boundary upwelling system, along the west coast.
 These regions display noticeable differences in seawater temperatures in comparison to each other, primarily due to the influences of the neighbouring ocean currents (Bolton et al.
@@
 Over the past three decades, climate change as a result of anthropogenic activities has established global concern and is shown to have negatively affected marine and terrestrial ecosystems.
-The consequences of these changes are also shown to have direct effects on the behaviour, survival and functioning of marine organisms, thus ultimately affecting species trophic levels and integration [@doney].
+The consequences of these changes are also shown to have direct effects on the behaviour and functioning of marine organisms, thus ultimately affecting species trophic levels and integration [@doney].
 According to the Bakun hypothesis, an increase in greenhouse gases, will result in an increase in day time warming and night time cooling, this ultimately causes an increase in temperature gradients and this further results in a strong pressure gradient.
@@
 To assess the changes within the upwelling events, various datasets would be used.
-These are satellite sea surface temperatures (SST), in situ collected seawater temperatures, wind data as well as satellite obtained chlorophyll-a data.
+These are satellite sea surface temperatures (broad scale), in situ collected seawater temperatures (local scale), wind data as well as satellite obtained chlorophyll-a data.
+Having assessed the eastern boundry upwelling system within a context that comprised of broad and local-scale features permitted us to assess how upwelling varied between sites located closer to the upwelling centres and furthur away from the upwelling centre.
+Specifically, we aim to assess whether or not it is possible to identify the upwelling signal throughout the various datasets and here we examine how far offshore the upwelling signal is present in the higher resolution data and how far offshore the SACTN signal extend.
+We hypothesized that (i) wind action may influence the intensity and duration of upwelling; (ii) Remotely sensed SST data will represent a similar event frequency and duration; (iii) Sites located closer to upwelling centres displays distinct variation in the intensity and duration of upwelling events compared to sites furthur away from the upwelling centeres.
 
@@
 In order to compare abiotic variables such as wind action and direction along the South African coastline, large historical datasets for temperature and wind were analysed and accessed.
-
-## Wind data
-
-Wind action was an important variable in this study as it was hypothesised to exhibit a direct
-influence on seawater temperatures (Sinnett and Feddersen 2014) and a direct influence on the intensity and duration of upwelling events; consequently, wind action was investigated for its impact on upwelling at specific sites along the west coast of South Africa.
-Wind data was obtained from the South African Weather Service (SAWS), and were provided at three-hour resolutions.
-Specific wind characteristics were measured, namely, wind direction (dirw) and wind speed (spw).
-A resolution of 200m was modelled at both 7 and 15m isobaths.
 
@@
 
+An obvious previledge exist when making using of in situ data over satellite data as they provide a more accurate representation of the thermal properties closer to the coast where satellite temperatures do not capture temperature properties well.
+Smit et al (2013) have shown that satellite data collected along the South African coastline represent a warm biases reaching as high as 6°C over in situ temperatures within the nearshore environment.
+
+Inorder to create a time series of each of the remotely sensed SST data representing and also comparing the broad-scale temperatures against the in situ time series, shore-normal transects were extended from the in situ sites a long the west coast of South Africa.
+Temperature values from the various satellite datasets were extracted at multiple distances from the coast.
+These distances are 15 x 15 km, 25 X 25 km and 50 X 50 km pixels along the transect (Figure 1).
+The daily sea surface temperatures obtained from the various remotely sensed datasets within each of the transects were initiated inorder to produce offshore, broad-scale time series that could accurately be used as a dataset to be compared against the in situ time series.
+
+## Wind data
+
+Wind action was an important variable in this study as it was hypothesised that wind direction and wind speed may have a direct influence on the intensity and duration of upwelling events; consequently, wind action was investigated for its impact on upwelling at specific sites along the west coast of South Africa.
+Wind data was obtained from the South African Weather Service (SAWS), and were provided at three-hour resolutions.
+Specific wind characteristics were measured, namely, wind direction (dirw) and wind speed (spw).
+
 ## Chlorophyll-a data
@@
 
-In order to compare the impact of wind action on the intensity and duration of upwelling, the appropriate sites along the west coast of South Africa (i.e.
+At a broad scale, the South African coastline exhibits a large variation in seawater temperature and is divided into four bioregions, each with contrasting temperatures [@smit, @mead].
+The western region of the coastline is dominated by the Benguela Current forming an Eastern Boundry Upwelling System (EBUS) [@hutchings].
+Seasonal upwelling is controlled by south-easterly trade winds and as a result reflect a distinct temperature variations representing much lower temperatures within the upwelling cells over a fairly narrow continental shelf found from the Cape Peninsula to Cape Columbine.
+In order to examine upwelling patterns between sites, the appropriate sites along the west coast of South Africa (i.e.
 sites influenced by the Benguela current) were selected.
-Since temperature data were not evenly recorded for each of the sites, Firstly the full dataset was narrowed to only those sites that could be adequately compared.
-To do this, the filter () function in R was used.
-This function selects all the sites with a time series of greater than 10 years.
-Once the sites were minimised, one years worth of data was selected from the SACTN data between the years 2014 – 2015.
-This yield a total of  seven sites along the west coast of South Africa.
-Hereafter, each of the remotely sensed SST datasets obtained were matched with the SACTN dataset.
-This was done for two reasons, to find the exact same points at which temperature was collected for the SACTN sites and secondly, to confirm that the dates and time that temperature were collected correspond accordingly.
+Since temperature data were not evenly recorded for each of the sites, the entire dataset was narrowed to only those sites that could be adequately compared.
+Once the sites were minimised, one years worth of data was selected from the SACTN data between the years 1992 – 2015.
+This yield a total of six sites along the west coast of South Africa.
+Hereafter, each of the remotely sensed SST datasets were matched to the SACTN dataset.
+This was done for two reasons, to find the same points at which temperature was collected for the SACTN sites and secondly, to confirm that the dates and time that temperature were collected correspond accordingly.
+
+## Detecting the intensity and duration of upwelling

### 2019-04-15 Paper1_results.Rmd
updated

@@
 # Loading libraries
-
 First I need to find, install and load various packages. These packages will be available on CRAN and can be accessed and installed in the usual way.
@@
 library(rgeos)
+library(ncdf4)
+library(reshape2)
+library(plyr) # Never load plyr when also loading the tidyverse. It causes a lot of conflicts.
+library(stringr)
+library(doMC); doMC::registerDoMC(cores = 4)
+library(fasttime)
 
@@
 source("functions/earthdist.R")
+source("functions/wind.rose.R")
 ```
@@
 Now we select only the sites occuring along the west coast.
-The purpose of this research task is to assess whether or not the intensity of upwelling varies at va range of distances from the coastline.
-The data used here is obtained from different satellites and thus possess different resolutions.
-Here I attempt to find an overlapping time series for all of the sites along the west coast.
-The sites here vary between upwelling centres and further away from the upwelling centre
+The purpose of this is to assess whether or not the intensity of upwelling varies at a range of distances from the coastline.
+The data used here is obtained from different satellites and thus represent different resolutions.
+I then find an overlapping time series for each of the sites along the west coast.
+The length of time series for each of the sites is greater than 10 years.
+The sites along the west coast vary between upwelling centres and further away from the upwelling centres.
+As a result it is hypothesised that sites located within the upwelling centre may resemble a longer more intense upwelling event when compared to sites located further away from the upwelling centre.
+For this analyses a long term time series of 30years is not required and so our time series starts in 1992 - 2015.
 
@@
 site_list_sub <- site_list_sub[c(-2, -3, -5, -6, -7,-9, -12, -13, -14),] # Here I keep all the sites with temperature obtained up to 2017
-save(site_list_sub, file = "Data_P1/site_list_sub.Rdata")
+# save(site_list_sub, file = "Data_P1/site_list_sub.Rdata")
 
@@
 
-In this thesis, different remotely-sensed SST datasets are used. Hence, these datasets represent different resolutions at which data was obtained.
+# Remotely sensed SST datasets
+Matching the satellite sea surface temperature data point to the in situ temperature collected data points.
 
@@
 # save(insitu_MUR, file = "Data_P1/insitu_MUR.RData")
+
 MUR_plot <- match_plot(df = insitu_MUR)
@@
 # save(insitu_G1SST, file = "Data_P1/insitu_G1SST.RData")
+
 G1SST_plot <- match_plot(df = insitu_G1SST )
@@
 # save(insitu_K10, file = "Data_P1/insitu_K10.RData")
+
 K10_plot <- match_plot(df = insitu_K10)
@@
 
-## Here I find the nearest SST pixels
-
+## The nearest SST pixels
 Now we apply the FNN (Fast Nearest Neighbor) package to determine the nearesr SST pixel to the insitu collected sites.
@@
 # save(insitu_OISST, file = "Data_P1/insitu_OISST.RData")
+
 OISST_plot <- match_plot(df = insitu_OISST)
@@
 
-## Wavelets
-
-Wavelet anlyses showing the temperature variation between sites along the west coast for the years 2014 and 2015.
-Thereafter, wavelet analyses ae created per season for each individual site along the west coast of South Africa.
-Working with remotely-sensed SST and *in-situ* collected coastal seawater temperature, in the analyses bellow I aim to show how the intensity and duration of upwelling events vary between sites along the west coast of South Africa.
-It is important to remember that each of the datasets used within this study obtained SSTs at different resolutions.
-
-### Load site list SACTN data
-Now to get to the data.
-The first step involves the loading of the site list.
-The statistical properties of the seawater temperature representing the South African coastline, such as the mean, minimum and maximum temperatures.
-These values vary among coastal sections due to the influence of the cold Benguala and warm Agulhas currents.
-Here we will only focus on the temperature data found along the west coast (wc) (i.e.
-sites influenced by the Benguela current, EBUS).
-
-```{r}
-load("Data_P1/site_list_v4.2.RData")
-load("Data_P1/SACTN_daily_v4.2.RData")
-```
-
-## Load in the rest of the datasets
+Loading in the matched satellite temperature data
 
@@
 
-# Time series of one year
+## Wavelets
+Wavelet anlyses showing the temperature variation between sites along the west coast for the years 2010 - 2014.
+Working with remotely-sensed SST and *in-situ* collected coastal seawater temperature, in the analyses bellow I aim to show how the intensity and duration of upwelling events vary between sites along the west coast of South Africa.
+It is important to remember that each of the datasets used within this study obtained SSTs at different resolutions.
 
-Examining the intensity and duration of upwelling for the years 2014 - 2015
+## Load in the datasets
+
+```{r}
+load("Data_P1/SACTN_overlap.RData")
+load("Data_P1/insitu_MUR.RData")
+load("Data_P1/insitu_OISST.RData")
+load("Data_P1/insitu_G1SST.RData")
+load("Data_P1/insitu_K10.RData")
+```
+
+# Time series
+
+Creating a time series from the year 2010 to 2014 for both the satellite and in situ collected SST data.
+This is done to provide an accurate comparison between datasets given that the different datasets have different starting and ending dates for which temperature was collected.
+E.g.
+Temperature for the MUR  only started in 2002
 
@@
 
-############################################################################################################################################
+#########################################################################################################################
 
 # Calculating a shore transect
+
 The aim of this is to find the temperature collected at different distances from the coast (distances = 15km, 25km, 35km and 50km).
 Then create wavelet analyses (or another analyses) where I compare how the intensity and duration of upwelling vary at different distances from the coastline.
-So would I have a dataset with temperature at 15km then at 25km then at 35km and then at 50km and so on...?
+So would I have a dataset with temperature at 15km then at 25km then at 50km and so on...?
 
@@
   coords <- data.frame(lon = site$lon, lat = site$lat)
-  # Find lon/ lats every X metres until the chosen isobath is reached
+  # Find lon/ lats every X metres 
   pixels <- data.frame()
@@
   }
-  pixels2 <- pixels[pixels$depth > -999,] # This may remove the only row of data
+  pixels2 <- pixels[pixels$depth > -999,] 
   if(nrow(pixels2) < 1){
@@
 
-# Upwelling correlated with wind patterns
-## Loading in the Wind data
+# Wind data
+
+Loading the Wind data.
+The wind dataset was supplied by the SAWS in .text format.
+Wind data was provided for the sites closes to the requested site as wind data may not be available for the 6 study sites in this research.
+The wind direction was collected every 2 hours and by using the circular function we calcuated the daily wind direction and mean speed.
+There is a numerically large gap between 360 and 2 where as in degrees its not as large.
+The circular mean function returns the mean direction of a vector of circular data.
+https://cran.r-project.org/web/packages/circular/circular.pdf (Pg 130).
+The circular function creates circular objects around the wind direction.
 
@@
 wind <- df%>% 
-  select(station_name, date, hour, dir, speed)
+  select(station_name, date, hour, dir, speed) #column names
 }
@@
   mutate(date = as.Date(date))
+
 # write.csv(wind_data, file = "Data_P1/wind_data.csv", row.names = T)
@@
              mean_speed = round(mean(speed),2))
-save(wind_daily, file = "data/wind_daily.RData")
 
+# save(wind_daily, file = "Data_P1/wind_daily.RData")
 # Loading in the daily wind data
 load("Data_P1/wind_daily.RData")
+
+# Subsetting to particular sites 
+selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay", "Yzerfontein", "Kommetjie")
+
+wind_daily_sub <- wind_daily %>% 
+  dplyr::rename(site = temp_sites) %>% 
+  filter(site %in% selected_sites) # ***Does not detect Saldanha Bay???****
+
+# save(wind_daily_sub, file = "Data_P1/wind_daily_sub.RData")
+# Loading in the daily wind data
+# load("Data_P1/wind_daily_sub.RData")
+```
+
+Matching wind data to the in situ collected data.
+This is done in order to plot the wind patterns for the same date at which temperature was collected.
+The satellite obtained SST data is also matched by date to the in situ temperature data and so the dates correspond.
+
+```{r}
+insitu_wind <- match_func(df = wind_daily_sub) %>%
+  drop_na()
+insitu_wind <- insitu_wind %>% 
+  filter(year(date) %in% seq(2010, 2014)) # The years 2010-2014 is the years of interest and the years for which wavelets are/will be created.
+
+# save(insitu_wind, file = "Data_P1/insitu_wind.RData")
+# Loading in the daily wind data
+# load("Data_P1/insitu_wind.RData")
+```
+
+# Plotting the wind data - Waverose diagrams
+
+```{r}
+wave_daily_renamed <- insitu_wind %>% 
+  dplyr::rename(spd = mean_speed) %>%
+  dplyr::rename(dir = dir_circ)
+
+p.wr2 <- plot.windrose(data = wave_daily_renamed,
+              spd = "spd",
+              dir = "dir")
+p.wr3 <- p.wr2 + facet_wrap(.~ site, ncol = 2, nrow = "") +
+  theme(strip.text.x = element_text(size = 25))
+p.wr3
+```
+
+## Chlorophyll-a data
+
+Extract the chlorophyll a netCDF files and convert it to CSV.
+The chlorophylla data were obtained from MODIS Aqua.
+The aim of working with these chlorophyll a data is to examine how the chlorophyll a concentration varies with upwelling.
+This function convert chlorophyll netCDF datasets to a single CSV dataset.
+
+```{r}
+# MODIS_chlor.dir <- "/home/amieroh/Documents/Data/Datasets/Chlorophyll_a"
+# MODIS_chlor.csv.dir <- "/home/amieroh/Documents/Data/Datasets"
+
+region <- "BC" # Benguela Current
+coords <- bbox[, region]
+coords <- c(-35, -20, 10, 20) # this is the BC
+
+ncList <- list.files(path = MODIS_chlor.dir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+strt.date <- str_sub(basename(ncList[1]), start = 2, end = 8)
+end.date <- str_sub(basename(ncList[length(ncList)]), start = 2, end = 8)
+nc.init <- nc_open(ncList[1])
+LatIdx <- which(nc.init$dim$lat$vals > coords[1] & nc.init$dim$lat$vals < coords[2])
+LonIdx <- which(nc.init$dim$lon$vals > coords[3] & nc.init$dim$lon$vals < coords[4])
+nc_close(nc.init)
+
+ncFun <- function(nc.file = nc.files, csv.dir = csv.dir) {
+nc <- nc_open(nc.file)
+  instrument <- ncatt_get(nc, 0, "instrument")$value
+  platform <- ncatt_get(nc, 0, "platform")$value
+  product_name <- ncatt_get(nc, 0, "product_name")$value
+  fNameStem <- substr(product_name, 17, 38)
+  timeStamp <- substr(product_name, 2, 8)
+  origin <- paste0(substr(timeStamp, 1, 4), "-01-01")
+  date <- as.Date(as.numeric(substr(timeStamp, 5, 7)), origin)
+  chl <- round(ncvar_get(nc,
+                   varid = "chlor_a",
+                   start = c(LonIdx[1], LatIdx[1]),
+                   count = c(length(LonIdx), length(LatIdx))),
+               3)
+  dimnames(chl) <- list(lon = nc$dim$lon$vals[LonIdx],
+                        lat =  nc$dim$lat$vals[LatIdx])
+  nc_close(nc)
+  chl <-
+    as.data.table(melt(chl, value.name = "chl"), row.names = NULL) %>%
+    mutate(t = date) %>%
+    na.omit()
+  fwrite(chl,
+         file = paste(csv.dir, "/", region, "-", instrument, ".",platform, ".",
+                      fNameStem, "-", strt.date, "-", end.date, ".csv", sep = ""),
+         append = TRUE, col.names = FALSE)
+  rm(chl)
+}
+
+llply(ncList, ncFun, csv.dir = MODIS_chlor.csv.dir, .parallel = TRUE)
+
+# MODIS_Chloro <- "~/Documents/Masters_2019/MastersProject/Data"
+# MODIS_Chloro <- fread(paste0(MODIS_Chloro, "/BC-MODIS.Aqua.L3m_8D_CHL_chlor_a_9km-2002185-2018345.csv"),
+#             col.names = c("lon", "lat", "chloro", "date"))
+
+# save(MODIS_Chloro, file = "Data/MODIS_Chloro.RData")
+```
+
+## Setting up the chlorophylla dataset
+
+```{r}
+
+# Loading the data
+load("Data_P1/MODIS_Chloro.RData") # Extracted chlorophyll data
+load("Data_P1/site_list_sub.Rdata") # Site list of all the sites along the west coast
+
+# Visualisation
+chloro_plot <- MODIS_Chloro %>%
+  filter(date == min(date)) %>%
+  ggplot(aes(x = lon, y = lat)) +
+  geom_raster(aes(fill = chloro))
+
+unique_pixel <- MODIS_Chloro %>% 
+  select(lon, lat) %>% 
+  unique()
+
+match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
+                          query = as.matrix(site_list_sub[,5:6]), k = 1)
+
+# Matching the chlorophylla-a lats and lons to the study sites
+pixel_match <- unique_pixel[match_index,] %>%
+  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
+  mutate(site = site_list_sub$site)
+
+chloro_match <- MODIS_Chloro %>%
+  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
+  filter(combi %in% pixel_match$combi)
+
+chloro_match %>%
+  filter(date == max(date)) %>%
+  ggplot(aes(x = lon, y = lat)) +
+  geom_raster(aes(fill = chloro))
+
+chloro_sites <- chloro_match %>%
+  left_join(pixel_match, by = c("combi", "lon", "lat")) %>% 
+  dplyr::mutate(date = as.Date(date))
+# save(chloro_sites, file = "Data_P1/chloro_sites.RData")
+
+# Visualisation
+Chloro_plot_complete <- ggplot(chloro_sites, aes(x = date, y = chloro)) +
+  geom_line() +
+  facet_wrap(~site, ncol = 1)
+Chloro_plot_complete
+
+# Study period from 2010 - 2014
+chloro_data <- chloro_sites %>% 
+  filter(year(date) %in% seq(2010, 2014))
+# save(chloro_data, file = "Data_P1/chloro_data.RData")
+
+# Plotting only years 2014-2015
+Chloro_plot_filtered <- ggplot(chloro_data, aes(x = date, y = chloro)) +
+  geom_line() +
+  facet_wrap(~site, ncol = 1)
+Chloro_plot_filtered
+
+# Creating daily temperatures
+load("Data_P1/chloro_data.RData") # Loading in the chlorophyll-a data
+
+daily_chloro_data <- chloro_data %>%
+  dplyr::mutate(date = as.Date(date)) %>%
+  dplyr::group_by(combi, lon, lat, date) %>%
+  dplyr::summarise(chloro = mean(chloro, na.rm = TRUE)) %>%
+  dplyr::group_by(combi, lon, lat, date) %>%
+  tidyr::nest() %>%
+  dplyr::mutate(clims = purrr::map(data, ts2clm, x = date, y = chloro,
+                     climatologyPeriod = c("2002-07-13", "2018-12-04"),
+                     maxPadLength = 8)) %>% # Adjust maxPadLength = X for longer interpolation
+  dplyr::select(-data) %>%
+  tidyr::unnest() %>%
+  select(-doy, -seas, -thresh)
 ```
@@
 - Positive indeces display a positive upwelling.
+
+Things to do
+
+- Plot seasonal detrend
+- chlorophyll a data
+- fix windrose plot
+- Stastical analyses

### 2019-04-15 Paper1_results.Rmd
Fixed

@@
 # Loading the data
-load("Data_P1/MODIS_Chloro.RData") # Extracted chlorophyll data
+load("Data_P1/MODIS_Chloro.RData") # Extracted/Raw chlorophyll data
 load("Data_P1/site_list_sub.Rdata") # Site list of all the sites along the west coast
@@
 
-## Upwelling indeces
+## EOF's
 

### 2019-04-15 Paper1_results.Rmd
updated

@@
 library(circular)
-wind_data <- read_csv("wind_data.csv", col_types = cols(X1 = col_number(), date = col_character(), dir = col_number(), hour = col_number(),  speed = col_number()))
 
-wind_daily <- wind_data %>% 
-  mutate(date = as.Date(date))%>%
-   group_by(temp_sites, date) %>%
-   summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees")),2),
-             mean_speed = round(mean(speed),2))
+# wind_data <- read_csv("wind_data.csv", col_types = cols(X1 = col_number(), date = col_character(), dir = col_number(), hour = col_number(),  speed = col_number())) # Dataset too large for Github but better data at line 638
+# 
+# wind_daily <- wind_data %>% 
+#   mutate(date = as.Date(date))%>%
+#    group_by(temp_sites, date) %>%
+#    summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees")),2),
+#              mean_speed = round(mean(speed),2))
 

### 2019-04-16 Paper1_results.Rmd
Marked off the changes I made with RWS

--- 2019-04-15 Paper1_results.Rmd
@@
 
-load("Data_P1/site_list_sub.RData") # Loading the SACTN_overlap dataset which allows me to match it to the OISST data
+load("Data_P1/site_list_sub.Rdata") # Loading the SACTN_overlap dataset which allows me to match it to the OISST data
 
@@
 
-PN_wave <- dlply(PN_prewhite, .(site), wl.fun)
+PN_wave <- dlply(PN_prewhite, .(site), wl.fun, .parallel = T) # RWS: Allow parallel processing
 
@@
 }
-ldply(PN_wave, plot_fun) 
+ldply(PN_wave, plot_fun, .parallel = T) # RWS: Allow parallel processing
 ```
@@
 
-transect.pixel.isobath <- function(site, distance = 25000, isobath = 999){ # Only at a distance of 25000km
+transect.pixel <- function(site, distances){
   # Extract coordinates
@@
   pixels <- data.frame()
-  deep <- 999
-  while(deep > isobath){
+  # deep <- 999
+  # distance_multiplier <- 1
+  # while(deep > isobath){
+  for(i in 1:length(distances)){
+    # Add index to show distance from coast
     # Find depth at distance step
-    coords2 <- as.data.frame(destPoint(p = coords, b = site$heading, d = distance))
+    coords2 <- as.data.frame(destPoint(p = coords, b = site$heading, d = distances[i]))
     sitesIdx <- knnx.index(sa_bathy[,1:2], as.matrix(coords2), k = 1)
@@
     bathy3 <- data.frame(site = site$site, lon = bathy2$lon, lat = bathy2$lat, 
-                         heading = site$heading, depth = bathy2$depth)
+                         heading = site$heading, 
+                         # depth = bathy2$depth,
+                         distance = distances[i])
     pixels <- rbind(pixels, bathy3)
     coords <- coords2
-    deep <- bathy2$depth
+    # distance_multiplier <- distance_multiplier +1 
+    # deep <- bathy2$depth
   }
-  pixels2 <- pixels[pixels$depth > -999,] 
-  if(nrow(pixels2) < 1){
-    pixels2 <- data.frame(site, depth = NA)
+  # pixels2 <- pixels[pixels$depth > -999,] 
+  if(nrow(pixels) < 1){
+    pixels <- data.frame(site, depth = NA)
   }else{
-    pixels2 <- pixels2
+    pixels <- pixels
   }
-  return(pixels2)
+  return(pixels)
 }
@@
   site <- site_transects[i,]
-  site_pixel <- transect.pixel.isobath(site, 25000, -900)
+  site_pixel <- transect.pixel(site, c(15000, 25000, 35000, 45000, 55000)) # RWS: The desired distances
   site_pixels <- rbind(site_pixels, site_pixel)
@@
 
-save(site_pixels, file = "Data_P1/site_pixels.RData")
+# save(site_pixels, file = "Data_P1/site_pixels.RData")
 load("Data_P1/site_pixels.RData")
@@
 ```{r}
-wind_1 <- read.delim("Data_P1/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings="") 
-colnames(wind_1) <- c("station_number", "station_name", "date", "hour", "sub", "speed", "dir")
+wind_1 <- read.delim("Data_P1/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings = "", 
+                     col.names = c("station_number", "station_name", "date", "hour", "sub", "speed", "dir"))
 
-wind_2 <- read.delim("Data_P1/Wind_data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings="")
-colnames(wind_2) <- c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir")
+wind_2 <- read.delim("Data_P1/Wind_data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings = "",
+                     col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
 
-wind_3 <- read.delim("Data_P1/Wind_data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings="")
-colnames(wind_3) <- c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir")
+wind_3 <- read.delim("Data_P1/Wind_data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings = "",
+                     col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
 
@@
 wind_fix <- function(df){
-wind <- df%>% 
-  select(station_name, date, hour, dir, speed) #column names
+wind <- df %>% 
+  select(station_name, date, hour, dir, speed) %>%  #column names
+  mutate(date = as.Date(as.character(date)),
+         hour = as.numeric(as.character(hour)), 
+         dir = as.numeric(as.character(dir)),
+         speed = as.numeric(as.character(speed)))
 }
+# RWS: We can see when we force the values to be numeric that there are some non-numeric values in the base data
 wind_1 <- wind_fix(df = wind_1)
@@
   sites <- df %>%
-    mutate(temp_sites = ifelse(station_name %in% c("CAPE COLUMBINE"), "St Helena Bay",        
-                           ifelse(station_name %in% c("KOINGNAAS"), "Hondeklipbaai",
-                                ifelse(station_name %in% c("PORT NOLLOTH"), "Port Nolloth",
-                                       ifelse(station_name %in% c("LAMBERTSBAAI NORTIER"), "Lamberts Bay",
-                                              ifelse(station_name %in% c("LANGEBAANWEG AWS"), "Saldanha Bay
-","Error"))))))
+      # RWS: An alternative way to replace values without having to use multiple ifelse() statements
+      mutate(temp_sites = case_when(station_name == "CAPE COLUMBINE" ~ "St Helena Bay",
+                                    station_name  == "KOINGNAAS" ~ "Hondeklipbaai",
+                                    station_name  == "PORT NOLLOTH" ~"Port Nolloth",
+                                    station_name  == "LAMBERTSBAAI NORTIER" ~ "Lamberts Bay",
+                                    station_name  == "LANGEBAANWEG AWS" ~ "Saldanha Bay"))
   return(sites)
@@
 wind_sitesmatched_1 <-  renaming_sites_1(df = wind_1)
+# RWS: You should never be removing values by specific call like this
+  # Rather you should be able to use some sort of conditional to screen out unwanted values
 wind_sitesmatched_1 <- wind_sitesmatched_1[-c(121352, 121353, 379892, 379893, 609324, 609325, 843506, 843507, 1014585), ]
@@
 wind_data <- wind_data %>% 
-  group_by(date,hour,temp_sites) %>% 
-  mutate(date = as.Date(date))
+  # mutate(date = as.Date(date)) %>%  
+  group_by(date, hour, temp_sites)
 
@@
 
-# LOading the wind data
+# Loading the wind data
 library(tidyverse)
 library(circular)
-wind_data <- read_csv("wind_data.csv", col_types = cols(X1 = col_number(), date = col_character(), dir = col_number(), hour = col_number(),  speed = col_number()))
+
+# wind_data <- read_csv("wind_data.csv", col_types = cols(X1 = col_number(), date = col_character(), dir = col_number(), hour = col_number(),  speed = col_number())) # Dataset too large for Github but better data at line 638
+
+# wind_data <- read_csv("Data_P1/wind_data.csv")
 
 wind_daily <- wind_data %>%
-  mutate(date = as.Date(date))%>%
-   group_by(temp_sites, date) %>%
-   summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees")),2),
+  # ungroup() %>%
+  # mutate(date = as.Date(date))%>%
+  dplyr::group_by(temp_sites, date) %>%
+  dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees")),2),
                    mean_speed = round(mean(speed),2))
@@
   dplyr::rename(site = temp_sites) %>% 
-  filter(site %in% selected_sites) # ***Does not detect Saldanha Bay???****
+  filter(site %in% selected_sites)
 
@@
   theme(strip.text.x = element_text(size = 25))
-p.wr3
+p.wr3 # RWS: You'll want to remove the NA part in the figure output
 ```
@@
 region <- "BC" # Benguela Current
-coords <- bbox[, region]
+# coords <- bbox[, region] # RWS: Where does bbox come from?
 coords <- c(-35, -20, 10, 20) # this is the BC
@@
 daily_chloro_data <- chloro_data %>%
-  dplyr::mutate(date = as.Date(date)) %>%
+  # dplyr::mutate(date = as.Date(date)) %>% # RWS: The values already are dates
   dplyr::group_by(combi, lon, lat, date) %>%
   dplyr::summarise(chloro = mean(chloro, na.rm = TRUE)) %>%
-  dplyr::group_by(combi, lon, lat, date) %>%
+  dplyr::group_by(combi, lon, lat) %>% # RWS: Deleted the date grouping variable being used here
   tidyr::nest() %>%
   dplyr::mutate(clims = purrr::map(data, ts2clm, x = date, y = chloro,
-                     climatologyPeriod = c("2002-07-13", "2018-12-04"),
+                     climatologyPeriod = c("2010-02-11", "2014-12-20"),
                      maxPadLength = 8)) %>% # Adjust maxPadLength = X for longer interpolation

### 2019-04-16 Paper1_results.Rmd
edited

--- 2019-04-15 Paper1_results.Rmd
@@
 library(ncdf4)
+# library(sinkr)
+library(spacetime)
 library(reshape2)
@@
 library(fasttime)
+
+# library(devtools)
+# install_github("marchtaylor/sinkr")
 

### 2019-04-16 Write_up/Papers/15_April_2019/upwelling_index comp and temp.r
Paper commit

@@ -1,0 +1,151 @@
+#Script to plot upwelling indices based on Widguru and SAWS wind data
+#upwelling index from Fielding and Davis 1989 paper
+#Updated: April 2019
+
+
+#upwelling index data visualisation
+
+setwd("~/Research/Data & Analyses/Upwelling Index")
+
+library(tidyverse)
+library(gridExtra)
+
+UI <- read.csv("upwelling.csv")
+
+#upwelling index calculated for data from one time (8am) each day
+
+head(UI) #just added this so that you can see what the data looks like
+#Date Day Temp   WG SAWS
+#01-Jan   1   12  7.8  6.3
+#02-Jan   2   12  5.7  3.2
+#03-Jan   3   11 11.4  7.5
+#04-Jan   4   10  8.8  5.4
+#05-Jan   5   10  9.4  5.7
+#06-Jan   6   10 10.0  9.6
+
+
+#add condition variable
+
+UI <- UI %>% mutate(WG.condition = ifelse(WG > 0, "upwelling", "downwelling")) %>% 
+        mutate(SAWS.condition = ifelse(SAWS > 0, "upwelling", "downwelling")) %>% 
+        mutate(ui.diff = WG - SAWS)
+
+#==============================================================================================
+
+#Windguru Data
+
+ggplot(UI, aes(x=Day, y=WG)) +
+  geom_area(aes(fill=WG.condition)) +
+  geom_line() +
+  geom_hline(yintercept=0)
+
+interp <- approx(UI$Day, UI$WG, n=100000)
+interp2 <- approx(UI$Day, UI$Temp, n=100000)
+
+UIi <- data.frame(Day=interp$x, Index=interp$y, Temp = interp2$y)
+UIi$WG_condition[UIi$Index >= 0] <- "upwelling"
+UIi$WG_condition[UIi$Index < 0] <- "downwelling"
+
+index_plot_wg <- ggplot(UIi, aes(x=Day, y=Index)) +
+                  theme_bw() +
+                  geom_area(aes(fill=WG_condition), alpha = .4) +
+                  geom_line() +
+                  geom_hline(yintercept=0) +
+                  annotate("text", x = 300, y = -20, label = "WindGuru Data") +
+                  #geom_vline(xintercept = c(116, 117, 133, 134, 244, 291), lty = 2, lwd = 1) +
+                  scale_fill_manual(values=c("salmon", "steelblue")) +
+                  scale_x_continuous(expand=c(0, 0)) +
[…]

### 2019-04-17 Paper1_results.Rmd
updated

@@
 library(geosphere)
-library(marmap)
+# library(marmap)
 library(PBSmapping)
@@
 library(mapproj)
-library(rgeos)
+# library(rgeos)
 library(ncdf4)
@@
 Matching the satellite sea surface temperature data point to the in situ temperature collected data points.
+The MUR data is not present within the data_P1 folder and so needs to be loaded into the data folder.
 
@@
 ```{r}
-MUR_Lamberts_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Lamberts Bay_SST_timeseries_5nearest.csv")
-MUR_Port_Nolloth <- read_csv("Data/MUR_nearest5pixels/MUR_Port Nolloth_SST_timeseries_5nearest.csv")
-MUR_Saldanha_Bay <- read_csv("Data/MUR_nearest5pixels/MUR_Saldanha Bay_SST_timeseries_5nearest.csv")
-MUR_Yzerfontein <- read_csv("Data/MUR_nearest5pixels/MUR_Yzerfontein_SST_timeseries_5nearest.csv")
-MUR_Kommetjie <- read_csv("Data/MUR_nearest5pixels/MUR_Kommetjie_SST_timeseries_5nearest.csv")
-MUR_Sea_Point <- read_csv("Data/MUR_nearest5pixels/MUR_Sea Point_SST_timeseries_5nearest.csv")
+MUR_Lamberts_Bay <- read_csv("Data_P1/MUR_nearest5pixels/MUR_Lamberts Bay_SST_timeseries_5nearest.csv")
+MUR_Port_Nolloth <- read_csv("Data_P1/MUR_nearest5pixels/MUR_Port Nolloth_SST_timeseries_5nearest.csv")
+MUR_Saldanha_Bay <- read_csv("Data_P1/MUR_nearest5pixels/MUR_Saldanha Bay_SST_timeseries_5nearest.csv")
+MUR_Yzerfontein <- read_csv("Data_P1/MUR_nearest5pixels/MUR_Yzerfontein_SST_timeseries_5nearest.csv")
+MUR_Kommetjie <- read_csv("Data_P1/MUR_nearest5pixels/MUR_Kommetjie_SST_timeseries_5nearest.csv")
+MUR_Sea_Point <- read_csv("Data_P1/MUR_nearest5pixels/MUR_Sea Point_SST_timeseries_5nearest.csv")
 
@@
 MUR_plot <- match_plot(df = insitu_MUR)
-MUR_plot <- temp_plot(df = insitu_MUR)
+MUR_temp <- temp_plot(df = insitu_MUR)
 ```
@@
 ```{r}
-Lamberts_Bay <- read_csv("Data/G1SST_sub/Lamberts Bay_SST_timeseries_5nearest.csv")
-Port_Nolloth <- read_csv("Data/G1SST_sub/Port Nolloth_SST_timeseries_5nearest.csv")
-Saldanha_Bay <- read_csv("Data/G1SST_sub/Saldanha Bay_SST_timeseries_5nearest.csv")
-Yzerfontein <- read_csv("Data/G1SST_sub/Yzerfontein_SST_timeseries_5nearest.csv")
-Kommetjie_SST <- read_csv("Data/G1SST_sub/Kommetjie_SST_timeseries_5nearest.csv")
-Sea_Point_SST<- read_csv("Data/G1SST_sub/Sea Point_SST_timeseries_5nearest.csv")
+Lamberts_Bay <- read_csv("Data_P1/G1SST_sub/Lamberts Bay_SST_timeseries_5nearest.csv")
+Port_Nolloth <- read_csv("Data_P1/G1SST_sub/Port Nolloth_SST_timeseries_5nearest.csv")
+Saldanha_Bay <- read_csv("Data_P1/G1SST_sub/Saldanha Bay_SST_timeseries_5nearest.csv")
+Yzerfontein <- read_csv("Data_P1/G1SST_sub/Yzerfontein_SST_timeseries_5nearest.csv")
+Kommetjie_SST <- read_csv("Data_P1/G1SST_sub/Kommetjie_SST_timeseries_5nearest.csv")
+Sea_Point_SST<- read_csv("Data_P1/G1SST_sub/Sea Point_SST_timeseries_5nearest.csv")
 
@@
 G1SST_plot <- match_plot(df = insitu_G1SST )
-G1SST_temp_plot <- temp_plot(df = insitu_G1SST)
+G1SST_temp<- temp_plot(df = insitu_G1SST)
 ```
@@
 ```{r}
-Lamberts_Bay <- read_csv("Data/K10/K10_Lamberts Bay_SST_timeseries_5nearest.csv")
-Port_Nolloth <- read_csv("Data/K10/K10_Port Nolloth_SST_timeseries_5nearest.csv")
-Saldanha_Bay <- read_csv("Data/K10/K10_Saldanha Bay_SST_timeseries_5nearest.csv")
-Yzerfontein <- read_csv("Data/K10/K10_Yzerfontein_SST_timeseries_5nearest.csv")
-Kommetjie <- read_csv("Data/K10/K10_Kommetjie_SST_timeseries_5nearest.csv")
-Sea_Point <- read_csv("Data/K10/K10_Sea Point_SST_timeseries_5nearest.csv")
+Lamberts_Bay <- read_csv("Data_P1/K10/K10_Lamberts Bay_SST_timeseries_5nearest.csv")
+Port_Nolloth <- read_csv("Data_P1/K10/K10_Port Nolloth_SST_timeseries_5nearest.csv")
+Saldanha_Bay <- read_csv("Data_P1/K10/K10_Saldanha Bay_SST_timeseries_5nearest.csv")
+Yzerfontein <- read_csv("Data_P1/K10/K10_Yzerfontein_SST_timeseries_5nearest.csv")
+Kommetjie <- read_csv("Data_P1/K10/K10_Kommetjie_SST_timeseries_5nearest.csv")
+Sea_Point <- read_csv("Data_P1/K10/K10_Sea Point_SST_timeseries_5nearest.csv")
 
@@
 K10_plot <- match_plot(df = insitu_K10)
-k10_temp_plot <- temp_plot(df = insitu_K10)
+k10_temp <- temp_plot(df = insitu_K10)
 ```
@@
 OISST_plot <- match_plot(df = insitu_OISST)
-OISST_temp_plot <- temp_plot(df = insitu_OISST)
+OISST_temp <- temp_plot(df = insitu_OISST)
 ```
@@
 The aim of this is to find the temperature collected at different distances from the coast (distances = 15km, 25km, 35km and 50km).
-Then create wavelet analyses (or another analyses) where I compare how the intensity and duration of upwelling vary at different distances from the coastline.
-So would I have a dataset with temperature at 15km then at 25km then at 50km and so on...?
 
@@
 site_transects$heading[5:6] <- 290 
-save(site_transects, file = "Data_P1/site_transects.RData")
+# save(site_transects, file = "Data_P1/site_transects.RData")
 load("Data_P1/site_transects.RData")
@@
   site <- site_transects[i,]
-  site_pixel <- transect.pixel(site, c(15000, 25000, 35000, 45000, 55000)) # RWS: The desired distances
+  site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: The desired distances
   site_pixels <- rbind(site_pixels, site_pixel)
@@
   # Only one is made in order to know how large the the geom_point() squares should be made to match
-test <- data.frame(xmin = destPoint(p = site_pixels[1,2:3], b = 270, d = 12500)[1],
+bbox <- data.frame(xmin = destPoint(p = site_pixels[1,2:3], b = 270, d = 12500)[1],
                    xmax = destPoint(p = site_pixels[1,2:3], b = 90, d = 12500)[1],
@@
                    ymax = destPoint(p = site_pixels[1,2:3], b = 0, d = 12500)[2])
+
+# Determining the temperature at the various distances from the coast
 ```
@@
 load("Data_P1/bathy/sa_bathy.RData")
+load("Data_P1/bathy/bathy.RData")
 names(south_africa_coast)[1] <- "lon"
@@
 
-# Plotting with pixel boxes
-plot1 <- ggplot() +
- geom_contour(data = bathy[bathy$depth >= -250,], aes(x = lon, y = lat, z = depth),
-               colour = "grey20", alpha = 0.7, size = 0.2, binwidth = 200, na.rm = TRUE, show.legend = FALSE) +
-  stat_contour(data = sa_bathy[sa_bathy$depth < -250,], aes(x = lon, y = lat, z = depth, alpha = ..level..),
-               colour = "grey20", size = 0.1, binwidth = 1000, na.rm = TRUE, show.legend = FALSE) +
+ plot1 <- ggplot() +#coord_equal() + 
+  geom_raster(data = sa_bathy, aes(x = lon, y = lat, fill = depth)) +
+  #geom_raster(data = isobath200, aes(x = lon, y = lat, fill = coast)) +
+  stat_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth, alpha = ..level..), 
+               colour = "black", size = 0.2, binwidth = 200, na.rm = TRUE, show_guide = FALSE) +
   geom_polygon(data = south_africa_coast, aes(x = lon, y = lat, group = group), 
                size = 0.5, colour = "black", fill = "white") +
-   # geom_text(data = west, aes(x = lon, y = lat, label = site),
-   #          size = 1.8, colour = "black") +
+  #geom_path(data = sa_provinces_new, aes(x = long, y = lat, group = group)) +
   # geom_polygon(data = wcBox, aes(x = lon, y = lat, group = group), alpha = 0.20, 
   #              colour = "bisque1", fill = "bisque1") +
-  geom_point(data = west, aes(x = lon, y = lat), alpha = 0.8, size = 1.6) +
-  geom_text(data = west, aes(x = lon, y = lat, label = site), 
-                     size = 6, box.padding = 2, nudge_x = -0.5, nudge_y = 0.2, 
-                     segment.alpha = 0.4, force = 0.1) +
-  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "black", shape = 0, alpha = 0.8, size = 2.1) +
+  geom_point(data = site_list_sub, aes(x = lon, y = lat), alpha = 0.8, size = 1.6) +
+  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "white", shape = 0, alpha = 0.8, size = 2.1) +
+  geom_rect(data = bbox, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), 
+            alpha = 0.4, colour = "red", size = 0.1, linetype = 1) +
   scale_alpha_continuous(breaks = c(-200, -1000, -2000, -3000, -4000, -5000),
@@
                       guide_legend(title = "depth (m)")) +
-   annotate("text", label = "INDIAN\nOCEAN", x = 34.00, y = -35.0,
-             size = 6.0, angle = 0, colour = "black") +
-    annotate("text", label = "ATLANTIC\nOCEAN", x = 14.00, y = -35.0,
-             size = 6.0, angle = 0, colour = "black") +
-    annotate("text", label = "Benguela", x = 15.0, y = -31.7,
-             size = 4.0, angle = 302, colour = "black") +
-    annotate("text", label = "Agulhas", x = 32.0, y = -31.7,
-             size = 4.0, angle = 50, colour = "black") +
   labs(title = NULL, x = NULL, y = NULL) +
+  guides(colour = guide_legend(override.aes = list(size = 3))) +
+  #guide_legend(ncol = 2) +
   theme(legend.key = element_rect(colour = NA, size = 0.2),
@@
         legend.background = element_blank()) +
+        #legend.justification = c(1,0), legend.position = c(0.5, 0.4)) +
   coord_cartesian(xlim = sa_lons, ylim = sa_lats)
  plot1
+
+ # Added in the bathy data but this is not needed for any of the analyses.
+ # This just shows the bounding boxes 
 ```
@@
       # RWS: An alternative way to replace values without having to use multiple ifelse() statements
-      mutate(temp_sites = case_when(station_name == "CAPE COLUMBINE" ~ "St Helena Bay",
-                                    station_name  == "KOINGNAAS" ~ "Hondeklipbaai",
+      mutate(temp_sites = case_when(station_name == "CAPE TOWN TABLE BAY" ~ "Sea Point",
+                                    station_name == "CAPE TOWN - ROYAL YACHT CLUB" ~ "Sea Point",
+                                    station_name  == "GEELBEK" ~ "Yzerfontein",
+                                    station_name  == "CAPE TOWN SLANGKOP" ~ "Kommetjie",
                                     station_name  == "PORT NOLLOTH" ~"Port Nolloth",
@@
 wind_sitesmatched_1 <-  renaming_sites_1(df = wind_1)
-# RWS: You should never be removing values by specific call like this
-  # Rather you should be able to use some sort of conditional to screen out unwanted values
-wind_sitesmatched_1 <- wind_sitesmatched_1[-c(121352, 121353, 379892, 379893, 609324, 609325, 843506, 843507, 1014585), ]
+wind_sitesmatched_2 <-  renaming_sites_1(df = wind_2)
+wind_sitesmatched_3 <-  renaming_sites_1(df = wind_3)
+wind_data <- rbind(wind_sitesmatched_3,wind_sitesmatched_2,wind_sitesmatched_1)
+# save(wind_data, file = "Data_P1/wind_data.RData")
 
-renaming_sites_2 <- function(df) {
-  sites <- df %>%
-    mutate(temp_sites = ifelse(station_name %in% c("ROBBENEILAND"), "Koeberg Basin",        
-                           ifelse(station_name %in% c("GEELBEK"), "Yzerfontein",
-                                ifelse(station_name %in% c("DASSEN ISLAND"), "Dassen Island",
-                                       ifelse(station_name %in% c("ATLANTIS"), "Koeberg Basin","Error")))))
-  return(sites)
-}
-
-wind_sitesmatched_2 <-  renaming_sites_2(df = wind_2)
-wind_sitesmatched_2 <- wind_sitesmatched_2[-c(228372, 228373, 313441, 313442, 364881, 364882, 557392), ]
-
-renaming_sites_3 <- function(df) {
-  sites <- df %>%
-    mutate(temp_sites = ifelse(station_name %in% c("CAPE TOWN SLANGKOP"), "Kommetjie",  
-                           ifelse(station_name %in% c("CAPE TOWN - ROYAL YACHT CLUB"), "Sea Point",
-                           ifelse(station_name %in% c("CAPE TOWN TABLE BAY"), "Sea Point","Error"))))
-  return(sites)
-}
-
-wind_sitesmatched_3 <-  renaming_sites_3(df = wind_3)
-wind_sitesmatched_3 <- wind_sitesmatched_3[-c(119317, 119318, 196551, 196552, 346759), ]
-
-## CAPE TOWN SLANGKOP may be used for Kommetjie and for Houtbay
-wind_3_HoutBay <- wind_3 %>% 
-  filter(station_name == "CAPE TOWN SLANGKOP") %>% 
-  mutate(temp_sites = ifelse(station_name %in% c("CAPE TOWN SLANGKOP"), "Hout Bay","Error"))
-
-wind_data <- rbind(wind_3_HoutBay,wind_sitesmatched_3,wind_sitesmatched_2,wind_sitesmatched_1)
-wind_data <- wind_data %>% 
-  na.omit()
-
-wind_data <- wind_data %>% 
-  # mutate(date = as.Date(date)) %>%  
-  group_by(date, hour, temp_sites)
-
-# write.csv(wind_data, file = "Data_P1/wind_data.csv", row.names = T)
-# save(wind_data, file = "Data_P1/wind_data.RData")
+# # RWS: You should never be removing values by specific call like this
+#   # Rather you should be able to use some sort of conditional to screen out unwanted values
+# wind_sitesmatched_1 <- wind_sitesmatched_1[-c(121352, 121353, 379892, 379893, 609324, 609325, 843506, 843507, 1014585), ]
+# renaming_sites_2 <- function(df) {
+#   sites <- df %>%
+#     mutate(temp_sites = ifelse(station_name %in% c("ROBBENEILAND"), "Koeberg Basin",        
+#                            ifelse(station_name %in% c("GEELBEK"), "Yzerfontein",
+#                                 ifelse(station_name %in% c("DASSEN ISLAND"), "Dassen Island",
+#                                        ifelse(station_name %in% c("ATLANTIS"), "Koeberg Basin","Error")))))
+#   return(sites)
+# }
+# 
+# wind_sitesmatched_2 <-  renaming_sites_2(df = wind_2)
+# wind_sitesmatched_2 <- wind_sitesmatched_2[-c(228372, 228373, 313441, 313442, 364881, 364882, 557392), ] 
+# renaming_sites_3 <- function(df) {
+#   sites <- df %>%
+#     mutate(temp_sites = ifelse(station_name %in% c("CAPE TOWN SLANGKOP"), "Kommetjie",  
+#                            ifelse(station_name %in% c("CAPE TOWN - ROYAL YACHT CLUB"), "Sea Point",
+#                            ifelse(station_name %in% c("CAPE TOWN TABLE BAY"), "Sea Point","Error"))))
+#   return(sites)
+# }
+# 
+# wind_sitesmatched_3 <-  renaming_sites_3(df = wind_3)
+# wind_sitesmatched_3 <- wind_sitesmatched_3[-c(119317, 119318, 196551, 196552, 346759), ] 
+### CAPE TOWN SLANGKOP may be used for Kommetjie and for Houtbay
+# wind_3_HoutBay <- wind_3 %>% 
+#   filter(station_name == "CAPE TOWN SLANGKOP") %>% 
+#   mutate(temp_sites = ifelse(station_name %in% c("CAPE TOWN SLANGKOP"), "Hout Bay","Error"))
+# wind_data <- rbind(wind_3_HoutBay,wind_sitesmatched_3,wind_sitesmatched_2,wind_sitesmatched_1)
+# wind_data <- wind_data %>% 
+#   na.omit()
+# 
+# wind_data <- wind_data %>% 
+#   # mutate(date = as.Date(date)) %>%  
+#   group_by(date, hour, temp_sites) 
+# 
+# # write.csv(wind_data, file = "Data_P1/wind_data.csv", row.names = T)
+# # save(wind_data, file = "Data_P1/wind_data.RData")
 
@@
 # wind_data <- read_csv("Data_P1/wind_data.csv")
+selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay", "Yzerfontein", "Kommetjie")
 
@@
   dplyr::group_by(temp_sites, date) %>%
+  filter(temp_sites %in% selected_sites) %>% 
   dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees")),2),
@@
 # save(wind_daily, file = "Data_P1/wind_daily.RData")
-# Loading in the daily wind data
+# # Loading in the daily wind data
+
+# Loading the subsetted wind data created from the text files above 
 load("Data_P1/wind_daily.RData") 
-
-# Subsetting to particular sites 
-selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay", "Yzerfontein", "Kommetjie")
-
-wind_daily_sub <- wind_daily %>% 
-  dplyr::rename(site = temp_sites) %>% 
-  filter(site %in% selected_sites)
-
-# save(wind_daily_sub, file = "Data_P1/wind_daily_sub.RData")
-# Loading in the daily wind data
-# load("Data_P1/wind_daily_sub.RData")
 ```
@@
 ```{r}
-insitu_wind <- match_func(df = wind_daily_sub) %>%
-  drop_na()
+load("Data_P1/SACTN_overlap.RData")
+load("Data_P1/wind_daily.RData")
+wind_daily <- wind_daily%>% 
+  dplyr::rename(site = temp_sites) # Renaming to match the `match_func` previously created
+
+insitu_wind <- match_func(df = wind_daily)  # Using `match_func`
+
 insitu_wind <- insitu_wind %>% 
@@
 # save(insitu_wind, file = "Data_P1/insitu_wind.RData")
-# Loading in the daily wind data
-# load("Data_P1/insitu_wind.RData")
 ```
@@
 ```{r}
+# Loading in the matched wind data
+load("Data_P1/insitu_wind.RData")
+
 wave_daily_renamed <- insitu_wind %>% 
@@
   dplyr::rename(dir = dir_circ) 
-
+source("functions/wind.rose.R")
 p.wr2 <- plot.windrose(data = wave_daily_renamed,
@@
 
-region <- "BC" # Benguela Current
-# coords <- bbox[, region] # RWS: Where does bbox come from?
-coords <- c(-35, -20, 10, 20) # this is the BC
-
-ncList <- list.files(path = MODIS_chlor.dir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
-strt.date <- str_sub(basename(ncList[1]), start = 2, end = 8)
-end.date <- str_sub(basename(ncList[length(ncList)]), start = 2, end = 8)
-nc.init <- nc_open(ncList[1])
-LatIdx <- which(nc.init$dim$lat$vals > coords[1] & nc.init$dim$lat$vals < coords[2])
-LonIdx <- which(nc.init$dim$lon$vals > coords[3] & nc.init$dim$lon$vals < coords[4])
-nc_close(nc.init)
-
-ncFun <- function(nc.file = nc.files, csv.dir = csv.dir) {
-nc <- nc_open(nc.file)
-  instrument <- ncatt_get(nc, 0, "instrument")$value
-  platform <- ncatt_get(nc, 0, "platform")$value
-  product_name <- ncatt_get(nc, 0, "product_name")$value
-  fNameStem <- substr(product_name, 17, 38)
-  timeStamp <- substr(product_name, 2, 8)
-  origin <- paste0(substr(timeStamp, 1, 4), "-01-01")
-  date <- as.Date(as.numeric(substr(timeStamp, 5, 7)), origin)
-  chl <- round(ncvar_get(nc,
-                   varid = "chlor_a",
-                   start = c(LonIdx[1], LatIdx[1]),
-                   count = c(length(LonIdx), length(LatIdx))),
-               3)
-  dimnames(chl) <- list(lon = nc$dim$lon$vals[LonIdx],
-                        lat =  nc$dim$lat$vals[LatIdx])
-  nc_close(nc)
-  chl <-
-    as.data.table(melt(chl, value.name = "chl"), row.names = NULL) %>%
-    mutate(t = date) %>%
-    na.omit()
-  fwrite(chl,
-         file = paste(csv.dir, "/", region, "-", instrument, ".",platform, ".",
-                      fNameStem, "-", strt.date, "-", end.date, ".csv", sep = ""),
-         append = TRUE, col.names = FALSE)
-  rm(chl)
-}
-
-llply(ncList, ncFun, csv.dir = MODIS_chlor.csv.dir, .parallel = TRUE)
+# region <- "BC" # Benguela Current
+# 
+# coords <- c(-35, -20, 10, 20) # this is the BC
+# 
+# ncList <- list.files(path = MODIS_chlor.dir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
+# strt.date <- str_sub(basename(ncList[1]), start = 2, end = 8)
+# end.date <- str_sub(basename(ncList[length(ncList)]), start = 2, end = 8)
+# nc.init <- nc_open(ncList[1])
+# LatIdx <- which(nc.init$dim$lat$vals > coords[1] & nc.init$dim$lat$vals < coords[2])
+# LonIdx <- which(nc.init$dim$lon$vals > coords[3] & nc.init$dim$lon$vals < coords[4])
+# nc_close(nc.init)
+# 
+# 
+# ncFun <- function(nc.file = nc.files, csv.dir = csv.dir) {
+# nc <- nc_open(nc.file)
+#   instrument <- ncatt_get(nc, 0, "instrument")$value
+#   platform <- ncatt_get(nc, 0, "platform")$value
+#   product_name <- ncatt_get(nc, 0, "product_name")$value
+#   fNameStem <- substr(product_name, 17, 38)
+#   timeStamp <- substr(product_name, 2, 8)
+#   origin <- paste0(substr(timeStamp, 1, 4), "-01-01")
+#   date <- as.Date(as.numeric(substr(timeStamp, 5, 7)), origin)
+#   chl <- round(ncvar_get(nc,
+#                    varid = "chlor_a",
+#                    start = c(LonIdx[1], LatIdx[1]),
+#                    count = c(length(LonIdx), length(LatIdx))),
+#                3)
+#   dimnames(chl) <- list(lon = nc$dim$lon$vals[LonIdx],
+#                         lat =  nc$dim$lat$vals[LatIdx])
+#   nc_close(nc)
+#   chl <-
+#     as.data.table(melt(chl, value.name = "chl"), row.names = NULL) %>%
+#     mutate(t = date) %>%
+#     na.omit()
+#   fwrite(chl,
+#          file = paste(csv.dir, "/", region, "-", instrument, ".",platform, ".",
+#                       fNameStem, "-", strt.date, "-", end.date, ".csv", sep = ""),
+#          append = TRUE, col.names = FALSE)
+#   rm(chl)
+# }
+# 
+# llply(ncList, ncFun, csv.dir = MODIS_chlor.csv.dir, .parallel = TRUE)
 
@@
 # Loading the data
-load("Data_P1/MODIS_Chloro.RData") # Extracted/Raw chlorophyll data
-load("Data_P1/site_list_sub.Rdata") # Site list of all the sites along the west coast
-
-# Visualisation
-chloro_plot <- MODIS_Chloro %>%
-  filter(date == min(date)) %>%
-  ggplot(aes(x = lon, y = lat)) +
-  geom_raster(aes(fill = chloro))
-
-unique_pixel <- MODIS_Chloro %>% 
-  select(lon, lat) %>% 
-  unique()
-
-match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
-                          query = as.matrix(site_list_sub[,5:6]), k = 1)
-
-# Matching the chlorophylla-a lats and lons to the study sites
-pixel_match <- unique_pixel[match_index,] %>%
-  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
-  mutate(site = site_list_sub$site)
-
-chloro_match <- MODIS_Chloro %>%
-  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
-  filter(combi %in% pixel_match$combi)
-
-chloro_match %>%
-  filter(date == max(date)) %>%
-  ggplot(aes(x = lon, y = lat)) +
-  geom_raster(aes(fill = chloro))
-
-chloro_sites <- chloro_match %>%
-  left_join(pixel_match, by = c("combi", "lon", "lat")) %>% 
-  dplyr::mutate(date = as.Date(date))
-# save(chloro_sites, file = "Data_P1/chloro_sites.RData")
-
-# Visualisation
-Chloro_plot_complete <- ggplot(chloro_sites, aes(x = date, y = chloro)) +
-  geom_line() +
-  facet_wrap(~site, ncol = 1)
-Chloro_plot_complete
-
-# Study period from 2010 - 2014
-chloro_data <- chloro_sites %>% 
-  filter(year(date) %in% seq(2010, 2014))
-# save(chloro_data, file = "Data_P1/chloro_data.RData")
-
-# Plotting only years 2014-2015
-Chloro_plot_filtered <- ggplot(chloro_data, aes(x = date, y = chloro)) +
-  geom_line() +
-  facet_wrap(~site, ncol = 1)
-Chloro_plot_filtered
+# load("Data_P1/MODIS_Chloro.RData") # Extracted/Raw chlorophyll data
+# load("Data_P1/site_list_sub.Rdata") # Site list of all the sites along the west coast
+# 
+# # Visualisation
+# chloro_plot <- MODIS_Chloro %>%
+#   filter(date == min(date)) %>%
+#   ggplot(aes(x = lon, y = lat)) +
+#   geom_raster(aes(fill = chloro))
+# 
+# unique_pixel <- MODIS_Chloro %>% 
+#   select(lon, lat) %>% 
+#   unique()
+# 
+# match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
+#                           query = as.matrix(site_list_sub[,5:6]), k = 1)
+# 
+# # Matching the chlorophylla-a lats and lons to the study sites
+# pixel_match <- unique_pixel[match_index,] %>%
+#   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
+#   mutate(site = site_list_sub$site)
+# 
+# chloro_match <- MODIS_Chloro %>%
+#   unite(col = combi, lon, lat, sep = "/", remove = F) %>%
+#   filter(combi %in% pixel_match$combi)
+# 
+# chloro_match %>%
+#   filter(date == max(date)) %>%
+#   ggplot(aes(x = lon, y = lat)) +
+#   geom_raster(aes(fill = chloro))
+# 
+# chloro_sites <- chloro_match %>%
+#   left_join(pixel_match, by = c("combi", "lon", "lat")) %>% 
+#   dplyr::mutate(date = as.Date(date))
+# # save(chloro_sites, file = "Data_P1/chloro_sites.RData")
+# 
+# # Visualisation
+# load("Data_P1/chloro_sites.RData")
+# Chloro_plot_complete <- ggplot(chloro_sites, aes(x = date, y = chloro)) +
+#   geom_line() +
+#   facet_wrap(~site, ncol = 1)
+# Chloro_plot_complete
+# 
+# # Study period from 2010 - 2014
+# chloro_data <- chloro_sites %>% 
+#   filter(year(date) %in% seq(2010, 2014))
+# # save(chloro_data, file = "Data_P1/chloro_data.RData")
+# 
+# # Plotting only years 2014-2015
+# load("Data_P1/chloro_data.RData")
+# Chloro_plot_filtered <- ggplot(chloro_data, aes(x = date, y = chloro)) +
+#   geom_line() +
+#   facet_wrap(~site, ncol = 1)
+# Chloro_plot_filtered
 
@@
 
-daily_chloro_data <- chloro_data %>%
-  # dplyr::mutate(date = as.Date(date)) %>% # RWS: The values already are dates
+daily_chloro <- chloro_data %>%
+  # dplyr::mutate(date = as.Date(date)) %>% 
   dplyr::group_by(combi, lon, lat, date) %>%
   dplyr::summarise(chloro = mean(chloro, na.rm = TRUE)) %>%
-  dplyr::group_by(combi, lon, lat) %>% # RWS: Deleted the date grouping variable being used here
+  dplyr::group_by(combi, lon, lat) %>% 
+  drop_na() %>% 
   tidyr::nest() %>%
@@
 
+To do list:
+- fix wind roses
+- fix chloro
+- Robs pixels
+- EOFs
+- fix pca
+
 ## EOF's
@@
 
+PCA_1 <- rda(tester)
+summary(PCA)
+
+# Proportion of variation explained
+# Values to be found opon running the code
+
+PCA <- cbind(scores(PCA_1, choices = c(1:4), display = "sites"))
+PCA_spe <- cbind(scores(PCA_1, choices = c(1:4), display = "species"))
+
+test1 <- ggplot(PCA, aes(x = lon, y = lat)) +
+  geom_raster(aes(fill = PC1)) +
+  geom_polygon(data = PC1, aes(long, lat, group = group),
+               fill = NA, colour = "blue", size = 0.3) +
+  geom_polygon(data = velocity, aes(long, lat, group = group),
+               fill = NA, colour = "green", size = 0.3)+ 
+         fill = guide_colourbar(title = "PC1")
+
+test2 <- ggplot(PCA, aes(x = lon, y = lat)) +
+  geom_raster(aes(fill = PC2)) +
+  geom_polygon(data = PC1, aes(long, lat, group = group),
+               fill = NA, colour = "blue", size = 0.3) +
+  geom_polygon(data = velocity, aes(long, lat, group = group),
+               fill = NA, colour = "green", size = 0.3)+ 
+         fill = guide_colourbar(title = "PC2")
+
+test3 <- ggplot(PCA, aes(x = lon, y = lat)) +
+  geom_raster(aes(fill = PC2)) +
+  geom_polygon(data = PC1, aes(long, lat, group = group),
+               fill = NA, colour = "blue", size = 0.3) +
+  geom_polygon(data = velocity, aes(long, lat, group = group),
+               fill = NA, colour = "green", size = 0.3)+ 
+         fill = guide_colourbar(title = "PC3")
+
+test4 <- ggplot(PCA, aes(x = lon, y = lat)) +
+  geom_raster(aes(fill = PC1)) +
+  geom_polygon(data = PC1, aes(long, lat, group = group),
+               fill = NA, colour = "blue", size = 0.3) +
+  geom_polygon(data = velocity, aes(long, lat, group = group),
+               fill = NA, colour = "green", size = 0.3)+ 
+         fill = guide_colourbar(title = "PC4")
+
+pc1_combined <- ggarrange( test1, test2, test3, test4)
+
 ```
 
-In particular, trends will be observed to be different at ocean and coastal locations and we hypothesize that
-those differences are due to changes in upwelling intensity.
-
-- Detemining the upwelling indeces in R
-- How does upwelling indeces vary at the different distances along the coast
-- Positive indeces display a positive upwelling.
-
-Things to do
-
-- Plot seasonal detrend
-- chlorophyll a data
-- fix windrose plot
-- Stastical analyses
+In particular, trends will be observed to be different at ocean and coastal locations and we hypothesize that those differences are due to changes in upwelling intensity.

### 2019-04-17 functions/theme.R
updated

@@ -1,0 +1,57 @@
+library(ggplot2)
+# library(extrafont)
+# loadfonts(device = "pdf", quiet = FALSE)
+
+# make a custom theme
+theme_set(theme_bw())
+bw_update <- theme_bw() +
+  theme(panel.border = element_rect(size = 1.0),
+        # panel.grid.major = element_line(colour = "black", size = 0.2, linetype = 2),
+        panel.grid.major = element_line(colour = NA),
+        panel.grid.minor = element_line(colour = NA),
+        axis.title = element_text(size = 12, face = "bold"),
+        axis.text = element_text(size = 12, colour = "black"),
+        plot.title = element_text(size = 12, hjust = 0),
+        legend.title = element_text(size = 10),
+        legend.text = element_text(size = 10),
+        legend.key = element_rect(size = 0.8, colour = NA),
+        strip.background = element_rect(colour = NA, fill = NA),
+        strip.text = element_text(size = 10))
+# theme_set(bw_update) # Set theme
+
+theme_set(theme_bw())
+bw_update2 <- theme_bw() +
+  theme(panel.border = element_rect(size = 1.0),
+        panel.grid.major = element_line(colour = "black", size = 0.2, linetype = 2),
+        panel.grid.minor = element_line(colour = NA),
+        axis.title = element_text(size = 12, face = "bold"),
+        axis.text = element_text(size = 12, colour = "black"),
+        plot.title = element_text(size = 12, hjust = 0),
+        legend.title = element_text(size = 10),
+        legend.text = element_text(size = 10),
+        legend.key = element_rect(size = 0.8, colour = NA),
+        legend.background = element_blank(),
+        strip.background = element_rect(colour = NA, fill = NA),
+        strip.text = element_text(size = 10))
+theme_set(bw_update2) # Set theme
+
+theme_set(theme_grey())
+grey_update <- theme_grey() +
+  theme(#panel.border = element_rect(colour = "black", fill = NA, size = 1.0),
+        panel.grid.major = element_line(size = 0.2, linetype = 2),
+        panel.grid.minor = element_line(colour = NA),
+        axis.title = element_text(size = 12, face = "bold"),
+        axis.text = element_text(size = 12, colour = "black"),
+        plot.title = element_text(size = 12, hjust = 0),
+        legend.title = element_text(size = 10),
+        legend.text = element_text(size = 10),
[…]

### 2019-04-24 Write_up/Writing/dissertation.Rmd
update

@@
 ---
-title: "Dynamical properties of the Benguela Current as seen through a cold-spell algorithm."
+title: "Dynamical properties of the Benguela Current"
 author: "Amieroh Abrahams"
@@
 output:
-  pdf_document: default
-  html_document: default
+  word_document:
+    fig_caption: yes
+    fig_height: 7
+    fig_width: 7
+    highlight: zenburn
+    toc: yes
+  html_document:
+    fig_caption: yes
+    fig_retina: 2
+    fig_width: 7
+    highlight: espresso
+    theme: sandstone
+    toc_float: no
+  bookdown::html_document2:
+    fig_caption: yes
+    fig_retina: 2
+    fig_width: 7
+    highlight: espresso
+    theme: sandstone
+    toc_float: no
+  pdf_document:
+    toc: yes
+    toc_depth: 4
+    highlight: default
+    keep_tex: yes
+    latex_engine: xelatex
 fontsize: 10pt
@@
 ---
-
 # Chapter 1: Introduction
@@
 Upwelling occuring along the south coastline of the of Southern Africa are particularly generated by local winds.
+Event-scale upwelling has been demonstrated to have substantial impacts of individual on fisheries [@roy; @van]
 
@@
 The process of upwelling is one such factor that is predicted to change with an increase in global warming.
-As prreviously mentioned, changes in temperature results in changes in pressure gradients, these pressure gradients control wind patterns which ultimately controls the intensity and duration of upwelling.
+As previously mentioned, changes in temperature results in changes in pressure gradients, these pressure gradients control wind patterns which ultimately controls the intensity and duration of upwelling.
 To assess the changes within the upwelling events, various datasets would be used.
@@
 
+```{r , echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Map representing the study region along the South African coastline.
+The black points represent the location of their *in situ* temperatures and approximations of the pixels used along the shore normal trasect from the satellite sea surface temperatures shown with black boxes", fig.height=8, fig.width=12, fig.pos="H"}
+
+# knitr::include_graphics("../figures/map_fixed.pdf")
+```
+
 ## Detecting the intensity and duration of upwelling
+
+## Results
+
+### Wind patterns during upwelling events
+
+```{r  echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Wind rose diagrams representing the the most predominant wind direction for each of the sites for the years 2010 - 2014.
+Each spoke is divided by color into wind speed ranges.
+The radial length of each spoke around the circle is the amount of times that the wind comes from that direction", fig.height=8, fig.width=12, fig.pos="H"}
+
+ knitr::include_graphics("../figures/wind.png")
+```
+
+## Local and broad scale upwelling

### 2019-04-27 Paper1_results.Rmd
working_bits

@@
 library(geosphere)
-# library(marmap)
 library(PBSmapping)
@@
 library(mapproj)
-# library(rgeos)
 library(ncdf4)
-# library(sinkr)
-library(spacetime)
 library(reshape2)
@@
 library(fasttime)
+library(xtable)
 
@@
 source("functions/wind.rose.R")
+source("functions/theme.R")
+source("functions/scale.bar.func.R")
 ```
@@
 
-Loading in the matched satellite temperature data
-
-```{r}
-load("Data_P1/SACTN_overlap.RData")
-load("Data_P1/insitu_MUR.RData")
-load("Data_P1/insitu_OISST.RData")
-load("Data_P1/insitu_G1SST.RData")
-load("Data_P1/insitu_K10.RData")
-```
-
 ## Wavelets
@@
 
-## Load in the datasets
+## Loading in the matched satellite temperature data
 
@@
 K10_fyears <- filtered_years(df = insitu_K10)
+
+# save(SACTN_fyears, file = "Data_P1/SACTN_fyears.RData")
+# save(MUR_fyears, file = "Data_P1/MUR_fyears.RData")
+# save(OISST_fyears, file = "Data_P1/OISST_fyears.RData")
+# save(G1SST_fyears, file = "Data_P1/G1SST_fyears.RData")
+# save(K10_fyears, file = "Data_P1/K10_fyears.RData")
 ```
@@
 
+## Load in the datasets
+
+```{r}
+load("Data_P1/SACTN_overlap.RData")
+load("Data_P1/insitu_MUR.RData")
+load("Data_P1/insitu_OISST.RData")
+load("Data_P1/insitu_G1SST.RData")
+load("Data_P1/insitu_K10.RData")
+```
+
 # Calculating a shore transect
@@
 load("Data_P1/site_list_sub.Rdata")
+xtable(site_list_sub, auto = TRUE)
 west <- site_list_sub
@@
 ## Create a visualisation of the sites at the different distances from the coastline
-****Fix: Fill to show the temperature variation at Benguela coast| also fix the site names*
 
 ```{r}
-
 load("Data_P1/south_africa_coast.RData")
@@
 load("Data_P1/bathy/bathy.RData")
+load("Data_P1/site_list_sub.Rdata")
+load("Data_P1/site_pixels.RData")
+load("Data_P1/MUR.RData")
 names(south_africa_coast)[1] <- "lon"
@@
 
+# Setting up the theme
+theme_set(theme_bw())
+limits <- c(12,28) # for colour bar
+#breaks <- seq(6, 30, 2) # Create breaks to be used for colour bar
+
 # Define plotting parameters
-sa_lats <- c(-37, -27); sa_lons <- c(14, 34)
+sa_lats <- c(-37, -27); sa_lons <- c(14, 50)
 
- plot1 <- ggplot() +#coord_equal() + 
-  geom_raster(data = sa_bathy, aes(x = lon, y = lat, fill = depth)) +
-  #geom_raster(data = isobath200, aes(x = lon, y = lat, fill = coast)) +
-  stat_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth, alpha = ..level..), 
-               colour = "black", size = 0.2, binwidth = 200, na.rm = TRUE, show_guide = FALSE) +
+site_map <- ggplot(data = south_africa_coast, aes(x = lon, y = lat)) + bw_update +
+  geom_raster(data = MUR, aes(x = lon, y = lat, fill = temp)) +
    geom_polygon(data = south_africa_coast, aes(x = lon, y = lat, group = group),
-               size = 0.5, colour = "black", fill = "white") +
-  #geom_path(data = sa_provinces_new, aes(x = long, y = lat, group = group)) +
-  # geom_polygon(data = wcBox, aes(x = lon, y = lat, group = group), alpha = 0.20, 
-  #              colour = "bisque1", fill = "bisque1") +
-  geom_point(data = site_list_sub, aes(x = lon, y = lat), alpha = 0.8, size = 1.6) +
+                fill = "grey", colour = "grey", size = 0.1, show.legend = FALSE) +
+  geom_point(data = site_list_sub, aes(x = lon, y = lat), alpha = 0.8, size = 3) +
   geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "white", shape = 0, alpha = 0.8, size = 2.1) +
-  geom_rect(data = bbox, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), 
-            alpha = 0.4, colour = "red", size = 0.1, linetype = 1) +
-  scale_alpha_continuous(breaks = c(-200, -1000, -2000, -3000, -4000, -5000),
-                         guide_legend(title = "depth (m)")) +
-  scale_fill_gradient(low = "dodgerblue4", high = "dodgerblue", na.value = "steelblue4", 
-                      breaks = c(-1000, -2000, -3000, -4000, -5000),
-                      guide_legend(title = "depth (m)")) +
-  labs(title = NULL, x = NULL, y = NULL) +
-  guides(colour = guide_legend(override.aes = list(size = 3))) +
-  #guide_legend(ncol = 2) +
-  theme(legend.key = element_rect(colour = NA, size = 0.2),
-        legend.key.height = unit(0.4, "cm"),
-        legend.background = element_blank()) +
-        #legend.justification = c(1,0), legend.position = c(0.5, 0.4)) +
-  coord_cartesian(xlim = sa_lons, ylim = sa_lats)
- plot1
+  # geom_rect(data = bbox, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), 
+  #           alpha = 0.4, colour = "red", size = 0.1, linetype = 1) +
+  # scale_x_continuous(limits = sa_lons, expand = c(0, 0), breaks = seq(15, 35, 5)) +
+  # scale_y_continuous(limits = sa_lats, expand = c(0, 0), breaks = seq(-35, -30, 5)) +
+  scale_fill_viridis(
+    option = "magma",
+    name = "Temperature (°C)",
+    alpha = 0.8) +
+  #scale_fill_viridis(limits = limits, expression(paste("Temperature (°C)"))) +
+  xlab("") + ylab("") +
+  annotate("text", label = "INDIAN\nOCEAN", x = 32.00, y = -35.0,
+             size = 5, angle = 0, colour = "black") +
+    annotate("text", label = "ATLANTIC\nOCEAN", x = 17.00, y = -35.0,
+             size = 5, angle = 0, colour = "black") +
+    annotate("text", label = "Benguela", x = 16.0, y = -31.7,
+             size = 6, angle = 302, colour = "black") +
+    annotate("text", label = "Agulhas", x = 31.7, y = -31.7,
+             size = 6, angle = 50, colour = "black") +
+  guides(fill = guide_colourbar()) +
+  scale_x_continuous(expand = c(0, 0),
+                       labels = scales::unit_format(unit = "°E", sep = "")) +
+    scale_y_continuous(expand = c(0, 0),
+                       labels = scales::unit_format(unit = "°S", sep = "")) +
+  theme(panel.background = element_rect(fill = "ivory", colour = NA),
+        panel.border = element_rect(colour = "black", size = 0.5),
+        panel.grid.minor = element_line(colour = "NA"),
+        panel.grid.major = element_line(colour = "ivory", size = 0.2, linetype = "dotted"),
+        legend.text = element_text(size = 8),
+        legend.title = element_text(size = 8),
+        legend.key = element_blank(),
+        legend.background = element_blank(),
+        axis.ticks = element_line(size = 0.5))
+site_map
 
- # Added in the bathy data but this is not needed for any of the analyses.
- # This just shows the bounding boxes 
+# png("Figures/site_map.png", width = 8, height = 5, pointsize = 6)
 ```
@@
 
+# load("Data_P1/insitu_wind.RData")
+# 
+# test1 <- insitu_wind %>% 
+#   mutate(lat = case_when(site == "Port Nolloth" ~ "-29.25241",
+#                                     site == "Saldanha Bay" ~ "-33.01054",
+#                                     site  == "Yzerfontein" ~ "-33.36068",
+#                                     site  == "Kommetjie" ~ "-34.13667",
+#                                     site  == "Lamberts Bay" ~ "-32.09185",
+#                                     site  == "Sea Point" ~ "-33.91847"))
+# 
+# test <- test1 %>% 
+#   mutate(lon = case_when(site == "Port Nolloth" ~ "16.86710",
+#                                     site == "Saldanha Bay" ~ "17.95063",
+#                                     site  == "Yzerfontein" ~ "18.15731",
+#                                     site  == "Kommetjie" ~ "18.32674",
+#                                     site  == "Lamberts Bay" ~ "18.30262",
+#                                     site  == "Sea Point" ~ "18.38217"))
+# 
+# test_final <- test %>% 
+#   select(date,lat,lon,dir_circ, mean_speed) %>% 
+#   dplyr::rename(speed = mean_speed) %>% 
+#   dplyr::rename(dir = dir_circ)
+
 # Loading the wind data
@@
 ```{r}
-load("Data_P1/SACTN_overlap.RData")
-load("Data_P1/wind_daily.RData")
-wind_daily <- wind_daily%>% 
-  dplyr::rename(site = temp_sites) # Renaming to match the `match_func` previously created
-
-insitu_wind <- match_func(df = wind_daily)  # Using `match_func`
-
-insitu_wind <- insitu_wind %>% 
-  filter(year(date) %in% seq(2010, 2014)) # The years 2010-2014 is the years of interest and the years for which wavelets are/will be created.
+# load("Data_P1/SACTN_overlap.RData")
+# load("Data_P1/wind_daily.RData")
+# wind_daily <- wind_daily%>% 
+#   dplyr::rename(site = temp_sites) # Renaming to match the `match_func` previously created
+# 
+# insitu_wind <- match_func(df = wind_daily)  # Using `match_func`
+# 
+# insitu_wind <- insitu_wind %>% 
+#   filter(year(date) %in% seq(2010, 2014)) # The years 2010-2014 is the years of interest and the years for which wavelets are/will be created.
 
@@
 load("Data_P1/insitu_wind.RData")
+
+insitu_wind <- insitu_wind %>% 
+  drop_na()
 
@@
 p.wr3 # RWS: You'll want to remove the NA part in the figure output
+
+#########################
+library(openair)
+print(
+with(insitu_wind,
+     windRose(data_frame(ws=mean_speed, wd=dir_circ, 
+                         date=date, station=factor(site)),
+
+              paddle=FALSE, type="station", width=2))
+)
+png("Figures/wind.png", width = 8, height = 5, pointsize = 6)
+?windRose
+# insitu_wind_2 <- insitu_wind %>% 
+#   filter(site == "Port Nolloth")
+# 
+# insitu_wind_3 <- insitu_wind %>% 
+#   filter(site == "Lamberts Bay")
+# 
+# plotPN <- plot(as.factor(format(insitu_wind_2$date, "%Y-%m")), insitu_wind_2$insitu_temp, col = "lightpink")
+# plotLB <- plot(as.factor(format(insitu_wind_3$date, "%Y-%m")), insitu_wind_3$insitu_temp, col = "lightpink")
+#        
+# ?plot
 ```
@@
 # MODIS_Chloro <- "~/Documents/Masters_2019/MastersProject/Data"
-# MODIS_Chloro <- fread(paste0(MODIS_Chloro, "/BC-MODIS.Aqua.L3m_8D_CHL_chlor_a_9km-2002185-2018345.csv"),
+# MODIS_Chloro <- fread(paste0(MODIS_Chloro, "/BC-M ODIS.Aqua.L3m_8D_CHL_chlor_a_9km-2002185-2018345.csv"),
 #             col.names = c("lon", "lat", "chloro", "date"))
@@
 
-To do list:
-- fix wind roses
-- fix chloro
-- Robs pixels
-- EOFs
-- fix pca
+Here I want to test using generalised linear hypotheses if there is a difference in the intensity of upwellling events between the insitu and satellite data.
+Do the in situ and satelllite derived data yield the same number of events.
+
+library(multcomp) # for generalized linear hypothesis test (glht)
+
+```{r}
+
+```
+
+Determining the R2 value between *in situ* and remotely sensed SST
+
+```{r}
+# Load time series for both datasets
+load("prep/SA_coastal_temps.RData")
+SA_coastal_temps$date <- as.Date(SA_coastal_temps$date)
+load("data/OISSTdaily.Rdata")
+
+resultsR2 <- data.frame()
+for(i in 1:length(levels(SA_coastal_temps$site))) {
+  x <- subset(SA_coastal_temps, site == levels(SA_coastal_temps$site)[i])
+  y <- subset(OISSTdaily, site == levels(SA_coastal_temps$site)[i])
+  x <- x[x$date %in% y$date,]
+  y <- y[y$date %in% x$date,]
+  z <- data.frame(site = as.character(x$site[1]), R2 = round(coef(lm(y$temp~x$temp))[2],2))
+  resultsR2 <- rbind(resultsR2, z)
+}
+
+# Add column for plotting
+row.names(resultsR2) <- NULL
+resultsR2$R22 <- paste0("R^2 ==", format(resultsR2$R2, digits=2))
+```
+
+```{r}
+# Loading the *in situ* and Remotely sensed SST
+
+load("Data_P1/SACTN_fyears.RData")
+load("Data_P1/MUR_fyears.RData")
+load("Data_P1/OISST_fyears.RData")
+load("Data_P1/G1SST_fyears.RData")
+load("Data_P1/K10_fyears.RData")
+
+SACTN_fyears <- SACTN_fyears %>% 
+  dplyr::rename(temp = insitu_temp)
+
+R2 <- data.frame()
+for(i in 1:length(levels(SACTN_fyears$site))) {
+  x <- subset(SACTN_fyears, site == levels(SACTN_fyears$site)[i])
+  y <- subset(OISST_fyears, site == levels(OISST_fyears$site)[i])
+  x <- x[x$date %in% y$date,]
+  y <- y[y$date %in% x$date,]
+  z <- data.frame(site = as.character(x$site[1]), R2 = coef(lm(y$temp~x$temp))[2],na.action=na.exclude)
+  R2 <- rbind(R2, z)
+}
+```
 
@@
 ```{r} 
-
 PCA_1 <- rda(tester)
@@
 In particular, trends will be observed to be different at ocean and coastal locations and we hypothesize that those differences are due to changes in upwelling intensity.
+
+Mond:
+- Try PCA
+- map it/plot
+- dave upwelling index
+
+Tuesday
+- Do exercises

### 2019-04-27 functions/scale.bar.func.R
updating

@@ -1,0 +1,177 @@
+#############################################################################
+###"func/scaleBarFunc.R"
+## This script does:
+# 1. Set the scale bar coordinates
+# 2. Obtain the coordinates of the North arrow
+# 3. Draw the elements in ggplot2
+## NB: This script was copied from: http://editerna.free.fr/wp/?p=76
+## DEPENDS ON:
+library(maps)
+library(maptools)
+library(ggplot2)
+library(grid)
+## USED BY:
+# "figures.R"
+## CREATES:
+# A wonderful scale bat in ggplot
+#############################################################################
+
+#############################################################################
+## 1. Set the scale bar coordinates
+#--------#
+# Return a list whose elements are :
+# 	- rectangle : a data.frame containing the coordinates to draw the first rectangle ;
+# 	- rectangle2 : a data.frame containing the coordinates to draw the second rectangle ;
+# 	- legend : a data.frame containing the coordinates of the legend texts, and the texts as well.
+#
+# Arguments : #
+#-------------#
+# lon, lat : longitude and latitude of the bottom left point of the first rectangle to draw ;
+# distanceLon : length of each rectangle ;
+# distanceLat : width of each rectangle ;
+# distanceLegend : distance between rectangles and legend texts ;
+# dist.units : units of distance "km" (kilometers) (default), "nm" (nautical miles), "mi" (statute miles).
+createScaleBar <- function(lon,lat,distanceLon,distanceLat,distanceLegend, dist.units = "km"){
+  # First rectangle
+  bottomRight <- gcDestination(lon = lon, lat = lat, bearing = 90, dist = distanceLon, dist.units = dist.units, model = "WGS84")
+  
+  topLeft <- gcDestination(lon = lon, lat = lat, bearing = 0, dist = distanceLat, dist.units = dist.units, model = "WGS84")
+  rectangle <- cbind(lon=c(lon, lon, bottomRight[1,"long"], bottomRight[1,"long"], lon),
+                     lat = c(lat, topLeft[1,"lat"], topLeft[1,"lat"],lat, lat))
+  rectangle <- data.frame(rectangle, stringsAsFactors = FALSE)
+  
+  # Second rectangle t right of the first rectangle
[…]

### 2019-04-27 Paper1_results.Rmd
fixed

@@
 ```{r}
-# Load time series for both datasets
-load("prep/SA_coastal_temps.RData")
-SA_coastal_temps$date <- as.Date(SA_coastal_temps$date)
-load("data/OISSTdaily.Rdata")
-
-resultsR2 <- data.frame()
-for(i in 1:length(levels(SA_coastal_temps$site))) {
-  x <- subset(SA_coastal_temps, site == levels(SA_coastal_temps$site)[i])
-  y <- subset(OISSTdaily, site == levels(SA_coastal_temps$site)[i])
-  x <- x[x$date %in% y$date,]
-  y <- y[y$date %in% x$date,]
-  z <- data.frame(site = as.character(x$site[1]), R2 = round(coef(lm(y$temp~x$temp))[2],2))
-  resultsR2 <- rbind(resultsR2, z)
-}
-
-# Add column for plotting
-row.names(resultsR2) <- NULL
-resultsR2$R22 <- paste0("R^2 ==", format(resultsR2$R2, digits=2))
-```
-
-```{r}
 # Loading the *in situ* and Remotely sensed SST

### 2019-04-27 Exploring/Detect_MCSs.Rmd
updating

@@
   ggtitle("MUR_PN") 
-
+library(tidyverse)
+library(heatwaveR)
 MUR_SST_LB <- MUR_SST %>% 

### 2019-05-01 Paper1_results.Rmd
updating
Updated

@@
 Wavelet anlyses showing the temperature variation between sites along the west coast for the years 2010 - 2014.
-Working with remotely-sensed SST and *in-situ* collected coastal seawater temperature, in the analyses bellow I aim to show how the intensity and duration of upwelling events vary between sites along the west coast of South Africa.
+Working with remotely-sensed SST and *in-situ* collected coastal seawater temperature, in the analyses below I aim to show how the intensity and duration of upwelling events vary between sites along the west coast of South Africa.
 It is important to remember that each of the datasets used within this study obtained SSTs at different resolutions.
@@
 
-# Wavelet analyses
+# Wavelet analyses.
+    # This creates a visualisation looking at the intensity and duration of upwelling. 
+    # Here we test how the intensity and duration of an upwelling event vary between sites along the west coast of SA
+    # With the hypotheses that sites closer to the upwelling centre will show a more intense, longer upwelling event
 
@@
 # When doing a wavelet analyses on the SACTN dataset the "insitu_temp" column needed to be renamed to "temp" for the rest of the code to run. 
-The rest of this code is kept constant for each of the datasets (Remotely-sensed SST and In-situ seawater temperature)
+# The rest of this code is kept constant for each of the datasets (Remotely-sensed SST and In-situ seawater temperature)
 
@@
 
-The aim of this is to find the temperature collected at different distances from the coast (distances = 15km, 25km, 35km and 50km).
+The aim of this is to find the temperature collected at different distances from the coast (distances = 10km, 20km, 30km, 40km and 50km).
 
@@
   for(i in 1:length(distances)){
-    # Add index to show distance from coast
-    # Find depth at distance step
     coords2 <- as.data.frame(destPoint(p = coords, b = site$heading, d = distances[i]))
@@
                          heading = site$heading, 
-                         # depth = bathy2$depth,
                          distance = distances[i])
@@
     coords <- coords2
-    # distance_multiplier <- distance_multiplier +1 
-    # deep <- bathy2$depth
-  }
-  # pixels2 <- pixels[pixels$depth > -999,] 
+  }
   if(nrow(pixels) < 1){
@@
   site <- site_transects[i,]
-  site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: The desired distances
+  site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: fixed error
   site_pixels <- rbind(site_pixels, site_pixel)
@@
 
-## Create a visualisation of the sites at the different distances from the coastline
+## Create a visualisation of the sites at the different distances from the coastline. Mapping the study area
 
@@
 load("Data_P1/south_africa_coast.RData")
-load("Data_P1/bathy/sa_bathy.RData")
-load("Data_P1/bathy/bathy.RData")
+# load("Data_P1/bathy/sa_bathy.RData")
+# load("Data_P1/bathy/bathy.RData")
 load("Data_P1/site_list_sub.Rdata")
@@
 # Define plotting parameters
-sa_lats <- c(-37, -27); sa_lons <- c(14, 50)
+sa_lats <- c(-37, -27); sa_lons <- c(14, 25)
 
@@
   geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "white", shape = 0, alpha = 0.8, size = 2.1) +
+  coord_equal(xlim = c(15, 25), ylim = c(-36,-26), expand = 0) +
   # geom_rect(data = bbox, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
@@
     alpha = 0.8) +
-  #scale_fill_viridis(limits = limits, expression(paste("Temperature (°C)"))) +
   xlab("") + ylab("") +
-  annotate("text", label = "INDIAN\nOCEAN", x = 32.00, y = -35.0,
-             size = 5, angle = 0, colour = "black") +
+  # annotate("text", label = "INDIAN\nOCEAN", x = 32.00, y = -35.0,
+  #            size = 4, angle = 0, colour = "black") +
     annotate("text", label = "ATLANTIC\nOCEAN", x = 17.00, y = -35.0,
-             size = 5, angle = 0, colour = "black") +
+             size =4, angle = 0, colour = "black") +
     annotate("text", label = "Benguela", x = 16.0, y = -31.7,
-             size = 6, angle = 302, colour = "black") +
-    annotate("text", label = "Agulhas", x = 31.7, y = -31.7,
-             size = 6, angle = 50, colour = "black") +
+             size = 5, angle = 302, colour = "black") +
+    # annotate("text", label = "Agulhas", x = 31.7, y = -31.7,
+    #          size = 5, angle = 50, colour = "black") +
   guides(fill = guide_colourbar()) +
@@
 
-# Plotting the wind data - Waverose diagrams
+# Plotting the wind data - windrose diagrams
+  # Here I produce a wind rose diagram showing wind action or wind patterns during the upwelling period
 
@@
 )
-png("Figures/wind.png", width = 8, height = 5, pointsize = 6)
+# png("Figures/wind.png", width = 8, height = 5, pointsize = 6)
 ?windRose
+
 # insitu_wind_2 <- insitu_wind %>% 
@@
 
-Here I want to test using generalised linear hypotheses if there is a difference in the intensity of upwellling events between the insitu and satellite data.
-Do the in situ and satelllite derived data yield the same number of events.
-
-library(multcomp) # for generalized linear hypothesis test (glht)
-
-```{r}
-
-```
-
-Determining the R2 value between *in situ* and remotely sensed SST
-
-```{r}
-# Loading the *in situ* and Remotely sensed SST
-
+First I calculate the distance between the sites along the wc using the deg2rad function which transforms degrees to radians: Transforms between angles in degrees and radians.
+Thereafter I run a three way ANOVA.
+This was primarily done to test whether a significant difference in temperature exist between sites on a seasonal, yearly bases.
+Here the distance between the sites could also play a role.
+But distance is mainly used when compaing the intensity and duration of an upwelling event given the distance between the sites and its location (location: at an upwelling centre or further away)
+
+```{r}
+load("Data_P1/sa_coast.Rdata")
+load("Data_P1/site_list_sub.Rdata")
 load("Data_P1/SACTN_fyears.RData")
@@
 load("Data_P1/K10_fyears.RData")
+SACTN_fyears <- SACTN_fyears %>% 
+  dplyr::rename(temp = insitu_temp)
+
+options(scipen=999) # Forces R to not use exponential notation
+sa_coast_prep <- sa_coast %>% 
+  mutate(site = 1:nrow(.),
+         X = deg2rad(X),
+         Y = deg2rad(Y)) %>% 
+  select(site, X, Y)
+
+sa_coast_dist <- as.data.frame(round(PairsDists(sa_coast_prep), 2)) %>% 
+  mutate(lon = sa_coast$X,
+         lat = sa_coast$Y) %>% 
+  dplyr::rename(dist = V1) %>%
+  select(lon, lat, dist) %>% 
+  mutate(cum_dist = cumsum(dist)) %>% 
+  mutate(dist = lag(dist),
+         cum_dist = lag(cum_dist))
+
+# Calculating the distances between sites.
+# This may be used to show how upwelling temperature varies between sites based on its location, i.e. closer 
+# to the upwelling centre and further from the upwelling centre
+
+SACTN_site_dis <- function(df) {
+  SACTN_dis <- data.frame()
+  df2 <- df %>%
+    select(-length) %>%
+    droplevels()
+  for (i in 1:length(levels(df2$index))) {  
+    # for (i in 1:10) {
+    SACTN_df_1 <- filter(df2, index == levels(index)[i])
+       # Find the cumulative distance of site 1 along the coast
+    site_dist_1 <- site_list_sub %>% 
+      filter(index == as.character(SACTN_df_1$index)[1])
+    site_dist_1 <- knnx.index(data = as.matrix(sa_coast_dist[,1:2]),
+                              query = as.matrix(site_dist_1[,5:6]), k = 1)
+    site_dist_1 <- sa_coast_dist$cum_dist[site_dist_1]
+
+    for (j in 1:length(levels(df2$index))) {
+      # for(j in 1:10) {
+      if (i == j) {
+        next
+      }
+      if (j < i) {
+        next
+      }
+      SACTN_df_2 <- filter(df2, index == levels(index)[j])
+
+      # Find the cumulative distance of site 2 along the coast
+      site_dist_2 <- site_list_sub %>% 
+        filter(index == as.character(SACTN_df_2$index)[1])
+      site_dist_2 <- knnx.index(data = as.matrix(sa_coast_dist[,1:2]),
+                                query = as.matrix(site_dist_2[,5:6]), k = 1)
+      site_dist_2 <- sa_coast_dist$cum_dist[site_dist_2]
+
+      # # The distance between the two sites
+      site_dist_3 <- round(abs(site_dist_1-site_dist_2), 0)
+
+      SACTN_df_3 <- left_join(SACTN_df_1, SACTN_df_2, by = "date") %>%
+        na.trim() %>% 
+      select(date, index.x, temp.x, index.y, temp.y) %>%
+        dplyr::rename(index_1 = index.x,
+          temp_1 = temp.x,
+          index_2 = index.y,
+          temp_2 = temp.y) %>% 
+        mutate(index_dist = site_dist_3)
+      SACTN_dis <- rbind(SACTN_dis, SACTN_df_3)
+    }
+  }
+  return(SACTN_dis)
+}
+SACTN_site_dist <- SACTN_site_dis(df = SACTN_fyears)
+
+# Adding seasonal column
+
+SACTN_monthly <- function(df) {
+  SACTN_split_monthly <- df %>%
+  mutate(month = month(date, abbr = T, label = T),
+         year = year(date)) %>%
+  group_by(index, month, year) %>%
+    mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",       
+                           ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
+                                ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
+                                       ifelse(month %in% c("Oct", "Nov", "Dec"), "Spring","Error")))))
+  return(SACTN_split_monthly)
+}
+
+# SACTN_season <- SACTN_monthly(df = SACTN_fyears)
+# MUR_fyears_season <- SACTN_monthly(df = MUR_fyears)
+# OISST_fyears_season <- SACTN_monthly(df = OISST_fyears)
+# G1SST_fyears_season <- SACTN_monthly(df = G1SST_fyears)
+# K10_fyears_season <- SACTN_monthly(df = K10_fyears)
+
+# save(SACTN_season, file = "SACTN_season.RData")
+# save(MUR_fyears_season, file = "MUR_fyears_season.RData")
+# save(OISST_fyears_season, file = "OISST_fyears_season.RData")
+# save(G1SST_fyears_season, file = "G1SST_fyears_season.RData")
+# save(K10_fyears_season, file = "K10_fyears_season.RData")
+
+load("Data_P1/SACTN_season.RData")
+load("Data_P1/MUR_fyears_season.RData")
+load("Data_P1/OISST_fyears_season.RData")
+load("Data_P1/G1SST_fyears_season.RData")
+load("Data_P1/K10_fyears_season.RData")
+
+# ANOVA analyses
+anova_func <- function(df){
+  sites_aov <- aov(temp ~ site * year * season, data = df)
+return(sites_aov)
+}
+
+SACTN_season <- SACTN_season %>% 
+  dplyr::rename(temp = insitu_temp)
+
+SACTN_season_func <- anova_func(df = SACTN_season)
+summary(SACTN_season_func)
+MUR_season_func <- anova_func(df = MUR_fyears_season)
+summary(MUR_season_func)
+OISST_season_func <- anova_func(df = OISST_fyears_season)
+summary(OISST_season_func)
+G1SST_season_func <- anova_func(df = G1SST_fyears_season)
+summary(G1SST_season_func)
+K10_season_func <- anova_func(df = K10_fyears_season)
+summary(K10_season_func)
+
+# Shows a significant different in temperatures between sites for each of the years and between seasons
+
+#    ? This is for all temperatures collected. not only temperatures collected during an upwelling event
+#    ? So now I need to find a way to restrict temperatures to only the periods where upwelling occured
+#    ? Can use the heatwaveR package to identify the count, intensity and duration but for this I require to know what the upwelling threshold is
+#     MHW: threshold- 90th percentile, MCS: Threshold - 10th percentile
+```
+
+# Here I want to test using generalised linear hypotheses if there is a difference in the intensity of upwellling events between the insitu and       satellite data.
+# Do the in situ and satelllite derived data yield the same number of events. First I add a seasonal column to test seasonal variation between 
+library(multcomp) # for generalized linear hypothesis test (glht)
+
+```{r}
+
+```
+
+Determining the R2 value between *in situ* and remotely sensed SST
+
+```{r}
+# Loading the *in situ* and Remotely sensed SST
+
+load("Data_P1/SACTN_fyears.RData")
+load("Data_P1/OISST_fyears.RData")
 
@@
   x <- subset(SACTN_fyears, site == levels(SACTN_fyears$site)[i])
-  y <- subset(OISST_fyears, site == levels(OISST_fyears$site)[i])
+  y <- subset(df, site == levels(df$site)[i])
   x <- x[x$date %in% y$date,]
   y <- y[y$date %in% x$date,]
-  z <- data.frame(site = as.character(x$site[1]), R2 = coef(lm(y$temp~x$temp))[2],na.action=na.exclude)
+  z <- data.frame(site = as.character(x$site[1]), R2 = coef(lm(y$temp~x$temp))[2])
   R2 <- rbind(R2, z)
@@
 
-## EOF's
+# Upwelling index
+## Determining upwelling index from wind data (WindGuru and SAWS)
+## Index Equation from Fielding & Davis 1989 paper
+
+```{r}
+load("Data_P1/insitu_wind.RData")
+load("Data_P1/SACTN_fyears.RData")
+load("Data_P1/MUR_fyears.RData")
+load("Data_P1/OISST_fyears.RData")
+load("Data_P1/G1SST_fyears.RData")
+load("Data_P1/K10_fyears.RData")
+
+# UPWELLING INDEX EQUATION  - FIELDING & DAVIS 1989
+# Index = Wind_speed*cos(wind_direction) - 160) [*160 refers to the angle (degrees) of the coast]
+
+#data wrangling - TEMPERATURE
+SACTN_fyears_UI <- SACTN_fyears %>% 
+  dplyr::rename(temp = insitu_temp) %>%
+  group_by(site) %>%
+  select("date", "temp")
+
+wind_UI <- insitu_wind %>%
+          group_by(site) %>%
+          mutate(ui.saws = mean_speed * (cos(dir_circ - 160)))
+
+index <- wind_UI %>%
+  group_by(site) %>% 
+  select("site", "date","ui.saws") %>% 
+  left_join(SACTN_fyears_UI, by = c("date", "site")) %>% 
+  mutate(ws.condition = ifelse(ui.saws > 0, "upwelling", "downwelling"))
+
+#Plotting the data
+#SAWS data
+ggplot(index, aes(x= date, y=ui.saws)) +
+  geom_area(aes(fill=ws.condition)) +
+  facet_wrap(~site) +
+  geom_line() +
+  geom_hline(yintercept=0)
+
+# Visualisation
+```
+
+### EOF's
 
@@
 
-Mond:
+To do:
 - Try PCA
-- map it/plot
+- plot
 - dave upwelling index
-
-Tuesday
-- Do exercises
+-

### 2019-05-02 Paper1_results.Rmd
daily_updated

@@
 library(xtable)
+library(ggpubr)
 
@@
 ```{r}
-
 OISSTDir <- "~/Documents/Masters_2019/Data_SST"
@@
 load("Data_P1/south_africa_coast.RData")
+load("Data_P1/africa_coast.RData")
+load("Data_P1/sa_provinces_new.RData")
 # load("Data_P1/bathy/sa_bathy.RData")
@@
 
+africa_map <- ggplot(africa_coast, aes(x = lon, y = lat)) + 
+  theme_bw() + 
+  coord_equal() + 
+  geom_polygon(aes(group = group), colour = "black", fill = "grey80") + 
+  geom_polygon(data = sa_provinces_new, (aes(group = group))) +
+  annotate("text", label = "Africa", x = 16.0, y = 15.0, size = 3) + 
+  theme(panel.border = element_rect(colour = "black", size = 0.4), 
+        plot.background = element_blank(), 
+        axis.ticks = element_blank(), 
+        axis.text = element_blank(), 
+        axis.title = element_blank(), 
+        panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank()) +
+  coord_map(xlim = c(-20, 53), ylim = c(-36, 38), projection = "mercator") 
+africa_map
+
+# More mapping - http://r-nold.blogspot.com/2014/06/creating-inset-map-with-ggplot2.html
+
+pdf("Figures/map_fixed.pdf", width = 8, height = 5, pointsize = 6) # Set PDF dimensions
+vp1 <- viewport(x = 0.70, y = 0.58, w = 0.25, h = 0.25, just = c("left", "bottom")) # Africa
+vp2 <- viewport(x = 1.0, y = 1.0, w = 1.00, h = 1.00, just = c("right", "top"))  # South Africa
+print(site_map, vp = vp2)
+print(africa_map, vp = vp1)
+dev.off()
+
 # png("Figures/site_map.png", width = 8, height = 5, pointsize = 6)
@@
   group_by(site) %>%
-          mutate(ui.saws = mean_speed * (cos(dir_circ - 160)))
+  mutate(ui.saws = mean_speed * (cos(dir_circ - 160))) %>% 
+  drop_na()
 
@@
   left_join(SACTN_fyears_UI, by = c("date", "site")) %>% 
-  mutate(ws.condition = ifelse(ui.saws > 0, "upwelling", "downwelling"))
+  mutate(saws.condition = ifelse(ui.saws > 0, "upwelling", "downwelling"))
 
@@
 ggplot(index, aes(x= date, y=ui.saws)) +
-  geom_area(aes(fill=ws.condition)) +
+  geom_area(aes(fill=saws.condition)) +
   facet_wrap(~site) +
@@
 
+index_SP <- index %>% 
+  filter(site == "Sea Point") %>% 
+  slice(-(c(34:70,435:632,686,1704)))
+
+index_KOM <- index %>% 
+  filter(site == "Kommetjie")
+
+index_PN <- index %>% 
+  filter(site == "Port Nolloth")
+
+index_YZ <- index %>% 
+  filter(site == "Yzerfontein")
+
+index_LB <- index %>% 
+  filter(site == "Lamberts Bay")
+
+index_SB <- index %>% 
+  filter(site == "Saldanha Bay")
+
 # Visualisation
+#Plotting the data
+#SAWS data
+
+upwell_plot <- function(df){
+  plot <- df %>% 
+  ggplot(aes(x= date, y=ui.saws)) +
+  geom_area(aes(fill=saws.condition)) +
+  geom_hline(yintercept=0)
+}
+
+upwell_SP <- upwell_plot(df = index_SP) 
+upwell_KOM <- upwell_plot(df = index_KOM) 
+upwell_PN <- upwell_plot(df = index_PN) 
+upwell_YZ <- upwell_plot(df = index_YZ) 
+upwell_SB <- upwell_plot(df = index_SB) 
+upwell_LB <- upwell_plot(df = index_LB)
+
+upwell_int <- function(df){
+  interp <- approx(df$date, df$ui.saws, n=100000)
+}
+upwell_int_SP <- upwell_int(df = index_SP) 
+upwell_int_KOM <- upwell_int(df = index_KOM) 
+upwell_int_PN <- upwell_int(df = index_PN) 
+upwell_int_YZ <- upwell_int(df = index_YZ) 
+upwell_int_LB <- upwell_int(df = index_LB) 
+upwell_int_SB<- upwell_int(df = index_SB)
+
+interp <- function(df){
+  interp <- approx(df$date, df$ui.saws, n=100000)
+}
+interp_SP <- interp(df = index_SP) 
+interp_KOM <- interp(df = index_KOM) 
+interp_PN <- interp(df = index_PN) 
+interp_YZ <- interp(df = index_YZ) 
+interp_LB <- interp(df = index_LB) 
+interp_SB<- interp(df = index_SB)
+
+interp2 <- function(df){
+  interp2 <- approx(df$date, df$ui.saws, n=100000)
+}
+
+interp2_SP <- interp2(df = index_SP) 
+interp2_KOM <- interp2(df = index_KOM) 
+interp2_PN <- interp2(df = index_PN) 
+interp2_YZ <- interp2(df = index_YZ) 
+interp2_LB <- interp2(df = index_LB) 
+interp2_SB<- interp2(df = index_SB)
+
+# Sea Point
+UIi_SP <- data.frame(Day=interp_SP$x, Index=interp_SP$y, Temp = interp2_SP$y)
+UIi_SP$saws_condition[UIi_SP$Index >= 0] <- "upwelling"
+UIi_SP$saws_condition[UIi_SP$Index < 0] <- "downwelling"
+
+index_plot_SP <- ggplot(UIi_SP, aes(x=Day, y=Index)) +
+  geom_area(aes(fill=saws_condition)) +
+  geom_hline(yintercept=0) +
+  theme_bw() +
+  # annotate("text", x = 300, y = -12.5, label = "SAWS Data") +
+  #geom_vline(xintercept = c(116, 117, 133, 134, 244, 291), lty = 2, lwd = 1) +
+  scale_fill_manual(values=c("salmon", "steelblue")) +
+  scale_x_continuous(expand=c(0, 0)) +
+  scale_y_reverse() +
+  labs(y= "Upwelling Index") +
+  guides(fill=guide_legend(title="Condition")) +
+  ggtitle("Sea Point")
+index_plot_SP
+
+# Kom
+UIi_KOM <- data.frame(Day=interp_KOM$x, Index=interp_KOM$y, Temp = interp2_KOM$y)
+UIi_KOM$saws_condition[UIi_KOM$Index >= 0] <- "upwelling"
+UIi_KOM$saws_condition[UIi_KOM$Index < 0] <- "downwelling"
+
+index_plot_KOM <- ggplot(UIi_KOM, aes(x=Day, y=Index)) +
+  geom_area(aes(fill=saws_condition)) +
+  geom_hline(yintercept=0) +
+  theme_bw() +
+  # annotate("text", x = 300, y = -12.5, label = "SAWS Data") +
+  #geom_vline(xintercept = c(116, 117, 133, 134, 244, 291), lty = 2, lwd = 1) +
+  scale_fill_manual(values=c("salmon", "steelblue")) +
+  scale_x_continuous(expand=c(0, 0)) +
+  scale_y_reverse() +
+  labs(y= "Upwelling Index") +
+  guides(fill=guide_legend(title="Condition")) +
+  ggtitle("Kommetjie")
+index_plot_KOM
+
+# Port Nolloth
+UIi_PN <- data.frame(Day=interp_PN$x, Index=interp_PN$y, Temp = interp2_PN$y)
+UIi_PN$saws_condition[UIi_PN$Index >= 0] <- "upwelling"
+UIi_PN$saws_condition[UIi_PN$Index < 0] <- "downwelling"
+
+index_plot_PN<- ggplot(UIi_PN, aes(x=Day, y=Index)) +
+  geom_area(aes(fill=saws_condition)) +
+  geom_hline(yintercept=0) +
+  theme_bw() +
+  # annotate("text", x = 300, y = -12.5, label = "SAWS Data") +
+  #geom_vline(xintercept = c(116, 117, 133, 134, 244, 291), lty = 2, lwd = 1) +
+  scale_fill_manual(values=c("salmon", "steelblue")) +
+  scale_x_continuous(expand=c(0, 0)) +
+  scale_y_reverse() +
+  labs(y= "Upwelling Index") +
+  guides(fill=guide_legend(title="Condition")) +
+  ggtitle("Port Nolloth")
+index_plot_PN
+
+# Yzerfontein
+UIi_YZ <- data.frame(Day=interp_YZ$x, Index=interp_YZ$y, Temp = interp2_YZ$y)
+UIi_YZ$saws_condition[UIi_YZ$Index >= 0] <- "upwelling"
+UIi_YZ$saws_condition[UIi_YZ$Index < 0] <- "downwelling"
+
+index_plot_YZ<- ggplot(UIi_YZ, aes(x=Day, y=Index)) +
+  geom_area(aes(fill=saws_condition)) +
+  geom_hline(yintercept=0) +
+  theme_bw() +
+  # annotate("text", x = 300, y = -12.5, label = "SAWS Data") +
+  #geom_vline(xintercept = c(116, 117, 133, 134, 244, 291), lty = 2, lwd = 1) +
+  scale_fill_manual(values=c("salmon", "steelblue")) +
+  scale_x_continuous(expand=c(0, 0)) +
+  scale_y_reverse() +
+  labs(y= "Upwelling Index") +
+  guides(fill=guide_legend(title="Condition")) +
+  ggtitle("Yzerfontein")
+index_plot_YZ
+
+# Lamberts Bay
+UIi_LB <- data.frame(Day=interp_LB$x, Index=interp_LB$y, Temp = interp2_LB$y)
+UIi_LB$saws_condition[UIi_LB$Index >= 0] <- "upwelling"
+UIi_LB$saws_condition[UIi_LB$Index < 0] <- "downwelling"
+
+index_plot_LB<- ggplot(UIi_LB, aes(x=Day, y=Index)) +
+  geom_area(aes(fill=saws_condition)) +
+  geom_hline(yintercept=0) +
+  theme_bw() +
+  # annotate("text", x = 300, y = -12.5, label = "SAWS Data") +
+  #geom_vline(xintercept = c(116, 117, 133, 134, 244, 291), lty = 2, lwd = 1) +
+  scale_fill_manual(values=c("salmon", "steelblue")) +
+  scale_x_continuous(expand=c(0, 0)) +
+  scale_y_reverse() +
+  labs(y= "Upwelling Index") +
+  guides(fill=guide_legend(title="Condition")) +
+  ggtitle("Lamberts Bay")
+index_plot_LB
+
+# Saldanha Bay
+UIi_SB <- data.frame(Day=interp_SB$x, Index=interp_SB$y, Temp = interp2_SB$y)
+UIi_SB$saws_condition[UIi_SB$Index >= 0] <- "upwelling"
+UIi_SB$saws_condition[UIi_SB$Index < 0] <- "downwelling"
+
+index_plot_SB<- ggplot(UIi_SB, aes(x=Day, y=Index)) +
+  geom_area(aes(fill=saws_condition)) +
+  geom_hline(yintercept=0) +
+  theme_bw() +
+  # annotate("text", x = 300, y = -12.5, label = "SAWS Data") +
+  #geom_vline(xintercept = c(116, 117, 133, 134, 244, 291), lty = 2, lwd = 1) +
+  scale_fill_manual(values=c("salmon", "steelblue")) +
+  scale_x_continuous(expand=c(0, 0)) +
+  scale_y_reverse() +
+  labs(y= "Upwelling Index") +
+  guides(fill=guide_legend(title="Condition")) +
+  ggtitle("Saldanha Bay")
+index_plot_SB
+
+upwelling_index <- ggarrange(index_plot_SP, index_plot_KOM, index_plot_PN,
+          index_plot_YZ,index_plot_LB, index_plot_SB, ncol = 2, nrow = 3)
 ```

### 2019-05-03 Paper1_results.Rmd
Updated

@@
         legend.background = element_blank(),
+        axis.ticks = element_line(size = 0.5)) +
+    guides(fill = guide_colourbar(barheight = 1.00, barwidth = 10)) +
+  theme(panel.background = element_rect(fill = "ivory", colour = NA),
+        panel.border = element_rect(colour = "black", size = 0.5),
+        panel.grid.minor = element_line(colour = "NA"),
+        panel.grid.major = element_line(colour = "ivory", size = 0.2, linetype = "dotted"),
+        legend.direction = "horizontal",
+        legend.justification = c(1,0),
+        legend.position = c(0.65, 0.80),
+        legend.text = element_text(size = 8),
+        legend.title = element_text(size = 8),
+        legend.key = element_blank(),
+        legend.background = element_blank(),
         axis.ticks = element_line(size = 0.5))
@@
 
-pdf("Figures/map_fixed.pdf", width = 8, height = 5, pointsize = 6) # Set PDF dimensions
-vp1 <- viewport(x = 0.70, y = 0.58, w = 0.25, h = 0.25, just = c("left", "bottom")) # Africa
+pdf("Figures/map_fixed.pdf", width = 10, height = 5, pointsize = 4) # Set PDF dimensions
+vp1 <- viewport(x = 0.70, y = 0.58, w = 0.25, h = 0.25) # Africa
 vp2 <- viewport(x = 1.0, y = 1.0, w = 1.00, h = 1.00, just = c("right", "top"))  # South Africa

### 2019-05-03 Write_up/Writing/dissertation.Rmd
working_bits
updating
daily_updated
house_keeping
Updated

@@
 
-# knitr::include_graphics("../figures/map_fixed.pdf")
+knitr::include_graphics("../Figures/map_fixed.png")
 ```
@@
 
- knitr::include_graphics("../figures/wind.png")
+ knitr::include_graphics("../Figures/wind.png")
 ```
 
-## Local and broad scale upwelling
+## Upwelling indeces
+
+In order to determine the upwelling condition amoung the various sites located along this EBUS, an upwelling index was calcukated by using the formula presented in Field and Davis (1989).
+This formula was developed for a local site; Oudekraal, located within this benguela upwelling system (Figure?
+- Shows where oudekraal is on the site map relative to the other).
+The upwelling index serves as an indication of the intensity of upwelling at the individual sites located within the benguela upwelling region and is calculated by using the follwoing equation:
+
+$$ Upwelling Index = μ{(Cos θ − 160)}$$
+
+In this equation μ represents the wind speed (m/s) and θ represents the wind direction which is measured in degrees.
+The 160 degrees is used as this refers to the angle of the coastline.
+This equation therefore is largely dependant on wind speed and direction data in order to determing the intensity of the upwelling event.
+Wind data were obtained daily from the South African Weather Service (SAWS).
+This wind data was then matched to the date at which temperature were collected.
+Wind this data the upwelling index was determined.
+By doing this we are able to examine how upwelling varied between sites located further away from the upwelling centres and sites more closeer to the upwelling centres.

### 2019-05-03 Paper1_results.Rmd
study_region

--- 2019-05-02 Paper1_results.Rmd
@@
     annotate("text", label = "Benguela", x = 16.0, y = -31.7,
-             size = 5, angle = 302, colour = "black") +
+             size = 4, angle = 302, colour = "black") +
     # annotate("text", label = "Agulhas", x = 31.7, y = -31.7,
@@
         legend.background = element_blank(),
+        axis.ticks = element_line(size = 0.5),
+        axis.text = element_text(size = 6)) +
+    guides(fill = guide_colourbar(barheight = 1.00, barwidth = 10)) +
+  theme(panel.background = element_rect(fill = "ivory", colour = NA),
+        panel.border = element_rect(colour = "black", size = 0.5),
+        panel.grid.minor = element_line(colour = "NA"),
+        panel.grid.major = element_line(colour = "ivory", size = 0.2, linetype = "dotted"),
+        legend.direction = "horizontal",
+        legend.justification = c(1,0),
+        legend.position = c(0.65, 0.80),
+        legend.text = element_text(size = 8),
+        legend.title = element_text(size = 8),
+        legend.key = element_blank(),
+        legend.background = element_blank(),
         axis.ticks = element_line(size = 0.5))
@@
 
-pdf("Figures/map_fixed.pdf", width = 8, height = 5, pointsize = 6) # Set PDF dimensions
-vp1 <- viewport(x = 0.70, y = 0.58, w = 0.25, h = 0.25, just = c("left", "bottom")) # Africa
+pdf("Figures/map_fixed.pdf", width = 10, height = 7, pointsize = 4) # Set PDF dimensions
+vp1 <- viewport(x = 0.70, y = 0.58, w = 0.25, h = 0.25) # Africa
 vp2 <- viewport(x = 1.0, y = 1.0, w = 1.00, h = 1.00, just = c("right", "top"))  # South Africa

### 2019-05-03 Paper1_results.Rmd
cleaned_script
progress

@@
   facet_wrap(~site) +
-  geom_line() +
   geom_hline(yintercept=0)
 
+# Seperating sites
 index_SP <- index %>% 
@@
 
+Calculating the temperature anomalies.
+This will be plotted on the same graph as the upwelling indeces.
+This allows for an observation on the relationship or patterns that exist as a result of temperature changes and he upwelling and downwelling conditions.
+
+```{r}
+
+```
+
 ### EOF's
@@
 In particular, trends will be observed to be different at ocean and coastal locations and we hypothesize that those differences are due to changes in upwelling intensity.
-
-To do:
-- Try PCA
-- plot
-- dave upwelling index
--

### 2019-05-09 Paper1_results.Rmd
exploring percentile bits

@@
 library(ggpubr)
-
-# library(devtools)
+library(ggpubr); theme_set(theme_pubr())# library(devtools)
 # install_github("marchtaylor/sinkr")
@@
 
-# png("Figures/site_map.png", width = 8, height = 5, pointsize = 6)
+# Site map
+# The insetting needs to be placed at appropriate location
 ```
@@
   select(-doy, -seas, -thresh)
+
+****Error: Evaluation error: sort(): detected NaN.*** 
 ```
@@
 
-# Here I want to test using generalised linear hypotheses if there is a difference in the intensity of upwellling events between the insitu and       satellite data.
-# Do the in situ and satelllite derived data yield the same number of events. First I add a seasonal column to test seasonal variation between 
-library(multcomp) # for generalized linear hypothesis test (glht)
-
-```{r}
-
-```
-
-Determining the R2 value between *in situ* and remotely sensed SST
-
-```{r}
-# Loading the *in situ* and Remotely sensed SST
-
-load("Data_P1/SACTN_fyears.RData")
-load("Data_P1/OISST_fyears.RData")
-
-SACTN_fyears <- SACTN_fyears %>% 
-  dplyr::rename(temp = insitu_temp)
-
-R2 <- data.frame()
-for(i in 1:length(levels(SACTN_fyears$site))) {
-  x <- subset(SACTN_fyears, site == levels(SACTN_fyears$site)[i])
-  y <- subset(df, site == levels(df$site)[i])
-  x <- x[x$date %in% y$date,]
-  y <- y[y$date %in% x$date,]
-  z <- data.frame(site = as.character(x$site[1]), R2 = coef(lm(y$temp~x$temp))[2])
-  R2 <- rbind(R2, z)
-}
-```
-
 # Upwelling index
-## Determining upwelling index from wind data (WindGuru and SAWS)
+## Determining upwelling index from wind data (SAWS)
 ## Index Equation from Fielding & Davis 1989 paper
+## This is only for the SACTN data
 
@@
   filter(site == "Sea Point") %>% 
-  slice(-(c(34:70,435:632,686,1704)))
+  ungroup()#%>% 
+  # slice(-(c(34:70,435:632,686,1704)))
 
@@
   theme_bw() +
-  # annotate("text", x = 300, y = -12.5, label = "SAWS Data") +
-  #geom_vline(xintercept = c(116, 117, 133, 134, 244, 291), lty = 2, lwd = 1) +
   scale_fill_manual(values=c("salmon", "steelblue")) +
@@
           index_plot_YZ,index_plot_LB, index_plot_SB, ncol = 2, nrow = 3)
-```
-
-Calculating the temperature anomalies.
-This will be plotted on the same graph as the upwelling indeces.
-This allows for an observation on the relationship or patterns that exist as a result of temperature changes and he upwelling and downwelling conditions.
-
-```{r}
-
-```
-
-### EOF's
-
-```{r} 
-PCA_1 <- rda(tester)
-summary(PCA)
-
-# Proportion of variation explained
-# Values to be found opon running the code
-
-PCA <- cbind(scores(PCA_1, choices = c(1:4), display = "sites"))
-PCA_spe <- cbind(scores(PCA_1, choices = c(1:4), display = "species"))
-
-test1 <- ggplot(PCA, aes(x = lon, y = lat)) +
-  geom_raster(aes(fill = PC1)) +
-  geom_polygon(data = PC1, aes(long, lat, group = group),
-               fill = NA, colour = "blue", size = 0.3) +
-  geom_polygon(data = velocity, aes(long, lat, group = group),
-               fill = NA, colour = "green", size = 0.3)+ 
-         fill = guide_colourbar(title = "PC1")
-
-test2 <- ggplot(PCA, aes(x = lon, y = lat)) +
-  geom_raster(aes(fill = PC2)) +
-  geom_polygon(data = PC1, aes(long, lat, group = group),
-               fill = NA, colour = "blue", size = 0.3) +
-  geom_polygon(data = velocity, aes(long, lat, group = group),
-               fill = NA, colour = "green", size = 0.3)+ 
-         fill = guide_colourbar(title = "PC2")
-
-test3 <- ggplot(PCA, aes(x = lon, y = lat)) +
-  geom_raster(aes(fill = PC2)) +
-  geom_polygon(data = PC1, aes(long, lat, group = group),
-               fill = NA, colour = "blue", size = 0.3) +
-  geom_polygon(data = velocity, aes(long, lat, group = group),
-               fill = NA, colour = "green", size = 0.3)+ 
-         fill = guide_colourbar(title = "PC3")
-
-test4 <- ggplot(PCA, aes(x = lon, y = lat)) +
-  geom_raster(aes(fill = PC1)) +
-  geom_polygon(data = PC1, aes(long, lat, group = group),
-               fill = NA, colour = "blue", size = 0.3) +
-  geom_polygon(data = velocity, aes(long, lat, group = group),
-               fill = NA, colour = "green", size = 0.3)+ 
-         fill = guide_colourbar(title = "PC4")
-
-pc1_combined <- ggarrange( test1, test2, test3, test4)
-
-```
-
-In particular, trends will be observed to be different at ocean and coastal locations and we hypothesize that those differences are due to changes in upwelling intensity.
+
+# See graph: ("Figures/upwelling_indeces.png")
+# Sea Point appears to have missing data
+# Lamberts bay appears to have missing data as well
+# Can we plot the graph where it ignores the missing data without interpolations? [AJ suggested]
+# Plot the temperature anomalies on the same graphs as the upwelling indeces for each of the sites
+```
+######################################
+
+# Question: How does the intensity and duration of an upwelling event vary at different distances from the coastline???
+  # Got the various pixels : load(Data_P1/site_pixels.RData)
+  # How do i determine where upwelling occured? - What is the threshold of an upwelling event? # How do i determine the threshold of this event
+  # Calculate event count, duration and mean intensity for each pixel
+
+Look at the percentile level that seawater temperature must decrease below when an upwelling event occurs
+  - To what exttent the temperature of upwelled water  is colder than the mean SST. 
+  - How the minumum temperature of upwelled water becomes to be that temperature
+  - Why temperature does not become any colder
+
+To do:
+- how count, duration, intensity of upwelling changed between 1990s and current times

### 2019-05-09 Write_up/Writing/dissertation.Rmd
exploring percentile bits

@@
 Global warming represents distinct warming and cooling periods throughout the 20th and 21st century (IPCC, 2007).
-Global warming however is not evenly distributed across the global ocean, and thus several regions are warming faster or slower than García-Reyes et al., 2015 the global average [@levitus; @santos].
+Global warming however is not evenly distributed across the global ocean, and thus several regions are warming faster or slower than the global average [@levitus; @santos].
 In particular, the Atlantic Ocean contributes most to the increase of heat content [@nerem; @levitus].
@@
 2002).
-In the past two decades, sea surface temperature has risen in many regions and this Seasonal rise in temperature can act as an important cue in species survival, reproduction and abundance (Cushing 1982; Cushing 1990).
+In the past two decades, sea surface temperature has risen in many regions and this seasonal rise in temperature can act as an important cue in species survival, reproduction and abundance (Cushing 1982; Cushing 1990).
 The different responses of species as a result of climate change are important to understand as these changes affect, the trophic dynamics and fisheries yield of several regional ecosystems among many things (Murawski 1993).
@@
 Northward, the Benguela Current Upwelling system is linked by the Angola Benguela Front which converge between the Benguela current and the Benguela coastal current as well as the warm saline waters of the Angola Current [@peterson].
-Research done by Yamagata and Iizuka (1995) discovered that coastal Kelvin waves, originating from equatorial Kelvin waves, carry the down welling process southward.
+Research done by Yamagata and Iizuka (1995) discovered that coastal Kelvin waves, originating from equatorial Kelvin waves, carry the downwelling process southward.
 Southward, the coastal upwelling system is bounded by the Agulhas Bank.

### 2019-05-09 upwelling_percentile.Rmd
exploring percentile bits

---
title: "upwelling_percentile"
author: "Amieroh Abrahams"
date: "08 May 2019"
output: html_document
---

# Upwelling points

* After analysing the data within the `upwelling_threshold.Rmd` script and plotting the data it is clear that upwelling occurs at a temperature less than or equal (<=) to 12.
* It is also evident that upwelling lasts for a total of 2-3 days

# Aims of this script:
  - Find a 30year time series
  - Then find the count of the number of temperatures found less than or equal to (<=) 12
  - Thereafter, divide that answer by the total number of number temperatures
  - Then we have the percentile

# Questions: How does the intensity and duration of an upwelling event vary at different distances from the coastline??? 
  # Here we can make use of the heatwaveR package to determine the count/durartion and intensity.
  # Got the various pixels : load(Data_P1/site_pixels.RData)
  # Calculate event count, duration and mean intensity once percentile has been determined

# Tips:
  - Look at the percentile level that seawater temperature must decrease below when an upwelling event occurs
  - To what extent the temperature of upwelled water  is colder than the mean SST. 
  - How the minumum temperature of upwelled water becomes to be that temperature
  - Why temperature does not become any colde

 # DETERMINING THE UPWELLING THRESHOLD
  # Plot the data
  # Find a value showing a clear drop in temperature
  # Determine from the dataset the amount of values falling bellow this defined temperature
  # Convert this to percentile
  # Determine how long (in days) upwelling events usually last 
  # Once we have the percentile we can make use of the heatwaveR package to determine the count, intensity, duration, etc.
  # Use one year of data maybe occuring in 1990s and one year of current data to observe how upwelling varied in past and present

# Loading packages

```{r prelim_opts, echo=FALSE}
knitr::opts_chunk$set(
  comment = "R>",
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(plyr)
library(lubridate)
library(ggpubr)
library(zoo)
library(lubridate)
library(FNN)
library(forecast)
library(astrochron)
library(WaveletComp)
library(data.table)
library(heatwaveR)
```

# Loading the data

```{r load_files1, include=TRUE}
load("Data_P1/site_list_v4.2.RData")
load("Data_P1/SACTN_daily_v4.2.RData")
```

Finding a time series of greater than 30years

```{r}
site_list_sub <- site_list %>% 
  filter(coast == "wc") %>% 
  filter(length > 10950)

SACTN_temps_US <- SACTN_daily_v4.2 %>% 
  left_join(site_list[,c(4,13)], by = "index") %>% 
  filter(index %in% site_list_sub$index)
# save(SACTN_temps_US, file = "Data/SACTN_temps_US.RData")

SACTN_split <- SACTN_temps_US %>% 
  separate(index, into = c("site", "src"), sep = "/", remove = FALSE) %>% 
  drop_na()

SACTN_temp <- SACTN_split %>% 
  filter(temp <= 13)

percentile <- 30840/60189*100 # Count values
# Percentile value of 50

###################################################Percentile classification#################################################
# percentile = 70
0.7*60189
# 42132
SACTN_TRY1 <- SACTN_split %>%
  arrange(temp) %>% 
  slice(1:42132) # Upwelling temp  = 14 degrees

# Percentile = 60
0.6*60189
# 36113
SACTN_TRY2 <- SACTN_split %>%
  arrange(temp) %>% 
  slice(1:36113) # Upwelling temp = 13.7 degrees

# Percentile = 50
0.5*60189
# 30094
SACTN_TRY3 <- SACTN_split %>%
  arrange(temp) %>% 
  slice(1:30094) # Upwelling temp = 13 degrees

# Percentile = 40
0.4*60189
# 24076
SACTN_TRY4 <- SACTN_split %>%
  arrange(temp) %>% 
  slice(1:24076) # Upwelling temp = 13 degrees

############################################################################################################################

SACTN_PN <- SACTN_split %>%
  filter(site == "Port Nolloth") %>%
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()

SACTN_LB <- SACTN_split %>%
  filter(site == "Lamberts Bay") %>%
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()

SACTN_SB <- SACTN_split %>%
  filter(site == "Saldanha Bay") %>%
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()

SACTN_SP <- SACTN_split %>%
  filter(site == "Sea Point") %>%
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()

SACTN_HB <- SACTN_split %>%
  filter(site == "Hout Bay") %>%
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()
```

```{r}
ts <- ts2clm(SACTN_PN, climatologyPeriod = c("1974-01-01", "2016-12-31"),
             pctile = 50)
res <- detect_event(ts)
upwelling <- res$climatology # Show a portion of the climatology
upwelling_events <- res$event[1:5,1:10] # Show some of the cold-spells
# MCS <- MCS[14408:15503,] # Get the years from 2013 - 2015

upwelling_SACTN_PN <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2010-01-01", end_date = "2015-12-31")
```

# MUR

```{r}
load("~/Documents/Masters_2019/MastersProject/Exploring/Data/MUR_SST.RData")
MUR_SST_PN <- MUR_SST %>%
  filter(site == "Port Nolloth") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(MUR_SST_PN, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 30)
res <- detect_event(ts)
upwelling <- event_line(res, spread = 500, metric = "intensity_cumulative",
           start_date = "2011-12-31", end_date = "2014-10-30") +
  ggtitle("MUR_PN") 
```
